### 2020-10-13 reports/rmarkdown/01_data_pre_processing.Rmd
adds data pre-processing and QC

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# Load datasets
dataset_names <- c("ASV_tissue_B2.csv",
                     "ASV_tissue_B1.csv",
                     "ASV_CVL_V2_B1.csv",
                     "ASV_CVL_V2_B2.csv",
                     "ASV_CVL_V2_C.csv",
                     "ASV_CVL_V3.csv")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
  not_complete <- grepl("_NA",sps)

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")

  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)

#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)

# datasets <- datasets[3] #in case you wanna test in 1 dataset

# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
```

# Merging, renaming and correcting batch effects on datasets

```{r, echo = FALSE}
#No batch correction needed in tissue_V3 dataset, only add it to the other samples
datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_B2.csv"]],datasets[["ASV_tissue_B1.csv"]] )

datasets <- datasets[-c(1,2)]
lapply(datasets,dim)

#Batch correction for the sample CVL_V2
batch1 <- cbind( datasets[["ASV_CVL_V2_B1.csv"]],datasets[["ASV_CVL_V2_B2.csv"]] )
batch1 <- t( log2( t(batch1) / colSums(batch1) * 1000 + 1 ))
batch2 <- datasets[["ASV_CVL_V2_C.csv"]]
batch2 <- t( log2( t(batch2) / colSums(batch2) * 1000 + 1 ))

common_samples <- sort(colnames(batch1) [ colnames(batch1) %in% colnames(batch2) ])
colnames(batch1) <- paste0("boston_",colnames(batch1))
colnames(batch2) <- paste0("CTMR_",colnames(batch2))

merged <- cbind(batch1,batch2)
batch <- factor(sub("_.*","",colnames(merged)))
common_samples2 <- sort(colnames(merged) [ sub(".*_","",colnames(merged)) %in% common_samples])

nonzeros <- rowSums(merged)>0
merged <- merged[nonzeros,]

PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
```

# Batch correction with different approaches

```{r, echo = FALSE}
mypar(3,3)
PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )

#COMBAT
mod0 <- model.matrix(~1,data = data.frame(colnames(merged),row.names = colnames(merged)))
corrected_combat <- ComBat(dat=merged, batch=batch, mod=mod0)

#mnnCorrect
corrected_mnnCorrect <- batchelor::mnnCorrect( batch2[nonzeros,],
                                               batch1[nonzeros,] ,
                                               cos.norm.out = F,cos.norm.in = F,
                                         restrict = list((1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ],
                                                         (1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ]))
corrected_mnnCorrect <- corrected_mnnCorrect@assays@data$corrected

#regressBatches
corrected_regressBatches <- batchelor::regressBatches( batch1[nonzeros,] ,
                                         batch2[nonzeros,],
                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
corrected_regressBatches <- corrected_regressBatches@assays@data$corrected

#rescaleBatches
corrected_rescaleBatches <- batchelor::rescaleBatches( batch1[nonzeros,] ,
                                         batch2[nonzeros,],
                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
corrected_rescaleBatches <- corrected_rescaleBatches@assays@data$corrected
```

# Batch correction with Autoencoder

```{r, echo = FALSE}
library(keras)
library(caret)

xtrain <- t(batch1[nonzeros,paste0("boston_",sort(common_samples))])
xtrain <- xtrain/max(xtrain)
ytrain <- t(batch2[nonzeros,paste0("CTMR_",sort(common_samples))])
ytrain <- ytrain/max(ytrain)
dim(xtrain)

input_size = dim(xtrain)[2]
latent_size = 10
print(input_size)

x_train = array_reshape(xtrain, dim=c(dim(xtrain)[1], input_size))
y_train = array_reshape(ytrain, dim=c(dim(ytrain)[1], input_size))

print(dim(x_train))
print(dim(y_train))

if(file.exists(paste0(PATH,"../results/encoder.h5")) & 
  file.exists(paste0(PATH,"../results/decoder.h5")) & 
  file.exists(paste0(PATH,"../results/aen.h5")) ){
  encoder <- load_model_hdf5(paste0(PATH,"../results/encoder.h5"))
  decoder <- load_model_hdf5(paste0(PATH,"../results/decoder.h5"))
  aen <- load_model_hdf5(paste0(PATH,"../results/aen.h5"))

} else {
  enc_input = layer_input(shape = input_size)
  enc_output = enc_input %>% 
    layer_dense(units=200, activation = "relu") %>% 
    layer_activation_leaky_relu() %>% 
    layer_dense(units=latent_size,activation = "sigmoid") %>% 
    layer_activation_leaky_relu()

  encoder = keras_model(enc_input, enc_output)
  summary(encoder)

  dec_input = layer_input(shape = latent_size)
  dec_output = dec_input %>% 
    layer_dense(units=200, activation = "relu") %>% 
    layer_activation_leaky_relu() %>% 
    layer_dense(units = input_size, activation = "sigmoid") %>% 
    layer_activation_leaky_relu()

  decoder = keras_model(dec_input, dec_output)
  summary(decoder)

  aen_input = layer_input(shape = input_size)
  aen_output = aen_input %>% 
    encoder() %>% 
    decoder()

  aen = keras_model(aen_input, aen_output)
  summary(aen)

  aen %>% compile(optimizer="rmsprop", loss="binary_crossentropy")

  aen %>% fit( y_train , x_train , epochs=2000, batch_size=round(dim(y_train)[1]/2) )

  save_model_hdf5(encoder, paste0(PATH,"../results/encoder.h5") )
  save_model_hdf5(decoder, paste0(PATH,"../results/decoder.h5"))
  save_model_hdf5(aen, paste0(PATH,"../results/aen.h5"))
}

encoded_imgs = encoder %>% predict(y_train)
decoded_imgs = decoder %>% predict(encoded_imgs)

# plot( x_train , decoded_imgs , cex=.3)
# abline(0,1,col="red")
# plot( x_train, y_train , cex=.3 )
# abline(0,1,col="red")
```

# Plotting the batch-corrected data for comparisson

```{r, echo = FALSE,fig.height=7*2,fig.width=7}
layout(matrix(c(1,1,1,1,     
                2,2,2,3,
                4,4,4,5,
                6,6,6,7,
                8,8,8,9,
                10,10,10,11,
                12,12,12,13
                ) , ncol= 4,byrow = T),heights = c(1))

fig_letters <- letters
par(mar=c(3,3,3,1))
pc_var <- PC_raw$sdev^2 / sum(PC_raw$sdev^2) * 100
barplot(pc_var,ylim=c(0,60),las=2,names.arg = paste0("PC",1:length(pc_var)),
        border = NA,col="grey20",ylab="% variance explained")
abline(h=0,lwd=2)
add_letter(fig_letters); fig_letters <- fig_letters[-1]
legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="uncorrected") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(batch2[nonzeros, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="ComBat") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_combat[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="mnnCorrect") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_mnnCorrect[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="regressBatches") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_regressBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="rescaleBatches") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_rescaleBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="AutoEncoder") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

```

As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits the data and learns the structure of the data.

```{r, echo = FALSE}
batch2_all <- t(batch2[nonzeros,])
batch2_all <- batch2_all/max(batch2_all)

batch2_encoded_data <- encoder %>% predict(batch2_all)
batch2_decoded_data <- decoder %>% predict(batch2_encoded_data)
batch2_decoded_data <- t(round(t(t(batch2_decoded_data*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4))
rownames(batch2_decoded_data) <- colnames(batch2_all) ; colnames(batch2_decoded_data) <- rownames(batch2_all)

mypar()
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", main="AutoEncoder",xlim=c(-20,15),ylim=c(-15,15),type="n")
b2_batch_projection <- scale(t(batch2_decoded_data), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
b1_projection <- scale(round(t(batch1[nonzeros,]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation

b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey20") }

points(b2_batch_projection[,1:2], cex=.8, pch=15, col= "navy")
points(b1_projection[,1:2], cex=.8, pch=16, col= "firebrick")
legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")

samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
datasets[["ASV_CVL_V2"]] <- cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] )
datasets <- datasets[-c(1,2,3)]
colnames(datasets[["ASV_CVL_V2"]]) <- sub(".*_","",colnames(datasets[["ASV_CVL_V2"]]))
lapply(datasets,dim)
lapply(datasets,colnames)

common_microbes <- unique(unlist(lapply(datasets,rownames)))
all_samples <- rownames(metadata)

datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)

datasets <- lapply(datasets,function(x){
  x <- x[,sort(colnames(x))]
  return(x)
})

datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)

write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )

```

```{r, echo = FALSE}
trx <- read.csv(paste0(PATH,"/TRX_rawcounts.csv"),row.names = 1)
trx <- rowsum(trx[,-c(1,2)],trx$symbol)
colnames(trx) <- sub("ML","",colnames(trx))
trx  <- trx[ rownames(trx) != "" , ]
colnames(trx) <- as.character(metadata$ID)[ match(colnames(trx) , as.character(metadata$PatID) ) ]
trx <- trx[ rowSums(trx > 5) >= 3 , ]
trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
write.csv( trx , paste0(PATH,"../results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
```

### 2020-10-13 reports/rmarkdown/02_data_quality_control.Rmd
adds data pre-processing and QC

--- 2020-10-13 reports/rmarkdown/01_data_pre_processing.Rmd
@@
 ```{r, echo = FALSE}
-PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data
+PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data
 
@@
 # Load metadata
-metadata <- read.csv(paste0(PATH,"/Clinical_visit_2_3_updatesept28.csv"))
+metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
 rownames(metadata) <- as.character(metadata$ID)
@@
 # Load datasets
-dataset_names <- c("ASV_tissue_B2.csv",
-                     "ASV_tissue_B1.csv",
-                     "ASV_CVL_V2_B1.csv",
-                     "ASV_CVL_V2_B2.csv",
-                     "ASV_CVL_V2_C.csv",
-                     "ASV_CVL_V3.csv")
+dataset_names <- c("Tissue_RNAseq_V3_normalized",
+                     "ASV_tissue_V3_normalized_batch_corrected",
+                     "ASV_CVL_V3_normalized_batch_corrected",
+                     "ASV_CVL_V2_normalized_batch_corrected")
 
@@
 datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"/",x))
-  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
-  temp[is.na(temp)] <- "NA"
-  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
-  not_complete <- grepl("_NA",sps)
+  x <- read.csv( paste0(PATH,"results/",x,".csv"),row.names = 1)
+  return(as.matrix(x)) })
+names(datasets) <- dataset_names
 
-  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
-  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
-  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
-  taxa[!not_complete] <- sps[!not_complete]
-  taxa <- sub("_"," ",taxa)
-  use <- !is.na(taxa) & (taxa != "NA")
-
-  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
-  temp <- rowsum(temp[use,],taxa[use])
-  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
-  return(temp)
-})
-names(datasets) <- dataset_names
-lapply(datasets,dim)
-lapply(datasets,colnames)
-
-#arrange all datasets to have the same rownames with all bacteria
-common_microbes <- unique(unlist(lapply(datasets,rownames)))
-datasets <- lapply(datasets, function(x) {
-  temp <- matrix(0,nrow = length(common_microbes),
-                 ncol = ncol(x),
-                 dimnames = list(common_microbes,colnames(x)))
+#fill all samples with 0s to make plotting easier
+datasets_all_samples <- lapply(datasets, function(x) {
+  temp <- matrix(0,nrow = nrow(x),
+                 ncol = nrow(metadata),
+                 dimnames = list(rownames(x),rownames(metadata)))
   temp[rownames(x),colnames(x)] <- x
@@
 })
-lapply(datasets,dim)
-
-# datasets <- datasets[3] #in case you wanna test in 1 dataset
-
-# Transform phyloseq data into normalized matricies ( log2[CP1K] )
-# datasets <- lapply(datasets,function(x){
-#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
-# })
-# lapply(datasets,dim)
-```
-
-# Merging, renaming and correcting batch effects on datasets
-
-```{r, echo = FALSE}
-#No batch correction needed in tissue_V3 dataset, only add it to the other samples
-datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_B2.csv"]],datasets[["ASV_tissue_B1.csv"]] )
-
-datasets <- datasets[-c(1,2)]
-lapply(datasets,dim)
-
-#Batch correction for the sample CVL_V2
-batch1 <- cbind( datasets[["ASV_CVL_V2_B1.csv"]],datasets[["ASV_CVL_V2_B2.csv"]] )
-batch1 <- t( log2( t(batch1) / colSums(batch1) * 1000 + 1 ))
-batch2 <- datasets[["ASV_CVL_V2_C.csv"]]
-batch2 <- t( log2( t(batch2) / colSums(batch2) * 1000 + 1 ))
-
-common_samples <- sort(colnames(batch1) [ colnames(batch1) %in% colnames(batch2) ])
-colnames(batch1) <- paste0("boston_",colnames(batch1))
-colnames(batch2) <- paste0("CTMR_",colnames(batch2))
-
-merged <- cbind(batch1,batch2)
-batch <- factor(sub("_.*","",colnames(merged)))
-common_samples2 <- sort(colnames(merged) [ sub(".*_","",colnames(merged)) %in% common_samples])
-
-nonzeros <- rowSums(merged)>0
-merged <- merged[nonzeros,]
-
-PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
-```
-
-# Batch correction with different approaches
-
-```{r, echo = FALSE}
-mypar(3,3)
-PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
-
-#COMBAT
-mod0 <- model.matrix(~1,data = data.frame(colnames(merged),row.names = colnames(merged)))
-corrected_combat <- ComBat(dat=merged, batch=batch, mod=mod0)
-
-#mnnCorrect
-corrected_mnnCorrect <- batchelor::mnnCorrect( batch2[nonzeros,],
-                                               batch1[nonzeros,] ,
-                                               cos.norm.out = F,cos.norm.in = F,
-                                         restrict = list((1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ],
-                                                         (1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ]))
-corrected_mnnCorrect <- corrected_mnnCorrect@assays@data$corrected
-
-#regressBatches
-corrected_regressBatches <- batchelor::regressBatches( batch1[nonzeros,] ,
-                                         batch2[nonzeros,],
-                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
-                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
-corrected_regressBatches <- corrected_regressBatches@assays@data$corrected
-
-#rescaleBatches
-corrected_rescaleBatches <- batchelor::rescaleBatches( batch1[nonzeros,] ,
-                                         batch2[nonzeros,],
-                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
-                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
-corrected_rescaleBatches <- corrected_rescaleBatches@assays@data$corrected
-```
-
-# Batch correction with Autoencoder
-
-```{r, echo = FALSE}
-library(keras)
-library(caret)
-
-xtrain <- t(batch1[nonzeros,paste0("boston_",sort(common_samples))])
-xtrain <- xtrain/max(xtrain)
-ytrain <- t(batch2[nonzeros,paste0("CTMR_",sort(common_samples))])
-ytrain <- ytrain/max(ytrain)
-dim(xtrain)
-
-input_size = dim(xtrain)[2]
-latent_size = 10
-print(input_size)
-
-x_train = array_reshape(xtrain, dim=c(dim(xtrain)[1], input_size))
-y_train = array_reshape(ytrain, dim=c(dim(ytrain)[1], input_size))
-
-print(dim(x_train))
-print(dim(y_train))
-
-if(file.exists(paste0(PATH,"../results/encoder.h5")) & 
-  file.exists(paste0(PATH,"../results/decoder.h5")) & 
-  file.exists(paste0(PATH,"../results/aen.h5")) ){
-  encoder <- load_model_hdf5(paste0(PATH,"../results/encoder.h5"))
-  decoder <- load_model_hdf5(paste0(PATH,"../results/decoder.h5"))
-  aen <- load_model_hdf5(paste0(PATH,"../results/aen.h5"))
-
-} else {
-  enc_input = layer_input(shape = input_size)
-  enc_output = enc_input %>% 
-    layer_dense(units=200, activation = "relu") %>% 
-    layer_activation_leaky_relu() %>% 
-    layer_dense(units=latent_size,activation = "sigmoid") %>% 
-    layer_activation_leaky_relu()
-
-  encoder = keras_model(enc_input, enc_output)
-  summary(encoder)
-
-  dec_input = layer_input(shape = latent_size)
-  dec_output = dec_input %>% 
-    layer_dense(units=200, activation = "relu") %>% 
-    layer_activation_leaky_relu() %>% 
-    layer_dense(units = input_size, activation = "sigmoid") %>% 
-    layer_activation_leaky_relu()
-
-  decoder = keras_model(dec_input, dec_output)
-  summary(decoder)
-
-  aen_input = layer_input(shape = input_size)
-  aen_output = aen_input %>% 
-    encoder() %>% 
-    decoder()
-
-  aen = keras_model(aen_input, aen_output)
-  summary(aen)
-
-  aen %>% compile(optimizer="rmsprop", loss="binary_crossentropy")
-
-  aen %>% fit( y_train , x_train , epochs=2000, batch_size=round(dim(y_train)[1]/2) )
-
-  save_model_hdf5(encoder, paste0(PATH,"../results/encoder.h5") )
-  save_model_hdf5(decoder, paste0(PATH,"../results/decoder.h5"))
-  save_model_hdf5(aen, paste0(PATH,"../results/aen.h5"))
-}
-
-encoded_imgs = encoder %>% predict(y_train)
-decoded_imgs = decoder %>% predict(encoded_imgs)
-
-# plot( x_train , decoded_imgs , cex=.3)
-# abline(0,1,col="red")
-# plot( x_train, y_train , cex=.3 )
-# abline(0,1,col="red")
-```
-
-# Plotting the batch-corrected data for comparisson
-
-```{r, echo = FALSE,fig.height=7*2,fig.width=7}
-layout(matrix(c(1,1,1,1,     
-                2,2,2,3,
-                4,4,4,5,
-                6,6,6,7,
-                8,8,8,9,
-                10,10,10,11,
-                12,12,12,13
-                ) , ncol= 4,byrow = T),heights = c(1))
-
-fig_letters <- letters
-par(mar=c(3,3,3,1))
-pc_var <- PC_raw$sdev^2 / sum(PC_raw$sdev^2) * 100
-barplot(pc_var,ylim=c(0,60),las=2,names.arg = paste0("PC",1:length(pc_var)),
-        border = NA,col="grey20",ylab="% variance explained")
-abline(h=0,lwd=2)
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="uncorrected") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(batch2[nonzeros, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="ComBat") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_combat[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="mnnCorrect") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_mnnCorrect[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="regressBatches") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_regressBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="rescaleBatches") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_rescaleBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="AutoEncoder") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
+lapply(datasets_all_samples,dim)
 
@@
 
-As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits the data and learns the structure of the data.
+# Calculate QC metrics
 
 ```{r, echo = FALSE}
-batch2_all <- t(batch2[nonzeros,])
-batch2_all <- batch2_all/max(batch2_all)
+counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
+counts[is.na(counts)] <- 0
+detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
+detected_counts[is.na(detected_counts)] <- 0
+rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
 
-batch2_encoded_data <- encoder %>% predict(batch2_all)
-batch2_decoded_data <- decoder %>% predict(batch2_encoded_data)
-batch2_decoded_data <- t(round(t(t(batch2_decoded_data*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4))
-rownames(batch2_decoded_data) <- colnames(batch2_all) ; colnames(batch2_decoded_data) <- rownames(batch2_all)
+mypar(mar=c(2,20,2,1))
+barlist( data = rbind( counts , detected_counts),
+         genes = c(rownames(counts),rownames(detected_counts)),
+         clustering = rep(1,nrow(metadata)),
+         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)
 
-mypar()
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", main="AutoEncoder",xlim=c(-20,15),ylim=c(-15,15),type="n")
-b2_batch_projection <- scale(t(batch2_decoded_data), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-b1_projection <- scale(round(t(batch1[nonzeros,]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-
-b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey20") }
-
-points(b2_batch_projection[,1:2], cex=.8, pch=15, col= "navy")
-points(b1_projection[,1:2], cex=.8, pch=16, col= "firebrick")
-legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")
-
-samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
-datasets[["ASV_CVL_V2"]] <- cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] )
-datasets <- datasets[-c(1,2,3)]
-colnames(datasets[["ASV_CVL_V2"]]) <- sub(".*_","",colnames(datasets[["ASV_CVL_V2"]]))
-lapply(datasets,dim)
-lapply(datasets,colnames)
-
-common_microbes <- unique(unlist(lapply(datasets,rownames)))
-all_samples <- rownames(metadata)
-
-datasets <- lapply(datasets, function(x) {
-  temp <- matrix(0,nrow = length(common_microbes),
-                 ncol = ncol(x),
-                 dimnames = list(common_microbes,colnames(x)))
-  temp[rownames(x),colnames(x)] <- x
-  return(temp)
-})
-lapply(datasets,dim)
-
-datasets <- lapply(datasets,function(x){
-  x <- x[,sort(colnames(x))]
-  return(x)
-})
-
-datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
-datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
-datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)
-
-write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
-
+sum((colSums(counts == 0)>=3)*1)
 ```
 
+# Calculate QC metrics
+
 ```{r, echo = FALSE}
-trx <- read.csv(paste0(PATH,"/TRX_rawcounts.csv"),row.names = 1)
-trx <- rowsum(trx[,-c(1,2)],trx$symbol)
-colnames(trx) <- sub("ML","",colnames(trx))
-trx  <- trx[ rownames(trx) != "" , ]
-colnames(trx) <- as.character(metadata$ID)[ match(colnames(trx) , as.character(metadata$PatID) ) ]
-trx <- trx[ rowSums(trx > 5) >= 3 , ]
-trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
-write.csv( trx , paste0(PATH,"../results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
+counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
+counts[is.na(counts)] <- 0
+detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
+detected_counts[is.na(detected_counts)] <- 0
+rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
+
+mypar(mar=c(2,20,2,1))
+barlist( data = rbind( counts , detected_counts),
+         genes = c(rownames(counts),rownames(detected_counts)),
+         clustering = rep(1,nrow(metadata)),
+         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)
+
+sum((colSums(counts == 0)>=3)*1)
 ```

### 2020-10-13 reports/rmarkdown/report.Rmd
adds data pre-processing and QC

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data

#color palettes
pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)

#Graph construction  
k <- 10 # k for KNN
P <- 0.4 # prunning threshold of SNN

#Graph construction  
fct <- .25 # FC threshold for differential expression
pvt <- 0.05 # Pvalue threshold for differential expression
min_pct <- .1 # minimun level of detected bacteria in each sample group
```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# a <- read.csv(paste0(PATH,"/ASV_tissue_B2.csv"))
# b <- read.csv(paste0(PATH,"/ASV_tissue_B1.csv"))
# rownames(a) <- paste0()
# temp <- as.matrix(a[,c((ncol(a)-6):ncol(a))])
# temp[is.na(temp)] <- "NA"
# taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
# taxa <- sub("_NA.*","_other",taxa)

# Load datasets
dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_batch_corrected.csv")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"/results/",x),row.names = 1)
  return(x)
})
names(datasets) <- dataset_names
lapply(datasets,dim)

#arrange all datasets to have the same rownames with all bacteria
# common_microbes <- unique(unlist(lapply(datasets,rownames)))
# datasets <- lapply(datasets, function(x) {
#   temp <- matrix(0,nrow = length(common_microbes),
#                  ncol = ncol(x),
#                  dimnames = list(common_microbes,colnames(x)))
#   temp[rownames(x),colnames(x)] <- x
#   return(temp)
# })
# lapply(datasets,dim)

# datasets <- datasets[2] #in case you wanna test in 1 dataset

# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
```

# Merging microbiome datasets

```{r, echo = FALSE}
# Define the common microbes
common_microbes <- sort(unique(unlist(lapply(datasets,rownames))))
sample_names <- sort(unique(unlist(lapply(datasets,colnames))))

# Create empty matrix for the full dataset
microbiome <- as.matrix(datasets[[3]])
microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]

dataset_origin <- factor(rep(sub("_norma.*","",names(datasets)[2]),ncol(datasets[[2]])))
```

# Organise the datasets

```{r, echo = FALSE}
zscore <- t(apply(microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 6 ] <- 6
zscore[ zscore < -6 ] <- -6
rownames(zscore) <- rownames(microbiome); colnames(zscore) <- colnames(microbiome)

eucl <- as.matrix(dist(t(zscore),method = "euclidean"))
h_bacs <- hclust(dist(zscore),method = "ward.D2")
h_samples <- hclust(dist(t(zscore)),method = "ward.D2")

microbiome <- microbiome[h_bacs$order,h_samples$order]
zscore <- zscore[h_bacs$order,h_samples$order]
```

# Organise the datasets

```{r, echo = FALSE,fig.width  = 10, fig.height=6}
mypar(1,2,mar=c(2,2,2,5))
image(t(microbiome[nrow(microbiome):1,]),
      col=heat_pal,
      axes=F,main="log2(CP1K+1)",
      xlab="samples",ylab="bacteria",
      line=.4,cex.main=1,font.main=1)
text(  1+.02 ,seq(1,0,length.out = nrow(zscore)) , labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)

image(t(zscore[nrow(zscore):1,]),
      col=heat_pal,
      axes=F,main="Z-score log2(CP1K+1)",
      xlab="samples",ylab="bacteria",
      line=.4,cex.main=1,font.main=1)
text(  1+.02 ,seq(1,0,length.out = nrow(zscore)) , labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
```

\clearpage

# Computing a SNN graph from sample correlations

```{r, echo = FALSE}
mypar(2,2)

#Compute sample-wise correlations
cors <- cor(zscore,method = "pearson")
eucl <- as.matrix(dist(t(zscore)))
mean_local <- quantile(eucl,probs = .9)
hist(eucl,breaks = seq(min(eucl),max(eucl),length.out = 50),
     border=NA,col="grey60",las=1,
     xaxs="i",yaxs="i",
     main="euclidean distance",xlim=c(min(eucl),max(eucl)))
abline(v=median(eucl),col="red",lty=1,xpd=F)
abline(v=mean_local,col="red",lty=2,xpd=F)

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Compute SNN graph from KNN
jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))

#Prune SNN graph based on the Jaccard index
pSNN <- jSNN
pSNN <- pSNN / max(pSNN)

hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
     border=NA,col="grey60",las=1,
     xaxs="i",yaxs="i",
     main="SNN jaccard",xlim=c(0,1))
abline(v=P,col="red",lty=2,xpd=F)

hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
     border=NA,col="grey60",las=1,
     xaxs="i",yaxs="i",
     main="correlation coeficient",xlim=c(-1,1))
abline(v=0,col="red",lty=1,xpd=F)

pSNN[ pSNN < P ] <- 0
pSNN[ pSNN > P ] <- cors[pSNN > P]
pSNN[ pSNN < 0 ] <- 0

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)

```

\clearpage

# Visualise the data

```{r, echo = FALSE,fig.width  = 10, fig.height=3}
mypar(2,2,mar=c(0,0,2,10))

set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[dataset_origin] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(dataset_origin),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("dataset",cex.main=1,font.main=1)

# plot( g , vertex.label.cex=0.000001 , 
#       vertex.color = hue_pal()(20)[factor(metadata[common_samples,"group_M9_CVLv3"])] ,
#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
#       vertex.size=10,
#       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
# legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
#        y = par("usr")[c(4)],
#        legend = levels(factor(metadata[common_samples,"group_M9_CVLv3"])),
#        bty = "n",pch = 21,pt.bg = hue_pal()(20),pt.cex = 1,xpd=T)
# title("group_M9_CVLv3",cex.main=1,font.main=1)

```

\clearpage

# Computing differential expression across clusters

```{r, echo = FALSE}

NN <- min(table(cl$membership))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in unique(cl$membership)){
  for(j in rownames(microbiome) ){

    set.seed(1)
    a <- c(sample(microbiome[j,cl$membership == i],NN))
    set.seed(1)
    b <- sample(microbiome[j,cl$membership != i],NN)

    perc1 <- sum(microbiome[j,cl$membership == i]>0) / sum(cl$membership == i)
    perc2 <- sum(microbiome[j,cl$membership != i]>0) / sum(cl$membership != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)
```

#Plotting the most significant bacteria across clusters

```{r, echo = FALSE, fig.width  = 10, fig.height=14}
library(dplyr)
res %>% group_by(cluster)  %>% top_n(-20, pvalue) %>% top_n(10, fc) -> top7
top7 <- top7[top7$fc > 0 , ]

groupings <- factor(cl$membership)
ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(microbiome, x, groupings)}))

mypar(2,2,mar=c(2,15,1,1))

plot_dots(microbiome, names(sort(ord)) ,
          clustering = cl$membership,
          show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)

barlist( microbiome , names(sort(ord)), clustering = cl$membership, col=pal,cex.axis = .8)

violist( microbiome , names(sort(ord)), clustering = cl$membership,cex.axis = .8,
         col=pal,smooth = 2,transparency = 60,pt.col = "black")

o <- order(cl$membership)
image(t(zscore[rev(names(sort(ord))),o]),col=heat_pal,axes=F,border=NA,
      main="",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)])) , seq(1,0,length.out = length(names(sort(ord)))),
      labels = names(sort(ord)), srt = 0, adj = c(1,.5), xpd = TRUE, cex=.8)
points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
        col=pal[factor(cl$membership)][o] ,xpd=T)

```

#Plotting the most significant bacteria across PREVIOUS ANNOTATION

```{r, echo = FALSE, fig.width  = 10, fig.height=11}
mypar(2,2,mar=c(2,15,1,1))

common_samples <- colnames(microbiome)[colnames(microbiome) %in% metadata$ID]

plot_dots(microbiome[,common_samples], names(sort(ord)) ,cex.row = .8,cex.col = .8,
          clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),
          show_grid = T,cex.main=1,font.main=1,
          srt = 90)

barlist( microbiome[,common_samples] , names(sort(ord)),cex.axis = .8,
         clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),
         col=hue_pal()(20),srt = 90)

violist( microbiome[,common_samples] , names(sort(ord)), 
         clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),cex.axis = .8,
         col=hue_pal()(20),smooth = 2,transparency = 60,pt.col = "black",srt = 45)

o <- order(factor(metadata[common_samples,"group_M9_CVLv3"]))
image(t(zscore[rev(names(sort(ord))),o]),col=heat_pal,axes=F,border=NA,
      main="",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)])) , seq(1,0,length.out = length(names(sort(ord)))),
      labels = names(sort(ord)), srt = 0, adj = c(1,.5), xpd = TRUE, cex=.8)
points( (0:(length(factor(metadata[common_samples,"group_M9_CVLv3"]))-1) )/(length(factor(metadata[common_samples,"group_M9_CVLv3"]))-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(factor(metadata[common_samples,"group_M9_CVLv3"]))), pch=16,
        col=hue_pal()(20)[factor(metadata[common_samples,"group_M9_CVLv3"])][o] ,xpd=T)

```

\clearpage

# Plotting the most significant bacteria across clusters

```{r, echo = FALSE, fig.width  = 10, fig.height=11}
mypar(6,6,mar=c(.5,.5,2,.5))

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",legend = levels(factor(cl$membership)),bty = "n",pch = 21,pt.bg = pal,pt.cex = .8)

# plot( l , bg=pal[factor(cl$membership)],
#       pch=21,main="clusters",cex.main=1,font.main=1,
#       frame=F,axes=F,cex=1,xlab="",ylab="")
# empty_plot()
# legend("topleft",legend = levels(factor(cl$membership)),bty = "n",pch = 21,pt.bg = pal)

res2 <- res[order(res$cluster),]
for( j in 1:nrow(res2) ){
  i <- res2$bacteria[j]
  temp <- microbiome[i,]
  temp <- temp / max(temp)
  o <- order(temp)
  plot( l[o,] , col=c("grey85", colorRampPalette(c("grey80","grey60","navy"))(50)) [temp[o]*49+1],
        pch=16,main=paste0(i,"\ncluster=",res2$cluster[j]," (",ifelse(res2$fc[j]>0,"up)","down)")),
        cex.main=1,font.main=1,frame=F,axes=F,cex=1,xlab="",ylab="")
}
```

\clearpage

#Plotting bacteria across clusters

```{r, echo = FALSE, fig.width  = 10, fig.height=6}
mypar(2,2,mar=c(12,6,2,2))
o <- order(cl$membership)
image(zscore[rev(names(sort(ord))),o][,ncol(zscore):1],col=heat_pal,axes=F,border=NA,
      main="bacteria",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
text( seq(1,0,length.out = length(names(sort(ord)))) ,
      par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]) , 
      labels = names(sort(ord)), srt = 60, adj = c(1,.5), xpd = TRUE, cex=.7)
points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl$membership)), pch=16,
    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
    col=pal[factor(cl$membership)][o] ,xpd=T)

par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)])

image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
      main="sample correlation matrix",xlab="samples",ylab="samples",line=.4,cex.main=1,font.main=1)
points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
        col=pal[factor(cl$membership)][o] ,xpd=T)
points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl$membership)), pch=16,
    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
    col=pal[factor(cl$membership)][o] ,xpd=T)
```

\clearpage

#Plotting bacteria across clusters

```{r, echo = FALSE, fig.width  = 10, fig.height=6}
mypar(2,2)
barlist( rbind( BacsDetected = colSums(microbiome > 0) ,
                TotalCounts = colSums(microbiome),
                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
         , c("BacsDetected","TotalCounts","CountsPerBac"),
         clustering = metadata[colnames(microbiome),"group_M4_CVLv3"],
         col=pal)

violist( rbind( BacsDetected = colSums(microbiome > 0) ,
                TotalCounts = colSums(microbiome),
                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
         , c("BacsDetected","TotalCounts","CountsPerBac"),
         clustering = metadata[colnames(microbiome),"group_M4_CVLv3"],
         col=pal,transparency = 50,pt.col = "black",smooth = 2)

barlist( rbind( BacsDetected = colSums(microbiome > 0) ,
                TotalCounts = colSums(microbiome),
                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
         , c("BacsDetected","TotalCounts","CountsPerBac"),
         clustering = cl$membership,
         col=pal)

violist( rbind( BacsDetected = colSums(microbiome > 0) ,
                TotalCounts = colSums(microbiome),
                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
         , c("BacsDetected","TotalCounts","CountsPerBac"),
         clustering = cl$membership,
         col=pal,transparency = 50,pt.col = "black",smooth = 2)

table(list(cl$membership,metadata[colnames(microbiome),"group_M4_CVLv3"]))
```

\clearpage

#Plotting bacteria across clusters

```{r, eval=F,echo = FALSE}
mypar(2,2)

zscore_bac <- apply(microbiome,2,function(i){scale(i,T,T)})
zscore_bac[ zscore_bac > 6 ] <- 6
zscore_bac[ zscore_bac < -6 ] <- -6
rownames(zscore_bac) <- rownames(microbiome); colnames(zscore_bac) <- colnames(microbiome)

#Compute sample-wise correlations
cors <- cor(t(microbiome),method = "pearson")

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Compute SNN graph from KNN
jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))

#Prune SNN graph based on the Jaccard index
pSNN <- jSNN
pSNN <- pSNN / max(pSNN)

hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
     border=NA,col="grey60",las=1,
     xaxs="i",yaxs="i",
     main="SNN jaccard",xlim=c(0,1))
abline(v=P,col="red",lty=2,xpd=F)

hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
     border=NA,col="grey60",las=1,
     xaxs="i",yaxs="i",
     main="correlation coeficient",xlim=c(-1,1))
abline(v=0,col="red",lty=1,xpd=F)

pSNN[ pSNN < .3 ] <- 0
pSNN[ pSNN > .3 ] <- cors[pSNN > .3]
pSNN[ pSNN < 0 ] <- 0

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)

mypar(2,2)

set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl_bac <- igraph::cluster_louvain(g)

plot( g , vertex.label.cex=0.000001  , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.color= hue_pal()(length(unique(cl_bac$membership)))[factor(cl_bac$membership)],
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )

legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl_bac$membership)),pt.bg=hue_pal()(length(unique(cl_bac$membership))),
       bty = "n",pch = 21,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)

o_bacs <- order(cl_bac$membership)
image( t(zscore[o_bacs,o][nrow(zscore):1,]),col=heat_pal,axes=F,border=NA,
      main="bacteria",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)

text(  par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]) , seq(1,0,length.out = nrow(zscore)),
      labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.7)
points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl_bac$membership)),
        (0:(length(cl_bac$membership)-1) )/(length(cl_bac$membership)-1), pch=16,
    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
    col=pal[factor(cl$membership)][o_bacs] ,xpd=T)

par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)])

points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
        col=pal[factor(cl$membership)][o] ,xpd=T)
points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl_bac$membership)), pch=16,
    rev((0:(length(cl_bac$membership)-1) )/(length(cl_bac$membership)-1)),
    col=pal[factor(cl_bac$membership)][o_bacs] ,xpd=T)

```

#Dataset integration

```{r, eval=F, echo = FALSE, fig.width  = 10, fig.height=6}
# Load the transcriptome datset
trx <- read.csv(paste0(PATH,"~/Downloads/TRX_rawcounts.csv"),row.names = 1)
trx <- rowsum(trx[,-c(1,2)],trx$symbol)
colnames(trx) <- sub("ML","",colnames(trx))
trx  <- trx[ rownames(trx) != "" , ]
trx <- trx[ rowSums(trx > 5) >= 3 , ]
trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )

# Identify highly variable genes in the transcriptomic dataset
vars <- log2(apply(trx,1,var)+2)
means <- apply(trx,1,mean)
mypar(1,2)
top_variable <- names( sort(vars,decreasing = T) ) [1:3000]
plot( means , vars , pch=16, las=1,xlab="logMeanExpression",ylab="variance",
      cex=ifelse( rownames(trx) %in% top_variable , .2,.1),
      col= ifelse( rownames(trx) %in% top_variable , "red","black"))

# Identify common samples between the transcriptome and the mirobiome dataset
common_samples <- colnames(microbiome)[colnames(microbiome) %in% colnames(trx)]
trx <- trx[,as.character(common_samples) ]
microbiome2 <- microbiome[,as.character(common_samples)]

# Combine datased and set all to same scale using Z-score transformation
mypar(1,3)
merged <- rbind(trx[top_variable,],microbiome2)
zscore_merged <- t(apply(merged,1,function(i){scale(i,T,T)}))
zscore_merged[ zscore_merged > 5 ] <- 5
zscore_merged[ zscore_merged < -5 ] <- -5
rownames(zscore_merged) <- rownames(zscore_merged)
colnames(zscore_merged) <- colnames(zscore_merged)
hist(zscore_merged,breaks = 100,main="all")
hist(zscore_merged[top_variable,],breaks = 100,main="trx")
hist(zscore_merged[rownames(microbiome2),],breaks = 100,main="microbiome")

PC <- prcomp(zscore_merged,center = F,scale. = F)
plot(PC$sdev,pch=21,bg="grey70")
abline(v=50.5,col="red",xpd=F,lty=2)

knn <- RANN::nn2( query = PC$x[rownames(microbiome2),1:50] , 
                data = PC$x[top_variable,1:50] ,k = 50)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, length(rownames(microbiome2)), length(top_variable))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
nn <- nn[ rowSums(nn)>0 , colSums(nn)>0 ]

mypar(1,2,mar=c(0,0,2,10))

g <- graph_from_adjacency_matrix(nn, mode = "undirected",diag = F,weighted = T)
set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
l <- layout_with_fr(g,niter=500,start.temp=100)

cl <- igraph::cluster_louvain(g)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)

```

```{r, eval=F,echo = FALSE, fig.width  = 10, fig.height=11}

```

### 2020-10-13 reports/rmarkdown/report_all_dataset.Rmd
adds data pre-processing and QC

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(phyloseq)
library(circlize)
library(igraph)
library(rafalib)
library(umap)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Downloads/" #Path to the data

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"Clinical_visit_2_3.csv"))
rownames(metadata) <- as.character(metadata$PatID)

# Load datasets
datasets <- lapply(c("phyloseq_Tv3.RDS",
                     "phyloseq_CVLv2.RDS",
                     "phyloseq_CVLv3.RDS"),
                   function(x){ readRDS( paste0(PATH,x)) })
names(datasets) <- c("Tv3","CVLv2","CVLv3")
datasets <- datasets[-1] #in case you wanna test in 1 dataset

# Transform phyloseq data into normalized matricies ( log2[CP1K] )
datasets <- lapply(datasets,function(x){
    x <- as.data.frame(otu_table(x))
    colnames(x) <- as.character( sub( ".*u","",sub("[.Tv_].*","",colnames(x)) )  )
    x <- t(log2( t(x) / colSums(x) * 1000 + 1))
    x <- x[rowSums(x)>0,]
})

zscores <- lapply(datasets,function(x){
  zs <- t(apply(x,1,function(z) scale(z,T,T)))
  colnames(zs) <- colnames(x) ; rownames(zs) <- rownames(x)
  zs[zs > 6] <- 6
  zs[zs < -6] <- -6
  return(zs)
})
```

# Merging microbiome datasets

```{r, echo = FALSE}
mypar()
circos.heatmap.initialize(t(zscores[[1]]))

#cluster per dataset1
eucl <- as.matrix(dist(t(zscores[[1]]),method = "euclidean"))
h_bacs <- hclust(dist(zscores[[1]]),method = "ward.D2")
h_samples <- hclust(dist(t(zscores[[1]])),method = "ward.D2")

#plot dataset1
circos.heatmap(t(zscores[[1]][h_bacs$order,h_samples$order]), col = colorRamp2(seq(-2,6,length.out = 7),c("#000000","grey5","grey30","orange3","yellow","yellow","white")))

#plot dataset2
circos.heatmap(t(zscores[[2]][h_bacs$order,h_samples$order]), col = colorRamp2(seq(-2,6,length.out = 7),c("#000000","grey5","grey30","orange3","yellow","yellow","white")),
               rownames.side = "outside")
circos.clear()

```

#

### 2020-10-15 reports/rmarkdown/01_data_pre_processing.Rmd
updates to export uncorrected data CVL2

@@
 
-As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits the data and learns the structure of the data.
+As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits and learns the structure of the data.
 
-```{r, echo = FALSE}
+```{r, echo = FALSE,fig.height=5,fig.width=5}
+datasets[["ASV_CVL_V2_raw"]] <- cbind( datasets[["ASV_CVL_V2_B2.csv"]] , datasets[["ASV_CVL_V2_B1.csv"]], datasets[["ASV_CVL_V2_C.csv"]][,!
+colnames(datasets[["ASV_CVL_V2_C.csv"]]) %in% common_samples] )
+
 batch2_all <- t(batch2[nonzeros,])
@@
 datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
+datasets[["ASV_CVL_V2_raw"]] <- round(t( log2( t(datasets[["ASV_CVL_V2_raw"]]) / colSums(datasets[["ASV_CVL_V2_raw"]]) * 1000 + 1 )),4)
 datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)
@@
 write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
 
+batches_CVL2 <- ifelse( colnames(datasets[["ASV_CVL_V2_raw"]]) %in% sub(".*_","",colnames(batch1)),"boston","CTMR" )
+batches_CVL2 <- t(t(setNames(batches_CVL2 , colnames(datasets[["ASV_CVL_V2_raw"]]))))
+write.csv( batches_CVL2 , paste0(PATH,"../results/batches_CVL2.csv"),row.names = T )
 ```

### 2020-10-15 reports/rmarkdown/02_data_quality_control.Rmd
updates to export uncorrected data CVL2

@@
 cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
+
+#Graph construction  
+fct <- .5 # FC threshold for differential expression
+pvt <- 0.01 # Pvalue threshold for differential expression
+min_pct <- .1 # minimun level of detected bacteria in each sample group
 ```
@@
                      "ASV_CVL_V3_normalized_batch_corrected",
-                     "ASV_CVL_V2_normalized_batch_corrected")
+                     "ASV_CVL_V2_normalized_batch_corrected",
+                     "ASV_CVL_V2_normalized_NOT_batch_corrected")
 
@@
 sum((colSums(counts == 0)>=3)*1)
+colnames(counts)[(colSums(counts == 0)>=3)]
+
+metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
 ```
 
-# Calculate QC metrics
+# Comparing CVL2 samples before and after batch correction
 
 ```{r, echo = FALSE}
-counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
-counts[is.na(counts)] <- 0
-detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
-detected_counts[is.na(detected_counts)] <- 0
-rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
+mypar(1,2,mar=c(2,15,1,1))
 
-mypar(mar=c(2,20,2,1))
-barlist( data = rbind( counts , detected_counts),
-         genes = c(rownames(counts),rownames(detected_counts)),
-         clustering = rep(1,nrow(metadata)),
-         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)
+for( dataset in c("ASV_CVL_V2_normalized_NOT_batch_corrected",
+            "ASV_CVL_V2_normalized_batch_corrected")){
 
-sum((colSums(counts == 0)>=3)*1)
+  # Computing differential expression across batches
+  all_microbiome <- cbind(datasets_all_samples[[dataset]])
+  datasets <- read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,2]
+
+  all_microbiome <- all_microbiome[,as.character(read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,1])]
+
+  NN <- min(table(datasets))
+
+  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+  for(i in levels(datasets)){
+    for(j in rownames(all_microbiome) ){
+
+      set.seed(1)
+      a <- c(sample(all_microbiome[j,datasets == i],NN))
+      set.seed(1)
+      b <- sample(all_microbiome[j,datasets != i],NN)
+
+      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+      temp <- wilcox.test(x=a, y=b,exact = F)
+
+      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                            c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+    }
+  }
+
+  res <- res[-1,]
+  res$pvalue <- as.numeric(res$pvalue)
+  res$perc.1 <- as.numeric(res$perc.1)
+  res$perc.2 <- as.numeric(res$perc.2)
+  res$perc.diff <- res$perc.1 - res$perc.2
+
+  res$fc <- as.numeric(res$fc)
+  res$FDR <- p.adjust(res$pvalue)
+  res <- res[order(res$pvalue),]
+  res <- res[res$fc > fct,]
+  res <- res[abs(res$pvalue) < 0.05,]
+  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+  dim(res)
+
+  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets, main = dataset,show_grid = T,cex.main=1,font.main=1,
+            cex.row = .8,cex.col = .8,srt = 0)
+}
 ```
+
+```{r, echo = FALSE}
+# Computing differential expression across datasets
+
+all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V2_normalized_batch_corrected"]])
+datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))
+
+NN <- min(table(datasets))
+
+res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+for(i in levels(datasets)){
+  for(j in rownames(all_microbiome) ){
+
+    set.seed(1)
+    a <- c(sample(all_microbiome[j,datasets == i],NN))
+    set.seed(1)
+    b <- sample(all_microbiome[j,datasets != i],NN)
+
+    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+    temp <- wilcox.test(x=a, y=b)
+
+    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+  }
+}
+res <- res[-1,]
+res$pvalue <- as.numeric(res$pvalue)
+res$perc.1 <- as.numeric(res$perc.1)
+res$perc.2 <- as.numeric(res$perc.2)
+res$perc.diff <- res$perc.1 - res$perc.2
+
+res$fc <- as.numeric(res$fc)
+res$FDR <- p.adjust(res$pvalue)
+res <- res[order(res$pvalue),]
+res <- res[abs(res$fc) > fct,]
+res <- res[abs(res$pvalue) < pvt,]
+res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+dim(res)
+
+ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+mypar(mar=c(2,15,1,1))
+plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+          cex.row = .8,cex.col = .8,srt = 0)
+
+```
+
+# Correlation CVL3 and RNAseq
+
+```{r, echo = FALSE}
+
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+CVL3 <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized","ASV_CVL_V3_normalized_batch_corrected"),] > 0)==2 ]
+
+#filter TRX dataset
+TRX <- TRX [ , samples_TRX_and_CVL3 ]
+TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
+top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:3000] )
+plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
+TRX <- TRX[ top_vars_TRX , ]
+dim(TRX)
+
+#filter CVL3 dataset
+CVL3 <- CVL3 [ , samples_TRX_and_CVL3 ]
+CVL3 <- CVL3 [ rowSums(CVL3>0)>= 2 , ]
+top_vars_CVL3 <- names(sort(apply(CVL3,1,var),decreasing = T)[1:100] )
+dim(CVL3)
+
+mypar()
+cors <- cor( t(rbind(TRX) ) ,  t(rbind(CVL3) ) )
+hist(cors,xlim=c(-1,1),breaks = 100)
+dim(cors)
+
+library(fgsea)
+bacteria_use <- colnames(cors)
+gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
+enrichments <- lapply(bacteria_use,function(x){
+  res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=1000)
+  return(res)
+})
+names(enrichments) <- bacteria_use
+
+pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
+pvalues <- as.data.frame(pvalues)
+pvalues <- -log10( pvalues )
+top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
+mypar(mar=c(4,15,2,2))
+boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
+
+pvalues[ pvalues < -log10(0.05)] <- 0
+sort(rowSums(pvalues>0),T)
+sort(colSums(pvalues>0),T)
+
+NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
+NESes <- t(as.data.frame(NESes))
+# 
+# o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+# o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
+o_kegg <- hclust( dist( t(NESes) ),"ward.D2")$order
+o_bacs <- hclust( dist( NESes ),"ward.D2")$order
+NESes <- NESes[o_bacs,o_kegg]
+
+mypar(mar=c(15,2,2,15))
+image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-4,4,length.out = 92),
+       axes=F,border=NA,main="KEGG pathways",xlab="",ylab="bacterias",line=.4,cex.main=1,font.main=1)
+text(  par("usr")[c(4)]+.02 , seq(1,0,length.out = nrow(NESes)),
+      labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.3)
+text( seq(0,1,length.out = nrow(NESes)) , par("usr")[c(1)]-.02,
+      labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.3)
+
+eee <- data.frame(x=c(sapply(colnames(NESes),function(x)rep(x,nrow(NESes)))),
+                  y=rep(rownames(NESes),ncol(NESes)),
+                  w=c(NESes),
+                  p=c(t(pvalues)),
+                  x1=0,
+                  y1=c(sapply(ncol(NESes):1,function(x)rep(x,nrow(NESes))))/ncol(NESes) ,
+                  x2=1,
+                  y2=rep(nrow(NESes):1,ncol(NESes))/nrow(NESes) )
+
+eee$w2 <- eee$w / max(eee$w)
+eee <- eee[eee$w > 1.5,]
+eee <- eee[eee$p > 2,]
+
+eee$y1 <- rank(eee$y1) / length(eee$y1)
+eee$y2 <- rank(eee$y2) / length(eee$y2)
+
+plot(c(0,1),c(0,1),type="n")
+for( i in 1:nrow(eee)){
+  lines(c(eee$x1[i],eee$x2[i]),c(eee$y1[i],eee$y2[i]),
+        col=colorRampPalette(c("grey95","grey","firebrick"))(99)[ eee$w2 * 98 + 1 ] )
+}
+
+```
+
+```{r, echo = FALSE}
+
+g <- graph_from_incidence_matrix( 1-cors, directed = F,weighted = T  )
+
+l <- layout_with_fr(g)
+plot( g , vertex.label.cex=0.000001 , layout=l,
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
+      vertex.size=1,vertex.color=ifelse(V(g)$name %in% colnames(CVL3),"red","blue") )
+
+```

### 2020-10-15 reports/rmarkdown/report.Rmd
updates to export uncorrected data CVL2

@@
 k <- 10 # k for KNN
-P <- 0.4 # prunning threshold of SNN
+P <- 0.35 # prunning threshold of SNN
 
@@
                    "ASV_CVL_V3_normalized_batch_corrected.csv",
-                   "ASV_CVL_V2_normalized_batch_corrected.csv")
+                   "ASV_CVL_V2_normalized_batch_corrected.csv",
+                   "ASV_CVL_V2_normalized_NOT_batch_corrected.csv")
 
@@
 # Create empty matrix for the full dataset
-microbiome <- as.matrix(datasets[[3]])
+microbiome <- as.matrix(datasets[[4]])
 microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]

### 2020-10-28 reports/rmarkdown/00_knit_all_reports.Rmd
adds the report for comparing microbiome datasets

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r, echo = T}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
```

```{r, echo = T}
rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/01_data_pre_processing.Rmd"),
                  output_format = "pdf_document")
```

```{r, echo = T}
rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/02_data_quality_control.Rmd"),
                  output_format = "pdf_document")
```

```{r, echo = T}
dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_NOT_batch_corrected.csv")

for(i in dataset_names){
  rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/03_individual_reports.Rmd"),
                    output_format = "pdf_document",
                    params = list(dataset=i),
                    output_file = paste0(PATH,"/reports/rmarkdown/report_",sub(".csv","",i)))
}
```

### 2020-10-28 reports/rmarkdown/02_individual_reports.Rmd
adds the report for comparing microbiome datasets

--- 2020-10-15 reports/rmarkdown/report.Rmd
@@
 params:
-  organisation: Karolinska Institutet
-  pi_mail: prin.cipal@ki.se
-  pi_name: Prin Cipal
-  project_title: Broliden_5325
-  redmine_issue: '#5325'
-  requester_mail: kristina.broliden@ki.se
-  requester_name: Kristina Broliden
-  staff_mail: supp.ortstaff@nbis.se
-  staff_name: Supp Ortstaff
-  staff_web: https://nbis.se/about/staff/supp-ortstaff/
+  dataset: 1
 ---
@@
 min_pct <- .1 # minimun level of detected bacteria in each sample group
+
+#dataset use
+dataset_use <- 3
 ```
@@
 rownames(metadata) <- as.character(metadata$ID)
-
-# a <- read.csv(paste0(PATH,"/ASV_tissue_B2.csv"))
-# b <- read.csv(paste0(PATH,"/ASV_tissue_B1.csv"))
-# rownames(a) <- paste0()
-# temp <- as.matrix(a[,c((ncol(a)-6):ncol(a))])
-# temp[is.na(temp)] <- "NA"
-# taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
-# taxa <- sub("_NA.*","_other",taxa)
 
@@
 lapply(datasets,dim)
-
-#arrange all datasets to have the same rownames with all bacteria
-# common_microbes <- unique(unlist(lapply(datasets,rownames)))
-# datasets <- lapply(datasets, function(x) {
-#   temp <- matrix(0,nrow = length(common_microbes),
-#                  ncol = ncol(x),
-#                  dimnames = list(common_microbes,colnames(x)))
-#   temp[rownames(x),colnames(x)] <- x
-#   return(temp)
-# })
-# lapply(datasets,dim)
-
-# datasets <- datasets[2] #in case you wanna test in 1 dataset
-
-# Transform phyloseq data into normalized matricies ( log2[CP1K] )
-# datasets <- lapply(datasets,function(x){
-#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
-# })
-# lapply(datasets,dim)
 ```
@@
 # Create empty matrix for the full dataset
-microbiome <- as.matrix(datasets[[4]])
+microbiome <- as.matrix(datasets[[ params$dataset ]])
 microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]
@@
 
-#Plotting bacteria across clusters
+# Saving clusters and differentially expressed bacteria
 
-```{r, eval=F,echo = FALSE}
-mypar(2,2)
+```{r, eval=T,echo = FALSE}
+write.csv2( data.frame(cl$membership,row.names = colnames(microbiome)) ,row.names = T,
+            paste0(PATH,"/results/Sample_Louvain_clusters_",params$dataset) )
 
-zscore_bac <- apply(microbiome,2,function(i){scale(i,T,T)})
-zscore_bac[ zscore_bac > 6 ] <- 6
-zscore_bac[ zscore_bac < -6 ] <- -6
-rownames(zscore_bac) <- rownames(microbiome); colnames(zscore_bac) <- colnames(microbiome)
+write.csv2( data.frame(l,row.names = colnames(microbiome)) ,row.names = T,
+            paste0(PATH,"/results/Sample_graph_layout_",params$dataset) )
 
-#Compute sample-wise correlations
-cors <- cor(t(microbiome),method = "pearson")
-
-#Compute KNN graph from correlation matrix
-knn <- RANN::nn2(cors,k = k)
-# knn <- RANN::nn2(t(zscore),k = k)
-nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
-for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
-
-#Compute SNN graph from KNN
-jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
-
-#Prune SNN graph based on the Jaccard index
-pSNN <- jSNN
-pSNN <- pSNN / max(pSNN)
-
-hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="SNN jaccard",xlim=c(0,1))
-abline(v=P,col="red",lty=2,xpd=F)
-
-hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="correlation coeficient",xlim=c(-1,1))
-abline(v=0,col="red",lty=1,xpd=F)
-
-pSNN[ pSNN < .3 ] <- 0
-pSNN[ pSNN > .3 ] <- cors[pSNN > .3]
-pSNN[ pSNN < 0 ] <- 0
-
-g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
-
-mypar(2,2)
-
-set.seed(1)
-l <- layout_with_fr(g,niter=3000,start.temp=30)
-
-cl_bac <- igraph::cluster_louvain(g)
-
-plot( g , vertex.label.cex=0.000001  , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.color= hue_pal()(length(unique(cl_bac$membership)))[factor(cl_bac$membership)],
-      vertex.size=10,
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
-       legend = levels(factor(cl_bac$membership)),pt.bg=hue_pal()(length(unique(cl_bac$membership))),
-       bty = "n",pch = 21,pt.cex = 1,xpd=T)
-title("Louvain clusters",cex.main=1,font.main=1)
-
-o_bacs <- order(cl_bac$membership)
-image( t(zscore[o_bacs,o][nrow(zscore):1,]),col=heat_pal,axes=F,border=NA,
-      main="bacteria",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
-
-text(  par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]) , seq(1,0,length.out = nrow(zscore)),
-      labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.7)
-points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl_bac$membership)),
-        (0:(length(cl_bac$membership)-1) )/(length(cl_bac$membership)-1), pch=16,
-    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
-    col=pal[factor(cl$membership)][o_bacs] ,xpd=T)
-
-par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)])
-
-points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
-        col=pal[factor(cl$membership)][o] ,xpd=T)
-points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl_bac$membership)), pch=16,
-    rev((0:(length(cl_bac$membership)-1) )/(length(cl_bac$membership)-1)),
-    col=pal[factor(cl_bac$membership)][o_bacs] ,xpd=T)
-
+write.csv2( res ,
+            paste0(PATH,"/results/DGE_",params$dataset),row.names = T )
 ```
-
-#Dataset integration
-
-```{r, eval=F, echo = FALSE, fig.width  = 10, fig.height=6}
-# Load the transcriptome datset
-trx <- read.csv(paste0(PATH,"~/Downloads/TRX_rawcounts.csv"),row.names = 1)
-trx <- rowsum(trx[,-c(1,2)],trx$symbol)
-colnames(trx) <- sub("ML","",colnames(trx))
-trx  <- trx[ rownames(trx) != "" , ]
-trx <- trx[ rowSums(trx > 5) >= 3 , ]
-trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
-
-# Identify highly variable genes in the transcriptomic dataset
-vars <- log2(apply(trx,1,var)+2)
-means <- apply(trx,1,mean)
-mypar(1,2)
-top_variable <- names( sort(vars,decreasing = T) ) [1:3000]
-plot( means , vars , pch=16, las=1,xlab="logMeanExpression",ylab="variance",
-      cex=ifelse( rownames(trx) %in% top_variable , .2,.1),
-      col= ifelse( rownames(trx) %in% top_variable , "red","black"))
-
-# Identify common samples between the transcriptome and the mirobiome dataset
-common_samples <- colnames(microbiome)[colnames(microbiome) %in% colnames(trx)]
-trx <- trx[,as.character(common_samples) ]
-microbiome2 <- microbiome[,as.character(common_samples)]
-
-# Combine datased and set all to same scale using Z-score transformation
-mypar(1,3)
-merged <- rbind(trx[top_variable,],microbiome2)
-zscore_merged <- t(apply(merged,1,function(i){scale(i,T,T)}))
-zscore_merged[ zscore_merged > 5 ] <- 5
-zscore_merged[ zscore_merged < -5 ] <- -5
-rownames(zscore_merged) <- rownames(zscore_merged)
-colnames(zscore_merged) <- colnames(zscore_merged)
-hist(zscore_merged,breaks = 100,main="all")
-hist(zscore_merged[top_variable,],breaks = 100,main="trx")
-hist(zscore_merged[rownames(microbiome2),],breaks = 100,main="microbiome")
-
-PC <- prcomp(zscore_merged,center = F,scale. = F)
-plot(PC$sdev,pch=21,bg="grey70")
-abline(v=50.5,col="red",xpd=F,lty=2)
-
-knn <- RANN::nn2( query = PC$x[rownames(microbiome2),1:50] , 
-                data = PC$x[top_variable,1:50] ,k = 50)
-# knn <- RANN::nn2(t(zscore),k = k)
-nn <- matrix(0, length(rownames(microbiome2)), length(top_variable))
-for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
-nn <- nn[ rowSums(nn)>0 , colSums(nn)>0 ]
-
-mypar(1,2,mar=c(0,0,2,10))
-
-g <- graph_from_adjacency_matrix(nn, mode = "undirected",diag = F,weighted = T)
-set.seed(1)
-# l <- layout_with_graphopt(g,niter = 1000)
-l <- layout_with_fr(g,niter=500,start.temp=100)
-
-cl <- igraph::cluster_louvain(g)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
-       legend = levels(factor(cl$membership)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
-title("Louvain clusters",cex.main=1,font.main=1)
-
-```
-
-```{r, eval=F,echo = FALSE, fig.width  = 10, fig.height=11}
-
-```

### 2020-10-28 reports/rmarkdown/02b_evaluating_batch_correction.Rmd
adds the report for comparing microbiome datasets

--- 2020-10-13 reports/rmarkdown/02_data_quality_control.Rmd
@@
 knitr::opts_chunk$set(fig.width  = 10,
-                      results    = "hold",
+                      results    = "hide",
                       message    = FALSE,
@@
 
-```{r, echo = FALSE}
+```{r}
 library(igraph)
@@
 
-```{r, echo = FALSE}
+```{r}
 PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data
@@
 
-```{r, echo = FALSE}
+```{r}
 # Load metadata
@@
 
-# Calculate QC metrics
+# Comparing CVL2 samples before and after batch correction
 
-```{r, echo = FALSE}
-counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
-counts[is.na(counts)] <- 0
-detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
-detected_counts[is.na(detected_counts)] <- 0
-rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
+```{r}
+mypar(1,2,mar=c(2,15,1,1))
 
-mypar(mar=c(2,20,2,1))
-barlist( data = rbind( counts , detected_counts),
-         genes = c(rownames(counts),rownames(detected_counts)),
-         clustering = rep(1,nrow(metadata)),
-         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)
+for( dataset in c("ASV_CVL_V2_normalized_NOT_batch_corrected",
+            "ASV_CVL_V2_normalized_batch_corrected")){
 
-sum((colSums(counts == 0)>=3)*1)
+  # Computing differential expression across batches
+  all_microbiome <- cbind(datasets_all_samples[[dataset]])
+  datasets <- read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,2]
+
+  all_microbiome <- all_microbiome[,as.character(read.csv(paste0(PATH,"/results/batches_CVL2.csv"))[,1])]
+
+  NN <- min(table(datasets))
+
+  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+  for(i in levels(datasets)){
+    for(j in rownames(all_microbiome) ){
+
+      set.seed(1)
+      a <- c(sample(all_microbiome[j,datasets == i],NN))
+      set.seed(1)
+      b <- sample(all_microbiome[j,datasets != i],NN)
+
+      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+      temp <- wilcox.test(x=a, y=b,exact = F)
+
+      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                            c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+    }
+  }
+
+  res <- res[-1,]
+  res$pvalue <- as.numeric(res$pvalue)
+  res$perc.1 <- as.numeric(res$perc.1)
+  res$perc.2 <- as.numeric(res$perc.2)
+  res$perc.diff <- res$perc.1 - res$perc.2
+
+  res$fc <- as.numeric(res$fc)
+  res$FDR <- p.adjust(res$pvalue)
+  res <- res[order(res$pvalue),]
+  res <- res[res$fc > fct,]
+  res <- res[abs(res$pvalue) < 0.05,]
+  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+  dim(res)
+
+  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets, main = dataset,show_grid = T,cex.main=1,font.main=1,
+            cex.row = .8,cex.col = .8,srt = 0)
+}
 ```
-
-# Calculate QC metrics
-
-```{r, echo = FALSE}
-counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
-counts[is.na(counts)] <- 0
-detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
-detected_counts[is.na(detected_counts)] <- 0
-rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
-
-mypar(mar=c(2,20,2,1))
-barlist( data = rbind( counts , detected_counts),
-         genes = c(rownames(counts),rownames(detected_counts)),
-         clustering = rep(1,nrow(metadata)),
-         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)
-
-sum((colSums(counts == 0)>=3)*1)
-```

### 2020-10-28 reports/rmarkdown/03_integrated_datasets.Rmd
adds the report for comparing microbiome datasets

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hide",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(fgsea)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)

#Graph construction  
fct <- .25 # FC threshold for differential expression
pvt <- 0.01 # Pvalue threshold for differential expression
min_pct <- .1 # minimun level of detected bacteria in each sample group

```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# Load datasets
dataset_names <- c("Tissue_RNAseq_V3_normalized",
                     "ASV_tissue_V3_normalized_batch_corrected",
                     "ASV_CVL_V3_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_NOT_batch_corrected")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"results/",x,".csv"),row.names = 1)
  return(as.matrix(x)) })
names(datasets) <- dataset_names

#fill all samples with 0s to make plotting easier
datasets_all_samples <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = nrow(x),
                 ncol = nrow(metadata),
                 dimnames = list(rownames(x),rownames(metadata)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets_all_samples,dim)
```

# Calculate QC metrics

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=5}
counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
counts[is.na(counts)] <- 0
detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
detected_counts[is.na(detected_counts)] <- 0
rownames(detected_counts) <- paste0(rownames(detected_counts),">0")

mypar(2,1,mar=c(2,10,2,1))
barlist( data = rbind( counts ),main = "total counts",
         genes = c(rownames(counts)), labels = sub("_norm.*","",rownames(counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
add_letter("a")

barlist( data = rbind( detected_counts),main = "detected features",
         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
add_letter("b")

sum((colSums(counts == 0)>=3)*1)
colnames(counts)[(colSums(counts == 0)==3)]
paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")

metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
```

**Figure 1.** Comparative barplot for the **a)** total counts and **b)** number of non-zero detected features (genes / bacteria) for each of the sequencing datasets. Samples are ordered alfabetically according to the patient ID. Samples that are present in exactly two datasets are shown in orange (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`). Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==2)],collapse = ",")`).

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,2,mar=c(2,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)
add_letter("a")
barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.axis = .8,srt = 0)
add_letter("b")
```

**Figure 2.** Differential bacterial abundance across microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

\clearpage

```{r, echo = FALSE}
groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
names(groupings) <- names(datasets_all_samples)[-1]

groupings_all_samples <- lapply(groupings,function(x){
  df <- data.frame(group=rep("NA",nrow(metadata)), row.names = rownames(metadata),stringsAsFactors = F)
  df[rownames(x),1] <- as.character(x[,1])
  return(df)
})
names(groupings_all_samples) <- names(groupings)

df <- as.data.frame(groupings_all_samples)
colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
metadata <- cbind(metadata,df)
```

# Computing differential expression across microbiome datasets

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) )
datasets <- paste0(datasets, "_",c(metadata$louvain_tissue_V3,metadata$louvain_CVL_V3,metadata$louvain_CVL_V2))

datasets <- factor( datasets[ colSums(all_microbiome)>0 ])
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))
```

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(2,2)
group_means <- sapply( levels(datasets), function(x) rowMeans( all_microbiome[,datasets==x] ) )
cors <- cor(group_means)
h <- hclust( as.dist( (1-cors)/2 ),method = "ward.D2" )
o <- h$order

par(mar=c(6,6,2,2))
image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
      main="correlation matrix",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(h$labels)),
        rev((0:(length(h$labels)-1) )/(length(h$labels)-1)),
            labels = h$labels[o], srt = 0, adj = c(1,.5), xpd = TRUE, cex=1)
text(  (0:(length(h$labels)-1) )/(length(h$labels)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(h$labels)),
            labels = h$labels[o], srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
add_letter("a")

par(mar=c(6,4,2,2))
plot(as.dendrogram(h),las=1,ylab="height")
add_letter("b")
```

**Figure 3.** Comparrisson among patient groups across datasets. **(a)** Correlation matrix across sample groups. **(b)** Hierachical clustering of sample groups.

```{r, echo = FALSE,fig.width=10,fig.height=14}
res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,1,mar=c(5,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .7,cex.col = .8,srt = 30)
add_letter("a")
```

**Figure 4.** Differential bacterial abundance across all groups and all microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}
sample_order <- order(df$louvain_CVL_V3)

# zscore_datasets_all_samples <- lapply(datasets_all_samples,function(x){
#   zscore <- t(apply(x,1,function(i){scale(i,T,T)}))
#   zscore[ zscore > 6 ] <- 6
#   zscore[ zscore < -6 ] <- -6
#   rownames(zscore) <- rownames(x); colnames(zscore) <- colnames(x)
#   zscore[is.nan(zscore)] <- -6
#   return(zscore)
# })

mypar(3,1,mar=c(1,2,1,0.5))

image(t(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_tissue_V3)), pch=16,
        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_tissue_V3)), pch=16,
        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)

image(t(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V3)), pch=16,
        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V3)), pch=16,
        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)

image(t(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V2)), pch=16,
        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V2)), pch=16,
        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
```

**Figure 5.** Comparisson of microbiome datasets, showing only the significant bacteria. Samples are ordered by the CVL3 groupings. The colors represent their respective bacterial groupings for each dataset.

\clearpage

```{r, echo = FALSE,fig.width=8,fig.height=8}
metadata_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
                         "blood_cervix_v2","blood_cervix_v3",
                         "cervical_inflammation_v2","cervical_inflammation_v3",
                         "HIVstatus","Contraception",
                         "CT_v2","CT_v3","regular_partner",
                         "vaginal_douching_v2","vaginal_douching_v3",
                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")
as.character(metadata$group_M9_CVLv3)

Ftest <- sapply(metadata_parameters,function(x){
  temp <- sapply(metadata_parameters,x=x,function(x,y){
    if (x==y){
      return( 1 )
    } else {
    f1 <- as.character(metadata[,x])
    f1[is.na(f1)] <- "NA"
    f1[f1==""] <- "NA"
    f2 <- as.character(metadata[,y])
    f2[f2==""] <- "NA"
    contT1  <- table(f1,f2)
    print(x)
    print(y)
    return( fisher.test(contT1,simulate.p.value = T)$p.value )
    # return( chisq.test(contT1,simulate.p.value = T)$p.value )
    }
  })
  return(temp)
})

clean_Ftest <- Ftest
# clean_Ftest <- ( clean_Ftest + t(clean_Ftest) ) / 2
clean_Ftest <- -log10(clean_Ftest)
clean_Ftest[clean_Ftest > 4] <- 4
# clean_Ftest[ clean_Ftest > 0.05 ] <- 1

mypar(1,1,mar=c(12,1,1,12))
image( t(clean_Ftest [nrow(Ftest):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(Ftest)) , 
       labels = rownames(Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(Ftest)) , par("usr")[3]-.02 , 
      labels = rownames(Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
```

**Figure 6.** Association analysis across several patient categorical metadata parameters, including patient groupings annotations from microbiome.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(6,4,mar=c(1,4,1,4) )
x <- as.character(metadata[,"Contraception"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
plot_sankey( data.frame(x,y),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "Contraception" , "Antibiotics" ),pos=3,xpd=T)

x <- as.character(metadata[,"BV_Monbilicus_v2"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
plot_sankey( data.frame(x,y),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "BV_Monbilicus_v2" , "Antibiotics" ),pos=3,xpd=T)

x <- as.character(metadata[,"HIVstatus"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
eee <- plot_sankey( data.frame(x,y),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "HIVstatus" , "Antibiotics_v2" ),pos=3,xpd=T)

x <- as.character(metadata[,"blood_cervix_v3"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"cervical_inflammation_v3"])
y[is.na(y)] <- "NA"
plot_sankey( data.frame(x,y),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(9))
text(c(0,1),c(1,1),labels = c( "blood_cervix_v3" , "cervical_inflammation_v3" ),pos=3,xpd=T)

```

**Figure 7.** A few examples of significant association between metadata parameters shown as sankey plots.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(min(length(metadata_parameters),6),4,mar=c(1,4,1,4))

# metadata$Contraception
#Antibiotics usage
for(i in metadata_parameters){
  empty_plot()
  text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
  for(j in c("louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")){
    x <- as.character(metadata[,i])
    x[is.na(x)] <- "NA"
    y <- as.character(metadata[,j])
    y[is.na(y)] <- "NA"
    plot_sankey( data.frame(x,y),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
    text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
  }
}
```

**Figure 8.** Sankey plots for all tested associations between the patient groups identified in in the microbiome datasets.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=13}
myletter <- letters

for(i in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected")){

  cat(paste0("Correlation between RNAseq and ",sub("_norm.*","",i)))

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- datasets_all_samples[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
  TRX <- TRX[ top_vars_TRX , ]
  dim(TRX)

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  # hist(cors,xlim=c(-1,1),breaks = 100)
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- bacteria_use

  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
  # mypar(mar=c(4,15,2,2))
  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  NESes <- NESes[rowSums(NESes!=0) >= 10 , colSums(NESes!=0) >= 10]

  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
  o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order

  NESes <- NESes[o_bacs,o_kegg]

  mypar(1,1,mar=c(12,2,2,12))
  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
  add_letter(myletter); myletter <- myletter[-1]
  # empty_plot()
  cat("\\clearpage")

}
```

**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown. Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}
# eee <- data.frame(x=c(sapply(colnames(NESes),function(x)rep(x,nrow(NESes)))),
#                   y=rep(rownames(NESes),ncol(NESes)),
#                   w=c(NESes),
#                   p=c(t(pvalues)),
#                   x1=0,
#                   y1=c(sapply(ncol(NESes):1,function(x)rep(x,nrow(NESes))))/ncol(NESes) ,
#                   x2=1,
#                   y2=rep(nrow(NESes):1,ncol(NESes))/nrow(NESes) )
# 
# eee$w2 <- eee$w / max(eee$w)
# eee <- eee[eee$w > 1.5,]
# eee <- eee[eee$p > 2,]
# 
# eee$y1 <- rank(eee$y1) / length(eee$y1)
# eee$y2 <- rank(eee$y2) / length(eee$y2)

# 
# plot(c(0,1),c(0,1),type="n")
# for( i in 1:nrow(eee)){
#   lines(c(eee$x1[i],eee$x2[i]),c(eee$y1[i],eee$y2[i]),
#         col=colorRampPalette(c("grey95","grey","firebrick"))(99)[ eee$w2 * 98 + 1 ] )
# }

```

```{r, eval=F,echo = FALSE}
g <- graph_from_incidence_matrix( 1-cors, directed = F,weighted = T  )

l <- layout_with_fr(g)
plot( g , vertex.label.cex=0.000001 , layout=l,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=1,vertex.color=ifelse(V(g)$name %in% colnames(CVL3),"red","blue") )
```

### 2020-10-30 reports/rmarkdown/00_knit_all_reports.Rmd
adds integrated analysis

@@
 ```{r, echo = T}
-rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/02_data_quality_control.Rmd"),
+rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/02b_data_quality_control.Rmd"),
                   output_format = "pdf_document")
@@
 for(i in dataset_names){
-  rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/03_individual_reports.Rmd"),
+  rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/02_individual_reports.Rmd"),
                     output_format = "pdf_document",
@@
 ```
+
+```{r, echo = T}
+rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/03_integrated_datasets.Rmd"),
+                  output_format = "pdf_document")
+```

### 2020-10-30 reports/rmarkdown/02_individual_reports.Rmd
adds integrated analysis

@@
 min_pct <- .1 # minimun level of detected bacteria in each sample group
-
-#dataset use
-dataset_use <- 3
 ```

### 2020-11-12 reports/rmarkdown/03_integrated_datasets.Rmd
removes NAs, creates bacterial communities

@@
 library(fgsea)
-source("~/repos/niceRplots/R/plotting_functions.R")
-source("~/repos/niceRplots/R/add_fig_label.R")
-source("~/repos/niceRplots/R/helper_functions.R")
+library(scales)
 
+remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
+library(niceRplots)
 ```
@@
 #color palettes
-pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
+pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
+         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
 heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
@@
 Samples are ordered alfabetically according to the patient ID.
-Samples that are present in exactly two datasets are shown in orange (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`).
-Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==2)],collapse = ",")`).
+Samples that are present in exactly two datasets are shown in orange (`r paste0(colnames(counts)[(colSums(counts == 0)==2)],collapse = ",")`).
+Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`).
 
@@
 **Figure 2.** Differential bacterial abundance across microbiome datasets.
+The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
+
+\clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=9}
+all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+datasets <- factor(c( rep("CVL_v3",121) , rep("CVL_v2",121) ))
+datasets <- datasets[ colSums(all_microbiome)>0 ]
+all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
+
+NN <- min(table(datasets))
+
+res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+for(i in levels(datasets)){
+  for(j in rownames(all_microbiome) ){
+
+    a <- all_microbiome[j,datasets == i]
+    b <- all_microbiome[j,datasets != i]
+
+    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+    temp <- wilcox.test(x=a, y=b)
+
+    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+  }
+}
+res <- res[-1,]
+res$pvalue <- as.numeric(res$pvalue)
+res$perc.1 <- as.numeric(res$perc.1)
+res$perc.2 <- as.numeric(res$perc.2)
+res$perc.diff <- res$perc.1 - res$perc.2
+
+res$fc <- as.numeric(res$fc)
+res$FDR <- p.adjust(res$pvalue)
+res <- res[order(res$pvalue),]
+res <- res[abs(res$fc) > fct,]
+res <- res[abs(res$pvalue) < pvt,]
+res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+dim(res)
+
+ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+mypar(1,2,mar=c(2,15,1,1))
+plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+          cex.row = .8,cex.col = .8,srt = 0)
+add_letter("a")
+barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+          cex.axis = .8,srt = 0)
+add_letter("b")
+```
+
+**Figure 2.** Differential bacterial abundance across CVL2 and CVL3 microbiome datasets.
+The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
+
+\clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=9}
+all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
+datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) ))
+datasets <- datasets[ colSums(all_microbiome)>0 ]
+all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
+
+NN <- min(table(datasets))
+
+res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+for(i in levels(datasets)){
+  for(j in rownames(all_microbiome) ){
+
+    a <- all_microbiome[j,datasets == i]
+    b <- all_microbiome[j,datasets != i]
+
+    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+    temp <- wilcox.test(x=a, y=b)
+
+    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+  }
+}
+res <- res[-1,]
+res$pvalue <- as.numeric(res$pvalue)
+res$perc.1 <- as.numeric(res$perc.1)
+res$perc.2 <- as.numeric(res$perc.2)
+res$perc.diff <- res$perc.1 - res$perc.2
+
+res$fc <- as.numeric(res$fc)
+res$FDR <- p.adjust(res$pvalue)
+res <- res[order(res$pvalue),]
+res <- res[abs(res$fc) > fct,]
+res <- res[abs(res$pvalue) < pvt,]
+res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+dim(res)
+
+ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+mypar(1,2,mar=c(2,15,1,1))
+plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+          cex.row = .8,cex.col = .8,srt = 0)
+add_letter("a")
+barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+          cex.axis = .8,srt = 0)
+add_letter("b")
+```
+
+**Figure 2b.** Differential bacterial abundance across tissue and CVL microbiome week3 datasets.
 The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
@@
                          "vaginal_douching_v2","vaginal_douching_v3",
-                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")
-as.character(metadata$group_M9_CVLv3)
+                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3",
+                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3")
 
@@
     f1[f1==""] <- "NA"
+
     f2 <- as.character(metadata[,y])
+    f2[is.na(f2)] <- "NA"
     f2[f2==""] <- "NA"
-    contT1  <- table(f1,f2)
+
+    f11 <- f1 [ (f1!="NA") & (f2!="NA") ]
+    f22 <- f2 [ (f1!="NA") & (f2!="NA") ]
+    contT1  <- table(f11,f22)
     print(x)
@@
 y[is.na(y)] <- "NA"
-plot_sankey( data.frame(x,y),color_by = 2,
+x1 <- x [ (x!="NA") & (y!="NA") ]
+y1 <- y [ (x!="NA") & (y!="NA") ]
+plot_sankey( data.frame(x1,y1),color_by = 2,
              plot_labels = T, plot_weights = F,order_1_by = "total",
@@
 y[is.na(y)] <- "NA"
-eee <- plot_sankey( data.frame(x,y),color_by = 2,
+x1 <- x [ (x!="NA") & (y!="NA") ]
+y1 <- y [ (x!="NA") & (y!="NA") ]
+plot_sankey( data.frame(x1,y1),color_by = 2,
              plot_labels = T, plot_weights = F,order_1_by = "total",
@@
 y[is.na(y)] <- "NA"
-plot_sankey( data.frame(x,y),
+x1 <- x [ (x!="NA") & (y!="NA") ]
+y1 <- y [ (x!="NA") & (y!="NA") ]
+plot_sankey( data.frame(x1,y1),
              plot_labels = T, plot_weights = F,order_1_by = "total",
@@
     y[is.na(y)] <- "NA"
-    plot_sankey( data.frame(x,y),
+    x1 <- x [ (x!="NA") & (y!="NA") ]
+    y1 <- y [ (x!="NA") & (y!="NA") ]
+    plot_sankey( data.frame(x1,y1),
              plot_labels = T, plot_weights = F,order_1_by = "total",
              order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(10))
+             pal=hue_pal()(13))
     text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
@@
 ```{r, echo = FALSE,fig.width=10,fig.height=10}
-# eee <- data.frame(x=c(sapply(colnames(NESes),function(x)rep(x,nrow(NESes)))),
-#                   y=rep(rownames(NESes),ncol(NESes)),
-#                   w=c(NESes),
-#                   p=c(t(pvalues)),
-#                   x1=0,
-#                   y1=c(sapply(ncol(NESes):1,function(x)rep(x,nrow(NESes))))/ncol(NESes) ,
-#                   x2=1,
-#                   y2=rep(nrow(NESes):1,ncol(NESes))/nrow(NESes) )
-# 
-# eee$w2 <- eee$w / max(eee$w)
-# eee <- eee[eee$w > 1.5,]
-# eee <- eee[eee$p > 2,]
-# 
-# eee$y1 <- rank(eee$y1) / length(eee$y1)
-# eee$y2 <- rank(eee$y2) / length(eee$y2)
+cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>10,]))
 
-# 
-# plot(c(0,1),c(0,1),type="n")
-# for( i in 1:nrow(eee)){
-#   lines(c(eee$x1[i],eee$x2[i]),c(eee$y1[i],eee$y2[i]),
-#         col=colorRampPalette(c("grey95","grey","firebrick"))(99)[ eee$w2 * 98 + 1 ] )
-# }
+#Graph construction  
+k <- 10 # k for KNN
+P <- 0.3 # prunning threshold of SNN
 
+knn <- RANN::nn2(cors,k = k)
+# knn <- RANN::nn2(t(zscore),k = k)
+nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
+for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
+
+#Compute SNN graph from KNN
+jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
+
+#Prune SNN graph based on the Jaccard index
+pSNN <- jSNN
+pSNN <- pSNN / max(pSNN)
+
+hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
+     border=NA,col="grey60",las=1,
+     xaxs="i",yaxs="i",
+     main="SNN jaccard",xlim=c(0,1))
+abline(v=P,col="red",lty=2,xpd=F)
+
+hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
+     border=NA,col="grey60",las=1,
+     xaxs="i",yaxs="i",
+     main="correlation coeficient",xlim=c(-1,1))
+abline(v=0,col="red",lty=1,xpd=F)
+
+pSNN[ pSNN < P ] <- 0
+pSNN[ pSNN > P ] <- cors[pSNN > P]
+pSNN[ pSNN < 0 ] <- 0
+
+g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
 ```
 
-```{r, eval=F,echo = FALSE}
-g <- graph_from_incidence_matrix( 1-cors, directed = F,weighted = T  )
+# Visualise the data
 
-l <- layout_with_fr(g)
-plot( g , vertex.label.cex=0.000001 , layout=l,
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
-      vertex.size=1,vertex.color=ifelse(V(g)$name %in% colnames(CVL3),"red","blue") )
+```{r, echo = FALSE,fig.width  = 10, fig.height=3}
+mypar(2,2,mar=c(0,0,2,10))
+
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+cl <- igraph::cluster_louvain(g)
+table(cl$membership)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
+       y = par("usr")[c(4)],
+       legend = levels(factor(cl$membership)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+title("Louvain clusters",cex.main=1,font.main=1)
 ```
+
+```{r, echo = FALSE,fig.width  = 10, fig.height=10}
+mypar(2,5,mar=c(1,1,2,1))
+for(j in sort(unique(cl$membership)) ){
+  empty_plot()
+  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
+  title(main=paste0("community",j))
+}
+```
+
+```{r, eval=T, echo = FALSE,fig.width=10, fig.height=10}
+sample_order <- order(df$louvain_CVL_V3)
+
+module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
+  return( rowsum( x[rownames(cors),] , group = cl$membership ) / c(table(cl$membership)) )
+})
+
+mypar(3,1,mar=c(1,2,1,0.5))
+sample_order <- order(df$louvain_tissue_V3)
+image(t(module_means_per_dataset[["ASV_tissue_V3_normalized_batch_corrected"]]
+        [ ,sample_order]),
+      col=heat_pal,axes=F,border=NA,
+      main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
+        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_tissue_V3)), pch=16,
+        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
+points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
+        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_tissue_V3)), pch=16,
+        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
+text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
+         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
+         labels= rev(sort(unique(cl$membership)) )    )
+
+sample_order <- order(df$louvain_CVL_V3)
+image(t(module_means_per_dataset[["ASV_CVL_V3_normalized_batch_corrected"]]
+        [ ,sample_order]),
+      col=heat_pal,axes=F,border=NA,
+      main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
+        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_CVL_V3)), pch=16,
+        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
+points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
+        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_CVL_V3)), pch=16,
+        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
+text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
+         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
+         labels= rev(sort(unique(cl$membership)) )    )
+
+sample_order <- order(df$louvain_CVL_V2)
+image(t(module_means_per_dataset[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
+        [ ,sample_order]),
+      col=heat_pal,axes=F,border=NA,
+      main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
+        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_CVL_V2)), pch=16,
+        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
+points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
+        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+            length(df$louvain_CVL_V2)), pch=16,
+        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
+text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
+         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
+         labels= rev(sort(unique(cl$membership)) )    )
+```
+
+**Figure 5.** Comparisson of microbiome datasets, showing only the distribution of different bacterial communities (1-8).
+Samples are ordered by the CVL3 groupings.
+The colors represent their respective bacterial groupings for each dataset.
+
+```{r, echo = FALSE,fig.width=10,fig.height=13}
+myletter <- letters
+
+for(i in c("ASV_tissue_V3_normalized_batch_corrected",
+           "ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+
+  cat(paste0("Correlation between RNAseq and ",sub("_norm.*","",i)))
+
+  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+  bac_dataset <- module_means_per_dataset[[i]]
+  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
+  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
+
+  # mypar(4,3,mar=c(4,2,2,2))
+  #filter TRX dataset
+  TRX <- TRX [ , samples_TRX_and_CVL3 ]
+  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
+  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:2000] )
+  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
+  TRX <- TRX[ top_vars_TRX , ]
+  dim(TRX)
+
+  #filter bac_dataset dataset
+  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
+  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
+  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
+  dim(bac_dataset)
+
+  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
+  # hist(cors,xlim=c(-1,1),breaks = 100)
+  dim(cors)
+
+  bacteria_use <- colnames(cors)
+  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
+  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
+  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
+  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
+    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
+    return(res)
+  })
+  parallel::stopCluster(cl)
+  names(enrichments) <- bacteria_use
+
+  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
+  pvalues <- t(as.data.frame(pvalues))
+  pvalues <- -log10( pvalues )
+  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
+  # mypar(mar=c(4,15,2,2))
+  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
+
+  # pvalues[ pvalues < -log10(0.05)] <- 0
+  # sort(rowSums(pvalues>0),T)
+  # sort(colSums(pvalues>0),T)
+  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
+  NESes <- t(as.data.frame(NESes))
+  NESes[is.na(NESes)] <- 0
+  NESes[ pvalues < -log10(0.1) ] <- 0
+  NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
+
+  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
+  o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+  o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
+
+  NESes <- NESes[o_bacs,o_kegg]
+
+  mypar(1,1,mar=c(12,2,2,12))
+  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
+         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
+  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
+        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
+  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
+        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
+  add_letter(myletter); myletter <- myletter[-1]
+  # empty_plot()
+  cat("\\clearpage")
+
+}
+```
+
+**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
+Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database.
+This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue.
+The heatmap shows the normalized enrichment score (NES).
+Only enrichments with pvalue below 0.05 are shown.
+Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.

### 2020-11-27 reports/rmarkdown/01_data_pre_processing.Rmd
updates integrated report

@@
 library(batchelor)
-source("~/repos/niceRplots/R/plotting_functions.R")
-source("~/repos/niceRplots/R/add_fig_label.R")
-source("~/repos/niceRplots/R/helper_functions.R")
+remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
+library(niceRplots)
 
@@
 # lapply(datasets,dim)
+
+# Extract taxonomic information to file
+taxa_info <- lapply(dataset_names, function(x){
+  x <- read.csv( paste0(PATH,"/",x))
+  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
+  temp[is.na(temp)] <- "NA"
+  return(t(temp))
+})
+taxa_info <- t(as.data.frame(taxa_info))
+rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
+rownames(taxa_info) <- sub("[.][0123456789","other ",rownames(taxa_info))
+sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
+not_complete <- grepl("_NA",sps)
+taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
+taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
+taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
+taxa[!not_complete] <- sps[!not_complete]
+taxa <- sub("_"," ",taxa)
+use <- !is.na(taxa) & (taxa != "NA")
+taxa[taxa == "NA"] <- "other"
+rownames(taxa_info) <- taxa
+taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
+write.csv2(taxa_info, paste0(PATH,"../results/taxonomy.csv"),row.names = T)
 ```

### 2020-11-27 reports/rmarkdown/02_individual_reports.Rmd
updates integrated report

@@
 #color palettes
-pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2")) #color pallete for plots
+pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
+         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
 heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))

### 2020-11-27 reports/rmarkdown/03_integrated_datasets.Rmd
updates integrated report

@@
 library(scales)
-
 remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
@@
 pvt <- 0.01 # Pvalue threshold for differential expression
-min_pct <- .1 # minimun level of detected bacteria in each sample group
+min_pct <- .05 # minimun level of detected bacteria in each sample group
 
@@
 Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`).
+
+```{r, echo = FALSE,fig.width=10,fig.height=9}
+all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
+
+zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
+zscore[ zscore > 6 ] <- 6
+zscore[ zscore < -6 ] <- -6
+rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
+
+#Compute sample-wise correlations
+cors <- cor(zscore, method = "pearson")
+
+#Graph construction  
+k <- 5 # k for KNN
+P <- 0.4# prunning threshold of SNN
+
+#Compute KNN graph from correlation matrix
+knn <- RANN::nn2(cors,k = k)
+# knn <- RANN::nn2(t(zscore),k = k)
+nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
+for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
+
+#Compute SNN graph from KNN
+jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
+
+#Prune SNN graph based on the Jaccard index
+pSNN <- jSNN
+pSNN <- pSNN / max(pSNN)
+
+# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
+#      border=NA,col="grey60",las=1,
+#      xaxs="i",yaxs="i",
+#      main="SNN jaccard",xlim=c(0,1))
+# abline(v=P,col="red",lty=2,xpd=F)
+# 
+# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
+#      border=NA,col="grey60",las=1,
+#      xaxs="i",yaxs="i",
+#      main="correlation coeficient",xlim=c(-1,1))
+# abline(v=0,col="red",lty=1,xpd=F)
+
+pSNN[ pSNN < P ] <- 0
+pSNN[ pSNN > P ] <- cors[pSNN > P]
+pSNN[ pSNN < 0 ] <- 0
+
+g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
+
+mypar(2,2,mar=c(0,0,2,10))
+
+set.seed(1)
+# l <- layout_with_graphopt(g,niter = 1000)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+cl <- igraph::cluster_louvain(g)
+# cl <- igraph::walktrap.community(g)
+metadata$joint_clustering <- cl$membership
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
+       y = par("usr")[c(4)],
+       legend = levels(factor(cl$membership)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+title("Louvain clusters",cex.main=1,font.main=1)
+
+```
+
+**Figure 1b.** Sample embedding of `r k`-SNN graph clustered using Louvain based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).
+
+\clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=4}
+myletters <- letters
+
+mypar(1,2,mar=c(2,15,1,1))
+
+for (m in c("ASV_tissue_V3_normalized_batch_corrected",
+            "ASV_CVL_V3_normalized_batch_corrected",
+            "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+
+  all_microbiome <- datasets_all_samples[[m]]
+
+  datasets <- factor(metadata$joint_clustering)
+  # datasets <- datasets[ colSums(all_microbiome)>0 ]
+  # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
+
+  NN <- min(table(datasets))
+
+  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
+  for(i in levels(datasets)){
+    for(j in rownames(all_microbiome) ){
+
+      set.seed(1)
+      a <- c(sample(all_microbiome[j,datasets == i],NN))
+      set.seed(1)
+      b <- sample(all_microbiome[j,datasets != i],NN)
+
+      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
+      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
+
+      temp <- wilcox.test(x=a, y=b)
+
+      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
+
+      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
+                            c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
+      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
+    }
+  }
+  res <- res[-1,]
+  res$pvalue <- as.numeric(res$pvalue)
+  res$perc.1 <- as.numeric(res$perc.1)
+  res$perc.2 <- as.numeric(res$perc.2)
+  res$perc.diff <- res$perc.1 - res$perc.2
+
+  res$fc <- as.numeric(res$fc)
+  res$FDR <- p.adjust(res$pvalue)
+  # res <- res[order(res$pvalue),]
+  res <- res[abs(res$fc) > fct,]
+  res <- res[abs(res$pvalue) < 0.05,]
+  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
+  dim(res)
+
+  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+
+  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+            cex.row = .8,cex.col = .8,srt = 0,main = m)
+  add_letter(myletters); myletters <- myletters[-1]
+
+  barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
+            cex.axis = .8,srt = 0,main = m)
+  add_letter(myletters); myletters <- myletters[-1]
+
+}
+```
+
+**Figure 2.** Differential bacterial abundance across joint clustered samples and compared across microbiome datasets.
+The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
 
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=14}
+```{r, echo = FALSE,fig.width=10,fig.height=12}
 res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
@@
         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-```
-
-**Figure 5.** Comparisson of microbiome datasets, showing only the significant bacteria.
+
+# **Figure 5.** Comparisson of microbiome datasets, showing only the significant bacteria.
 Samples are ordered by the CVL3 groupings.
 The colors represent their respective bacterial groupings for each dataset.
+
+```
 
@@
                          "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3",
-                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3")
+                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3","joint_clustering")
 
@@
 clean_Ftest <- -log10(clean_Ftest)
-clean_Ftest[clean_Ftest > 4] <- 4
+clean_Ftest[clean_Ftest > 5] <- 5
 # clean_Ftest[ clean_Ftest > 0.05 ] <- 1
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=13}
+```{r, eval=F, echo = FALSE,fig.width=10,fig.height=13}
 myletter <- letters
@@
 }
-```
-
-**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+
+# **Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
 Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
@@
 
+```
+
 \clearpage
@@
 ```{r, echo = FALSE,fig.width=10,fig.height=10}
-cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>10,]))
+
+all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+
+cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))
 
 #Graph construction  
-k <- 10 # k for KNN
-P <- 0.3 # prunning threshold of SNN
+k <- 5 # k for KNN
+P <- 0.2 # prunning threshold of SNN
 
@@
 
-hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="SNN jaccard",xlim=c(0,1))
-abline(v=P,col="red",lty=2,xpd=F)
-
-hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="correlation coeficient",xlim=c(-1,1))
-abline(v=0,col="red",lty=1,xpd=F)
+# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
+#      border=NA,col="grey60",las=1,
+#      xaxs="i",yaxs="i",
+#      main="SNN jaccard",xlim=c(0,1))
+# abline(v=P,col="red",lty=2,xpd=F)
+# 
+# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
+#      border=NA,col="grey60",las=1,
+#      xaxs="i",yaxs="i",
+#      main="correlation coeficient",xlim=c(-1,1))
+# abline(v=0,col="red",lty=1,xpd=F)
 
@@
 
-```{r, echo = FALSE,fig.width  = 10, fig.height=3}
+```{r, echo = FALSE,fig.width=10,fig.height=9}
 mypar(2,2,mar=c(0,0,2,10))
@@
 set.seed(1)
-l <- layout_with_fr(g,niter=3000,start.temp=30)
-
-cl <- igraph::cluster_louvain(g)
+l <- layout_with_fr(g,niter=3000,start.temp=1)
+
+# cl <- igraph::cluster_louvain(g)
+cl <- igraph::walktrap.community(g)
+
 table(cl$membership)
@@
 
-```{r, echo = FALSE,fig.width  = 10, fig.height=10}
-mypar(2,5,mar=c(1,1,2,1))
+**Figure 9a.** Bacterial community embedding of `r k`-SNN graph clustered using walktrap community detection algorithm based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).
+
+```{r, echo = FALSE,fig.width  = 10, fig.height=12}
+mypar(3,6,mar=c(1,1,2,1))
 for(j in sort(unique(cl$membership)) ){
@@
 
+**Figure 9b.** List of bacterial belongin to each community.
+
+```{r, echo = FALSE,fig.width  = 10, fig.height=5}
+taxa <- read.csv2("../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+taxa[is.na(taxa)] <- "NA"
+rownames(cors) %in% rownames(taxa)
+
+mypar(2,2,mar=c(1,10,2,10))
+x <- taxa[match(rownames(cors),rownames(taxa)),2]
+y <- taxa[match(rownames(cors),rownames(taxa)),3]
+plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
+       plot_labels = T, plot_weights = F,order_1_by = "total",
+       order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+       pal=hue_pal()(length(unique(x))))
+
+plot_sankey( data.frame( y[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
+     plot_labels = T, plot_weights = F,order_1_by = "total",
+     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+     pal=hue_pal()(length(unique(y))))
+
+plot_sankey( data.frame( x[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
+     plot_labels = T, plot_weights = F,order_1_by = "total",
+     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+     pal=hue_pal()(length(unique(x))))
+
+```
+
+**Figure 9b.** Taxonomic annotation of bacterial communities.
+
 ```{r, eval=T, echo = FALSE,fig.width=10, fig.height=10}
@@
 
-mypar(3,1,mar=c(1,2,1,0.5))
-sample_order <- order(df$louvain_tissue_V3)
-image(t(module_means_per_dataset[["ASV_tissue_V3_normalized_batch_corrected"]]
-        [ ,sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_tissue_V3)), pch=16,
-        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_tissue_V3)), pch=16,
-        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
-         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
-         labels= rev(sort(unique(cl$membership)) )    )
-
-sample_order <- order(df$louvain_CVL_V3)
-image(t(module_means_per_dataset[["ASV_CVL_V3_normalized_batch_corrected"]]
-        [ ,sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V3)), pch=16,
-        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V3)), pch=16,
-        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
-         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
-         labels= rev(sort(unique(cl$membership)) )    )
-
-sample_order <- order(df$louvain_CVL_V2)
-image(t(module_means_per_dataset[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
-        [ ,sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V2)), pch=16,
-        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V2)), pch=16,
-        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-text( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]), length(sort(unique(cl$membership))) ),
-         (0:(length( sort(unique(cl$membership)) )-1) )/(length(sort(unique(cl$membership)))-1),
-         labels= rev(sort(unique(cl$membership)) )    )
-```
-
-**Figure 5.** Comparisson of microbiome datasets, showing only the distribution of different bacterial communities (1-8).
-Samples are ordered by the CVL3 groupings.
-The colors represent their respective bacterial groupings for each dataset.
-
-```{r, echo = FALSE,fig.width=10,fig.height=13}
-myletter <- letters
-
-for(i in c("ASV_tissue_V3_normalized_batch_corrected",
+# mypar(3,1,mar=c(1,2,1,3))
+# sample_order <- order(df$louvain_tissue_V3)
+# image(t(module_means_per_dataset[["ASV_tissue_V3_normalized_batch_corrected"]]
+#         [ ,sample_order]),
+#       col=heat_pal,axes=F,border=NA,
+#       main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
+#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_tissue_V3)), pch=16,
+#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
+# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
+#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_tissue_V3)), pch=16,
+#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
+# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
+#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+# 
+# 
+# sample_order <- order(df$louvain_CVL_V3)
+# image(t(module_means_per_dataset[["ASV_CVL_V3_normalized_batch_corrected"]]
+#         [ ,sample_order]),
+#       col=heat_pal,axes=F,border=NA,
+#       main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
+#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_CVL_V3)), pch=16,
+#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
+# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
+#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_CVL_V3)), pch=16,
+#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
+# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
+#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+# 
+# 
+# sample_order <- order(df$louvain_CVL_V2)
+# image(t(module_means_per_dataset[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
+#         [ ,sample_order]),
+#       col=heat_pal,axes=F,border=NA,
+#       main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
+# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
+#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_CVL_V2)), pch=16,
+#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
+# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
+#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
+#             length(df$louvain_CVL_V2)), pch=16,
+#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
+# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
+#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+mypar(3,3,mar=c(3,5,2,2))
+
+for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
+  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-
-  cat(paste0("Correlation between RNAseq and ",sub("_norm.*","",i)))
+    temp_group <- as.character(df[,i])
+    temp_group[ is.na(temp_group) ] <- "NA"
+    temp_group[ temp_group == "" ] <- "NA"
+    temp_group <- factor(temp_group)
+
+    barlist(data=module_means_per_dataset[[j]],
+            genes=rownames(module_means_per_dataset[[j]]),
+            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
+            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
+  }
+}
+
+```
+
+```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}
+mypar(7,3,mar=c(3,5,2,2))
+
+for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
+  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+          "ASV_CVL_V3_normalized_batch_corrected",
+          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+    temp_group <- as.character(df[,i])
+    temp_group[ is.na(temp_group) ] <- "NA"
+    temp_group[ temp_group == "" ] <- "NA"
+    temp_group <- factor(temp_group)
+
+    barlist(data=module_means_per_dataset[[j]],
+            genes=rownames(module_means_per_dataset[[j]]),
+            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
+            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
+  }
+}
+
+for( i in c("group_M4_CVLv3","group_M6_CVLv3","group_M9_CVLv3","joint_clustering")){
+  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+          "ASV_CVL_V3_normalized_batch_corrected",
+          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+
+    temp_group <- as.character(metadata[,i])
+    temp_group[ is.na(temp_group) ] <- "NA"
+    temp_group[ temp_group == "" ] <- "NA"
+    temp_group <- factor(temp_group)
+
+    barlist(data=module_means_per_dataset[[j]],
+            genes=rownames(module_means_per_dataset[[j]]),
+            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
+            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
+  }
+}
+```
+
+**Figure 5.** Comparisson of microbiome communities across all datasets and all clustering methods.
+
+```{r, echo = FALSE,fig.width=10,fig.height=8}
+myletter <- letters
+
+NESse_list <- list()
+
+for(i in c("ASV_tissue_V3_normalized_batch_corrected",
+           "ASV_CVL_V3_normalized_batch_corrected")){
+
+  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))
 
@@
   TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
-  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:2000] )
+  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
+  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
+  TRX <- TRX[ top_vars_TRX , ]
+  dim(TRX)
+
+  #filter bac_dataset dataset
+  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
+  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
+  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
+  dim(bac_dataset)
+
+  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
+  # hist(cors,xlim=c(-1,1),breaks = 100)
+  dim(cors)
+
+  bacteria_use <- colnames(cors)
+  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
+  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
+  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
+  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
+    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
+    return(res)
+  })
+  parallel::stopCluster(cl)
+  names(enrichments) <- paste0( "bac.com.", bacteria_use )
+
+  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
+  pvalues <- t(as.data.frame(pvalues))
+  pvalues <- -log10( pvalues )
+  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
+  # mypar(mar=c(4,15,2,2))
+  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
+
+  # pvalues[ pvalues < -log10(0.05)] <- 0
+  # sort(rowSums(pvalues>0),T)
+  # sort(colSums(pvalues>0),T)
+  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
+  NESes <- t(as.data.frame(NESes))
+  NESes[is.na(NESes)] <- 0
+  NESes[ pvalues < -log10(0.05) ] <- 0
+  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
+
+  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
+  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
+  # NESes <- NESes[o_bacs,o_kegg]
+
+  NESse_list[[i]] <- NESes
+
+  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
+  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
+  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
+  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
+  # add_letter(myletter); myletter <- myletter[-1]
+  # empty_plot()
+  # cat("\\clearpage")
+
+}
+
+common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[colSums(x!=0)>=1] ) ) )
+
+layout(matrix(c(1,2,3,4,4),nrow = 1))
+par(mar=c(6,1,2,0))
+for( i in names(NESse_list)){
+  x <- NESse_list[[i]][,common_pathways]
+  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+  # add_letter(myletter); myletter <- myletter[-1]
+  text(   seq(1,0,length.out = nrow(x)) , par("usr")[c(3)] ,
+      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+}
+empty_plot(xaxs="i",yaxs="i",
+           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
+text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
+      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+cat("\\clearpage")
+
+#Combined representation of the two datasets
+layout(matrix(c(1,2,2,2,2),nrow = 1))
+par(mar=c(6,1,3,0))
+x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
+x[NESse_list[[1]][,common_pathways] == 0] <- 0
+x[NESse_list[[2]][,common_pathways] == 0] <- 0
+x <- x[, colSums(x!=0)>0]
+image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# add_letter(myletter); myletter <- myletter[-1]
+text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+empty_plot(xaxs="i",yaxs="i",
+           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
+text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
+      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+#Combined representation of the two datasets
+layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
+par(mar=c(6,30,3,0))
+x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
+x[NESse_list[[1]][,common_pathways] == 0] <- 0
+x[NESse_list[[2]][,common_pathways] == 0] <- 0
+x <- x[, colSums(x!=0)>0]
+
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+non_zeros <- colSums(TRX)!=0
+TRX <- TRX [ top_vars_TRX , non_zeros ]
+pathway_means <- t(sapply( colnames(x) , function(i){
+  i <- gmt[[i]]
+  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
+  if(ncol(i)>1){
+    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
+    i <- colMeans(i)
+  } else {
+    i <- (i - min(i)) / (max(i) - min(i))
+  }
+  return( i - min(i) )
+} ))
+plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
+        genes=rownames(pathway_means),
+        clustering=metadata$joint_clustering[non_zeros],
+        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
+
+par(mar=c(6,2,3,2))
+image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# add_letter(myletter); myletter <- myletter[-1]
+text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+# empty_plot(xaxs="i",yaxs="i",
+#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
+# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
+      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+```
+
+**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
+Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the Hallmark MSigDB gene annotation database.
+This, in turn, will result in a matrix associating every bacteria with every Hallmark MSigDB process in the tissue.
+The heatmap shows the normalized enrichment score (NES).
+Only enrichments with pvalue below 0.05 are shown.
+
+\clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=8}
+myletter <- letters
+
+NESse_list <- list()
+
+for(i in c("ASV_tissue_V3_normalized_batch_corrected",
+           "ASV_CVL_V3_normalized_batch_corrected")){
+
+  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))
+
+  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+  bac_dataset <- module_means_per_dataset[[i]]
+  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
+  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
+
+  # mypar(4,3,mar=c(4,2,2,2))
+  #filter TRX dataset
+  TRX <- TRX [ , samples_TRX_and_CVL3 ]
+  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
+  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
   # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
@@
   parallel::stopCluster(cl)
-  names(enrichments) <- bacteria_use
+  names(enrichments) <- paste0( "bac.com.", bacteria_use )
 
@@
   NESes[is.na(NESes)] <- 0
-  NESes[ pvalues < -log10(0.1) ] <- 0
-  NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
+  NESes[ pvalues < -log10(0.05) ] <- 0
+  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
 
@@
   # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
-  o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
-
-  NESes <- NESes[o_bacs,o_kegg]
-
-  mypar(1,1,mar=c(12,2,2,12))
-  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
-         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
-  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
-        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
-  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
-        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
-  add_letter(myletter); myletter <- myletter[-1]
+  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
+  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
+  # NESes <- NESes[o_bacs,o_kegg]
+
+  NESse_list[[i]] <- NESes
+
+  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
+  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
+  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
+  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
+  # add_letter(myletter); myletter <- myletter[-1]
   # empty_plot()
+  # cat("\\clearpage")
+
+}
+
+common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[ colSums(x!=0)>=2 ] ) ) )
+
+layout(matrix(c(1,2,3,3,3),nrow = 1))
+par(mar=c(6,1,3,0))
+
+for( i in names(NESse_list)){
+  x <- NESse_list[[i]][,common_pathways]
+  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+  # add_letter(myletter); myletter <- myletter[-1]
+  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+}
+empty_plot(xaxs="i",yaxs="i",
+           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
+text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
+      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
 cat("\\clearpage")
 
-}
+#Combined representation of the two datasets
+layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
+par(mar=c(6,30,3,0))
+x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
+x[NESse_list[[1]][,common_pathways] == 0] <- 0
+x[NESse_list[[2]][,common_pathways] == 0] <- 0
+x <- x[, colSums(x!=0)>0]
+
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+non_zeros <- colSums(TRX)!=0
+TRX <- TRX [ top_vars_TRX , non_zeros ]
+pathway_means <- t(sapply( colnames(x) , function(i){
+  i <- gmt[[i]]
+  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
+  if(ncol(i)>1){
+    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
+    i <- colMeans(i)
+  } else {
+    i <- (i - min(i)) / (max(i) - min(i))
+  }
+  return( i - min(i) )
+} ))
+plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
+        genes=rownames(pathway_means),
+        clustering=metadata$joint_clustering[non_zeros],
+        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
+
+par(mar=c(6,2,3,2))
+image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# add_letter(myletter); myletter <- myletter[-1]
+text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+# empty_plot(xaxs="i",yaxs="i",
+#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
+# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
+      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
 ```
@@
 Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.
+
+\clearpage

### 2020-12-11 reports/rmarkdown/03_integrated_datasets.Rmd
updates with manuscript template

@@
 metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
+
+samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
+datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
+metadata <- metadata[ samples_to_use , ]
 ```
@@
 ```{r, echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
 all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
+dim(all_microbiome)
 
 zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
-zscore[ zscore > 6 ] <- 6
-zscore[ zscore < -6 ] <- -6
+zscore[ zscore > 4 ] <- 4
+zscore[ zscore < -4 ] <- -4
 rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
@@
 #Compute sample-wise correlations
-cors <- cor(zscore, method = "pearson")
+cors <- cor(all_microbiome, method = "pearson")
+adj <- (1-cors)/2
+# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
+
+# cors <- cor(all_microbiome, method = "pearson")
+# adj <- (1-cors)/2
+# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
 
 #Graph construction  
-k <- 5 # k for KNN
-P <- 0.4# prunning threshold of SNN
+k <- 10 # k for KNN
+P <- 0.5# prunning threshold of SNN
 
 #Compute KNN graph from correlation matrix
-knn <- RANN::nn2(cors,k = k)
+knn <- RANN::nn2(adj,k = k)
 # knn <- RANN::nn2(t(zscore),k = k)
@@
 jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
+
+jSNN[1:10,1:10]
+
+k=20
+P=0.1
+snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
+snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
+snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
+
+# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
+for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
+
+for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
+
+snnm[ snnj <= ceiling(k*P) ] <- 0
 
@@
 
-g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
+g <- graph_from_adjacency_matrix(snnm,mode = "undirected",diag = F,weighted = T)
 
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=4}
+```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}
+
+all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+
+datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) ,
+                      rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ,
+                      rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
+
+groups <- factor( rep(metadata$joint_clustering,3) )
+age <- rep(metadata$age,3)
+work_time <- rep(metadata$sexwork_months,3)
+
+datasets <- datasets[ colSums(all_microbiome)>0 ]
+groups <- groups[ colSums(all_microbiome)>0 ]
+age <- age[ colSums(all_microbiome)>0 ]
+all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
+
+df <- data.frame(groups,datasets,age)
+
+library(edgeR)
+design <- model.matrix( ~ datasets * groups + age , data=df )
+rownames(design) <- paste0(datasets,"_",colnames(all_microbiome))
+colnames(all_microbiome) <- rownames(design)
+
+y <- DGEList(counts = round(((2^all_microbiome)-1)*1000) , remove.zeros = T)
+y <- calcNormFactors(y,method = "TMM")
+y <- estimateGLMCommonDisp(y,design)
+y <- estimateGLMTagwiseDisp(y,design)
+fit <- glmFit(y, design)
+
+#age
+lrt <- glmLRT(fit,coef = 8)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+plot(top$logFC,-log10(top$PValue))
+
+#age
+lrt <- glmLRT(fit,coef = 4:7)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+
+#age
+lrt <- glmLRT(fit,coef = 9:12)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+
+#age
+lrt <- glmLRT(fit,coef = 13:16)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+
+```
+
+```{r, echo = FALSE,fig.width=5,fig.height=5}
+x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
+xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
+xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
+
+mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
+mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
+mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
+
+mypar(3,1,mar=c(2,6,2,2))
+barlist( data = rbind(  BV=mL ,
+                        counts_16S=xL),
+         main = "Lactobacillus",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+barlist( data = rbind(  BV=mM ,
+                        counts_16S=xM),
+         main = "Monbilicus",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+barlist( data = rbind(  BV=mG ,
+                        counts_16S=xG),
+         main = "Gardanerella",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+```
+
+\clearpage
+
+```{r, eval=F,echo = FALSE,fig.width=10,fig.height=4}
 myletters <- letters
@@
   # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
   NN <- min(table(datasets))
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=9}
+```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
@@
                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))
+datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
 datasets <- datasets[ colSums(all_microbiome)>0 ]
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=9}
+```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- factor(c( rep("CVL_v3",121) , rep("CVL_v2",121) ))
+datasets <- factor(c( rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
 datasets <- datasets[ colSums(all_microbiome)>0 ]
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=9}
+```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
-datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) ))
+datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
 datasets <- datasets[ colSums(all_microbiome)>0 ]
@@
 names(groupings) <- names(datasets_all_samples)[-1]
+all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))
 
 groupings_all_samples <- lapply(groupings,function(x){
-  df <- data.frame(group=rep("NA",nrow(metadata)), row.names = rownames(metadata),stringsAsFactors = F)
+  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
   df[rownames(x),1] <- as.character(x[,1])
@@
 colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
-metadata <- cbind(metadata,df)
+metadata <- cbind(metadata,df[rownames(metadata),])
 ```
@@
                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) )
-datasets <- paste0(datasets, "_",c(metadata$louvain_tissue_V3,metadata$louvain_CVL_V3,metadata$louvain_CVL_V2))
-
-datasets <- factor( datasets[ colSums(all_microbiome)>0 ])
+datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
+
+datasets <- datasets[ colSums(all_microbiome)>0 ]
 all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
@@
 
-```{r, echo = FALSE,fig.width=10,fig.height=10}
-mypar(2,2)
-group_means <- sapply( levels(datasets), function(x) rowMeans( all_microbiome[,datasets==x] ) )
-cors <- cor(group_means)
-h <- hclust( as.dist( (1-cors)/2 ),method = "ward.D2" )
-o <- h$order
-
-par(mar=c(6,6,2,2))
-image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
-      main="correlation matrix",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(h$labels)),
-        rev((0:(length(h$labels)-1) )/(length(h$labels)-1)),
-            labels = h$labels[o], srt = 0, adj = c(1,.5), xpd = TRUE, cex=1)
-text(  (0:(length(h$labels)-1) )/(length(h$labels)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(h$labels)),
-            labels = h$labels[o], srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-add_letter("a")
-
-par(mar=c(6,4,2,2))
-plot(as.dendrogram(h),las=1,ylab="height")
-add_letter("b")
-```
-
-**Figure 3.** Comparrisson among patient groups across datasets. **(a)** Correlation matrix across sample groups. **(b)** Hierachical clustering of sample groups.
-
-```{r, echo = FALSE,fig.width=10,fig.height=12}
+```{r, eval=F,echo = FALSE,fig.width=10,fig.height=12}
 res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
@@
                          "CT_v2","CT_v3","regular_partner",
-                         "vaginal_douching_v2","vaginal_douching_v3",
-                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3",
-                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3","joint_clustering")
+                         "vaginal_douching_v2","vaginal_douching_v3","joint_clustering")
 
@@
   text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
-  for(j in c("louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")){
+  for(j in c("joint_clustering")){
     x <- as.character(metadata[,i])
@@
   enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
+    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T))
     return(res)
@@
 
-cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))
+cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=5,]))
 
@@
 k <- 5 # k for KNN
-P <- 0.2 # prunning threshold of SNN
+P <- 0.1 # prunning threshold of SNN
 
@@
 for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
+
+# k=10
+# P=0.1
+snn <- dbscan::sNN(x = as.dist( (1-cors)/2 ), k = k, jp = FALSE, approx = TRUE)
+snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
+snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
+
+# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
+for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
+for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
+snnm[ snnj <= ceiling(k*P) ] <- 0
 
@@
 cl <- igraph::walktrap.community(g)
-
+# 
 table(cl$membership)
@@
 }
+write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))
+
 ```
@@
 
-```{r, eval=T, echo = FALSE,fig.width=10, fig.height=10}
-sample_order <- order(df$louvain_CVL_V3)
+```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}
 
@@
 })
+
+mypar(4,3,mar=c(3,5,2,2))
 
@@
 ```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}
-mypar(7,3,mar=c(3,5,2,2))
-
-for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
-  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-          "ASV_CVL_V3_normalized_batch_corrected",
-          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-    temp_group <- as.character(df[,i])
-    temp_group[ is.na(temp_group) ] <- "NA"
-    temp_group[ temp_group == "" ] <- "NA"
-    temp_group <- factor(temp_group)
-
-    barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),
-            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
-            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
-  }
-}
+
+module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
+  return( rowsum( x[rownames(cors),] , group = cl$membership ) / c(table(cl$membership)) )
+})
+
+mypar(4,3,mar=c(3,5,2,2))
+
+# for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
+#   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+#           "ASV_CVL_V3_normalized_batch_corrected",
+#           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+#     temp_group <- as.character(df[,i])
+#     temp_group[ is.na(temp_group) ] <- "NA"
+#     temp_group[ temp_group == "" ] <- "NA"
+#     temp_group <- factor(temp_group)
+#     
+#     barlist(data=module_means_per_dataset[[j]],
+#             genes=rownames(module_means_per_dataset[[j]]),
+#             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
+#             clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
+#   }
+# }
 
@@
   enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
+    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
     return(res)
@@
 text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
 empty_plot(xaxs="i",yaxs="i",
@@
 text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
-      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+      labels = rev(sub("HALLMARK_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
 
@@
 text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
 # empty_plot(xaxs="i",yaxs="i",
@@
 text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
-      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+      labels = rev(sub("KEGG_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
 
@@
 \clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=8}
+write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
+```

### 2020-12-11 reports/rmarkdown/manuscript/manuscript_template.Rmd
updates with manuscript template

---
title: "MANUSCRIPT TITLE"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
# library(ggthemes)
library(dplyr)
library(stringr)
remotes::install_github("czarnewski/niceRplots")
library(niceRplots)
```

Abstract
===============================================================================

Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract

\clearpage

Introduction
================================================================================

Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction ([@Johnston1972]), Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction **Introduction Introduction** *Introduction* Introduction ([@Darwin1859]^,^[@Bumpus1898]) .

Problem / question to answer

\clearpage

Results
===============================================================================

```{r, echo = FALSE, eval = TRUE}
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
patient_groups <- metadata$joint_clustering
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
```

**Joint analysis of vaginal microbiome reveals XXXXXXXXXXXXXX**

To understand the longitudinal and tissue-specific microbiome profile in vaginal samples, `r nrow(metadata)` adult female sex workers were enrolled in [...]. Among those, `r sum(metadata$HIVstatus=="pos")` were previously tested positive for HIV during the cohort's sampling procedure. [Describe here what was done and when, which samples, which tissues].

To be able to better undertand the differences in microbiome profile across all datasets collected, we  performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see "Methods" section for details). A total of `r length(unique(bac_communities))` bacterial communities were identified.

Noticebly, bacterial community `r bac_communities["Lactobacillus iners"]` consisted only of Lactobacillus species (*`r names(bac_communities)[bac_communities %in% bac_communities["Lactobacillus iners"]]`*).

Patients were thus subdivided into `r length(unique(patient_groups))` groups,

\clearpage

Discussion
================================================================================

I have analysed data collected by Herman Bumpus [@Bumpus1898] on the relationship between sparrow (*Passer domesticus*) total length and surival following an unusually severe storm. I found that sparrows that died in the storm were longer than sparrows that survived, which suggests that higher sparrow body length decreased survival. Of course, it is not possible to definitively conclude a causal relationship between any aspect of body size and sparrow survival<!--- BD Note: maybe explain this better --->, and even the available data collected by Bumpus would permit a more thoughtful analysis than that conducted in this study (see [Appendix Table 1](#appendix)).

<!---

Here is one way to add some more detailed comments into the manuscript itself, though this can also be done within the text (see above).

--->

Overall, this document demonstrates how high quality, professional looking documents can be written using Rmarkdown. The [underlying code](https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/ms.Rmd) for this manuscript is publicly available, along with [accompanying notes](https://stirlingcodingclub.github.io/Manuscripts_in_Rmarkdown/Rmarkdown_notes.html) to understand how it was written. By using Rmarkdown to write manuscripts, authors can more easily use version control (e.g., git) throughout the writing process. The ability to easily integrate citations though BibTeX, LaTeX tools, and dynamic R code can also make writing much more efficient and more enjoyable. Further, obtaining the benefits of using Rmarkdown does not need to come with the cost of isolating colleagues who prefer to work with Word or LaTeX because Rmarkdown can easily be converted to these formats (in the case of Word, with the push of a button). By learning all of the tools used in this manuscript, readers should have all of the necessary knowledge to get started writing and collaborating in Rmarkdown.

\clearpage

Methods
================================================================================

Bumpus focused his study on the House Sparrow (*Passer domesticus*; see Figure 1), which has a very wide global distribution. It is native to Europe and Asia, but not the Americas where Bumpus collected his original study [@Bumpus1898]. In addition to measuring total length and survival for 136 sparrows, Bumpus measured sparrow sex, wingspan, and mass, and also the length of each sparrow's head, humerus, tibiotarsus, skull, and sternum. While modern ornithologists believe that the total body length measurement that I will use today is subject to high observational error [@Johnston1972], it will be more than sufficient for demonstrating Rmarkdown.

<!--- Note that the image below is in the public domain --->
<!--- https://commons.wikimedia.org/wiki/File:Lto-tpbo-passer-domesticus-05.JPG --->
<center>

![*Passer domesticus*](images/sparrow.jpg)

</center>

I performed an independent two-sample student's t-test on sparrow total body length to test whether or not sparrows that died in the 1898 storm were larger than sparrows that survived. I assume that both groups of sparrows (dead and living) have equal variances, so the test statistic $t$ is calculated as follows,

$$t = \frac{\bar{X}_{1} - \bar{X}_{2}} {s_{p} \times \sqrt{\frac{1}{n_{1}} + \frac{1}{n_{2}}}}.$$

In the above, $\bar{X}_{1}$ and $\bar{X}_{2}$ are the mean of the samples of sparrows that died and lived, respectively. Similarly, $n_{1}$ and $n_{2}$ are the sample sizes of sparrows that died and lived, and $s_{p}$ is the pooled sample mean, which is calculated as follows,

$$s_{p} = \sqrt{\frac{s^{2}_{X_{1}} + s^{2}_{X_{2}}}{2}}.$$

In the above, the $s^{2}_{X_{1}}$ and $s^{2}_{X_{2}}$ are the sample standard deviations for sparrows that died and lived, respectively. I conduceted the two sample t-test using the `t.test` function in R.

\clearpage

References
================================================================================

<div id="refs"></div>

\clearpage

<a name="appendix">Appendix Table 1</a>
================================================================================

An example table is shown below, which includes all of the variables collected by @Bumpus1898 for the first 10 measured sparrows. The full data set can be found online in [GitHub](https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/data/Bumpus_data.csv).

```{r, echo = FALSE, fig.pos='H',fig.show='hold'}
library(knitr)
```

### 2020-12-11 reports/rmarkdown/manuscript/manuscript_template.tex
updates with manuscript template

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={MANUSCRIPT TITLE},
  pdfkeywords={pandoc, r markdown, knitr},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[L]{XXXXXXXXXXX et al.}
\fancyhead[R]{MANUSCRIPT SHORT TITLE}
\usepackage{lineno}
\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\title{MANUSCRIPT TITLE}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A demonstration of Rmarkdown using Herman Bumpus' data}
\author{Author One\textsuperscript{1},
Author Two\textsuperscript{2},
Author Three\textsuperscript{1,2}}
\date{December 04, 2020}

\begin{document}
\maketitle

\setstretch{1.5}
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract

\clearpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction (\textsuperscript{\protect\hyperlink{ref-Johnston1972}{1}}), Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction \textbf{Introduction Introduction} \emph{Introduction} Introduction (\textsuperscript{\protect\hyperlink{ref-Darwin1859}{2}}\textsuperscript{,}\textsuperscript{\protect\hyperlink{ref-Bumpus1898}{3}}) .

Problem / question to answer

\clearpage

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{Joint analysis of vaginal microbiome reveals XXXXXXXXXXXXXX}

To understand the longitudinal and tissue-specific microbiome profile in vaginal samples, 113 adult female sex workers were enrolled in {[}\ldots{]}. Among those, 14 were previously tested positive for HIV during the cohort's sampling procedure. {[}Describe here what was done and when, which samples, which tissues{]}.

To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis at both patient and baterial level (see ``Methods'' section for details). Patients were thus subdivided into 7 groups, while 17 bacterial communities were identified.

Noticebly, bacterial community 14 presented several Lactobacillus species, such as \emph{Lactobacillus coleohominis, Lactobacillus crispatus/acidophilus, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus reuteri/oris/frumenti/antri, other Lactobacillales}.

\clearpage

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

I have analysed data collected by Herman Bumpus\textsuperscript{\protect\hyperlink{ref-Bumpus1898}{3}} on the relationship between sparrow (\emph{Passer domesticus}) total length and surival following an unusually severe storm. I found that sparrows that died in the storm were longer than sparrows that survived, which suggests that higher sparrow body length decreased survival. Of course, it is not possible to definitively conclude a causal relationship between any aspect of body size and sparrow survival, and even the available data collected by Bumpus would permit a more thoughtful analysis than that conducted in this study (see \protect\hyperlink{appendix}{Appendix Table 1}).

Overall, this document demonstrates how high quality, professional looking documents can be written using Rmarkdown. The \href{https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/ms.Rmd}{underlying code} for this manuscript is publicly available, along with \href{https://stirlingcodingclub.github.io/Manuscripts_in_Rmarkdown/Rmarkdown_notes.html}{accompanying notes} to understand how it was written. By using Rmarkdown to write manuscripts, authors can more easily use version control (e.g., git) throughout the writing process. The ability to easily integrate citations though BibTeX, LaTeX tools, and dynamic R code can also make writing much more efficient and more enjoyable. Further, obtaining the benefits of using Rmarkdown does not need to come with the cost of isolating colleagues who prefer to work with Word or LaTeX because Rmarkdown can easily be converted to these formats (in the case of Word, with the push of a button). By learning all of the tools used in this manuscript, readers should have all of the necessary knowledge to get started writing and collaborating in Rmarkdown.

\clearpage

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Bumpus focused his study on the House Sparrow (\emph{Passer domesticus}; see Figure 1), which has a very wide global distribution. It is native to Europe and Asia, but not the Americas where Bumpus collected his original study.\textsuperscript{\protect\hyperlink{ref-Bumpus1898}{3}} In addition to measuring total length and survival for 136 sparrows, Bumpus measured sparrow sex, wingspan, and mass, and also the length of each sparrow's head, humerus, tibiotarsus, skull, and sternum. While modern ornithologists believe that the total body length measurement that I will use today is subject to high observational error,\textsuperscript{\protect\hyperlink{ref-Johnston1972}{1}} it will be more than sufficient for demonstrating Rmarkdown.

\begin{figure}
\centering
\includegraphics{images/sparrow.jpg}
\caption{\emph{Passer domesticus}}
\end{figure}

I performed an independent two-sample student's t-test on sparrow total body length to test whether or not sparrows that died in the 1898 storm were larger than sparrows that survived. I assume that both groups of sparrows (dead and living) have equal variances, so the test statistic \(t\) is calculated as follows,

\[t = \frac{\bar{X}_{1} - \bar{X}_{2}} {s_{p} \times \sqrt{\frac{1}{n_{1}} + \frac{1}{n_{2}}}}.\]

In the above, \(\bar{X}_{1}\) and \(\bar{X}_{2}\) are the mean of the samples of sparrows that died and lived, respectively. Similarly, \(n_{1}\) and \(n_{2}\) are the sample sizes of sparrows that died and lived, and \(s_{p}\) is the pooled sample mean, which is calculated as follows,

\[s_{p} = \sqrt{\frac{s^{2}_{X_{1}} + s^{2}_{X_{2}}}{2}}.\]

In the above, the \(s^{2}_{X_{1}}\) and \(s^{2}_{X_{2}}\) are the sample standard deviations for sparrows that died and lived, respectively. I conduceted the two sample t-test using the \texttt{t.test} function in R.

\clearpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Johnston1972}{}%
1. Johnston, R. F., Niles, D. M. \& Rohwer, S. A. Hermon bumpus and natural selection in the house sparrow \emph{Passer domesticus}. \emph{Evolution} \textbf{26}, 20--31 (1972).

\leavevmode\hypertarget{ref-Darwin1859}{}%
2. Darwin, C. \emph{The origin of species}. 495 (Penguin, 1859).

\leavevmode\hypertarget{ref-Bumpus1898}{}%
3. Bumpus, H. C. Eleventh lecture. The elimination of the unfit as illustrated by the introduced sparrow, \emph{Passer domesticus}. (A fourth contribution to the study of variation.). \emph{Biological Lectures: Woods Hole Marine Biological Laboratory} 209--225 (1898).
\end{cslreferences}

\clearpage

\hypertarget{appendix-table-1}{%
\section{Appendix Table 1}\label{appendix-table-1}}

An example table is shown below, which includes all of the variables collected by \protect\hyperlink{ref-Bumpus1898}{3} for the first 10 measured sparrows. The full data set can be found online in \href{https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/data/Bumpus_data.csv}{GitHub}.

\end{document}

### 2020-12-17 reports/rmarkdown/03_integrated_datasets.Rmd
adds figure1 in manuscript

@@
 metadata <- metadata[ samples_to_use , ]
+saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
 ```
@@
 all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+# all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
 all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
@@
 rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
-
+mypar(mar=c(2,15,2,2))
+top_var <- apply(all_microbiome,1,var)
+top_var_t <- sort(top_var,decreasing = T)[100]
+is_top_var <- top_var > top_var_t 
+barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
 #Compute sample-wise correlations
-cors <- cor(all_microbiome, method = "pearson")
+cors <- cor(all_microbiome[is_top_var,], method = "pearson")
 adj <- (1-cors)/2
@@
 #Graph construction  
-k <- 10 # k for KNN
-P <- 0.5# prunning threshold of SNN
+k <- 15 # k for KNN
+P <- 0.1# prunning threshold of SNN
 
@@
 
-k=20
-P=0.1
+# k=10
+# P=0.1
 snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
@@
 
+N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
+P <- t(nn) %*% nn
+B <- N - nn
+C <- t(B) %*% B
+Q <- nrow(nn)*N - C
+J <- P/Q
+J[is.nan(J)] <- 0
+J[J < 0.2] <- 0
+colnames(J) <- colnames(all_microbiome)
+rownames(J) <- colnames(all_microbiome)
+# image(J[,nrow(nn):1],col = colorRampPalette(c("grey95","black"))(90),axes=F )
+# set.seed(1)
+# g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
+# plot( g , vertex.label.cex=0.000001 , vertex.color = "grey" , 
+#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
+#       vertex.size=10 )
+
 pSNN[ pSNN < P ] <- 0
@@
 
-g <- graph_from_adjacency_matrix(snnm,mode = "undirected",diag = F,weighted = T)
-
-mypar(2,2,mar=c(0,0,2,10))
+g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
+write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)
+
+mypar(3,3,mar=c(0,0,2,0))
 
@@
 # l <- layout_with_graphopt(g,niter = 1000)
+set.seed(1)
 l <- layout_with_fr(g,niter=3000,start.temp=30)
@@
 cl <- igraph::cluster_louvain(g)
-# cl <- igraph::walktrap.community(g)
+cl <- igraph::walktrap.community(g)
 metadata$joint_clustering <- cl$membership
+set.seed(1)
+U <- uwot::umap(adj,n_neighbors = 30)
+h <- hclust(as.dist(adj),method = "ward.D2")
+# metadata$joint_clustering <- cutree(h,k = 5)
+write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
 
@@
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
+      vertex.size=10,main="FR embedding\nlouvain",
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups",
        legend = levels(factor(cl$membership)),
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
-title("Louvain clusters",cex.main=1,font.main=1)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,main="FR embedding",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
+       legend = levels(factor(metadata$group_M4_CVLv3)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+# title("Louvain clusters",cex.main=1,font.main=1)
 
@@
 **Figure 1b.** Sample embedding of `r k`-SNN graph clustered using Louvain based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).
+
+```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}
+
+```
 
@@
 clean_Ftest <- -log10(clean_Ftest)
-clean_Ftest[clean_Ftest > 5] <- 5
+clean_Ftest[clean_Ftest > 5.5] <- 5.5
 # clean_Ftest[ clean_Ftest > 0.05 ] <- 1
+
+write.csv2(clean_Ftest , "../../results/metadata_association_results.csv",row.names = T)
 
@@
 #Graph construction 
-k <- 5 # k for KNN
-P <- 0.1 # prunning threshold of SNN
-
-knn <- RANN::nn2(cors,k = k)
-# knn <- RANN::nn2(t(zscore),k = k)
+k_nearest_neighbours <- 10 # k for KNN
+snn_threshold <- 0.1 # prunning threshold of SNN
+
+#Compute kNN graph
+knn <- RANN::nn2( (1-cors)/2 ,k = k_nearest_neighbours)
 nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
@@
 
-# k=10
-# P=0.1
-snn <- dbscan::sNN(x = as.dist( (1-cors)/2 ), k = k, jp = FALSE, approx = TRUE)
-snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
-snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
-
-# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
-for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
-for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
-snnm[ snnj <= ceiling(k*P) ] <- 0
-
-#Compute SNN graph from KNN
-jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
-
-#Prune SNN graph based on the Jaccard index
-pSNN <- jSNN
-pSNN <- pSNN / max(pSNN)
-
-# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
-#      border=NA,col="grey60",las=1,
-#      xaxs="i",yaxs="i",
-#      main="SNN jaccard",xlim=c(0,1))
-# abline(v=P,col="red",lty=2,xpd=F)
-# 
-# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
-#      border=NA,col="grey60",las=1,
-#      xaxs="i",yaxs="i",
-#      main="correlation coeficient",xlim=c(-1,1))
-# abline(v=0,col="red",lty=1,xpd=F)
-
-pSNN[ pSNN < P ] <- 0
-pSNN[ pSNN > P ] <- cors[pSNN > P]
-pSNN[ pSNN < 0 ] <- 0
-
-g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
-```
-
-# Visualise the data
-
-```{r, echo = FALSE,fig.width=10,fig.height=9}
-mypar(2,2,mar=c(0,0,2,10))
-
+#Calculate SNN and prune based on Jaccard index
+N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
+P <- t(nn) %*% nn
+B <- N - nn
+C <- t(B) %*% B
+Q <- nrow(nn)*N - C
+J <- P/Q
+J[is.nan(J)] <- 0
+J[J < snn_threshold] <- 0
+rownames(J) <- rownames(cors)
+colnames(J) <- colnames(cors)
+
+#Build a graph object, perform clustering and define the force-directed layout
+g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
+write.csv(J,"../../results/bacteria_SNN_graph.csv",row.names = T)
+
+cl <- igraph::cluster_louvain(g)
+# cl <- igraph::walktrap.community(g)
 set.seed(1)
@@
 
-# cl <- igraph::cluster_louvain(g)
-cl <- igraph::walktrap.community(g)
-# 
-table(cl$membership)
+#Define layout with UMAP
+mU <- uwot::umap((1-cors)/2, n_neighbors = 10)
+
+#Plot
+mypar(3,3,mar=c(0,0,2,0))
 
@@
     edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
+    vertex.size=10, main="graph layout FR",
     edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
+    vertex.size=10, main="UMAP",
+    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+par(mar=c(0,0,0,0))
+empty_plot()
+legend("topleft",title = "bact. communities",
      legend = levels(factor(cl$membership)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
-title("Louvain clusters",cex.main=1,font.main=1)
+     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
+
 ```
@@
     barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),
+            genes=rownames(module_means_per_dataset[[j]]),col = pal,
             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
@@
     barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),
+            genes=rownames(module_means_per_dataset[[j]]),col=pal,
             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),

### 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds figure1 in manuscript

@@
 ---
-title: "MANUSCRIPT TITLE"
+title: "Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue"
 subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
@@
 options(tinytex.verbose = TRUE)
-knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
-knitr::opts_chunk$set(out.width = '100%', dpi=300)
+knitr::opts_chunk$set(out.width = '100%',
+                      dpi=600,
+                      fig.width=10,
+                      echo = FALSE,
+                      # fig.pos = 'p', # Places figures on their own pages
+                      eval = TRUE, 
+                      results = 'hide',
+                      warning = FALSE,
+                      message = FALSE )
 
@@
 library(ggplot2)
-# library(ggthemes)
+library(igraph)
 library(dplyr)
 library(stringr)
+library(rafalib)
 remotes::install_github("czarnewski/niceRplots")
 library(niceRplots)
+
+#color palettes
+pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
+         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
+heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
+cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
 ```
@@
 ```{r, echo = FALSE, eval = TRUE}
+# Load reasult from the analysis
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
-patient_groups <- metadata$joint_clustering
+patient_groups <- metadata$group_M4_CVLv3
 bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
+SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
+SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
+clean_Ftest <- read.csv2("../../../results/metadata_association_results.csv",row.names = 1)
+datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
+
 ```
 
-**Joint analysis of vaginal microbiome reveals XXXXXXXXXXXXXX**
+**Joint analysis of vaginal microbiome reveals distinct patient subgroups**
 
@@
 Patients were thus subdivided into `r length(unique(patient_groups))` groups,
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
+
+\clearpage
+
+**Indentification of bacterial communities metabolic processes linked to Lactobacilli**
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
+
+\clearpage
+
+**Indentification of bacterial communities metabolic processes linked to Lactobacilli**
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
 
@@
 
-Bumpus focused his study on the House Sparrow (*Passer domesticus*; see Figure 1), which has a very wide global distribution.
-It is native to Europe and Asia, but not the Americas where Bumpus collected his original study [@Bumpus1898].
-In addition to measuring total length and survival for 136 sparrows, Bumpus measured sparrow sex, wingspan, and mass, and also the length of each sparrow's head, humerus, tibiotarsus, skull, and sternum.
-While modern ornithologists believe that the total body length measurement that I will use today is subject to high observational error [@Johnston1972], it will be more than sufficient for demonstrating Rmarkdown.
-
-<!--- Note that the image below is in the public domain --->
-<!--- https://commons.wikimedia.org/wiki/File:Lto-tpbo-passer-domesticus-05.JPG --->
-<center>
-
-![*Passer domesticus*](images/sparrow.jpg)
-
-</center>
-
-I performed an independent two-sample student's t-test on sparrow total body length to test whether or not sparrows that died in the 1898 storm were larger than sparrows that survived.
-I assume that both groups of sparrows (dead and living) have equal variances, so the test statistic $t$ is calculated as follows,
-
-$$t = \frac{\bar{X}_{1} - \bar{X}_{2}} {s_{p} \times \sqrt{\frac{1}{n_{1}} + \frac{1}{n_{2}}}}.$$
-
-In the above, $\bar{X}_{1}$ and $\bar{X}_{2}$ are the mean of the samples of sparrows that died and lived, respectively.
-Similarly, $n_{1}$ and $n_{2}$ are the sample sizes of sparrows that died and lived, and $s_{p}$ is the pooled sample mean, which is calculated as follows,
-
-$$s_{p} = \sqrt{\frac{s^{2}_{X_{1}} + s^{2}_{X_{2}}}{2}}.$$
-
-In the above, the $s^{2}_{X_{1}}$ and $s^{2}_{X_{2}}$ are the sample standard deviations for sparrows that died and lived, respectively.
-I conduceted the two sample t-test using the `t.test` function in R.
-
 \clearpage
@@
 ```
+
+\clearpage
+
+FIGURES (MAIN)
+================================================================================
+
+## Figure 1
+
+```{r, echo = FALSE, eval = TRUE , fig.asp=6/10}
+# Define Layout and set labels
+layout(matrix(c(1,1,1,1,      3,3,3,3,          4,4,4,4,
+                2,2,2,2,      3,3,3,3,          5,5,5,5,
+                6,6,6,6,      9,9,9,9,          9,9,9,9,
+                7,7,7,7,     10,10,10,10,     10,10,10,10,
+                8,8,8,8,     10,10,10,10,     10,10,10,10 ) , ncol= 12,byrow = T),
+       widths = 1,heights = c(1,1,0.5,0.5,0.5))
+# layout.show(29)
+figlabels <- letters
+
+#################################
+# COHORT SCHEMATIC ILLUSTRATION #
+#################################
+par(mar=c(0,0,0,0))
+fig1a <- magick::image_read("schematics.pdf",density = 900)
+plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
+rasterImage(fig1a, 0, 0, 1, 1)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+####################
+# PATIENT GROUPING #
+####################
+g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+par(mar=c(0,0,0,0))
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$joint_clustering)] ,
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
+      vertex.size=10,main="", 
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+legend("topright",title.adj = 0,
+       legend = levels(factor(metadata$joint_clustering)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+##############################
+# METADATA CROSS-ASSOCIATION #
+##############################
+par(mar=c(12,1,1,12))
+image( t(clean_Ftest [nrow(clean_Ftest):1,]),axes=F,
+       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
+text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(clean_Ftest)) ,
+       labels = rownames(clean_Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+text( seq(0,1,length.out = nrow(clean_Ftest)) , par("usr")[3]-.02 ,
+      labels = rownames(clean_Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+par(mar=c(0,0,0,0))
+empty_plot(frame = T)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+empty_plot(frame = T)
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+par(mar=c(1,5,2,1))
+x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
+xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
+xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
+
+mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
+mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
+mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
+
+barlist( data = rbind(  BV=mL ,
+                        counts_16S=xL),
+         main = "Lactobacillus",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+barlist( data = rbind(  BV=mM ,
+                        counts_16S=xM),
+         main = "Monbilicus",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+barlist( data = rbind(  BV=mG ,
+                        counts_16S=xG),
+         main = "Gardanerella",
+         genes = c("BV","counts_16S"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+par(mar=c(1,1,1,1))
+
+shann <- vegan::diversity(t(x),index = "shannon")
+shann <- t(as.matrix(shann))
+rownames(shann) <- "Shannon"
+barlist( data = shann,
+         main = "Shannon",
+         labels = "",
+         genes = c("Shannon"),
+         clustering = metadata$joint_clustering,
+         draw_mean_lines=F,col = pal)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+par(mar=c(4,4,1,1))
+temp <- t(t(2^x-1)/colSums(2^x-1))
+taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+taxa[is.na(taxa)] <- "NA"
+taxa <- taxa[rownames(temp),]
+temp <- rowsum(temp , taxa$Class)
+o <- order( rowSums(temp) , decreasing = F)
+sample_o <- order( as.numeric(metadata$group_M4_CVLv3)+(shann/max(shann)/2) )
+temp <- temp[o,sample_o]*100
+barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i", col = pal[factor(rownames(temp))] )
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+```
+
+**Figure 1. Identification of patient groups.**
+(**a**) Schematic representation of ########## .
+(**b**) Schematic representation of ########## .
+(**c**) Schematic representation of ########## .
+(**d**) Schematic representation of ########## .
+
+\clearpage
+
+## Figure 2
+
+```{r, echo = FALSE, eval = TRUE}
+# Define Layout and set labels
+layout(matrix(c(1) , ncol= 1,byrow = T),
+       heights = c(1))
+figlabels <- letters
+
+```
+
+**Figure 2. Identification and characterization of vaginal bacterial communities.**
+(**a**) Schematic representation of ########## .
+(**b**) Schematic representation of ########## .
+(**c**) Schematic representation of ########## .
+(**d**) Schematic representation of ########## .
+
+\clearpage
+
+## Figure 3
+
+```{r, echo = FALSE, eval = TRUE}
+# Define Layout and set labels
+layout(matrix(c(1) , ncol= 1,byrow = T),
+       heights = c(1))
+figlabels <- letters
+
+```
+
+**Figure 1. Identification of patient groups.**
+(**a**) Schematic representation of ########## .
+(**b**) Schematic representation of ########## .
+(**c**) Schematic representation of ########## .
+(**d**) Schematic representation of ########## .
+
+\clearpage
+
+FIGURES (SUPPL)
+================================================================================
+
+## Figure S1
+
+## Figure S2
+
+## Figure S3
+
+## Figure S4
+
+TABLES (MAIN)
+================================================================================
+
+## Table 1
+
+## Table 2
+
+## Table 3
+
+TABLES (SUPPL)
+================================================================================
+
+## Table S1
+
+## Table S2
+
+## Table S3

### 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.tex
adds figure1 in manuscript

@@
 \hypersetup{
-  pdftitle={MANUSCRIPT TITLE},
+  pdftitle={Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue},
   pdfkeywords={pandoc, r markdown, knitr},
@@
 
-\title{MANUSCRIPT TITLE}
+\title{Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue}
 \usepackage{etoolbox}
@@
 Author Three\textsuperscript{1,2}}
-\date{December 04, 2020}
+\date{December 16, 2020}
 
@@
 
-\textbf{Joint analysis of vaginal microbiome reveals XXXXXXXXXXXXXX}
+\textbf{Joint analysis of vaginal microbiome reveals distinct patient subgroups}
 
@@
 
-To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis at both patient and baterial level (see ``Methods'' section for details).
-Patients were thus subdivided into 7 groups, while 17 bacterial communities were identified.
+To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see ``Methods'' section for details).
+A total of 11 bacterial communities were identified.
 
-Noticebly, bacterial community 14 presented several Lactobacillus species, such as \emph{Lactobacillus coleohominis, Lactobacillus crispatus/acidophilus, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus reuteri/oris/frumenti/antri, other Lactobacillales}.
+Noticebly, bacterial community 11 consisted only of Lactobacillus species (\emph{Anaerococcus hydrogenalis, Atopobium deltae, Bifidobacterium adolescentis, Bifidobacterium bifidum, Bifidobacterium longum/breve, Brevibacterium paucivorans, Brevibacterium ravenspurgense, Campylobacter hominis, Chlamydia trachomatis, Corynebacterium aurimucosum, Corynebacterium genitalium, Corynebacterium imitans, Corynebacterium jeikeium, Corynebacterium pseudogenitalium, Corynebacterium simulans/striatum, Corynebacterium sp/amycolatum, Corynebacterium spheniscorum, Corynebacterium sundsvallense/thomssenii, Dermabacter hominis, Dorea formicigenerans, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/dispar/durans, Gemmiger formicilis, Lactobacillus coleohominis, Lactobacillus crispatus, Lactobacillus crispatus/acidophilus, Lactobacillus gasseri/johnsonii/taiwanensis, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus mucosae, Lactobacillus reuteri, Lactobacillus reuteri/oris/frumenti/antri, Lactobacillus salivarius, Megasphaera elsdenii, other Bacilli, other Bifidobacteriaceae, other Clostridia, other Corynebacterium, other Ezakiella, other Gemella, other Gordonia, other Lactobacillales, other Lactobacillus, other Listeria, other Mobiluncus, other Mogibacterium, other Murdochiella, other Nesterenkonia, other Paraeggerthella, other Rhizobium, Peptoniphilus asaccharolyticus, Prevotella denticola, Staphylococcus saccharolyticus/capitis/hominis/caprae/lugdunensis/epidermidis/simiae, Streptococcus dysgalactiae, Streptococcus lutetiensis/equinus, Ureaplasma parvum, Brevundimonas mediterranea/intermedia/poindexterae, Corynebacterium appendicis/imitans, Corynebacterium freneyi/sp/amycolatum, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/saccharolyticus/ratti/dispar/gallinarum/durans, other Bacteroides, other Gammaproteobacteria}).
+
+Patients were thus subdivided into 6 groups,
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
+
+\clearpage
+
+\textbf{Indentification of bacterial communities metabolic processes linked to Lactobacilli}
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
+
+\clearpage
+
+\textbf{Indentification of bacterial communities metabolic processes linked to Lactobacilli}
+
+Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results
 
@@
 \section{Methods}\label{methods}}
-
-Bumpus focused his study on the House Sparrow (\emph{Passer domesticus}; see Figure 1), which has a very wide global distribution.
-It is native to Europe and Asia, but not the Americas where Bumpus collected his original study.\textsuperscript{\protect\hyperlink{ref-Bumpus1898}{3}} In addition to measuring total length and survival for 136 sparrows, Bumpus measured sparrow sex, wingspan, and mass, and also the length of each sparrow's head, humerus, tibiotarsus, skull, and sternum.
-While modern ornithologists believe that the total body length measurement that I will use today is subject to high observational error,\textsuperscript{\protect\hyperlink{ref-Johnston1972}{1}} it will be more than sufficient for demonstrating Rmarkdown.
-
-\begin{figure}
-\centering
-\includegraphics{images/sparrow.jpg}
-\caption{\emph{Passer domesticus}}
-\end{figure}
-
-I performed an independent two-sample student's t-test on sparrow total body length to test whether or not sparrows that died in the 1898 storm were larger than sparrows that survived.
-I assume that both groups of sparrows (dead and living) have equal variances, so the test statistic \(t\) is calculated as follows,
-
-\[t = \frac{\bar{X}_{1} - \bar{X}_{2}} {s_{p} \times \sqrt{\frac{1}{n_{1}} + \frac{1}{n_{2}}}}.\]
-
-In the above, \(\bar{X}_{1}\) and \(\bar{X}_{2}\) are the mean of the samples of sparrows that died and lived, respectively.
-Similarly, \(n_{1}\) and \(n_{2}\) are the sample sizes of sparrows that died and lived, and \(s_{p}\) is the pooled sample mean, which is calculated as follows,
-
-\[s_{p} = \sqrt{\frac{s^{2}_{X_{1}} + s^{2}_{X_{2}}}{2}}.\]
-
-In the above, the \(s^{2}_{X_{1}}\) and \(s^{2}_{X_{2}}\) are the sample standard deviations for sparrows that died and lived, respectively.
-I conduceted the two sample t-test using the \texttt{t.test} function in R.
 
@@
 
+\clearpage
+
+\hypertarget{figures-main}{%
+\section{FIGURES (MAIN)}\label{figures-main}}
+
+\hypertarget{figure-1}{%
+\subsection{Figure 1}\label{figure-1}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-3-1}
+
+\textbf{Figure 1. Identification of patient groups.}
+(\textbf{a}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{b}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{c}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{d}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+
+\clearpage
+
+\hypertarget{figure-2}{%
+\subsection{Figure 2}\label{figure-2}}
+
+\textbf{Figure 2. Identification and characterization of vaginal bacterial communities.}
+(\textbf{a}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{b}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{c}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{d}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+
+\clearpage
+
+\hypertarget{figure-3}{%
+\subsection{Figure 3}\label{figure-3}}
+
+\textbf{Figure 1. Identification of patient groups.}
+(\textbf{a}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{b}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{c}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+(\textbf{d}) Schematic representation of \#\#\#\#\#\#\#\#\#\# .
+
+\clearpage
+
+\hypertarget{figures-suppl}{%
+\section{FIGURES (SUPPL)}\label{figures-suppl}}
+
+\hypertarget{figure-s1}{%
+\subsection{Figure S1}\label{figure-s1}}
+
+\hypertarget{figure-s2}{%
+\subsection{Figure S2}\label{figure-s2}}
+
+\hypertarget{figure-s3}{%
+\subsection{Figure S3}\label{figure-s3}}
+
+\hypertarget{figure-s4}{%
+\subsection{Figure S4}\label{figure-s4}}
+
+\hypertarget{tables-main}{%
+\section{TABLES (MAIN)}\label{tables-main}}
+
+\hypertarget{table-1}{%
+\subsection{Table 1}\label{table-1}}
+
+\hypertarget{table-2}{%
+\subsection{Table 2}\label{table-2}}
+
+\hypertarget{table-3}{%
+\subsection{Table 3}\label{table-3}}
+
+\hypertarget{tables-suppl}{%
+\section{TABLES (SUPPL)}\label{tables-suppl}}
+
+\hypertarget{table-s1}{%
+\subsection{Table S1}\label{table-s1}}
+
+\hypertarget{table-s2}{%
+\subsection{Table S2}\label{table-s2}}
+
+\hypertarget{table-s3}{%
+\subsection{Table S3}\label{table-s3}}
+
 \end{document}

### 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds figure1 in manuscript

@@
 
-```{r, echo = FALSE, eval = TRUE , fig.asp=6/10}
+```{r, echo = FALSE, eval = TRUE , fig.asp=9/10}
 # Define Layout and set labels
@@
                 8,8,8,8,     10,10,10,10,     10,10,10,10 ) , ncol= 12,byrow = T),
-       widths = 1,heights = c(1,1,0.5,0.5,0.5))
+       widths = 1,heights = c(1,1,.7,.7,.7))
 # layout.show(29)
@@
 ###################################
-par(mar=c(1,1,1,1))
+par(mar=c(1,4,1,10))
 
@@
 ###################################
-par(mar=c(4,4,1,1))
+par(mar=c(4,4,1,10))
 temp <- t(t(2^x-1)/colSums(2^x-1))
@@
 temp <- rowsum(temp , taxa$Class)
-o <- order( rowSums(temp) , decreasing = F)
-sample_o <- order( as.numeric(metadata$group_M4_CVLv3)+(shann/max(shann)/2) )
+temp[is.nan(temp)] <- 0
+o <- order( rowSums(temp) , decreasing = T)
+sample_o <- order( metadata$joint_clustering )
 temp <- temp[o,sample_o]*100
-barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i", col = pal[factor(rownames(temp))] )
+temp <- temp[-grep("NA",rownames(temp)),]
+temp <- rbind(temp[1:15,], Other = colSums(temp[16:nrow(temp),]) )
+
+taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"))
+barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",
+         col = taxa_pal[factor(rownames(temp))] )
+legend(x = par("usr")[c(2)],
+       y = par("usr")[c(4)],title.adj = 0,
+       legend = rownames(temp),xjust = 0,yjust = 1,
+       bty = "n",pch = 22,pt.cex = 2,cex = .8,xpd=T,
+       pt.bg = taxa_pal)
 

### 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
updates figure1 in manuscript

--- 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 library(rafalib)
-remotes::install_github("czarnewski/niceRplots")
+# remotes::install_github("czarnewski/niceRplots")
 library(niceRplots)
@@
 
-```{r, echo = FALSE, eval = TRUE , fig.asp=6/10}
+```{r, echo = FALSE, eval = TRUE , fig.asp=9/10}
 # Define Layout and set labels
-layout(matrix(c(1,1,1,1,      3,3,3,3,          4,4,4,4,
-                2,2,2,2,      3,3,3,3,          5,5,5,5,
+layout(matrix(c(1,1,1,1,      3,3,3,3,          3,4,4,4,
+                2,2,2,2,      3,3,3,3,          3,5,5,5,
                 6,6,6,6,      9,9,9,9,          9,9,9,9,
@@
 ###################################
-par(mar=c(0,0,0,0))
+par(mar=c(1,1,1,1))
 empty_plot(frame = T)
@@
 ###################################
-par(mar=c(1,1,1,1))
+par(mar=c(1,5,1,12))
 
-shann <- vegan::diversity(t(x),index = "shannon")
-shann <- t(as.matrix(shann))
-rownames(shann) <- "Shannon"
-barlist( data = shann,
-         main = "Shannon",
-         labels = "",
-         genes = c("Shannon"),
+shann <- vegan::diversity(t(x),index = c("shannon") )
+simp <- vegan::diversity(t(x),index = c("simpson") )
+invsimp <- vegan::diversity(t(x),index = c("invsimpson") )
+divers <- as.matrix(rbind(shann,simp,invsimp) )
+divers[is.infinite(divers)] <- 0
+rownames(divers) <- c("Shannon","simpson","Inv Simpson")
+barlist( data = divers,
+         main = "",
+         genes = c("Shannon","simpson","Inv Simpson"),
          clustering = metadata$joint_clustering,
@@
 ###################################
-par(mar=c(4,4,1,1))
+par(mar=c(4,5,1,12))
 temp <- t(t(2^x-1)/colSums(2^x-1))
@@
 temp <- rowsum(temp , taxa$Class)
-o <- order( rowSums(temp) , decreasing = F)
-sample_o <- order( as.numeric(metadata$group_M4_CVLv3)+(shann/max(shann)/2) )
+temp[is.nan(temp)] <- 0
+o <- order( rowSums(temp) , decreasing = T)
+sample_o <- order( metadata$joint_clustering )
 temp <- temp[o,sample_o]*100
-barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i", col = pal[factor(rownames(temp))] )
+temp <- temp[-grep("NA",rownames(temp)),]
+temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
+
+taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
+barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
+         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+legend(x = par("usr")[c(2)],
+       y = par("usr")[c(4)],title.adj = 0,
+       legend = rownames(temp),xjust = 0,yjust = 1,
+       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+       pt.bg = taxa_pal)
 

### 2021-01-04 reports/rmarkdown/manuscript/manuscript_template.Rmd
test git push

@@
 Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+test test test
 

### 2021-01-06 reports/rmarkdown/manuscript/manuscript_template.Rmd
updates manuscript

--- 2020-12-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
-patient_groups <- metadata$group_M4_CVLv3
+patient_groups <- metadata$joint_clustering
 bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
@@
        widths = 1,heights = c(1,1,.7,.7,.7))
-# layout.show(29)
+# layout.show(12)
 figlabels <- letters
@@
 par(mar=c(0,0,0,0))
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$joint_clustering)] ,
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(patient_groups)] ,
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
@@
 legend("topright",title.adj = 0,
-       legend = levels(factor(metadata$joint_clustering)),xjust = 0,yjust = 1,
+       legend = levels(factor(patient_groups)),xjust = 0,yjust = 1,
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
@@
          genes = c("BV","counts_16S"),
-         clustering = metadata$joint_clustering,
+         clustering = patient_groups,
          draw_mean_lines=F,col = pal)
@@
          genes = c("Shannon","simpson","Inv Simpson"),
-         clustering = metadata$joint_clustering,
+         clustering = patient_groups,
          draw_mean_lines=F,col = pal)
@@
 o <- order( rowSums(temp) , decreasing = T)
-sample_o <- order( metadata$joint_clustering )
+sample_o <- order( patient_groups )
 temp <- temp[o,sample_o]*100

### 2021-01-15 reports/rmarkdown/03_integrated_datasets.Rmd
adds fig2 and figS1-2

@@
 k <- 15 # k for KNN
-P <- 0.1# prunning threshold of SNN
+P <- 0.2# prunning threshold of SNN
 
@@
 #Compute SNN graph from KNN
-jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
-
-jSNN[1:10,1:10]
-
-# k=10
-# P=0.1
-snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
-snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
-snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
-
-# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
-for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
-
-for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
-
-snnm[ snnj <= ceiling(k*P) ] <- 0
+# jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
+# 
+# jSNN[1:10,1:10]
+# 
+# # k=10
+# # P=0.1
+# snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
+# snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
+# snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
+# 
+# # for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
+# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
+# 
+# for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
+# 
+# snnm[ snnj <= ceiling(k*P) ] <- 0
 
 #Prune SNN graph based on the Jaccard index
-pSNN <- jSNN
-pSNN <- pSNN / max(pSNN)
+# pSNN <- jSNN
+# pSNN <- pSNN / max(pSNN)
 
@@
 #Graph construction 
-k_nearest_neighbours <- 10 # k for KNN
+k_nearest_neighbours <- 5 # k for KNN
 snn_threshold <- 0.1 # prunning threshold of SNN

### 2021-01-15 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds fig2 and figS1-2

@@
 library(rafalib)
+library(scales)
 # remotes::install_github("czarnewski/niceRplots")
@@
 
-###################################
-# BACTERIAL VAGINOSIS COMPARISSON #
-###################################
+#######################
+# BACTERIAL ABUNDANCE #
+#######################
 par(mar=c(4,5,1,12))
@@
 # Define Layout and set labels
-layout(matrix(c(1) , ncol= 1,byrow = T),
-       heights = c(1))
+layout(matrix(c(2 ,3,4,5,      6,7,20,20,          20,20,20,20,
+                19,1,1,1,      1,8,20,20,          20,20,20,20,
+                18,1,1,1,      1,9,21,21,         21,21,24,24,
+                17,1,1,1,      1,10,22,22,         22,22,24,24,
+                16,15,14,13,  12,11,23,23,         23,23,24,24 ) , ncol= 12,byrow = T),
+       widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
+# layout.show(12)
 figlabels <- letters
+
+#########################
+# BACTERIAL COMMUNITIES #
+#########################
+gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+lB <- layout_with_fr(gB,niter=3000,start.temp=30)
+
+par(mar=c(0,0,0,0))
+plot( gB , vertex.label.cex=0.000001 , vertex.color = pal[factor(bac_communities)] ,
+      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
+      vertex.size=10,main="", 
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] )
+legend("topright",title.adj = 0,
+       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+par(mar=c(1,1,1,1))
+for(j in sort(unique(bac_communities)) ){
+  empty_plot()
+  text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
+  title(main=paste0("community",j),cex.main=.7)
+}
+for(i in 1:(18-length(unique(bac_communities)))){empty_plot()}
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+taxa[is.na(taxa)] <- "NA"
+
+par(mar=c(1,7,2,3))
+x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
+y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
+# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
+#        plot_labels = T, plot_weights = F,order_1_by = "total",
+#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+#        pal=hue_pal()(length(unique(x))))
+
+# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
+#      plot_labels = T, plot_weights = F,order_1_by = "total",
+#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+#      pal=hue_pal()(length(unique(y))))
+
+plot_sankey( data.frame( x[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
+     plot_labels = T, plot_weights = F,order_1_by = "total",
+     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+     pal=hue_pal()(length(unique(x))))
+
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+##############################################
+# BACTERIAL COMMUNITY DIFFERENTIAL ABUNDANCE #
+##############################################
+par(mar=c(1,7,1,3))
+
+module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
+  return( rowsum( x[rownames(SNN_bacteria),] , group = bac_communities ) / c(table(bac_communities)) )
+})
+module_means_per_dataset <- lapply(module_means_per_dataset,function(i){
+  rownames(i) <- paste0("BC",rownames(i))
+  return(i)
+})
+
+for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+        "ASV_CVL_V3_normalized_batch_corrected",
+        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+  temp_group <- as.character(patient_groups)
+  temp_group[ is.na(temp_group) ] <- "NA"
+  temp_group[ temp_group == "" ] <- "NA"
+  temp_group <- factor(temp_group)
+
+  barlist(data=module_means_per_dataset[[j]],
+          genes=rownames(module_means_per_dataset[[j]]),col = pal,
+          clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
+}
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+empty_plot()
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+# 
+# figlabels <- letters
+# temp_legend <- NULL
+# 
+# layout(matrix(c(1,4,
+#                 2,4,
+#                 3,4) , ncol= 2,byrow = T),widths = c(4,1),heights = c(1,1,1))
+# 
+# for(i in c("ASV_CVL_V3_normalized_batch_corrected",
+#            "ASV_CVL_V2_normalized_NOT_batch_corrected",
+#            "ASV_tissue_V3_normalized_batch_corrected")){
+#   x <- datasets_all_samples[[i]]
+#   
+#   par(mar=c(3,5,1,1))
+#   temp <- t(t(2^x-1)/colSums(2^x-1))
+#   taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+#   taxa[is.na(taxa)] <- "NA"
+#   taxa <- taxa[rownames(temp),]
+#   temp <- rowsum(temp , taxa$Class)
+#   temp[is.nan(temp)] <- 0
+#   
+#   if(is.null(temp_legend)){ 
+#     temp_legend <- temp   #Use the same legend from the 1st dataset
+#   }
+#   
+#   o <- order( rowSums(temp_legend) , decreasing = T)
+#   sample_o <- order( patient_groups )
+#   temp <- temp[o,sample_o]*100
+#   temp <- temp[-grep("NA",rownames(temp)),]
+#   temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
+#   
+#   taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
+#   barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
+#            col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+#   # legend(x = par("usr")[c(2)],
+#   #        y = par("usr")[c(4)],title.adj = 0,
+#   #        legend = rownames(temp),xjust = 0,yjust = 1,
+#   #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+#   #        pt.bg = taxa_pal)
+#   
+#   #add label
+#   add_letter(figlabels[1]); figlabels <- figlabels[-1]
+# }
+# 
+# 
+# par(mar=c(4,0,1,0))
+# empty_plot()
+# legend(x = par("usr")[c(1)],
+#          y = par("usr")[c(4)],title.adj = 0,
+#          legend = rownames(temp),xjust = 0,yjust = 1,
+#          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+#          pt.bg = taxa_pal)
+
+figlabels <- letters
+temp_legend <- NULL
+
+layout(matrix(c(1,7,
+                2,7,
+                3,7,
+                4,7,
+                5,7,
+                6,7) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))
+
+for(i in c("ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected",
+           "ASV_tissue_V3_normalized_batch_corrected")){
+  x <- datasets_all_samples[[i]]
+
+  par(mar=c(1,5,1,1))
+  temp <- t(t(2^x-1)/colSums(2^x-1))
+
+  shann <- vegan::diversity(t(temp),index = c("shannon") )
+  simp <- vegan::diversity(t(temp),index = c("simpson") )
+  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
+  divers <- as.matrix(rbind(shann,simp,invsimp) )
+  divers[is.infinite(divers)] <- 0
+  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
+  barlist( data = divers,
+         main = "",
+         genes = c("Shannon","simpson","Inv Simpson"),
+         clustering = patient_groups,
+         draw_mean_lines=F,col = pal)
+
+  par(mar=c(3,5,1,1))
+  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+  # taxa[is.na(taxa)] <- "NA"
+  # taxa <- taxa[rownames(temp),]
+  # temp <- rowsum(temp , taxa$Class)
+  temp[is.nan(temp)] <- 0
+
+  if(is.null(temp_legend)){ 
+    temp_legend <- temp   #Use the same legend from the 1st dataset
+  }
+
+  o <- order( rowSums(temp_legend) , decreasing = T)
+  sample_o <- order( patient_groups )
+  temp <- temp[o,sample_o]*100
+  # temp <- temp[-grep("NA",rownames(temp)),]
+  temp <- rbind(temp[1:20,], Other = colSums(temp[21:nrow(temp),]) )
+
+  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
+  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
+           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+  # legend(x = par("usr")[c(2)],
+  #        y = par("usr")[c(4)],title.adj = 0,
+  #        legend = rownames(temp),xjust = 0,yjust = 1,
+  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+  #        pt.bg = taxa_pal)
+
+  #add label
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+}
+
+par(mar=c(4,0,1,0))
+empty_plot()
+legend(x = par("usr")[c(1)],
+         y = par("usr")[c(4)],title.adj = 0,
+         legend = rownames(temp),xjust = 0,yjust = 1,
+         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+         pt.bg = taxa_pal)
+```
+
+\clearpage
+
 ## Figure S2
+
+```{r, echo = FALSE, eval = TRUE}
+
+figlabels <- letters
+temp_legend <- NULL
+
+mypar(1,4,mar=c(12,1,1,0))
+
+for(i in c("ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected",
+           "ASV_tissue_V3_normalized_batch_corrected")){
+
+  continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+  continuous_metadata <- metadata[,continuous_parameters]
+  continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )
+
+  metadata_cor <- cor(continuous_metadata,use = "pairwise.complete.obs")
+
+  #TO ORDER THE HEATMAP
+  o <- hclust( as.dist(1-metadata_cor),method = "complete")$order
+  metadata_cor <- metadata_cor[o,o]
+
+  image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
+         breaks = seq(-1,1,length.out = 90),main=sub("_nor.*","",i),font.main=1,
+         col=colorRampPalette(c("navy","grey95","firebrick"))(89) )
+  # text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(metadata_cor)) , 
+  #        labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+  text( seq(0,1,length.out = nrow(metadata_cor)) , par("usr")[3]-.02 , 
+        labels = rownames(metadata_cor), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+
+  #add label
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+}
+
+par(mar=c(12,0,1,0))
+image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
+         breaks = seq(-1,1,length.out = 90),main="",
+         col=colorRampPalette(c("white"))(89) )
+text(  0 ,seq(1,0,length.out = nrow(metadata_cor)) , 
+  labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+```
+
+\clearpage
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
+
 ## Figure S4
+
+```{r, echo = FALSE, eval = TRUE}
+```
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
+
 ## Table 2
 
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
+
 ## Table 3
+
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
+
 ## Table S2
 
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage
+
 ## Table S3
+
+```{r, echo = FALSE, eval = TRUE}
+```
+
+\clearpage

### 2021-01-15 reports/rmarkdown/manuscript/manuscript_template.tex
adds figure1 in manuscript
updates manuscript
adds fig2 and figS1-2

@@
 Author Three\textsuperscript{1,2}}
-\date{December 16, 2020}
+\date{January 15, 2021}
 
@@
 To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see ``Methods'' section for details).
-A total of 11 bacterial communities were identified.
+A total of 15 bacterial communities were identified.
 
-Noticebly, bacterial community 11 consisted only of Lactobacillus species (\emph{Anaerococcus hydrogenalis, Atopobium deltae, Bifidobacterium adolescentis, Bifidobacterium bifidum, Bifidobacterium longum/breve, Brevibacterium paucivorans, Brevibacterium ravenspurgense, Campylobacter hominis, Chlamydia trachomatis, Corynebacterium aurimucosum, Corynebacterium genitalium, Corynebacterium imitans, Corynebacterium jeikeium, Corynebacterium pseudogenitalium, Corynebacterium simulans/striatum, Corynebacterium sp/amycolatum, Corynebacterium spheniscorum, Corynebacterium sundsvallense/thomssenii, Dermabacter hominis, Dorea formicigenerans, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/dispar/durans, Gemmiger formicilis, Lactobacillus coleohominis, Lactobacillus crispatus, Lactobacillus crispatus/acidophilus, Lactobacillus gasseri/johnsonii/taiwanensis, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus mucosae, Lactobacillus reuteri, Lactobacillus reuteri/oris/frumenti/antri, Lactobacillus salivarius, Megasphaera elsdenii, other Bacilli, other Bifidobacteriaceae, other Clostridia, other Corynebacterium, other Ezakiella, other Gemella, other Gordonia, other Lactobacillales, other Lactobacillus, other Listeria, other Mobiluncus, other Mogibacterium, other Murdochiella, other Nesterenkonia, other Paraeggerthella, other Rhizobium, Peptoniphilus asaccharolyticus, Prevotella denticola, Staphylococcus saccharolyticus/capitis/hominis/caprae/lugdunensis/epidermidis/simiae, Streptococcus dysgalactiae, Streptococcus lutetiensis/equinus, Ureaplasma parvum, Brevundimonas mediterranea/intermedia/poindexterae, Corynebacterium appendicis/imitans, Corynebacterium freneyi/sp/amycolatum, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/saccharolyticus/ratti/dispar/gallinarum/durans, other Bacteroides, other Gammaproteobacteria}).
+Noticebly, bacterial community 15 consisted only of Lactobacillus species (\emph{Lactobacillus coleohominis, Lactobacillus crispatus/acidophilus, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus reuteri/oris/frumenti/antri, other Lactobacillales, Ureaplasma parvum}).
 
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-4-1}
+
 \textbf{Figure 2. Identification and characterization of vaginal bacterial communities.}
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-1}
+
+\clearpage
+
 \hypertarget{figure-s2}{%
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
+
+\clearpage
+
 \hypertarget{figure-s3}{%
 \subsection{Figure S3}\label{figure-s3}}
+
+\clearpage
 
@@
 
+\clearpage
+
 \hypertarget{table-2}{%
@@
 
+\clearpage
+
 \hypertarget{table-3}{%
 \subsection{Table 3}\label{table-3}}
+
+\clearpage
 
@@
 
+\clearpage
+
 \hypertarget{table-s2}{%
 \subsection{Table S2}\label{table-s2}}
+
+\clearpage
 
@@
 
+\clearpage
+
 \end{document}

### 2021-01-15 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds fig2 and figS1-2

@@
   # temp <- temp[-grep("NA",rownames(temp)),]
-  temp <- rbind(temp[1:20,], Other = colSums(temp[21:nrow(temp),]) )
+  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
 

### 2021-02-03 reports/rmarkdown/03_integrated_datasets.Rmd
updates clustering and samples used

@@
 ```{r, echo = FALSE}
-PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data
+PATH <- "~/Desktop/Desktop_stuff/NBIS/SMS_Projects/broliden_5325/" #Path to the data
 
@@
 
-samples_to_use <- colnames(counts)[(colSums(counts == 0)<3)]
+samples_to_use <- colnames(counts)[counts[c(3),] != 0]
 datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
 metadata <- metadata[ samples_to_use , ]
+
+detected_counts <- detected_counts[,samples_to_use]
+barlist( data = rbind( detected_counts),main = "detected features",
+         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
+         clustering = rep(1,nrow(metadata)),
+         col="grey",draw_mean_lines=F)
+add_letter("b")
+
 saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
@@
 # all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
-all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
-all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]
+all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
+# all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
 dim(all_microbiome)
@@
 barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
+
 #Compute sample-wise correlations
-cors <- cor(all_microbiome[is_top_var,], method = "pearson")
+cors <- cor(all_microbiome[,], method = "pearson")
 adj <- (1-cors)/2
@@
 #Graph construction  
-k <- 15 # k for KNN
-P <- 0.2# prunning threshold of SNN
+k <- 10 # k for KNN
+p <- 0.1 # prunning threshold of SNN
 
@@
 knn <- RANN::nn2(adj,k = k)
-# knn <- RANN::nn2(t(zscore),k = k)
 nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
@@
 
-#Compute SNN graph from KNN
-# jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
-# 
-# jSNN[1:10,1:10]
-# 
-# # k=10
-# # P=0.1
-# snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
-# snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
-# snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
-# 
-# # for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
-# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
-# 
-# for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
-# 
-# snnm[ snnj <= ceiling(k*P) ] <- 0
-
-#Prune SNN graph based on the Jaccard index
-# pSNN <- jSNN
-# pSNN <- pSNN / max(pSNN)
-
-# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
-#      border=NA,col="grey60",las=1,
-#      xaxs="i",yaxs="i",
-#      main="SNN jaccard",xlim=c(0,1))
-# abline(v=P,col="red",lty=2,xpd=F)
-# 
-# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
-#      border=NA,col="grey60",las=1,
-#      xaxs="i",yaxs="i",
-#      main="correlation coeficient",xlim=c(-1,1))
-# abline(v=0,col="red",lty=1,xpd=F)
-
+#Computing SNN matrix
 N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
@@
 J[is.nan(J)] <- 0
-J[J < 0.2] <- 0
+J[J < p] <- 0
 colnames(J) <- colnames(all_microbiome)
 rownames(J) <- colnames(all_microbiome)
-# image(J[,nrow(nn):1],col = colorRampPalette(c("grey95","black"))(90),axes=F )
-# set.seed(1)
-# g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
-# plot( g , vertex.label.cex=0.000001 , vertex.color = "grey" , 
-#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
-#       vertex.size=10 )
-
-pSNN[ pSNN < P ] <- 0
-pSNN[ pSNN > P ] <- cors[pSNN > P]
-pSNN[ pSNN < 0 ] <- 0
-
+
+#Transforming SNN into an igraph object
 g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
@@
 set.seed(1)
-# l <- layout_with_graphopt(g,niter = 1000)
 set.seed(1)
@@
 cl <- igraph::cluster_louvain(g)
-cl <- igraph::walktrap.community(g)
+# cl <- igraph::walktrap.community(g)
+
 metadata$joint_clustering <- cl$membership
@@
        legend = levels(factor(cl$membership)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,main="FR embedding",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups\ngroup_M9_CVLv3",
+       legend = levels(factor(metadata$group_M9_CVLv3)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
+# title("Louvain clusters",cex.main=1,font.main=1)
 
@@
        legend = levels(factor(metadata$group_M4_CVLv3)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
 # title("Louvain clusters",cex.main=1,font.main=1)
@@
 
-```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}
+```{r}
+mypar(mar=c(2,15,2,2))
+
+barlist( data = rbind( detected_counts),main = "detected features",
+         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
+         clustering = metadata$joint_clustering,
+         col=pal,draw_mean_lines=F)
+add_letter("b")
+
+barlist( data = all_microbiome, main = "detected features",
+         genes = c("Gardnerella vaginalis",
+                   "Lactobacillus iners",
+                   "Lactobacillus jensenii",
+                   "Lactobacillus crispatus/acidophilus",
+                   "Prevotella amnii",
+                   "Prevotella timonensis",
+                   "Atopobium vaginae"),
+         clustering = metadata$joint_clustering,
+         col=pal,draw_mean_lines=F)
 
@@
 \clearpage
-
-```{r, eval=F, echo = FALSE,fig.width=10,fig.height=4}
-
-all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-
-datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) ,
-                      rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ,
-                      rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
-
-groups <- factor( rep(metadata$joint_clustering,3) )
-age <- rep(metadata$age,3)
-work_time <- rep(metadata$sexwork_months,3)
-
-datasets <- datasets[ colSums(all_microbiome)>0 ]
-groups <- groups[ colSums(all_microbiome)>0 ]
-age <- age[ colSums(all_microbiome)>0 ]
-all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
-df <- data.frame(groups,datasets,age)
-
-library(edgeR)
-design <- model.matrix( ~ datasets * groups + age , data=df )
-rownames(design) <- paste0(datasets,"_",colnames(all_microbiome))
-colnames(all_microbiome) <- rownames(design)
-
-y <- DGEList(counts = round(((2^all_microbiome)-1)*1000) , remove.zeros = T)
-y <- calcNormFactors(y,method = "TMM")
-y <- estimateGLMCommonDisp(y,design)
-y <- estimateGLMTagwiseDisp(y,design)
-fit <- glmFit(y, design)
-
-#age
-lrt <- glmLRT(fit,coef = 8)
-top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
-plot(top$logFC,-log10(top$PValue))
-
-#age
-lrt <- glmLRT(fit,coef = 4:7)
-top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
-
-#age
-lrt <- glmLRT(fit,coef = 9:12)
-top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
-
-#age
-lrt <- glmLRT(fit,coef = 13:16)
-top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
-
-```
 
@@
 
-cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=5,]))
+cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))
 
@@
      bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
-
 ```
@@
 write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))
-
 ```
@@
 
-mypar(4,3,mar=c(3,5,2,2))
-
-# for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
-#   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-#           "ASV_CVL_V3_normalized_batch_corrected",
-#           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-#     temp_group <- as.character(df[,i])
-#     temp_group[ is.na(temp_group) ] <- "NA"
-#     temp_group[ temp_group == "" ] <- "NA"
-#     temp_group <- factor(temp_group)
-#     
-#     barlist(data=module_means_per_dataset[[j]],
-#             genes=rownames(module_means_per_dataset[[j]]),
-#             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
-#             clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
-#   }
-# }
-
-for( i in c("group_M4_CVLv3","group_M6_CVLv3","group_M9_CVLv3","joint_clustering")){
+mypar(3,3,mar=c(3,5,2,2))
+
+for( i in c("group_M4_CVLv3","group_M9_CVLv3","joint_clustering")){
   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
@@
 
-  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
   bac_dataset <- module_means_per_dataset[[i]]
-  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
-  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
+  samples_TRX_and_CVL3 <- colnames(bac_dataset)
 
@@
   TRX <- TRX [ , samples_TRX_and_CVL3 ]
-  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
-  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
-  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
-  TRX <- TRX[ top_vars_TRX , ]
-  dim(TRX)
+  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]
 
@@
   bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
-  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
-  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
   dim(bac_dataset)
@@
   cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
-  # hist(cors,xlim=c(-1,1),breaks = 100)
   dim(cors)
@@
   enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
+    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
     return(res)
@@
 
-  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
+  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
+  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)
+
+  pvalues <- lapply(enrichments,function(x){
+    temp <- allpvals
+    temp[x$pathway] <- x$pval
+    return(temp)
+  })
   pvalues <- t(as.data.frame(pvalues))
   pvalues <- -log10( pvalues )
-  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
-  # mypar(mar=c(4,15,2,2))
-  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
-
-  # pvalues[ pvalues < -log10(0.05)] <- 0
-  # sort(rowSums(pvalues>0),T)
-  # sort(colSums(pvalues>0),T)
-  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
+  pvalues[is.na(pvalues)] <- 1
+
+  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
+  NESes <- lapply(enrichments,function(x) {
+    temp <- allNES
+    temp[x$pathway] <- x$NES
+    return(temp)
+  })
   NESes <- t(as.data.frame(NESes))
   NESes[is.na(NESes)] <- 0
-  NESes[ pvalues < -log10(0.05) ] <- 0
-  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
-
-  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
-  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
-  # NESes <- NESes[o_bacs,o_kegg]
+  NESes[ pvalues < -log10(0.1) ] <- 0
 
   NESse_list[[i]] <- NESes
-
-  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
-  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
-  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
-  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
-  # add_letter(myletter); myletter <- myletter[-1]
-  # empty_plot()
-  # cat("\\clearpage")
 
@@
   # add_letter(myletter); myletter <- myletter[-1]
-  text(   seq(1,0,length.out = nrow(x)) , par("usr")[c(3)] ,
+  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
       labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
@@
            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
-text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
+text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
       labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
@@
 non_zeros <- colSums(TRX)!=0
-TRX <- TRX [ top_vars_TRX , non_zeros ]
+TRX <- TRX [  , non_zeros ]
 pathway_means <- t(sapply( colnames(x) , function(i){
@@
         genes=rownames(pathway_means),
-        clustering=metadata$joint_clustering[non_zeros],
-        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
-
-par(mar=c(6,2,3,2))
-image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
-       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-# add_letter(myletter); myletter <- myletter[-1]
-text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-# empty_plot(xaxs="i",yaxs="i",
-#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
-# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
-      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-
-```
-
-**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
-Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
-Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the Hallmark MSigDB gene annotation database.
-This, in turn, will result in a matrix associating every bacteria with every Hallmark MSigDB process in the tissue.
-The heatmap shows the normalized enrichment score (NES).
-Only enrichments with pvalue below 0.05 are shown.
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=8}
-myletter <- letters
-
-NESse_list <- list()
-
-for(i in c("ASV_tissue_V3_normalized_batch_corrected",
-           "ASV_CVL_V3_normalized_batch_corrected")){
-
-  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))
-
-  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-  bac_dataset <- module_means_per_dataset[[i]]
-  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
-  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
-
-  # mypar(4,3,mar=c(4,2,2,2))
-  #filter TRX dataset
-  TRX <- TRX [ , samples_TRX_and_CVL3 ]
-  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
-  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
-  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
-  TRX <- TRX[ top_vars_TRX , ]
-  dim(TRX)
-
-  #filter bac_dataset dataset
-  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
-  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
-  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
-  dim(bac_dataset)
-
-  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
-  # hist(cors,xlim=c(-1,1),breaks = 100)
-  dim(cors)
-
-  bacteria_use <- colnames(cors)
-  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
-  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
-  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
-  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgseaMultilevel(pathways = gmt,stats = sort(cors[,x],decreasing = T))
-    return(res)
-  })
-  parallel::stopCluster(cl)
-  names(enrichments) <- paste0( "bac.com.", bacteria_use )
-
-  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
-  pvalues <- t(as.data.frame(pvalues))
-  pvalues <- -log10( pvalues )
-  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
-  # mypar(mar=c(4,15,2,2))
-  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
-
-  # pvalues[ pvalues < -log10(0.05)] <- 0
-  # sort(rowSums(pvalues>0),T)
-  # sort(colSums(pvalues>0),T)
-  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
-  NESes <- t(as.data.frame(NESes))
-  NESes[is.na(NESes)] <- 0
-  NESes[ pvalues < -log10(0.05) ] <- 0
-  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]
-
-  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
-  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
-  # NESes <- NESes[o_bacs,o_kegg]
-
-  NESse_list[[i]] <- NESes
-
-  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
-  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
-  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
-  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
-  # add_letter(myletter); myletter <- myletter[-1]
-  # empty_plot()
-  # cat("\\clearpage")
-
-}
-
-common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[ colSums(x!=0)>=2 ] ) ) )
-
-layout(matrix(c(1,2,3,3,3),nrow = 1))
-par(mar=c(6,1,3,0))
-
-for( i in names(NESse_list)){
-  x <- NESse_list[[i]][,common_pathways]
-  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
-         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-  # add_letter(myletter); myletter <- myletter[-1]
-  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-}
-empty_plot(xaxs="i",yaxs="i",
-           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
-text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
-      labels = rev(sub("KEGG_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-
-cat("\\clearpage")
-
-#Combined representation of the two datasets
-layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
-par(mar=c(6,30,3,0))
-x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
-x[NESse_list[[1]][,common_pathways] == 0] <- 0
-x[NESse_list[[2]][,common_pathways] == 0] <- 0
-x <- x[, colSums(x!=0)>0]
-
-TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-non_zeros <- colSums(TRX)!=0
-TRX <- TRX [ top_vars_TRX , non_zeros ]
-pathway_means <- t(sapply( colnames(x) , function(i){
-  i <- gmt[[i]]
-  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
-  if(ncol(i)>1){
-    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
-    i <- colMeans(i)
-  } else {
-    i <- (i - min(i)) / (max(i) - min(i))
-  }
-  return( i - min(i) )
-} ))
-plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
-        genes=rownames(pathway_means),
-        clustering=metadata$joint_clustering[non_zeros],
+        clustering=factor(metadata$joint_clustering[non_zeros]),
         srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
@@
 Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
+Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the Hallmark MSigDB gene annotation database.
+This, in turn, will result in a matrix associating every bacteria with every Hallmark MSigDB process in the tissue.
+The heatmap shows the normalized enrichment score (NES).
+Only enrichments with pvalue below 0.05 are shown.
+
+\clearpage
+
+```{r, echo = FALSE,fig.width=10,fig.height=8}
+myletter <- letters
+
+NESse_list <- list()
+
+for(i in c("ASV_tissue_V3_normalized_batch_corrected",
+           "ASV_CVL_V3_normalized_batch_corrected")){
+
+  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))
+
+  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
+  bac_dataset <- module_means_per_dataset[[i]]
+  samples_TRX_and_CVL3 <- colnames(bac_dataset)
+
+  # mypar(4,3,mar=c(4,2,2,2))
+  #filter TRX dataset
+  TRX <- TRX [ , samples_TRX_and_CVL3 ]
+  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]
+
+  #filter bac_dataset dataset
+  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
+  dim(bac_dataset)
+
+  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
+  dim(cors)
+
+  bacteria_use <- colnames(cors)
+  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
+  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
+  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
+  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
+    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
+    return(res)
+  })
+  parallel::stopCluster(cl)
+  names(enrichments) <- paste0( "bac.com.", bacteria_use )
+
+  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
+  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)
+
+  pvalues <- lapply(enrichments,function(x){
+    temp <- allpvals
+    temp[x$pathway] <- x$pval
+    return(temp)
+  })
+  pvalues <- t(as.data.frame(pvalues))
+  pvalues <- -log10( pvalues )
+  pvalues[is.na(pvalues)] <- 1
+
+  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
+  NESes <- lapply(enrichments,function(x) {
+    temp <- allNES
+    temp[x$pathway] <- x$NES
+    return(temp)
+  })
+  NESes <- t(as.data.frame(NESes))
+  NESes[is.na(NESes)] <- 0
+  NESes[ pvalues < -log10(0.1) ] <- 0
+
+  NESse_list[[i]] <- NESes
+
+}
+
+common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[colSums(x!=0)>=1] ) ) )
+
+layout(matrix(c(1,2,3,4,4),nrow = 1))
+par(mar=c(6,1,2,0))
+for( i in names(NESse_list)){
+  x <- NESse_list[[i]][,common_pathways]
+  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+  # add_letter(myletter); myletter <- myletter[-1]
+  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+}
+empty_plot(xaxs="i",yaxs="i",
+           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
+text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
+      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+cat("\\clearpage")
+
+#Combined representation of the two datasets
+layout(matrix(c(1,2,2,2,2),nrow = 1))
+par(mar=c(6,1,3,0))
+x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
+x[NESse_list[[1]][,common_pathways] == 0] <- 0
+x[NESse_list[[2]][,common_pathways] == 0] <- 0
+x <- x[, colSums(x!=0)>0]
+image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# add_letter(myletter); myletter <- myletter[-1]
+text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+empty_plot(xaxs="i",yaxs="i",
+           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
+text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
+      labels = rev(sub("KEGG_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+
+#Combined representation of the two datasets
+layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
+par(mar=c(6,30,3,0))
+x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
+x[NESse_list[[1]][,common_pathways] == 0] <- 0
+x[NESse_list[[2]][,common_pathways] == 0] <- 0
+x <- x[, colSums(x!=0)>0]
+
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+non_zeros <- colSums(TRX)!=0
+TRX <- TRX [ , non_zeros ]
+pathway_means <- t(sapply( colnames(x) , function(i){
+  i <- gmt[[i]]
+  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
+  if(ncol(i)>1){
+    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
+    i <- colMeans(i)
+  } else {
+    i <- (i - min(i)) / (max(i) - min(i))
+  }
+  return( i - min(i) )
+} ))
+plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
+        genes=rownames(pathway_means),
+        clustering=factor(metadata$joint_clustering[non_zeros]),
+        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
+
+par(mar=c(6,2,3,2))
+image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# add_letter(myletter); myletter <- myletter[-1]
+text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
+    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+# empty_plot(xaxs="i",yaxs="i",
+#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
+# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
+      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+```
+
+**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
 Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database.

### 2021-02-15 reports/rmarkdown/03_integrated_datasets.Rmd
fixes plots and adds picrust

@@
 ```{r, echo = FALSE}
-PATH <- "~/Desktop/Desktop_stuff/NBIS/SMS_Projects/broliden_5325/" #Path to the data
+PATH <- "../../" #Path to the data
 
@@
   pvalues <- -log10( pvalues )
-  pvalues[is.na(pvalues)] <- 1
+  pvalues[is.na(pvalues)] <- 0
 
@@
   NESes <- t(as.data.frame(NESes))
-  NESes[is.na(NESes)] <- 0
-  NESes[ pvalues < -log10(0.1) ] <- 0
+  # NESes[is.na(NESes)] <- 0
+  # NESes[ pvalues < -log10(0.1) ] <- 0
 
@@
 }
+
+write.csv2(NESse_list,"../../results/Hallmark_NESse_list.csv")
 
@@
 }
+
+write.csv2(NESse_list,"../../results/Kegg_NESse_list.csv")
 

### 2021-02-15 reports/rmarkdown/manuscript/manuscript_template.Rmd
fixes plots and adds picrust

--- 2021-01-06 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 library(rafalib)
-# remotes::install_github("czarnewski/niceRplots")
+library(scales)
+# remotes::install_github("czarnewski/niceRplots",force=T)
 library(niceRplots)
@@
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
-patient_groups <- metadata$joint_clustering
+patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
 bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
+
 SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
@@
 
+picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
+picrust_dge <- read.csv("../../../data/Picrust2_top_functions.csv",header = T,row.names = 1)
+```
+
+```{r, echo = FALSE, eval = TRUE}
+#Assign names to patient groups
+big_groups <- c(
+  L.inners=    "5",
+  L.inners2=   "4",
+  L.crispatus= "2",
+  Gard=        "3",
+  Gard.Prev=   "6",
+  Gard.Prev2=  "8",
+  Gard.Lacto=  "7",
+  Patho=       "1"
+)
+
+group_annotation <- as.character(patient_groups)
+group_annotation <- names(big_groups)[match(group_annotation,big_groups)]
+group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))
+
+metadata$group_annotation <- group_annotation
+```
+
+```{r, echo = FALSE, eval = TRUE}
+bac_communities <- factor(paste0("BC",sprintf("%02d",bac_communities)))
+bacs <- names(bac_communities)
+
+#Assign names to bacterial communities
+BC_groups <- c(
+  BC01= "BC01_Bifidobacterium",
+  BC02= "BC02_Mycoplasma",
+  BC03= "BC03_Lactobacillus.spp",
+  BC04= "BC04_Escherichia.sp",
+  BC05= "BC05_Other1",
+  BC06= "BC06_Other2",
+  BC07= "BC07_Corynebacterium.spp",
+  BC08= "BC08_Other3",
+  BC09= "BC09_Ureaplasma.sp",
+  BC10= "BC10_Campylobacter",
+  BC11= "BC11_Mobiluncus_BVAB",
+  BC12= "BC12_Faecalibacterium",
+  BC13= "BC13_Dialister",
+  BC14= "BC14_P.timonesis",
+  BC15= "BC15_T.scotoductus"
+)
+
+bac_communities <- as.character(bac_communities)
+bac_communities <- BC_groups[match(bac_communities,names(BC_groups))]
+bac_communities <- as.factor(setNames(bac_communities,bacs))
 ```
@@
 
-Patients were thus subdivided into `r length(unique(patient_groups))` groups,
+Patients were thus subdivided into `r length(unique(group_annotation))` groups,
 
@@
 
-par(mar=c(0,0,0,0))
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(patient_groups)] ,
+par(mar=c(0,0,0,5))
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
@@
 legend("topright",title.adj = 0,
-       legend = levels(factor(patient_groups)),xjust = 0,yjust = 1,
+       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
@@
          genes = c("BV","counts_16S"),
-         clustering = patient_groups,
+         clustering = group_annotation,
          draw_mean_lines=F,col = pal)
@@
          genes = c("Shannon","simpson","Inv Simpson"),
-         clustering = patient_groups,
+         clustering = group_annotation,
          draw_mean_lines=F,col = pal)
@@
 o <- order( rowSums(temp) , decreasing = T)
-sample_o <- order( patient_groups )
+sample_o <- order( group_annotation )
 temp <- temp[o,sample_o]*100
@@
 # Define Layout and set labels
-layout(matrix(c(1) , ncol= 1,byrow = T),
-       heights = c(1))
+layout(matrix(c(2 ,3,4,5,      6,7,17,17,          19,19,19,19,
+                1,1,1,1,      1,8,17,17,          20,20,20,20,
+                1,1,1,1,      1,9,18,18,         21,21,24,24,
+                1,1,1,1,      1,10,18,18,         22,22,24,24,
+                16,15,14,13,  12,11,23,23,         23,23,24,24 ) , ncol= 12,byrow = T),
+       widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
+# layout.show(12)
 figlabels <- letters
 
+#########################
+# BACTERIAL COMMUNITIES #
+#########################
+gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+lB <- layout_with_fr(gB,niter=3000,start.temp=30)
+
+par(mar=c(0,0,0,5))
+plot( gB , vertex.label.cex=0.000001 , vertex.color = pal[factor(bac_communities)] ,
+      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
+      vertex.size=10,main="", 
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] )
+legend("topright",title.adj = 0,
+       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+par(mar=c(1,1,1,1))
+for(j in sort(unique(bac_communities)) ){
+  empty_plot()
+  text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
+  title(main=paste0("community",j),cex.main=.7)
+}
+for(i in 1:(18-length(unique(bac_communities)))){empty_plot()}
+
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+taxa[is.na(taxa)] <- "NA"
+
+par(mar=c(1,7,2,3))
+x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
+y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
+# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
+#        plot_labels = T, plot_weights = F,order_1_by = "total",
+#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+#        pal=hue_pal()(length(unique(x))))
+
+# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
+#      plot_labels = T, plot_weights = F,order_1_by = "total",
+#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+#      pal=hue_pal()(length(unique(y))))
+
+plot_sankey( data.frame( x[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
+     plot_labels = T, plot_weights = F,order_1_by = "total",
+     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
+     pal=hue_pal()(length(unique(x))))
+
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+##############################################
+# BACTERIAL COMMUNITY DIFFERENTIAL ABUNDANCE #
+##############################################
+par(mar=c(1,7,1,3))
+
+module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
+  return( rowsum( x[rownames(SNN_bacteria),] , group = bac_communities ) / c(table(bac_communities)) )
+})
+
+for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+        "ASV_CVL_V3_normalized_batch_corrected",
+        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+  temp_group <- as.character(group_annotation)
+  temp_group[ is.na(temp_group) ] <- "NA"
+  temp_group[ temp_group == "" ] <- "NA"
+  temp_group <- factor(temp_group)
+
+  barlist(data=module_means_per_dataset[[j]],
+          genes=rownames(module_means_per_dataset[[j]]),col = pal,
+          clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
+}
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+empty_plot()
+
+#####################
+# BACTERIAL PICRUST #
+#####################
+par(mar=c(1,7,1,3))
+# 
+# library(keggorthology)
+# data("KOgraph")
+# annot <- as.data.frame(unlist(nodeData(KOgraph,attr="tag")))
+# colnames(annot)[1] <- "KEGGID"
+# annot$pathway <- rownames(annot)
+# annot$KEGGID <- paste0("K",annot$KEGGID)
+# 
+# library(KEGGREST)
+
+download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/ko_info.tsv.gz","../../../data/KEGG_pathways_to_KO.tsv")
+download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/pathway_mapfiles/KEGG_pathways_to_KO.tsv","../../../data/KEGG_pathways_to_KO.tsv")
+download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/KEGG_pathways_info.tsv.gz","../../../data/KEGG_pathways_info.tsv.gz")
+system("gunzip ../../../data/*.gz")
+
+kegg_pathway_annot <- fgsea::gmtPathways('../../../data/KEGG_pathways_to_KO.tsv')
+
+kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
+
+ko_names <- readLines('../../../data/ko_info.tsv')
+ko_names <- strsplit(ko_names,"\t")
+
+path_trend <- t(sapply(kegg_pathway_annot,function(x){
+  x <- rownames(picrust_data)[rownames(picrust_data) %in% x]
+  if(length(x)>=2){
+    return(colSums(picrust_data [ x , ]))
+  } else {
+    return(rep(0, ncol(picrust_data)))
+  }
+}))
+
+path_trend <- path_trend[rowSums(path_trend)!=0 , ]
+rownames(path_trend) <- kegg_names[ match(rownames(path_trend),kegg_names[,1]),2]
+
+library(edgeR)
+design <- model.matrix(~  group_annotation, data=metadata)
+
+y <- DGEList(counts=floor(path_trend), remove.zeros = T)
+y <- calcNormFactors(y,method = "TMM")
+y <- estimateGLMCommonDisp(y,design)
+y <- estimateGLMTagwiseDisp(y,design)
+fit <- glmFit(y, design)
+
+lrt <- glmLRT(fit,coef=1:8)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+colnames(top) <- sub("group_annotation","",colnames(top))
+top <- cbind(logFC.Gard=0,top)
+
+top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+top_dge <- rownames(top)[ top_dge ]
+
+norm_path_trend <- log2( t( t(path_trend) / colSums(path_trend) ) * 1000 + 1 )
+norm_path_trend <- norm_path_trend[top_dge, ]
+norm_path_trend <- t(apply(norm_path_trend,1,function(x){scale(x,T,T)}))
+norm_path_trend[norm_path_trend > 5] <- 5
+colnames(norm_path_trend) <- colnames(path_trend)
+sample_use <- colnames(norm_path_trend)
+
+mypar(2,4,mar=c(1,.2,2,.2))
+h <- hclust( as.dist( (1- cor(t(norm_path_trend)))/2 ), method = "ward.D2")
+plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
+abline(v=1,xpd=F,col="red",lty=2)
+gene_module <- cutree(h, h = 1)
+
+points( rep(0,length(gene_module)),
+        seq(length(gene_module),1,length.out = length(gene_module)),
+        col=taxa_pal[factor(gene_module[h$order])],
+        pch=16,cex=.5)
+
+image( t(norm_path_trend[h$order,][nrow(norm_path_trend):1,order(group_annotation[sample_use])]),
+       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
+       breaks = seq(-5,5,length.out = 100),axes=F)
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
+        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
+        pch=16)
+
+par(mar=c(4,4,2,2))
+set.seed(1)
+UMAP_picrust <- uwot::umap(t(norm_path_trend),n_neighbors = 20,
+                 metric = "correlation",min_dist = .2,spread = .4,
+                 repulsion_strength = .4,negative_sample_rate = 3)
+plot(UMAP_picrust,bg=pal[group_annotation[sample_use]],main="UMAP on top functional profile",
+     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
 ```
@@
 
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+TRX_counts <- round(2^TRX - 1)
+TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
+sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
+TRX_counts <- TRX_counts[,sample_use]
+
+library(edgeR)
+
+design <- model.matrix(~  group_annotation, data=metadata[sample_use,])
+
+y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
+y <- calcNormFactors(y,method = "TMM")
+y <- estimateGLMCommonDisp(y,design)
+y <- estimateGLMTagwiseDisp(y,design)
+fit <- glmFit(y, design)
+
+lrt <- glmLRT(fit,coef=2:8)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+colnames(top) <- sub("group_annotation","",colnames(top))
+top <- cbind(logFC.Gard=0,top)
+
+top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+top_dge <- rownames(top)[ top_dge ]
+
+top_TRX_counts <- cpm(y)[top_dge, ]
+top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
+top_TRX_counts[top_TRX_counts > 5] <- 5
+colnames(top_TRX_counts) <- colnames(TRX_counts)
+
+mypar(2,4,mar=c(1,.2,2,.2))
+h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
+plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
+abline(v=2,xpd=F,col="red",lty=2)
+gene_module <- cutree(h, h = 2)
+
+points( rep(0,length(gene_module)),
+        seq(length(gene_module),1,length.out = length(gene_module)),
+        col=taxa_pal[factor(gene_module[h$order])],
+        pch=16,cex=.5)
+
+image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(group_annotation[sample_use])]),
+       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
+       breaks = seq(-5,5,length.out = 100),axes=F)
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
+        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
+        pch=16)
+
+set.seed(1)
+UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
+                 metric = "correlation",min_dist = 0.1,spread = 5,
+                 negative_sample_rate = 10)
+plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="UMAP on top DGE RNAseq",
+     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
 ```
@@
 
+```{r, echo = FALSE, eval = TRUE, fig.asp=1}
+figlabels <- letters
+temp_legend <- NULL
+
+layout(matrix(c(1,7,
+                2,7,
+                3,7,
+                4,8,
+                5,8,
+                6,8) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))
+
+for(i in c("ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected",
+           "ASV_tissue_V3_normalized_batch_corrected")){
+  x <- datasets_all_samples[[i]]
+
+  par(mar=c(1,5,1,1))
+  temp <- t(t(2^x-1)/colSums(2^x-1))
+
+  shann <- vegan::diversity(t(temp),index = c("shannon") )
+  simp <- vegan::diversity(t(temp),index = c("simpson") )
+  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
+  divers <- as.matrix(rbind(shann,simp,invsimp) )
+  divers[is.infinite(divers)] <- 0
+  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
+  barlist( data = divers[,order(group_annotation)],
+         main = "",
+         genes = c("Shannon","simpson","Inv Simpson"),
+         # clustering = group_annotation,
+         draw_mean_lines=F,col = pal[factor(group_annotation[order(group_annotation)])])
+
+  #add label
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+  par(mar=c(3,5,1,1))
+  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+  # taxa[is.na(taxa)] <- "NA"
+  # taxa <- taxa[rownames(temp),]
+  # temp <- rowsum(temp , taxa$Class)
+  temp[is.nan(temp)] <- 0
+
+  if(is.null(temp_legend)){ 
+    temp_legend <- temp   #Use the same legend from the 1st dataset
+  }
+
+  o <- order( rowSums(temp_legend) , decreasing = T)
+  sample_o <- names(group_annotation)[ order( group_annotation )]
+  temp <- temp[o,sample_o]*100
+  # temp <- temp[-grep("NA",rownames(temp)),]
+  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
+
+  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
+  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
+           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+  # legend(x = par("usr")[c(2)],
+  #        y = par("usr")[c(4)],title.adj = 0,
+  #        legend = rownames(temp),xjust = 0,yjust = 1,
+  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+  #        pt.bg = taxa_pal)
+}
+
+par(mar=c(4,0,1,0))
+empty_plot()
+legend(x = par("usr")[c(1)],
+         y = par("usr")[c(4)],title.adj = 0,
+         legend = levels(group_annotation),xjust = 0,yjust = 1,
+         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
+         pt.bg = pal)
+
+empty_plot()
+legend(x = par("usr")[c(1)],
+         y = par("usr")[c(4)],title.adj = 0,
+         legend = rownames(temp),xjust = 0,yjust = 1,
+         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+          title="top abundant bacteria",
+         pt.bg = taxa_pal)
+
+```
+
+\clearpage
+
 ## Figure S2
+
+```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
+figlabels <- letters
+temp_legend <- NULL
+
+mypar(1,4,mar=c(12,1,1,0))
+
+for(i in c("ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected",
+           "ASV_tissue_V3_normalized_batch_corrected")){
+
+  continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+  continuous_metadata <- metadata[,continuous_parameters]
+  # continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )
+
+  metadata_cor <- cor(t(module_means_per_dataset[[i]]),continuous_metadata,use = "pairwise.complete.obs")
+  # metadata_cor_test <- cor.test(t(module_means_per_dataset[[i]]),continuous_metadata)
+
+  image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
+         breaks = seq(-1,1,length.out = 90),main=sub("_nor.*","",i),font.main=1,
+         col=colorRampPalette(c("navy","grey95","firebrick"))(89) )
+  # text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(metadata_cor)) , 
+  #        labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+  text( seq(0,1,length.out = ncol(metadata_cor)) , par("usr")[3]-.02 , 
+        labels = colnames(metadata_cor), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+
+  #add label
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+}
+
+par(mar=c(12,0,1,0))
+image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
+         breaks = seq(-1,1,length.out = 90),main="",
+         col=colorRampPalette(c("white"))(89) )
+text(  0 ,seq(1,0,length.out = nrow(metadata_cor)) , 
+  labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+```
+
+\clearpage
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+figlabels <- letters
+mypar(1,2,mar=c(3,3,3,10))
+
+datasets <- datasets_all_samples[grep("ASV",names(datasets_all_samples))]
+datasets <- lapply(datasets,t)
+datasets <- t( as.data.frame(datasets) )
+
+group_means <- sapply(levels(group_annotation), function(x){
+  rowMeans( datasets[ , group_annotation == x ] )
+})
+
+d <- as.dist( (1 - cor(group_means)) / 2 )
+h <- hclust( d , method = "ward.D2")
+plot( as.dendrogram(h) , horiz=T, xlab="height")
+points( rep(0 , length(h$labels)) , 1:length(h$labels) , pch=21, cex=1.5,
+        bg=pal[factor(h$labels[ h$order ],levels = h$labels)] )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+datasets <- datasets_all_samples[grep("ASV",names(datasets_all_samples))]
+datasets <- as.data.frame(datasets)
+
+group_means <- sapply(levels(factor(bac_communities)), function(x){
+  colMeans( datasets[ bac_communities == x , ] )
+})
+
+d <- as.dist( (1 - cor(group_means)) / 2 )
+h <- hclust( d , method = "ward.D2")
+plot( as.dendrogram(h) , horiz=T, xlab="height")
+points( rep(0 , length(h$labels)) , 1:length(h$labels) , pch=21, cex=1.5,
+        bg=taxa_pal[factor(h$labels[ h$order ],levels = h$labels)] )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+```
+
+\clearpage
+
 ## Figure S4
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
+figlabels <- letters
+for(j in names(module_means_per_dataset)){
+  mypar(2,8,mar=c(0.5,0.5,4,0.5))
+
+  plot(UMAP_picrust,bg=pal[group_annotation],main="patient groups",
+     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+  for(i in 1:nrow(module_means_per_dataset[[j]])){
+
+    xx <- module_means_per_dataset[[j]][i,]
+    xx <- xx / max(xx)
+    o <- order(xx)
+    plot(UMAP_picrust[o,],
+         col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
+         main=rownames(module_means_per_dataset[[j]])[i],
+         pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
+  }
+}
+```
+
+\clearpage
+
+## Figure S5
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
+figlabels <- letters
+for(j in names(module_means_per_dataset)){
+  mypar(2,8,mar=c(0.5,0.5,4,0.5))
+
+  plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="patient groups",
+     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+  for(i in 1:nrow(module_means_per_dataset[[j]])){
+
+    xx <- module_means_per_dataset[[j]][i,][sample_use]
+    xx <- xx / max(xx)
+    o <- order(xx)
+    plot(UMAP_TRX[o,],
+         col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
+         main=rownames(module_means_per_dataset[[j]])[i],
+         pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
+  }
+}
+```
+
+\clearpage
 

### 2021-02-15 reports/rmarkdown/manuscript/manuscript_template.tex
fixes plots and adds picrust

@@
 Author Three\textsuperscript{1,2}}
-\date{January 15, 2021}
+\date{February 13, 2021}
 
@@
 
-To understand the longitudinal and tissue-specific microbiome profile in vaginal samples, 113 adult female sex workers were enrolled in {[}\ldots{]}.
+To understand the longitudinal and tissue-specific microbiome profile in vaginal samples, 111 adult female sex workers were enrolled in {[}\ldots{]}.
 Among those, 14 were previously tested positive for HIV during the cohort's sampling procedure.
@@
 
-Noticebly, bacterial community 15 consisted only of Lactobacillus species (\emph{Lactobacillus coleohominis, Lactobacillus crispatus/acidophilus, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus reuteri/oris/frumenti/antri, other Lactobacillales, Ureaplasma parvum}).
+Noticebly, bacterial community NA consisted only of Lactobacillus species (**).
 
-Patients were thus subdivided into 6 groups,
+Patients were thus subdivided into 8 groups,
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-3-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-5-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-4-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-2}
 
@@
 \subsection{Figure 3}\label{figure-3}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-9-1}
 
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-10-1}
+
 \clearpage
@@
 \subsection{Figure S4}\label{figure-s4}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-3}
+
+\clearpage
+
+\hypertarget{figure-s5}{%
+\subsection{Figure S5}\label{figure-s5}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-3}
 

### 2021-02-20 reports/rmarkdown/03_integrated_datasets.Rmd
adds picrust and RNAseq to manuscript

@@
 NESse_list <- list()
+pvalues_list <- list()
 
@@
   NESes <- t(as.data.frame(NESes))
-  # NESes[is.na(NESes)] <- 0
+  NESes[is.na(NESes)] <- 0
   # NESes[ pvalues < -log10(0.1) ] <- 0
@@
   NESse_list[[i]] <- NESes
+  pvalues_list[[i]] <- pvalues
 
@@
 write.csv2(NESse_list,"../../results/Hallmark_NESse_list.csv")
+write.csv2(pvalues_list,"../../results/Hallmark_pvalues_list.csv")
 
@@
   NESse_list[[i]] <- NESes
-
+  pvalues_list[[i]] <- pvalues
 }
@@
 write.csv2(NESse_list,"../../results/Kegg_NESse_list.csv")
+write.csv2(pvalues_list,"../../results/Kegg_pvalues_list.csv")
 

### 2021-02-20 reports/rmarkdown/05_picrust_analysis.Rmd
adds picrust and RNAseq to manuscript

---
title: "Piecrust_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r }
picrust_data <- read.table("../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
picrust_dge <- read.csv("../../data/Picrust2_top_functions.csv",header = T,row.names = 1)
metadata <- read.csv2("../../results/metadata_integration.csv",row.names = 1)

download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/ko_info.tsv.gz","../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/pathway_mapfiles/KEGG_pathways_to_KO.tsv","../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/KEGG_pathways_info.tsv.gz","../../data/KEGG_pathways_info.tsv.gz")
system("gunzip ../../data/*.gz")

kegg_pathway_annot <- fgsea::gmtPathways('../../data/KEGG_pathways_to_KO.tsv')

kegg_names <- read.delim('../../data/KEGG_pathways_info.tsv',header = F)

ko_names <- readLines('../../data/ko_info.tsv')
ko_names <- strsplit(ko_names,"\t")
```

```{r }

#Sum counts to pathway level
path_trend <- t(sapply(kegg_pathway_annot,function(x){
  x <- rownames(picrust_data)[rownames(picrust_data) %in% x]
  if(length(x)>=2){
    return(colSums(picrust_data [ x , ]))
  } else {
    return(rep(0, ncol(picrust_data)))
  }
}))
path_trend <- path_trend[rowSums(path_trend)!=0 , ]
rownames(path_trend) <- kegg_names[ match(rownames(path_trend),kegg_names[,1]),2]

library(edgeR)

#GLM fit across all patient groups
metadata$joint_clustering <- factor(metadata$joint_clustering)
design <- model.matrix(~  joint_clustering, data=metadata)
y <- DGEList(counts=floor(path_trend), remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

#Perform the test
lrt <- glmLRT(fit,coef=2:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("joint_clustering","",colnames(top))
write.csv2(top,"../../results/Picrust_DGE_table.csv",row.names = T)

#Filter significant bacteria
top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

#Log Normalize
norm_path_trend <- log2( t( t(path_trend) / colSums(path_trend) ) * 1000 + 1 )
norm_path_trend <- norm_path_trend[top_dge, ]

#Z-score transform
norm_path_trend <- t(apply(norm_path_trend,1,function(x){scale(x,T,T)}))
norm_path_trend[norm_path_trend > 5] <- 5
colnames(norm_path_trend) <- colnames(path_trend)

write.csv2(norm_path_trend,"../../results/Picrust_norm_path_trend.csv",row.names = T)
```

### 2021-02-20 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds picrust and RNAseq to manuscript

@@
 library(scales)
-# remotes::install_github("czarnewski/niceRplots",force=T)
+library(dendextend)
+library(edgeR)
+remotes::install_github("czarnewski/niceRplots",force=T)
 library(niceRplots)
@@
 picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
-picrust_dge <- read.csv("../../../data/Picrust2_top_functions.csv",header = T,row.names = 1)
 ```
@@
 metadata$group_annotation <- group_annotation
+write.csv2(metadata,"../../../results/metadata_integration.csv",row.names = T)
+
 ```
@@
 
-par(mar=c(0,0,0,5))
+par(mar=c(0,0,0,6))
 plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
-      vertex.size=10,main="", 
+      vertex.size=10,main="", layout=l,
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend("topright",title.adj = 0,
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
        legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
@@
 
-```{r, echo = FALSE, eval = TRUE}
+```{r, echo = FALSE, eval = TRUE,fig.asp=.6}
 # Define Layout and set labels
-layout(matrix(c(2 ,3,4,5,      6,7,17,17,          19,19,19,19,
-                1,1,1,1,      1,8,17,17,          20,20,20,20,
-                1,1,1,1,      1,9,18,18,         21,21,24,24,
-                1,1,1,1,      1,10,18,18,         22,22,24,24,
-                16,15,14,13,  12,11,23,23,         23,23,24,24 ) , ncol= 12,byrow = T),
-       widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
+layout(matrix(c(2 ,3,4,5,      6,7,17,17,          17,17,17,17,
+                1,1,1,1,      1,8,17,17,          17,17,17,17,
+                1,1,1,1,      1,9,18,18,         18,18,18,18,
+                1,1,1,1,      1,10,19,19,         19,19,19,19,
+                16,15,14,13,  12,11,20,20,         20,20,20,20 ) , ncol= 12,byrow = T),widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
 # layout.show(12)
@@
 gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
-set.seed(1)
-lB <- layout_with_fr(gB,niter=3000,start.temp=30)
+# set.seed(1)
+lB <- layout_nicely(gB,niter=3000,start.temp=30)
 
-par(mar=c(0,0,0,5))
+par(mar=c(0,0,0,10))
 plot( gB , vertex.label.cex=0.000001 , vertex.color = pal[factor(bac_communities)] ,
@@
       vertex.size=10,main="", 
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] )
-legend("topright",title.adj = 0,
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
        legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
@@
 }
-for(i in 1:(18-length(unique(bac_communities)))){empty_plot()}
 
@@
 
-par(mar=c(1,7,2,3))
+par(mar=c(1,10,1,15))
 x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
@@
      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-     pal=hue_pal()(length(unique(x))))
+     pal=hue_pal()(length(unique(x))),xaxs="i",yaxs="i")
 
@@
 ##############################################
-par(mar=c(1,7,1,3))
-
 module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
@@
 })
+# 
+par(mar=c(1,10,1,6))
+# for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+#         "ASV_CVL_V3_normalized_batch_corrected",
+#         "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+#   temp_group <- as.character(group_annotation)
+#   temp_group[ is.na(temp_group) ] <- "NA"
+#   temp_group[ temp_group == "" ] <- "NA"
+#   temp_group <- factor(temp_group)
+#   
+#   plot_dots(data=module_means_per_dataset[[j]],
+#           genes=rownames(module_means_per_dataset[[j]]),col = pal,
+#           clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
+# }
+# add_letter(figlabels[1]); figlabels <- figlabels[-1]
 
-for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-        "ASV_CVL_V3_normalized_batch_corrected",
-        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-  temp_group <- as.character(group_annotation)
-  temp_group[ is.na(temp_group) ] <- "NA"
-  temp_group[ temp_group == "" ] <- "NA"
-  temp_group <- factor(temp_group)
+plot_dots_difference(data = module_means_per_dataset[[2]],
+                     data2 = module_means_per_dataset[[3]],
+                     clustering = as.character(group_annotation),
+                     clustering2 = as.character(group_annotation),
+                     genes = rownames(module_means_per_dataset[[1]]),
+                     main = "CVL3 vs CVL2",cex.row = .8)
 
-  barlist(data=module_means_per_dataset[[j]],
-          genes=rownames(module_means_per_dataset[[j]]),col = pal,
-          clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
-}
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
-empty_plot()
+plot_dots_difference(data = module_means_per_dataset[[2]],
+                     data2 = module_means_per_dataset[[1]],
+                     clustering = as.character(group_annotation),
+                     clustering2 = as.character(group_annotation),
+                     genes = rownames(module_means_per_dataset[[1]]),
+                     main = "CVL3 vs tissue_V3",cex.row = .8)
+# mypar(mar=c(2,2,2,6))
+# plot(c(-2,10),c(-2,10),xaxs="i",yaxs="i")
+# rrr <- rev(as.raster(colorRampPalette(c("blue","navy","grey95","firebrick","red"))(99)))
+# rasterImage(as.raster(rrr),
+#             xleft = par("usr")[2]+diff(par("usr")[1:2])/40,
+#             xright = par("usr")[2]+diff(par("usr")[1:2])/20,
+#             ybottom = par("usr")[4]-diff(par("usr")[1:2])/4,
+#             ytop = par("usr")[4], 
+#             interpolate = FALSE )
+# text( c(par("usr")[2]+diff(par("usr")[1:2])/20,
+#         par("usr")[2]+diff(par("usr")[1:2])/20) ,
+#       c(par("usr")[4]-diff(par("usr")[3:4])/4,
+#         par("usr")[4]),labels = c("min","max"),pos=4 )
+# 
+# 
+# text( c(par("usr")[2]+diff(par("usr")[1:2])/20,
+#         par("usr")[2]+diff(par("usr")[1:2])/20,
+#         par("usr")[2]+diff(par("usr")[1:2])/20) ,
+#       c(par("usr")[4]-diff(par("usr")[1:2])/4,
+#         par("usr")[4]-diff(par("usr")[1:2])/8,
+#         par("usr")[4]),labels = c("-1","0","+1"),pos=4 )
+# text( c(par("usr")[2]+diff(par("usr")[1:2])/40) ,xpd=T,
+#       c(par("usr")[4]+diff(par("usr")[3:4])/40),labels = c("Z-score"),adj=0 )
 
@@
 #####################
-par(mar=c(1,7,1,3))
-# 
-# library(keggorthology)
-# data("KOgraph")
-# annot <- as.data.frame(unlist(nodeData(KOgraph,attr="tag")))
-# colnames(annot)[1] <- "KEGGID"
-# annot$pathway <- rownames(annot)
-# annot$KEGGID <- paste0("K",annot$KEGGID)
-# 
-# library(KEGGREST)
-
-download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/ko_info.tsv.gz","../../../data/KEGG_pathways_to_KO.tsv")
-download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/pathway_mapfiles/KEGG_pathways_to_KO.tsv","../../../data/KEGG_pathways_to_KO.tsv")
-download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/KEGG_pathways_info.tsv.gz","../../../data/KEGG_pathways_info.tsv.gz")
-system("gunzip ../../../data/*.gz")
-
-kegg_pathway_annot <- fgsea::gmtPathways('../../../data/KEGG_pathways_to_KO.tsv')
-
-kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
+picrust_data <- round(picrust_data[ rowSums(picrust_data>5)>=10, ])
+sample_use <- colnames(picrust_data)
 
@@
 ko_names <- strsplit(ko_names,"\t")
+ko_names <- setNames(unlist(lapply(ko_names,function(x){x[2]})), unlist(lapply(ko_names,function(x){x[1]})))
+ko_names <- sub(" [[].*","",ko_names)
+ko_names <- sub(".*; ","",ko_names)
 
-path_trend <- t(sapply(kegg_pathway_annot,function(x){
-  x <- rownames(picrust_data)[rownames(picrust_data) %in% x]
-  if(length(x)>=2){
-    return(colSums(picrust_data [ x , ]))
-  } else {
-    return(rep(0, ncol(picrust_data)))
-  }
-}))
+named_picrust_data <- rowsum( picrust_data , group = ko_names[ rownames(picrust_data) ] )
+named_picrust_data[1:20,1:3]
+dim(named_picrust_data)
+named_picrust_data <- round(named_picrust_data[ rowSums(named_picrust_data>5)>=5, ])
 
-path_trend <- path_trend[rowSums(path_trend)!=0 , ]
-rownames(path_trend) <- kegg_names[ match(rownames(path_trend),kegg_names[,1]),2]
+design <- model.matrix(~group_annotation, data=metadata[sample_use,])
 
-library(edgeR)
-design <- model.matrix(~  group_annotation, data=metadata)
-
-y <- DGEList(counts=floor(path_trend), remove.zeros = T)
+y <- DGEList(counts=named_picrust_data[,sample_use])
 y <- calcNormFactors(y,method = "TMM")
-y <- estimateGLMCommonDisp(y,design)
-y <- estimateGLMTagwiseDisp(y,design)
+y <- estimateGLMCommonDisp(y)
+y <- estimateGLMTagwiseDisp(y)
 fit <- glmFit(y, design)
 
-lrt <- glmLRT(fit,coef=1:8)
+lrt <- glmLRT(fit,coef=2:8)
 top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
@@
 
-top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+top_dge <- (top$FDR < 1e-5) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
 top_dge <- rownames(top)[ top_dge ]
 
-norm_path_trend <- log2( t( t(path_trend) / colSums(path_trend) ) * 1000 + 1 )
-norm_path_trend <- norm_path_trend[top_dge, ]
-norm_path_trend <- t(apply(norm_path_trend,1,function(x){scale(x,T,T)}))
-norm_path_trend[norm_path_trend > 5] <- 5
-colnames(norm_path_trend) <- colnames(path_trend)
-sample_use <- colnames(norm_path_trend)
+top_picrust_data <- cpm(y)[top_dge, ]
+top_picrust_data <- t(apply(top_picrust_data,1,function(x){scale(x,T,T)}))
+top_picrust_data[top_picrust_data > 5] <- 5
+colnames(top_picrust_data) <- colnames(picrust_data)
 
-mypar(2,4,mar=c(1,.2,2,.2))
-h <- hclust( as.dist( (1- cor(t(norm_path_trend)))/2 ), method = "ward.D2")
-plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
-abline(v=1,xpd=F,col="red",lty=2)
-gene_module <- cutree(h, h = 1)
+layout(matrix(c(1,2,3,4,
+              1,2,3,4),
+              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
+par(mar=c(4,.1,2,.1))
 
-points( rep(0,length(gene_module)),
+h <- hclust( as.dist( (1- cor(t(top_picrust_data)))/2 ), method = "ward.D2")
+plot( rev(as.dendrogram(h)) ,xlim=c(max(h$height),-1), horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
+cutoff <- 4
+abline(v=cutoff,xpd=F,col="red",lty=2)
+gene_module <- cutree(h, h = cutoff)
+
+points( rep(-.5,length(gene_module)),
         seq(length(gene_module),1,length.out = length(gene_module)),
         col=taxa_pal[factor(gene_module[h$order])],
-        pch=16,cex=.5)
+        pch=16,cex=1,xpd=T)
 
-image( t(norm_path_trend[h$order,][nrow(norm_path_trend):1,order(group_annotation[sample_use])]),
+image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
        col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
@@
 
-par(mar=c(4,4,2,2))
+source("../enrichment_function.R")
+gmt_list <- fgsea::gmtPathways("../../../data/KEGG_pathways_to_KO.tsv")
+kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
+names(gmt_list) <- kegg_names [ match ( names(gmt_list), as.character(kegg_names[,1]) ) , 2 ]
+gmt_list <- lapply(gmt_list , function(x){
+  na.omit(ko_names[x])
+})
+res_list <- lapply( unique( gene_module[h$order] ),function(x){
+  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
+                            gmt_list = gmt_list,
+                            min_terms_pathway = 10,
+                            max_terms_pathway = 300,
+                            min_overlap = 3,
+                            sort_by_pvalue = T)
+  temp <- temp[!grepl("REGULATION",rownames(temp)),]
+  return(temp)
+} )
+names(res_list) <- unique( gene_module[h$order] )
+
+pvalues <- unlist(lapply(res_list,function(x){ -log10(as.numeric(x$pvalue)) [1:5] }))
+pvalues[is.na(pvalues)] <- 0
+terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
+terms[is.na(terms)] <- ""
+genes <- unlist(lapply(res_list,function(x){ x$genes [1:5] }))
+genes[is.na(genes)] <- ""
+
+pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
+module_color <- unlist(lapply(names(res_list),function(x){ rep(x,5) }))
+
+temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
+plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))] , use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)
+
+par(mar=c(4,.1,2,30))
+barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
+         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",xlab="-log10(pvalue)",las=1,names.arg = "")
+points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,xpd=T,
+      labels = rev(names(pvalues)),xpd=T)
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,
+      labels = rev(sub(" [(].*","",names(pvalues)) ),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
+# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
+#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
+
+mypar(2,3,mar=c(2,2,2,6))
 set.seed(1)
-UMAP_picrust <- uwot::umap(t(norm_path_trend),n_neighbors = 20,
+UMAP_picrust <- uwot::umap(t(named_picrust_data),n_neighbors = 20,
                  metric = "correlation",min_dist = .2,spread = .4,
@@
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
 ```
@@
 
-mypar(2,4,mar=c(1,.2,2,.2))
 h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
-plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
-abline(v=2,xpd=F,col="red",lty=2)
-gene_module <- cutree(h, h = 2)
+cuttoff <- 2
+gene_module <- cutree(h, h=cuttoff)
 
-points( rep(0,length(gene_module)),
+layout(matrix(c(1,2,3,4,
+              1,2,3,4),
+              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
+par(mar=c(4,.1,2,.1))
+plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
+# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
+abline(v=cuttoff,xpd=F,col="red",lty=2)
+points( rep(-.1,length(gene_module)),
         seq(length(gene_module),1,length.out = length(gene_module)),
         col=taxa_pal[factor(gene_module[h$order])],
-        pch=16,cex=.5)
+        pch=16,cex=.5,xpd=T)
 
@@
        col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
-       breaks = seq(-5,5,length.out = 100),axes=F)
-points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
+       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .01,length(sample_use)),
         col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
-        pch=16)
+        pch=16,xpd=T)
 
+source("../enrichment_function.R")
+gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
+res_list <- lapply( unique( gene_module[h$order] ),function(x){
+  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
+                            gmt_list = gmt_list,
+                            min_terms_pathway = 10,
+                            max_terms_pathway = 300,
+                            min_overlap = 3,
+                            sort_by_pvalue = T)
+  temp <- temp[!grepl("REGULATION",rownames(temp)),]
+  return(temp)
+} )
+names(res_list) <- unique( gene_module[h$order] )
+
+pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:5] }))
+pvalues[is.na(pvalues)] <- 0
+terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
+terms[is.na(terms)] <- ""
+genes <- unlist(lapply(res_list,function(x){ x$genes [1:5] }))
+genes[is.na(genes)] <- ""
+
+pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
+module_color <- unlist(lapply(names(res_list),function(x){ rep(x,5) }))
+
+temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
+plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))] , use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)
+
+par(mar=c(4,.1,2,30))
+
+barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
+         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",xlab="-log10(pvalue)",las=1,names.arg = "")
+points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,xpd=T,
+      labels = rev(names(pvalues)))
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,
+      labels = rev(sub(" [(].*","",names(pvalues)) ),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
+# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
+#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
+
+mypar(2,3,mar=c(2,2,2,6))
 set.seed(1)
@@
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+
+NESse_list <- read.csv2("../../../results/Hallmark_NESse_list.csv",row.names = 1)
+NESse_list <- lapply(unique(sub("[.].*","",colnames(NESse_list))),function(x){
+  temp <- NESse_list[ grep( x, colnames(NESse_list)) ]
+  colnames(temp) <- sub(".*[.]","",colnames(temp))
+  temp[is.na(temp)]<-0
+  return(t(temp))
+})
+pvalue_list <- read.csv2("../../../results/Hallmark_pvalues_list.csv",row.names = 1)
+pvalue_list <- lapply(unique(sub("[.].*","",colnames(pvalue_list))),function(x){
+  temp <- pvalue_list[ grep( x, colnames(pvalue_list)) ]
+  colnames(temp) <- sub(".*[.]","",colnames(temp))
+  temp[is.na(temp)]<-0
+  return(t(temp))
+})
+x <- pvalue_list[[1]]
+df <- data.frame( pathway=c(sapply(1:ncol(x),x=x,function(i,x){rownames(x)} )),
+                  bc=c(sapply(colnames(x),x=x,function(i,x){ rep(i,nrow(x)) }) ),
+                  NES=c(NESse_list[[1]]),
+                  abs_NES=abs(c(NESse_list[[1]])),
+                  pvalue=c(x))
+df <- df[df$pvalue!=0,]
+df <- df[df$pvalue<0.05,]
+df <- df[df$NES!=0,]
+
+df %>% group_by(bc)  %>% top_n(3,NES) -> top_per_bc
+df %>% group_by(bc)  %>% top_n(-3,NES) ->  bottom_per_bc
+
+as.data.frame(top_per_bc)
+
+mypar()
+x <- NESse_list[[1]] - rowMeans( NESse_list[[1]] )
+x [ pvalue_list[[1]] < 0.05 ] <- 0
+x <- x[unique(top_per_bc$pathway),]
+h <- hclust( as.dist( (1-cor(t(x)))/2 ) , method = "ward.D2" )
+image( x[h$order,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),
+       breaks=seq(-5,5,length.out = 92),
+       axes=F,border=NA,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
 ```
@@
          pt.bg = taxa_pal)
-
 ```
@@
 
+## Figure S6
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+combined_bc <- rbind( top_per_bc  , bottom_per_bc )
+
+mypar(8,2,mar=c(2,25,1,1))
+for(i in sort(levels(combined_bc$bc)) ){
+  barplot( c(combined_bc [combined_bc$bc == i,"NES" ])[[1]] , horiz = T,las=1,
+           main=i,
+           names.arg = as.character(combined_bc$pathway)[combined_bc$bc == i])
+}
+```
+
+\clearpage
+
 TABLES (MAIN)

### 2021-02-20 reports/rmarkdown/manuscript/manuscript_template.tex
adds picrust and RNAseq to manuscript

@@
 Author Three\textsuperscript{1,2}}
-\date{February 13, 2021}
+\date{February 20, 2021}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-3}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-3}
 
@@
 
+\clearpage
+
+\hypertarget{figure-s6}{%
+\subsection{Figure S6}\label{figure-s6}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
+
+\clearpage
+
 \hypertarget{tables-main}{%

### 2021-02-20 reports/rmarkdown/enrichment_function.R

@@ -1,0 +1,64 @@
+
+compute_enrichment <- function(genes , gmt_list , min_overlap = NULL, min_terms_pathway=NULL, max_terms_pathway=NULL, sort_by_pvalue=F){
+  
+  allgenes <- unique( c(unlist(gmt_list) , genes) )
+  
+  res <- t(sapply( gmt_list ,
+                   genes=genes,
+                   Total=length(allgenes),
+                   inselection=length(genes),
+                   function(x , genes,Total, inselection){
+                     Overlap <- sum(x%in%genes)
+                     inpathway <- length(x)
+                     return( c(overlap=Overlap,
+                               in_input_list=inselection,
+                               in_pathway=inpathway,
+                               all_gene=Total,
+                               percent_overlap=Overlap/inselection,
+                               overlap_coeficient=Overlap/min(inpathway,inselection),
+                               jaccard_index=Overlap / (inpathway+inselection-Overlap),
+                               pvalue=phyper(Overlap-1, 
+                                             inpathway, 
+                                             Total-inpathway, 
+                                             inselection, 
+                                             lower.tail= FALSE)) )
+                   }))
+  res <- as.data.frame(res)
+  res$FDR <- p.adjust(res$pvalue,method = "BH" )
+  res$genes <- sapply(gmt_list,genes=genes,function(x,genes) paste0(x[x%in%genes],collapse = ";") )
+  
+  if(sort_by_pvalue){ res <- res[order(res$pvalue),] }
+  if(!is.null(min_overlap)){ res <- res[res$overlap >= min_overlap, ] }
+  if(!is.null(min_terms_pathway)){ res <- res[res$in_pathway >= min_terms_pathway, ] }
+  if(!is.null(max_terms_pathway)){ res <- res[res$in_pathway <= max_terms_pathway, ] }
+  
+  return(res)
+}
+
+# 
+# mypar()
+# sel <- c("grey80","red")[ (res$in_pathway > 1000) + 1 ]
+# plot(res$jaccard_index, res$percent_overlap,cex=.2,col= sel )
+# plot(res$jaccard_index, res$overlap_coeficient,cex=.2,col= sel )
[…]

### 2021-02-23 reports/rmarkdown/03_integrated_datasets.Rmd
add metadata association and KEGG for TRX

@@
 Ftest <- sapply(metadata_parameters,function(x){
+  cat(paste0("\n",x,"\n"))
   temp <- sapply(metadata_parameters,x=x,function(x,y){
@@
     } else {
+      # if( !is.na(sum(as.numeric(metadata[,x]))) ){ f1 <- cut(as.numeric(metadata[,x]),breaks = 5 ) }
+      if( is.numeric(metadata[,x]) ){ f1 <- cut(metadata[,x],breaks = 5 ) }
     f1 <- as.character(metadata[,x])
@@
 
+    # if( !is.na(sum(as.numeric(metadata[,y]))) ){ f1 <- cut(as.numeric(metadata[,y]),breaks = 5 ) }
+    if( is.numeric(metadata[,y]) ){ f2 <- cut(metadata[,y],breaks = 5 ) }
     f2 <- as.character(metadata[,y])
@@
     f22 <- f2 [ (f1!="NA") & (f2!="NA") ]
-    contT1  <- table(f11,f22)
-    print(x)
-    print(y)
-    return( fisher.test(contT1,simulate.p.value = T)$p.value )
+    contT1  <- table(list(factor(f11,levels=unique(f1)),factor(f22,levels=unique(f2))))
+    cat(".")
+    f <- NULL
+    try(f <- fisher.test(contT1,simulate.p.value = T),silent = T)
+    if( is.null(f) ){ 
+      return( 1  ) 
+      } else {
+      return( c(f$p.value)  ) 
+    }
     # return( chisq.test(contT1,simulate.p.value = T)$p.value )
@@
   })
-  return(temp)
+
+  return( temp )
 })
@@
 clean_Ftest <- -log10(clean_Ftest)
-clean_Ftest[clean_Ftest > 5.5] <- 5.5
+clean_Ftest[clean_Ftest > 6] <- 6
 # clean_Ftest[ clean_Ftest > 0.05 ] <- 1
 
-write.csv2(clean_Ftest , "../../results/metadata_association_results.csv",row.names = T)
+write.csv2(Ftest , "../../results/metadata_association_results_complete.csv",row.names = T)
+write.csv2(clean_Ftest , "../../results/metadata_association_results_filtered.csv",row.names = T)
 

### 2021-02-23 reports/rmarkdown/manuscript/manuscript_template.Rmd
add metadata association and KEGG for TRX

@@
 SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
-clean_Ftest <- read.csv2("../../../results/metadata_association_results.csv",row.names = 1)
+clean_Ftest <- read.csv2("../../../results/metadata_association_results_complete.csv",row.names = 1)
 datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
@@
 figlabels <- letters
+# 
+# m <- matrix(as.character(c(1,2,3,4)),nrow = 2,ncol = 2)
+# m
+# m2 <- matrix(as.character(c(1,2,3,"m")),nrow = 2,ncol = 2)
+# m2
+# 
+# matrix(c(1,2,j,3),nrow = 2,ncol = 2)
+# layout_matrix <- function(data, nrow, ncol){
+#   return( matrix(as.character(data), nrow = nrow, ncol = ncol)  )
+# }
+# layout_matrix2 <- function(data, nrow, ncol, widths=c(1,2), heights=c(2,3), res=res){
+#   temp <- matrix(as.character(data), nrow = nrow, ncol = ncol)
+#   widths <- rep(widths,ncol(temp))[1:ncol(temp)]
+#   heights <- rep(heights,nrow(temp))[1:nrow(temp)]
+#   
+#   cumsum(widths)*100/sum(widths)
+#   cumsum(heights)*100/sum(heights)
+#   
+#   temp2 <- matrix(1,nrow = nrow(data),ncol = ncol(data))
+#   temp3 <- temp2 * heights
+#   temp4 <- t( t( temp2 ) * widths ) 
+#   temp2 <- matrix( "" , ncol = res , nrow = res )
+#   
+#   
+#   for(i in 1:nrow(temp)){
+#     for(j in 1:ncol(temp)){
+#       temp2 <- 
+#     }
+#   }
+#   temp <- apply(temp,1,function(x){rep( x , widths/ )})
+#   return(   )
+# }
+# 
+# m <- layout_matrix(c(1,2,3,4),2,2)
+# m2 <- layout_matrix(c(1,2,3,"m"),2,2)
 
@@
   text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
-  title(main=paste0("community",j),cex.main=.7)
+  title(main=j,cex.main=.5)
 }
@@
 # 
-par(mar=c(1,10,1,6))
-# for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-#         "ASV_CVL_V3_normalized_batch_corrected",
-#         "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-#   temp_group <- as.character(group_annotation)
-#   temp_group[ is.na(temp_group) ] <- "NA"
-#   temp_group[ temp_group == "" ] <- "NA"
-#   temp_group <- factor(temp_group)
-#   
-#   plot_dots(data=module_means_per_dataset[[j]],
-#           genes=rownames(module_means_per_dataset[[j]]),col = pal,
-#           clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
-# }
-# add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
+par(mar=c(2,10,2,6))
 plot_dots_difference(data = module_means_per_dataset[[2]],
@@
                      genes = rownames(module_means_per_dataset[[1]]),
-                     main = "CVL3 vs CVL2",cex.row = .8)
+                     main = "CVL3 vs CVL2",cex.row = .8,srt=30)
 
@@
                      genes = rownames(module_means_per_dataset[[1]]),
-                     main = "CVL3 vs tissue_V3",cex.row = .8)
-# mypar(mar=c(2,2,2,6))
-# plot(c(-2,10),c(-2,10),xaxs="i",yaxs="i")
-# rrr <- rev(as.raster(colorRampPalette(c("blue","navy","grey95","firebrick","red"))(99)))
-# rasterImage(as.raster(rrr),
-#             xleft = par("usr")[2]+diff(par("usr")[1:2])/40,
-#             xright = par("usr")[2]+diff(par("usr")[1:2])/20,
-#             ybottom = par("usr")[4]-diff(par("usr")[1:2])/4,
-#             ytop = par("usr")[4], 
-#             interpolate = FALSE )
-# text( c(par("usr")[2]+diff(par("usr")[1:2])/20,
-#         par("usr")[2]+diff(par("usr")[1:2])/20) ,
-#       c(par("usr")[4]-diff(par("usr")[3:4])/4,
-#         par("usr")[4]),labels = c("min","max"),pos=4 )
-# 
-# 
-# text( c(par("usr")[2]+diff(par("usr")[1:2])/20,
-#         par("usr")[2]+diff(par("usr")[1:2])/20,
-#         par("usr")[2]+diff(par("usr")[1:2])/20) ,
-#       c(par("usr")[4]-diff(par("usr")[1:2])/4,
-#         par("usr")[4]-diff(par("usr")[1:2])/8,
-#         par("usr")[4]),labels = c("-1","0","+1"),pos=4 )
-# text( c(par("usr")[2]+diff(par("usr")[1:2])/40) ,xpd=T,
-#       c(par("usr")[4]+diff(par("usr")[3:4])/40),labels = c("Z-score"),adj=0 )
+                     main = "CVL3 vs tissue_V3",cex.row = .8,srt=30)
 
@@
 top <- cbind(logFC.Gard=0,top)
+
+saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
+write.csv2( top , "../../../results/Picrust_EdgeR_top_dge.csv" )
 
@@
 
-par(mar=c(4,.1,2,30))
+par(mar=c(4,.1,2,20))
 barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
@@
 
+saveRDS( y , "../../../results/TRX_EdgeR_estimations.rds" )
+write.csv2( top , "../../../results/TRX_EdgeR_top_dge.csv" )
+
 top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
@@
 
-top_TRX_counts <- cpm(y)[top_dge, ]
+top_TRX_counts <- cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
 top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
@@
 gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
+res_list <- lapply( unique( gene_module[h$order] ),function(x){
+  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
+                            gmt_list = gmt_list,
+                            min_terms_pathway = 10,
+                            max_terms_pathway = 300,
+                            min_overlap = 3,
+                            sort_by_pvalue = T)
+  temp <- temp[!grepl("REGULATION",rownames(temp)),]
+  return(temp)
+} )
+names(res_list) <- unique( gene_module[h$order] )
+
+pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:5] }))
+pvalues[is.na(pvalues)] <- 0
+terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
+terms[is.na(terms)] <- ""
+genes <- unlist(lapply(res_list,function(x){ x$genes [1:5] }))
+genes[is.na(genes)] <- ""
+
+pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
+module_color <- unlist(lapply(names(res_list),function(x){ rep(x,5) }))
+
+temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
+plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))] , use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)
+
+par(mar=c(4,.1,2,20))
+
+barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
+         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",xlab="-log10(pvalue)",las=1,names.arg = "")
+points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,xpd=T,
+      labels = rev(names(pvalues)))
+text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,
+      labels = rev(sub(" [(].*","",names(pvalues)) ),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
+# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
+#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
+
+layout(matrix(c(1,2,3,4,
+              1,2,3,4),
+              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
+par(mar=c(4,.1,2,.1))
+
+plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
+# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
+abline(v=cuttoff,xpd=F,col="red",lty=2)
+points( rep(-.1,length(gene_module)),
+        seq(length(gene_module),1,length.out = length(gene_module)),
+        col=taxa_pal[factor(gene_module[h$order])],
+        pch=16,cex=.5,xpd=T)
+
+image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(group_annotation[sample_use])]),
+       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
+       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .01,length(sample_use)),
+        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
+        pch=16,xpd=T)
+
+source("../enrichment_function.R")
+gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
 res_list <- lapply( unique( gene_module[h$order] ),function(x){
@@
 })
-x <- pvalue_list[[1]]
+x <- pvalue_list[[2]]
 df <- data.frame( pathway=c(sapply(1:ncol(x),x=x,function(i,x){rownames(x)} )),
@@
 
-mypar()
-x <- NESse_list[[1]] - rowMeans( NESse_list[[1]] )
-x [ pvalue_list[[1]] < 0.05 ] <- 0
+mypar(1,2,mar=c(4,10,2,4))
+for(i in 1:2){
+x <- NESse_list[[i]]
+x [ pvalue_list[[i]] < 1 ] <- 0
 x <- x[unique(top_per_bc$pathway),]
-h <- hclust( as.dist( (1-cor(t(x)))/2 ) , method = "ward.D2" )
-image( x[h$order,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),
-       breaks=seq(-5,5,length.out = 92),
-       axes=F,border=NA,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+# h <- hclust( as.dist( (1-cor(t(x)))/2 ) , method = "ward.D2" )
+# image( x[h$order,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),
+#        breaks=seq(-5,5,length.out = 92),
+#        axes=F,border=NA,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
+
+sizes <- pvalue_list[[i]]
+sizes[sizes < 2] <- 0
+sizes[sizes > 10] <- 10
+sizes <- sizes / max(sizes)
+
+plot( sort(rep(1:ncol(x),nrow(x))),
+      rep(nrow(x):1,ncol(x)),
+        col=colorRampPalette(c("navy","grey95","firebrick"))(91)[cut( c(x) , breaks = seq(-5,5,length.out = 92) )],pch=16 ,
+      cex= sizes*2 + 0.2,
+      axes=F,xlab="",ylab="", frame=T)
+srt=30
+text(1:ncol(x), par("usr")[3] - (par("usr")[4])/200, labels = colnames(x), srt = srt,
+     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
+     xpd = TRUE, cex=.6)
+text(par("usr")[1] - (par("usr")[2])/200, 1:nrow(x) , labels = rownames(x), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=.6)
+
+add_scale_legend(pal = colorRampPalette(c("navy","grey95","firebrick"))(91),labels = c("-5","0","5"))
+add_size_legend(labels = c("> 0.01","< 1e-6","< 1e-10"),max.cex = 2.2, min.cex = .2)
+}
+
 ```
@@
 
+continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+continuous_metadata <- metadata[,continuous_parameters]
+# continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )
+
 mypar(1,4,mar=c(12,1,1,0))
@@
            "ASV_tissue_V3_normalized_batch_corrected")){
-
-  continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
-  continuous_metadata <- metadata[,continuous_parameters]
-  # continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )
-
   metadata_cor <- cor(t(module_means_per_dataset[[i]]),continuous_metadata,use = "pairwise.complete.obs")
@@
 }
-
 par(mar=c(12,0,1,0))
@@
   labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+```
+
+\clearpage
+
+## Figure S2 v2
+
+```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
+figlabels <- letters
+temp_legend <- NULL
+
+mypar(1,4,mar=c(12,1,1,0))
+
+for(k in c("ASV_CVL_V3_normalized_batch_corrected",
+           "ASV_CVL_V2_normalized_NOT_batch_corrected",
+           "ASV_tissue_V3_normalized_batch_corrected")){
+res2 <- lapply( 1:nrow(module_means_per_dataset[[k]]),k=k, function(i,k){
+  res <- lapply( c( colnames(clean_Ftest) , continuous_parameters ) , i=i,k=k,function(j,i,k) {
+    model <- paste0("module_means_per_dataset[['",k,"']][",i,",] ~ ",j)
+    m <- summary(lm( eval(parse(text=model)), data = metadata ))
+    f <- m$fstatistic
+    p <- pf(f[1],f[2],f[3],lower.tail=F)
+    temp <- c(m$r.squared,m$adj.r.squared,m$fstatistic,p)
+    return(temp)
+  })
+  res <- t(as.data.frame(res))
+  rownames(res) <- c( colnames(clean_Ftest) , continuous_parameters )
+  colnames(res) <- c("r.squared","adj.r.squared","F-statistic","df1","df2","pvalue")
+  return(res)
+})
+names(res2) <- rownames(module_means_per_dataset[[k]])
+
+pvals <- sapply(res2, function(x){x[,"pvalue"]})
+rs <- sapply(res2, function(x){x[,"r.squared"]})
+
+sizes <- -log10( pvals )
+sizes[sizes < 2] <- 0
+sizes[sizes > 11] <- 10
+sizes <- sizes / max(sizes)
+
+plot(0,0, las=1,xlim=c(.5,ncol(rs)+.5),ylim=c(.5,nrow(rs)+.5),
+       axes=F,ylab="",xlab="",main=gsub("^ASV_|_norm.*","",k),xaxs="i",yaxs="i", frame=T)
+for(i in 1:ncol(rs)){
+  lines(c(i,i),c(nrow(rs)+.5,0),col="grey95",lwd=.5,xpd=F)
+}
+for(i in 1:nrow(rs) ){
+  lines(c(0,ncol(rs)+0.5),c(i,i),col="grey95",lwd=.5,xpd=F)
+}
+points(sort(rep(1:ncol(rs),nrow(rs))),
+      rep(nrow(rs):1,ncol(rs)),
+        col=colorRampPalette(c("navy","grey95","firebrick"))(91)[cut( c(rs) , breaks = seq(-1,1,length.out = 92) )],pch=16 ,
+      cex= sizes*2 + 0.2)
+srt=90
+text(1:ncol(rs), par("usr")[3] - (par("usr")[4])/200, labels = colnames(rs), srt = srt,
+     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
+     xpd = TRUE, cex=1)
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+}
+
+empty_plot(xlim=c(.5,ncol(rs)+.5),ylim=c(.5,nrow(rs)+.5),xaxs="i",yaxs="i")
+text(par("usr")[1] , 1:nrow(rs) , labels = rownames(rs), srt = 0, adj = c(0,0.5), xpd = TRUE, cex=1)
+
+add_scale_legend(pal = colorRampPalette(c("grey90","firebrick"))(91),labels = c("0","0.5","1"))
+add_size_legend(labels = c("> 0.01","< 1e-5","< 1e-8","< 1e-11"),max.cex = 2.2, min.cex = .2)
+
+```
+
+\clearpage
+
+## Figure S2 v3
+
+```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
+metadata_parameters <- colnames(metadata)[-c(1:2)]
+
+set.seed(1)
+noise <- matrix(rnorm(prod(dim(clean_Ftest)),mean = 0,sd = 1e-10),nrow(clean_Ftest),ncol(clean_Ftest))
+o <- hclust(as.dist( (1-cor(clean_Ftest+noise)/2) ),method = "ward.D2")$order
+
+sizes <- as.matrix(-log10(clean_Ftest))
+# sizes[sizes < 2] <- 0
+sizes[sizes > 7] <- 7
+sizes <- sizes / max(sizes)
+sizes <- sizes * lower.tri(sizes)
+
+mypar(1,1,mar=c(10,10,2,6))
+plot(0,0, las=1,xlim=c(.5,ncol(sizes)+.5),ylim=c(.5,nrow(sizes)+.5),
+       axes=F,ylab="",xlab="",main="",xaxs="i",yaxs="i", frame=T)
+for(i in 1:ncol(sizes)){
+  lines(c(i,i),c(nrow(sizes)+.5,0),col="grey95",lwd=.5,xpd=F)
+}
+for(i in 1:nrow(sizes) ){
+  lines(c(0,ncol(sizes)+0.5),c(i,i),col="grey95",lwd=.5,xpd=F)
+}
+points(sort(rep(1:ncol(sizes),nrow(sizes))),
+      rep(nrow(sizes):1,ncol(sizes)),
+        col=colorRampPalette(c("grey95","orange3","firebrick","black"))(91)[cut( c(sizes) , breaks = seq(0,1,length.out = 92) )],pch=16 ,
+      cex= sizes*1 + 0.2)
+srt=90
+text(1:ncol(sizes), par("usr")[3] - (par("usr")[4])/200, labels = colnames(sizes), srt = srt,
+     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
+     xpd = TRUE, cex=1)
+text(par("usr")[1] , 1:nrow(sizes) , labels = rownames(sizes), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=1)
+
+add_scale_legend(pal = colorRampPalette(c("grey95","orange3","firebrick","red"))(91),labels = c("> 0.01","< 1e-5","< 1e-8","< 1e-11"))
+add_size_legend(labels = c("> 0.01","< 1e-5","< 1e-8","< 1e-11"),max.cex = 1.2, min.cex = .2)
 ```
@@
 
+## Figure S7
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+figlabels <- letters
+mypar(3,3,mar=c(4,10,2,2))
+for( j in c("ASV_tissue_V3_normalized_batch_corrected",
+        "ASV_CVL_V3_normalized_batch_corrected",
+        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
+  temp_group <- as.character(group_annotation)
+  temp_group[ is.na(temp_group) ] <- "NA"
+  temp_group[ temp_group == "" ] <- "NA"
+  temp_group <- factor(temp_group)
+
+  plot_dots(data=module_means_per_dataset[[j]],srt=45,
+          genes=rownames(module_means_per_dataset[[j]]),col = pal,
+          clustering=temp_group, main=paste0(sub("_norm.*","",j)))
+  add_letter(figlabels[1]); figlabels <- figlabels[-1]
+  empty_plot()
+  empty_plot()
+}
+```
+
+\clearpage
+
+## Figure S8
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+mypar(mar=c(10,6,1,1))
+
+top_all <- lapply( datasets_all_samples[2:4] , function(x){
+   sort(apply(x,1,var),decreasing = T)[1:20]} )
+top <- unique( sub( ".*[.]" , "" , names(unlist(top_all)) ) )
+
+x <- t(sapply(datasets_all_samples[2:4],function(x){
+  c(t(x))
+}))
+colnames(x) <- c(sapply(rownames(datasets_all_samples[[2]]),
+                        function(i){rep(i,ncol(datasets_all_samples[[2]]))}))
+rownames(x) <- gsub("ASV_|_norm.*", "",rownames(x))
+
+top <- sort(sapply( unique(colnames(x)),
+               function(i) { mean( x[,colnames(x)==i ] ) } ),decreasing = T)
+top <- names(top)[1:30]
+
+violist(data = x[,colnames(x)%in%top],
+        genes = rownames(x),
+        clustering = factor( colnames(x)[colnames(x)%in%top],levels = top) ,
+        srt=45,
+        transparency = 50,col = pal)
+```
+
+\clearpage
+
+## Figure S9
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+
+```
+
+\clearpage
+
 TABLES (MAIN)

### 2021-02-23 reports/rmarkdown/manuscript/manuscript_template.tex
add metadata association and KEGG for TRX

@@
 Author Three\textsuperscript{1,2}}
-\date{February 20, 2021}
+\date{February 23, 2021}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-4}
 
@@
 
+\hypertarget{figure-s2-v2}{%
+\subsection{Figure S2 v2}\label{figure-s2-v2}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-10-1}
+
+\clearpage
+
+\hypertarget{figure-s2-v3}{%
+\subsection{Figure S2 v3}\label{figure-s2-v3}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-1}
+
+\clearpage
+
 \hypertarget{figure-s3}{%
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-10-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-3}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-3}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-15-1}
+
+\clearpage
+
+\hypertarget{figure-s7}{%
+\subsection{Figure S7}\label{figure-s7}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-16-1}
+
+\clearpage
+
+\hypertarget{figure-s8}{%
+\subsection{Figure S8}\label{figure-s8}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-17-1}
+
+\clearpage
+
+\hypertarget{figure-s9}{%
+\subsection{Figure S9}\label{figure-s9}}
 

### 2021-02-23 reports/rmarkdown/enrichment_function.R

@@ -1,4 +1,3 @@
-
 compute_enrichment <- function(genes , gmt_list , min_overlap = NULL, min_terms_pathway=NULL, max_terms_pathway=NULL, sort_by_pvalue=F){
   
   allgenes <- unique( c(unlist(gmt_list) , genes) )

### 2021-03-03 reports/rmarkdown/manuscript/manuscript_template.Rmd

@@
 library(kableExtra)
-library(ggplot2)
+library(tidyverse)
 library(igraph)
-library(dplyr)
-library(stringr)
+library(phyloseq)
+library(vegan)
+library(gridExtra)
 library(rafalib)
@@
 
+## Figure S10
+
+```{r, echo = FALSE, eval = TRUE, fig.width=6, fig.height=10}
+paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3
+             '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
+             '../../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region
+
+n <- c("boston_r1", "boston_r2", "Sthlm")
+ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+
+# remove controls and duplicates
+ps_filt <- map(ps_list, ~ .x %>%
+  subset_samples(., Include_final == "yes"))
+
+# separate tissue from CVLv2 samples in the second Boston run
+CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == 2) 
+tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == 3)
+
+# merge the lists
+ps_types <- c(ps_filt, "CVLv2"=CVLv2, "tissue"=tissue) 
+# Get ASV tables
+seqtabs <- map(ps_types, ~data.frame(.x@otu_table@.Data))
+
+# Get rarefaction curves
+raremax <- map(seqtabs, ~min(rowSums(.x)))
+rare_runs <- map2(seqtabs, raremax, ~ .x %>%
+             rarecurve(., step = 20, sample = .y, 
+                       label = FALSE#,col = col, lty = lty,
+          )
+)
+
+ggplot_df_fun <- function(rare_obj, count) {
+  # Add sample names 
+  names(rare_obj) <- c(paste0("sample",sprintf("%03.0f", 1:length(rare_obj))))
+
+  # Coerce data into "long" form.
+  rare_df <- imap_dfr(rare_obj, ~ .x %>%
+              as_tibble(., rownames = "Sample size") %>%
+              mutate(Sample = .y) %>%
+              mutate(`Sample size` = as.numeric(gsub("N","",`Sample size`)))
+  )
+  # Get nr. of counts in each sample 
+  c <- formatC(count, format="f", big.mark = " ", digits=0)
+  rare_df_col <- rare_df %>%
+    rename(ASV = "value") %>%
+    group_by(Sample) %>% 
+    nest() %>%
+    mutate(Counts = map(data, ~max(.$`Sample size`))) %>%
+    unnest(c(data, Counts)) %>%
+    mutate(Counts = ifelse(Counts > count, paste0(">", c), paste0("<", c)))
+
+  return(rare_df_col)
+}
+
+# Define some plot variables
+count <- c(33000, 10000, 33000, 20000, 2500)
+limit <- c(65000, 65000, 65000, 65000, 10000)
+title <- c("Run B1. CVLv3", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B2. CVLv2", "Run B2. TISSUEv3")
+
+# Extract data frame from the rarecurve object
+rare_df_col<- map2(rare_runs, count, ~ggplot_df_fun(.x, .y))
+# Create Plot 
+rare_plot <- map2(rare_df_col, raremax, 
+              ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
+              theme_bw() +
+              scale_color_manual(values=c('tomato', alpha('black', 0.1))) + 
+              geom_line(aes(color = Counts, group = Sample)) +
+              theme(plot.title = element_text(hjust = 0.5))
+            )
+
+r_plot <- pmap(list(rare_plot, raremax, limit, title),
+               ~ ..1 + xlim(c(0,..3)) +
+               labs(title=paste0("Rarefaction Curve ", ..4)) +
+               geom_vline(xintercept = ..2, 
+                          linetype='dashed', 
+                          color='slategray' )
+)
+
+# Save png
+#imap(r_plot, ~ggsave(filename = paste0('Rarefaction_plot_', .y, '.png'), plot = .x, 
+#        path = paste0("../../../results/"), 
+#        width = 10, height = 5,)
+#)
+
+grid.arrange(
+  r_plot[[1]],
+  r_plot[[3]],
+  r_plot[[4]],
+  r_plot[[5]],
+  nrow = 4,
+  #height = c(1, 1, 1, 1),
+  clip = FALSE
+)
+```
+
+\clearpage
+
 TABLES (MAIN)

### 2021-03-03 reports/rmarkdown/manuscript/manuscript_template.Rmd
updates KEGG and correlation matricies

--- 2021-02-23 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 par(mar=c(12,1,1,12))
-image( t(clean_Ftest [nrow(clean_Ftest):1,]),axes=F,
+set.seed(1)
+noise <- matrix(rnorm(prod(dim(clean_Ftest)),mean = 0,sd = 1e-10),nrow(clean_Ftest),ncol(clean_Ftest))
+o <- hclust(as.dist( (1-cor(clean_Ftest+noise)/2) ),method = "ward.D2")$order
+
+sizes <- as.matrix(clean_Ftest[o,o])
+sizes[sizes > 7] <- 7
+sizes <- sizes / max(sizes)
+image( t(sizes [nrow(sizes):1,]),axes=F,
        col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
-text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(clean_Ftest)) ,
-       labels = rownames(clean_Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-text( seq(0,1,length.out = nrow(clean_Ftest)) , par("usr")[3]-.02 ,
-      labels = rownames(clean_Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
+text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(sizes)) ,
+       labels = rownames(sizes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
+text( seq(0,1,length.out = nrow(sizes)) , par("usr")[3]-.02 ,
+      labels = rownames(sizes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
 
@@
 
-```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
+```{r, echo = FALSE, eval = TRUE, fig.asp=1}
 figlabels <- letters
@@
 rs <- sapply(res2, function(x){x[,"r.squared"]})
+rs[is.nan(rs)] <- 0
+pvals[is.nan(pvals)] <- 1
 
@@
      adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
-     xpd = TRUE, cex=1)
+     xpd = TRUE, cex=.8)
 add_letter(figlabels[1]); figlabels <- figlabels[-1]
@@
 empty_plot(xlim=c(.5,ncol(rs)+.5),ylim=c(.5,nrow(rs)+.5),xaxs="i",yaxs="i")
-text(par("usr")[1] , 1:nrow(rs) , labels = rownames(rs), srt = 0, adj = c(0,0.5), xpd = TRUE, cex=1)
+text(par("usr")[1] , 1:nrow(rs) , labels = rownames(rs), srt = 0, adj = c(0,0.5), xpd = TRUE, cex=.8)
 
@@
 
-```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
+```{r, echo = FALSE, eval = TRUE, fig.asp=.9}
 metadata_parameters <- colnames(metadata)[-c(1:2)]
@@
 
-sizes <- as.matrix(-log10(clean_Ftest))
+sizes <- as.matrix(clean_Ftest[o,o])
 # sizes[sizes < 2] <- 0
@@
      xpd = TRUE, cex=1)
-text(par("usr")[1] , 1:nrow(sizes) , labels = rownames(sizes), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=1)
+text(par("usr")[1] , nrow(sizes):1 , labels = rownames(sizes), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=1)
 
@@
 
-```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.3}
+layout( matrix(1:(16*8), 16,8,byrow = F) )
+par(mar=c(.5,.5,1,.5))
+
 figlabels <- letters
 for(j in names(module_means_per_dataset)){
-  mypar(2,8,mar=c(0.5,0.5,4,0.5))
-
-  plot(UMAP_picrust,bg=pal[group_annotation],main="patient groups",
+
+  plot(UMAP_picrust,bg=pal[group_annotation],main=gsub("ASV_|_norm.*","",j),
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
@@
          col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
-         main=rownames(module_means_per_dataset[[j]])[i],
+         main="",
          pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
@@
 }
+empty_plot()
+
+text( par("usr")[1], mean(par("usr")[3:4]),
+        adj=c(0,0.5),xpd=T,
+        labels = "patient groups" )
+for(i in 1:nrow(module_means_per_dataset[[j]])){
+  empty_plot()
+  text( par("usr")[1], mean(par("usr")[3:4]),
+        adj=c(0,0.5),xpd=T,
+        labels = sub( "_","\n",rownames(module_means_per_dataset[[j]])[i]) )
+}
+
 ```
@@
 
-```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.3}
+layout( matrix(1:(16*8), 16,8,byrow = F) )
+par(mar=c(.5,.5,1,.5))
+
 figlabels <- letters
 for(j in names(module_means_per_dataset)){
-  mypar(2,8,mar=c(0.5,0.5,4,0.5))
-
-  plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="patient groups",
+
+  plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main=gsub("ASV_|_norm.*","",j),
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
@@
          col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
-         main=rownames(module_means_per_dataset[[j]])[i],
+         main="",xpd=T,
          pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
@@
 }
+
+empty_plot()
+text( par("usr")[1], mean(par("usr")[3:4]),
+        adj=c(0,0.5),
+        labels = "patient groups" )
+for(i in 1:nrow(module_means_per_dataset[[j]])){
+  empty_plot()
+  text( par("usr")[1], mean(par("usr")[3:4]),
+        adj=c(0,0.5),xpd=T,
+        labels = sub( "_","\n",rownames(module_means_per_dataset[[j]])[i]) )
+}
+
 ```
@@
 
-```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+```{r, echo = FALSE, eval = TRUE,fig.asp=.8}
 figlabels <- letters
-mypar(3,3,mar=c(4,10,2,2))
+mypar(3,3,mar=c(4,12,2,4))
 for( j in c("ASV_tissue_V3_normalized_batch_corrected",
@@
 
-```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+```{r, echo = FALSE, eval = TRUE,fig.asp=.5}
 mypar(mar=c(10,6,1,1))

### 2021-03-03 reports/rmarkdown/manuscript/manuscript_template.tex
updates KEGG and correlation matricies

@@
 Author Three\textsuperscript{1,2}}
-\date{February 23, 2021}
+\date{March 03, 2021}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-3}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-3}
 

### 2021-03-11 reports/rmarkdown/01_data_pre_processing.Rmd
Changes to run on Vildes computer

@@
 output:
-  pdf_document:
-    toc: yes
+  #pdf_document:
+    #toc: yes
   html_document:
@@
 PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data
+PATH <- "../../data/" #Path to the data
 
@@
 rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
-rownames(taxa_info) <- sub("[.][0123456789","other ",rownames(taxa_info))
+rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
 sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
@@
 trx  <- trx[ rownames(trx) != "" , ]
+trx  <- trx[ rownames(trx) != "NA" , ]
 colnames(trx) <- as.character(metadata$ID)[ match(colnames(trx) , as.character(metadata$PatID) ) ]

### 2021-03-11 reports/rmarkdown/03_integrated_datasets.Rmd
Changes to run on Vildes computer

@@
 
-```{r, echo = FALSE}
+```{r eval=FALSE, include=FALSE}
 groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){

### 2021-03-11 reports/rmarkdown/manuscript/manuscript_template.Rmd
Changes to run on Vildes computer

@@
 par(mar=c(0,0,0,0))
-fig1a <- magick::image_read("schematics.pdf",density = 900)
+#fig1a <- magick::image_read("schematics.pdf",density = 900)
 plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
-rasterImage(fig1a, 0, 0, 1, 1)
+#rasterImage(fig1a, 0, 0, 1, 1)
 

### 2021-03-11 reports/rmarkdown/02_individual_reports.Rmd
Changes to run on Vildes computer

@@
 library(batchelor)
+library(scales)
 remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
@@
 PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
+PATH <- "../.." #Path to the data
 

### 2021-03-11 reports/rmarkdown/03_integrated_datasets.Rmd
Changes to run on Vildes computer

@@
 # Load metadata
-metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
+metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatefeb22.csv"))
 rownames(metadata) <- as.character(metadata$ID)

### 2021-03-11 reports/rmarkdown/manuscript/manuscript_template.Rmd
Changes to run on Vildes computer

--- 2021-03-03 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 library(kableExtra)
-library(ggplot2)
+library(dplyr)
+library(tidyverse)
 library(igraph)
-library(dplyr)
-library(stringr)
+library(phyloseq)
+library(vegan)
+library(ggpubr)
 library(rafalib)
@@
 picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
+picrust_KO_seq <- read.table("../../../data/KO_predicted.tsv",header = T,row.names = 1)
+picrust_taxa <- read.table("../../../data/SeqID_to_taxa.csv",header = T,row.names = 1,sep = ",")
 ```
@@
 big_groups <- c(
-  L.inners=    "5",
-  L.inners2=   "4",
-  L.crispatus= "2",
-  Gard=        "3",
-  Gard.Prev=   "6",
-  Gard.Prev2=  "8",
-  Gard.Lacto=  "7",
-  Patho=       "1"
+  g2_L.iners=    "5",
+  g3_L.iners2=   "4",
+  g1_L.crispatus= "2",
+  g5_Gard=        "3",
+  g6_Gard.Prev=   "6",
+  g7_Gard.Prev2=  "8",
+  g4_Gard.Lacto=  "7",
+  g8_Patho=       "1"
 )
@@
 sizes <- as.matrix(clean_Ftest[o,o])
-sizes[sizes > 7] <- 7
+sizes[sizes > 15] <- 15
 sizes <- sizes / max(sizes)
@@
 ###################################
-par(mar=c(1,5,2,1))
+par(mar=c(3,5,2,1))
 x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
@@
          clustering = group_annotation,
-         draw_mean_lines=F,col = pal)
+         draw_mean_lines=F,col = pal,srt=20)
 
@@
 ###################################
-par(mar=c(1,5,1,12))
+par(mar=c(2,5,1,12))
 
@@
 rownames(divers) <- c("Shannon","simpson","Inv Simpson")
-barlist( data = divers,
+# barlist( data = divers,
+#          main = "",
+#          genes = c("Shannon","simpson","Inv Simpson"),
+#          clustering = group_annotation,
+#          draw_mean_lines=F,col = pal,srt=30)
+  barlist( data = divers[,order(group_annotation)],
          main = "",
          genes = c("Shannon","simpson","Inv Simpson"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal)
+         # clustering = group_annotation,
+         draw_mean_lines=F,col = pal[factor(group_annotation[order(group_annotation)])])
 
@@
 barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
-         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
 legend(x = par("usr")[c(2)],
@@
        pt.bg = taxa_pal)
+# 
+# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-1,
+#         col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][sample_o],
+#         pch=16,cex=.5,xpd=T)
+# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
+#         col= c("grey","red")[as.numeric(metadata$HIVstatus)][sample_o],
+#         pch=16,cex=.5,xpd=T)
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
+        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
+        pch=15,cex=.8,xpd=T)
+  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)
+
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-6,
+        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
+        pch=15,cex=.8,xpd=T)
+  text(par("usr")[1],par("usr")[3]-6,labels = "HIVstatus",pos=2,cex=.8,xpd=T)
 
@@
         pch=16)
+points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.005,
+        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
+        pch=16,xpd=T)
+points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.01,
+        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
+        pch=16,xpd=T)
 
@@
 
+```{r, echo = FALSE, eval = TRUE}
+taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+dim(picrust_taxa)
+dim(picrust_KO_seq)
+
+#Convert picrust to the same bacterial names in the manuscript
+temp <- as.matrix(picrust_taxa)
+temp[is.na(temp)] <- "NA"
+sps <- apply( temp[,8:9] ,1,function(x) paste(x,collapse ="_"))
+not_complete <- grepl("_NA",sps)
+taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
+taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
+taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
+taxa[!not_complete] <- sps[!not_complete]
+taxa <- sub("_"," ",taxa)
+length(taxa)
+picrust_taxa$final_name <- taxa
+
+dim(picrust_data)
+dim(picrust_KO_seq)
+
+common <- rownames(picrust_data)[rownames(picrust_data) %in% colnames(picrust_KO_seq)]
+res <- t( t(as.matrix(picrust_data[common,])) %*% t(as.matrix(picrust_KO_seq[,common])) )
+picrust_KO_seq <- picrust_KO_seq[,common]
+
+picrust_KO_seq[1:100,1:100]
+
+picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , as.character(picrust_taxa$Seq_ID)  ),"final_name"])
+picrust_KO_seq <- as.matrix(picrust_KO_seq)
+
+picrust_KO_seq[,1]
+
+sub(" [(].*","",names(pvalues))[1]
+
+ress <- sapply( sub(" [(].*","",names(pvalues)) , function(x){
+  inn <- colnames(picrust_KO_seq) [ colnames(picrust_KO_seq) %in% names(gmt_list[[x]] ) ]
+  temp <- picrust_KO_seq[,inn]
+  temp <- temp[ order(rowMeans(temp),decreasing = T) ,  ]
+  common_bacs <- rownames(SNN_bacteria)[rownames(SNN_bacteria) %in% rownames(temp)]
+  temp <- rowsum(temp[common_bacs , ] , group = bac_communities [ rownames(SNN_bacteria) %in% common_bacs ] )
+  return( temp )
+} )
+
+ress
+
+mypar(1,2,mar=c(4,10,2,4))
+x <- sapply(ress,function(x){
+  rowMeans(x)
+})
+x [ is.nan(x) ] <- 0
+x <- t(x)
+
+sizes <- sapply(ress,function(x){
+  return( rowSums( x > 0) / ncol(x) )
+})
+sizes [ is.nan(sizes) ] <- 0
+sizes <- t(sizes)
+
+plot( sort(rep(1:ncol(x),nrow(x))),
+      rep(nrow(x):1,ncol(x)),
+        col=colorRampPalette(c("grey95","firebrick"))(91)[cut( c(x) , breaks = seq(0,15,length.out = 92) )],pch=16 ,
+      cex= sizes*2 + 0.2, ylim=c(0.5,nrow(x)+.5),yaxs="i",
+      axes=F,xlab="",ylab="", frame=T)
+srt=30
+text(1:ncol(x), par("usr")[3] - (par("usr")[4])/200, labels = colnames(x), srt = srt,
+     adj = c(ifelse(srt==0,.5,ifelse(srt==90,1,1)),ifelse(srt==0,1,ifelse(srt==90,.5,1))),
+     xpd = TRUE, cex=.6)
+text(par("usr")[1] - (par("usr")[2])/200, nrow(x):1 , labels = rownames(x), srt = 0, adj = c(1,0.5), xpd = TRUE, cex=.6)
+
+add_scale_legend(pal = colorRampPalette(c("grey95","firebrick"))(91),labels = c(min(x),round(max(x)) ))
+add_size_legend(labels = c("0%","50%","100%"),max.cex = 2.2, min.cex = .2)
+
+```
+
 \clearpage
@@
 
-image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(group_annotation[sample_use])]),
+image( t(top_TRX_counts[h$order,order(group_annotation[sample_use])][nrow(top_TRX_counts):1,]),
        col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
@@
         col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
+        pch=16,xpd=T)
+
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
+        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
+        pch=16,xpd=T)
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.01,
+        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
         pch=16,xpd=T)
@@
                 2,7,
-                3,7,
+                3,8,
                 4,8,
@@
   barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
-           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
+           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
   # legend(x = par("usr")[c(2)],
@@
   #        pt.bg = taxa_pal)
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
+        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
+        pch=15,cex=1,xpd=T)
+  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)
+
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-6,
+        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
+        pch=15,cex=1,xpd=T)
+  text(par("usr")[1],par("usr")[3]-6,labels = "HIVstatus",pos=2,cex=.8,xpd=T)
+
 }
@@
 
-continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_ng_mL_v3","Plasma_S_Estradiol_pg_mL_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
 continuous_metadata <- metadata[,continuous_parameters]
@@
 # sizes[sizes < 2] <- 0
-sizes[sizes > 7] <- 7
+sizes[sizes > 10] <- 10
 sizes <- sizes / max(sizes)
@@
       rep(nrow(sizes):1,ncol(sizes)),
-        col=colorRampPalette(c("grey95","orange3","firebrick","black"))(91)[cut( c(sizes) , breaks = seq(0,1,length.out = 92) )],pch=16 ,
+        col=colorRampPalette(c("grey95","orange3","firebrick","red"))(91)[cut( c(sizes) , breaks = seq(0,1,length.out = 92) )],pch=16 ,
       cex= sizes*1 + 0.2)
@@
 
+## Figure S10
+
+```{r, echo = FALSE, eval = TRUE, fig.show='hide'}
+paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3
+             '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
+             '../../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region
+
+n <- c("boston_r1", "boston_r2", "Sthlm")
+ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+
+# remove controls and duplicates
+ps_filt <- map(ps_list, ~ .x %>%
+  subset_samples(., Include_final == "yes"))
+
+# separate tissue from CVLv2 samples in the second Boston run
+CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v2") 
+tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == "v3")
+
+# merge the lists
+ps_types <- c(ps_filt, "CVLv2"=CVLv2, "tissue"=tissue) 
+# Get ASV tables
+seqtabs <- map(ps_types, ~data.frame(.x@otu_table@.Data))
+
+# Get rarefaction curves
+rare_runs <- map(seqtabs, ~ .x %>%
+             rarecurve(., step = 20, label = FALSE)
+)
+
+ggplot_df_fun <- function(rare_obj, count) {
+  # Add sample names 
+  names(rare_obj) <- c(paste0("sample",sprintf("%03.0f", 1:length(rare_obj))))
+
+  # Coerce data into "long" form.
+  rare_df <- imap_dfr(rare_obj, ~ .x %>%
+              as_tibble(., rownames = "Sample size") %>%
+              mutate(Sample = .y) %>%
+              mutate(`Sample size` = as.numeric(gsub("N","",`Sample size`)))
+  )
+  # Get nr. of counts in each sample 
+  c <- formatC(count, format="f", big.mark = " ", digits=0)
+  rare_df_col <- rare_df %>%
+    dplyr::rename(ASV = "value") %>%
+    group_by(Sample) %>% 
+    nest() %>%
+    mutate(Counts = map(data, ~max(.$`Sample size`))) %>%
+    unnest(c(data, Counts)) %>%
+    mutate(Counts = ifelse(Counts > count, paste0(">", c), paste0("<", c)))
+
+  return(rare_df_col)
+}
+
+# Define some plot variables
+count <- c(33000, 10000, 33000, 20000, 2500)
+limit <- c(65000, 65000, 65000, 65000, 10000)
+title <- c("Run B1. CVLv3", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B2. CVLv2", "Run B2. TISSUEv3")
+
+# Extract data frame from the rarecurve object
+rare_df_col<- map2(rare_runs, count, ~ggplot_df_fun(.x, .y))
+# Create Plot 
+rare_plot <- map(rare_df_col, 
+              ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
+              theme_bw() +
+              scale_color_manual(values=c('tomato', alpha('black', 0.1))) + 
+              geom_line(aes(color = Counts, group = Sample)) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.margin=unit(c(0,0,0.5,0),"cm"))
+            )
+raremax <- map(seqtabs, ~min(rowSums(.x)))
+r_plot <- pmap(list(rare_plot, raremax, limit, title),
+               ~ ..1 + xlim(c(0,..3)) +
+               labs(title=paste0("Sequencing ", ..4)) +
+               geom_vline(xintercept = ..2, 
+                          linetype='dashed', 
+                          color='slategray' )
+)
+
+# Save png
+#imap(r_plot, ~ggsave(filename = paste0('Rarefaction_plot_', .y, '.png'), plot = .x, 
+#        path = paste0("../../../results/"), 
+#        width = 10, height = 5,)
+#)
+
+```
+
+```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
+ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),
+          r_plot[[3]] + theme(axis.title.x=element_blank()),
+          r_plot[[4]] + theme(axis.title.x=element_blank()),
+          r_plot[[5]],
+          ncol = 1, 
+          labels = c("A","B","C","D")
+          )
+```
+
+**Supp.
+Figure 10.
+Rarefraction curves**.
+**A** Boston 1 sequencing run, **B** Sthlm sequencing run and **C** and **D** shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
+
+\clearpage
+
 TABLES (MAIN)

### 2021-03-12 reports/rmarkdown/manuscript/manuscript_template.Rmd
updates in figures

@@
 library(vegan)
-library(ggpubr)
+library(gridExtra)
 library(rafalib)
@@
 par(mar=c(0,0,0,0))
-#fig1a <- magick::image_read("schematics.pdf",density = 900)
+fig1a <- magick::image_read("schematics.pdf",density = 900)
 plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
-#rasterImage(fig1a, 0, 0, 1, 1)
+rasterImage(fig1a, 0, 0, 1, 1)
 
@@
 
-```{r, echo = FALSE, eval = TRUE, fig.show='hide'}
+```{r, echo = FALSE, eval = TRUE, fig.width=6, fig.height=10}
 paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3
@@
 # separate tissue from CVLv2 samples in the second Boston run
-CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v2") 
-tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == "v3")
+CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == 2) 
+tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == 3)
 
@@
 # Get rarefaction curves
-rare_runs <- map(seqtabs, ~ .x %>%
-             rarecurve(., step = 20, label = FALSE)
+raremax <- map(seqtabs, ~min(rowSums(.x)))
+rare_runs <- map2(seqtabs, raremax, ~ .x %>%
+             rarecurve(., step = 20, sample = .y, 
+                       label = FALSE#,col = col, lty = lty,
+          )
 )
@@
   rare_df_col <- rare_df %>%
-    dplyr::rename(ASV = "value") %>%
+    rename(ASV = "value") %>%
     group_by(Sample) %>% 
@@
 # Create Plot 
-rare_plot <- map(rare_df_col, 
+rare_plot <- map2(rare_df_col, raremax, 
               ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
@@
               geom_line(aes(color = Counts, group = Sample)) +
-              theme(plot.title = element_text(hjust = 0.5),
-                    plot.margin=unit(c(0,0,0.5,0),"cm"))
+              theme(plot.title = element_text(hjust = 0.5))
             )
-raremax <- map(seqtabs, ~min(rowSums(.x)))
+
 r_plot <- pmap(list(rare_plot, raremax, limit, title),
                ~ ..1 + xlim(c(0,..3)) +
-               labs(title=paste0("Sequencing ", ..4)) +
+               labs(title=paste0("Rarefaction Curve ", ..4)) +
                geom_vline(xintercept = ..2, 
@@
 
-```
-
-```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
-ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),
-          r_plot[[3]] + theme(axis.title.x=element_blank()),
-          r_plot[[4]] + theme(axis.title.x=element_blank()),
+grid.arrange(
+  r_plot[[1]],
+  r_plot[[3]],
+  r_plot[[4]],
   r_plot[[5]],
-          ncol = 1, 
-          labels = c("A","B","C","D")
+  nrow = 4,
+  #height = c(1, 1, 1, 1),
+  clip = FALSE
 )
 ```
-
-**Supp.
-Figure 10.
-Rarefraction curves**.
-**A** Boston 1 sequencing run, **B** Sthlm sequencing run and **C** and **D** shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
 

### 2021-03-12 reports/rmarkdown/manuscript/manuscript_template.tex
updates in figures

@@
 Author Three\textsuperscript{1,2}}
-\date{March 03, 2021}
+\date{March 12, 2021}
 
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
+
 \clearpage
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-3}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-4}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-4}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-8-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-9-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-9-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-10-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-10-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-11-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-12-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-13-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-14-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-15-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-15-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-16-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-16-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-17-1}
 
@@
 
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-17-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-18-1}
 
@@
 \subsection{Figure S9}\label{figure-s9}}
+
+\clearpage
+
+\hypertarget{figure-s10}{%
+\subsection{Figure S10}\label{figure-s10}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-4}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-5}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-6}
 

### 2021-03-12 reports/rmarkdown/manuscript/manuscript_template.Rmd
Hide base r plot rarefaction curves

--- 2021-03-11 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
               geom_line(aes(color = Counts, group = Sample)) +
-              theme(plot.title = element_text(hjust = 0.5),
-                    plot.margin=unit(c(0,0,0.5,0),"cm"))
+              theme(plot.title = element_text(hjust = 0.5))
             )
@@
                ~ ..1 + xlim(c(0,..3)) +
-               labs(title=paste0("Sequencing ", ..4)) +
+               labs(title=paste0("Rarefaction Curve ", ..4)) +
                geom_vline(xintercept = ..2, 
@@
 
-```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
-ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),
-          r_plot[[3]] + theme(axis.title.x=element_blank()),
-          r_plot[[4]] + theme(axis.title.x=element_blank()),
+```{r Print rare curves, fig.width=6, fig.height=10, message=F}
+grid.arrange(
+  r_plot[[1]],
+  r_plot[[3]],
+  r_plot[[4]],
   r_plot[[5]],
-          ncol = 1, 
-          labels = c("A","B","C","D")
+  nrow = 4,
+  clip = FALSE
 )
 ```
-
-**Supp.
-Figure 10.
-Rarefraction curves**.
-**A** Boston 1 sequencing run, **B** Sthlm sequencing run and **C** and **D** shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
 

### 2021-03-16 reports/rmarkdown/01_data_pre_processing.Rmd
run on my computer update

--- 2021-03-11 reports/rmarkdown/01_data_pre_processing.Rmd
@@
 output:
-  #pdf_document:
-    #toc: yes
+  pdf_document:
+    toc: yes
   html_document:

### 2021-03-16 reports/rmarkdown/03_integrated_datasets.Rmd
Small buggs

--- 2021-03-12 reports/rmarkdown/03_integrated_datasets.Rmd
@@
 
-```{r, echo = FALSE}
+```{r, eval=FALSE, include=FALSE}
 groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){

### 2021-03-18 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds metadata on umap and cytokine and IF plots

--- 2021-03-12 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 
+```{r, echo = FALSE, eval = TRUE,fig.asp=.5}
+figlabels <- letters
+
+cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
+cytokine_data$ID <- metadata [ match( cytokine_data$PatID , as.character(metadata$PatID)  ) ,"ID" ]
+cytokine_data <- cytokine_data[ !is.na(cytokine_data$ID) , ]
+rownames(cytokine_data) <- cytokine_data$ID
+cytokine_data <- cytokine_data[,grepl("CVL",colnames(cytokine_data))]
+colnames(cytokine_data) <- sub("CVL_","",colnames(cytokine_data))
+cytokine_data <- t(cytokine_data)
+cytokine_data[is.na(cytokine_data)] <- 0.35
+cytokine_data <- log2(cytokine_data+1)
+
+groupings <- metadata$group_annotation [ match( colnames(cytokine_data) , as.character(metadata$ID)  ) ]
+mypar(1,3,mar=c(6,5,1,5))
+
+plot_dots(data=cytokine_data,
+          srt=45,
+          genes=rownames(cytokine_data),
+          col = pal,
+          clustering=groupings )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+barlist(data=cytokine_data,
+          srt=45,
+          genes=rownames(cytokine_data),
+          col = pal,
+          clustering=groupings )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+violist(data=cytokine_data,
+          srt=45,
+          genes=rownames(cytokine_data),
+          col = pal, transparency = 50,
+          clustering=groupings,smooth = 2 )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+```
+
+\clearpage
+
+## Figure S10
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=.8}
+figlabels <- letters
+
+IF_data <- read.csv("../../../data/200105_CD4Ecad_Visit3LAMIQ.csv")
+IF_data$ID <- metadata [ match( IF_data$PatID , as.character(metadata$PatID)  ) ,"ID" ]
+IF_data <- IF_data[ !is.na(IF_data$ID) , ]
+rownames(IF_data) <- IF_data$ID
+IF_data <- IF_data [ , -c(1:3)]
+IF_data <- IF_data [ , -ncol(IF_data)]
+IF_data <- t(IF_data)
+# IF_data[is.na(IF_data)] <- 0
+# IF_data <- log2(IF_data+1)
+
+groupings <- metadata$group_annotation [ match( colnames(IF_data) , as.character(metadata$ID)  ) ]
+
+mypar(1,3,mar=c(6,15,1,4))
+plot_dots(data=IF_data,
+          srt=45,
+          genes=rownames(IF_data),
+          col = pal,
+          clustering=groupings )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+barlist(data=IF_data,
+          srt=45,
+          genes=rownames(IF_data),
+          col = pal,
+          clustering=groupings )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+violist(data=IF_data,
+          srt=45,
+          genes=rownames(IF_data),
+          col = pal, transparency = 50,
+          clustering=groupings,smooth = 2 )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+```
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=.5}
+
+continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_ng_mL_v3","Plasma_S_Estradiol_pg_mL_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+
+mypar(1,3,mar=c(6,15,1,4))
+plot_dots(data=as.matrix(t(metadata[,continuous_parameters])),
+          srt=45,
+          genes=continuous_parameters,
+          col = pal,
+          clustering=metadata$group_annotation )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+barlist(data=as.matrix(t(metadata[,continuous_parameters])),
+          srt=45,
+          genes=continuous_parameters,
+          col = pal,
+          clustering=metadata$group_annotation )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+violist(data=as.matrix(t(metadata[,continuous_parameters])),
+          srt=45,
+          genes=continuous_parameters,
+          col = pal, transparency = 50,
+          clustering=metadata$group_annotation,smooth = 2 )
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+```
+
+\clearpage
+
+## Figure S11
+
 ```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
-
-```
-
-\clearpage
-
-## Figure S10
+figlabels <- letters
+
+g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_ng_mL_v3","Plasma_S_Estradiol_pg_mL_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+
+mypar(6,6,mar=c(1,1,1,1))
+
+plot(l,
+       col=pal[metadata$group_annotation],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+empty_plot()
+legend("topleft",title.adj = 0,
+       legend = levels(metadata$group_annotation),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+
+for(i in continuous_parameters){
+  xx <- metadata[,i]
+  xx[is.na(xx)] <- 0
+  xx <- xx / max(xx)
+  o <- order(xx)
+  plot(l[o,],
+       col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+
+}
+
+```
+
+\clearpage
+
+## Figure S12
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+figlabels <- letters
+
+g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+categorical_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
+                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
+                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
+                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
+                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
+                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
+                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
+                         "blood_cervix_v2","blood_cervix_v3",
+                         "cervical_inflammation_v2","cervical_inflammation_v3",
+                         "HIVstatus","Contraception",
+                         "CT_v2","CT_v3","regular_partner",
+                         "vaginal_douching_v2","vaginal_douching_v3")
+
+mypar(6,5,mar=c(1,1,1,8))
+plot(l,
+       col=pal[metadata$group_annotation],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(metadata$group_annotation),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+
+for(i in categorical_parameters){
+  xx <- factor(metadata[,i])
+  plot(l[,],
+       col=pal[xx],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(xx),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+}
+```
+
+\clearpage
+
+## Figure S13
 

### 2021-03-18 reports/rmarkdown/manuscript/manuscript_template.tex
adds metadata on umap and cytokine and IF plots

@@
 Author Three\textsuperscript{1,2}}
-\date{March 12, 2021}
+\date{March 18, 2021}
 
@@
 
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-19-1}
+
 \clearpage
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-3}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-4}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-5}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-20-6}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-21-1}
+
+\clearpage
+
+\hypertarget{figure-s11}{%
+\subsection{Figure S11}\label{figure-s11}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-22-1}
+
+\clearpage
+
+\hypertarget{figure-s12}{%
+\subsection{Figure S12}\label{figure-s12}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-23-1}
+
+\clearpage
+
+\hypertarget{figure-s13}{%
+\subsection{Figure S13}\label{figure-s13}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-4}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-5}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-6}
 

### 2021-03-18 reports/rmarkdown/01_taxonomy_improved.Rmd

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: krisina.broliden@ki.se
  pi_name: Kristina Broliden
  project_title: Broliden_5325
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)

```

#Load libraries and functions

```{r, echo = FALSE}
library(tidyverse)
library(phyloseq)
library(Biostrings)
```

```{r function}
######################## FASTA Function ########################
write.fasta <- function(df, file.out, open = "w"){
  outfile <- file(description = file.out, open = "w")
  write.oneseq <- function(sequence, name){
    writeLines(paste(">", name, sep = ""), outfile)
    writeLines(sequence, outfile)
    return(name)
  }
  map2_chr(df$Sequence, df$Seq_ID, ~write.oneseq(.x, .y))
  close(outfile)
}

```

# Loading data and metadata

```{r Get ASV and tax tables, message=FALSE}
paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3
             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region

n <- c("boston_r1", "boston_r2", "Sthlm")
ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

seqtabs <- map(ps_list, ~data.frame(.x@otu_table@.Data))
taxtabs <- map(ps_list, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
sample_info <- map(ps_list, ~as_tibble(sample_data(.x), rownames = 'SampleID')) 
```

## Reshape seqtab

```{r reshape seqtabs}
seqtab.nochim.reshaped <- map(seqtabs, ~.x %>%
                            rownames_to_column(var = "Sample_ID") %>%
                            pivot_longer(., -Sample_ID) %>%
                            dplyr::rename(Sequence="name", Reads="value")
)
seqtab.taxa <- seqtab.nochim.reshaped %>%
                map2(., taxtabs, ~left_join(.x, .y, by="Sequence"))
```

## Creat Short ID For Sequences across runs

```{r SeqID dict, include=FALSE, results='hide'}
############### Create dictonary for Sequence id ####
Seq_ID_key <-seqtab.taxa %>% 
  bind_rows(.id = "Sequencing_run") %>%
  # Get only uniqe sequences by grouping on all taxonomy
  group_by(Sequence,Kingdom,Phylum,Class,Order,Family,Genus,Species) %>%
  #summarize(n()) %>% filter(`n()`>1) %>% .$`n()` %>% sum(.)
  summarise(Reads_per_taxa = sum(Reads),
            Sequencing_run = paste(unique(Sequencing_run), collapse="; "),
            .groups = "drop") %>%
  bind_cols(Seq_ID = c(paste0("Seq_",sprintf("%05.0f", 1:nrow(.))))) %>%
  {. ->> seqtab.taxa.all } %>% 
  select(Seq_ID, Sequence)

write.csv(Seq_ID_key, file=paste0("../../results/Seq_ID_key.csv"), quote=FALSE, row.names=FALSE)

# Check that all sequences are unique:
#S <- Seq_ID_key %>% group_by(Sequence) %>% summarize(n()) %>% filter(`n()`>1) %>% .$`n()` %>% sum(.)
```

## Create Query files

```{r Querys, echo=T, results='hide'}
# Identify sequences not assigned by RDP at species level
unnasigned_0 <- seqtab.taxa.all %>%
  filter(., is.na(Species)) #%>%
  #filter(., Reads != 0)

# Create Query files
write.fasta(Seq_ID_key, paste0("../../results/BVAB_Query.fasta"))
write.fasta(unnasigned_0, paste0("../../results/optivag_Query.fasta"))
```

# BLAST
## Download databases

```{r Download data base}
db_dir <- "../../data/database/"
if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }

url <- c( "https://raw.githubusercontent.com/ctmrbio/BVAB-and-Lac-sequences/master/BVAB_rRNA_database.fa",
          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_db.fasta.gz",
          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_seqinfo.csv")

walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
system("gunzip ../../data/database/*.gz")

OptiVag <- read_csv("../../data/database/optivag_seqinfo.csv", col_names = T)
```

## BLAST seraches

```{r BLASTn search}
# Make uniqe identifyers in the BVAB file
fa = readDNAStringSet(paste0(db_dir,"BVAB_rRNA_database.fa"))
ID <-  make.unique(paste0(str_match(names(fa), "^BVAB._"),"0"), sep = "")
names(fa) = paste(ID, names(fa))
writeXStringSet(fa, "../../data/database/BVAB_db.fasta")

for(name in c("BVAB", "optivag")) { 
  db_dir     = "../../data/database/"
  result_dir = "../../results/"
  if (isFALSE(is_empty(list.files(path = db_dir, pattern = paste0(name,".+\\.nhr"))))) {
    message("using database provided in path:")
    message(paste0(db_dir, name, "_reference.db"))
  } else {
    system2(command = "makeblastdb", 
                          args = c("-in", paste0(db_dir, name, "_db.fasta"), 
                                  "-out", paste0(db_dir, name, "_reference.db"), 
                                  "-parse_seqids", 
                                  "-dbtype", "nucl"),
                         wait = TRUE,
                         stdout = TRUE)
  }
  blast_out <- system2(command = "blastn", 
                       args = c("-db", paste0(db_dir, name, "_reference.db"), 
                                "-query", paste0(result_dir, name, "_Query.fasta"), 
                                #"-out", paste0(result_dir, name, "_hits.tab"),
                                "-outfmt", 6, 
                                "-perc_identity", 99.5,
                                "-qcov_hsp_perc", 90),
                       wait = TRUE,
                       stdout = TRUE) %>%
  as_tibble() %>% 
  separate(col = value, sep = "\t", convert = TRUE,
           into = c("qseqid","sseqid","pident","length","mismatch","gapopen",
                    "qstart","qend","sstart","send","evalue","bitscore") )

  write_tsv(blast_out, paste0("../../results/", name, "_hits.tsv"))
}

```

## Load BLAST results

```{r Load BLAST hits, message=FALSE}
blast_OptiVag <- read_table2(paste0("../../results/OptiVag_hits.tsv"))
blast_BVAB <- read_table2(paste0("../../results/BVAB_hits.tsv"))

```

## Retrive Genus and Species hits from BLAST search

```{r Retrive hits}
# Resolve ambiguously asigned genus and Lacto spp.
OptiVag_hits <- blast_OptiVag %>%
                left_join(select(OptiVag, Full_taxonomy, Species, Seq_ID), 
                          by=c(sseqid='Seq_ID')) %>%
                select(Seq_ID=qseqid, Full_taxonomy, Species, Ref_iD=sseqid, 
                       Identity=pident, Mismatch=mismatch) %>%
                group_by(Seq_ID) %>% 
# removes hits with mismatch=1, for sequences with multiple hits where on of the hits has mismatch=0 
                filter(!(n()>1 & any(grepl('0', Mismatch)) & Mismatch == 1)) %>%
                ungroup() %>%
                separate(Species, into=c('Genus', 'Species'), sep = '_') %>% 
                separate(Full_taxonomy, into=c("Full_taxonomy", NA), sep = '(;)(?:.(?!;))+$') %>%
                select(Seq_ID, Full_taxonomy, Genus, Species) %>%
                unique() %>%
# the last group is dropped after summarize (.groups argument)
                group_by(Seq_ID, Full_taxonomy, Genus) %>% 
                summarise(Species = paste(Species, collapse="/"), 
                          .groups='drop_last') %>%
                mutate(Genus = paste(Genus, collapse="/")) %>%
                left_join(select(unnasigned_0, Seq_ID, Reads_per_taxa), 
                          by='Seq_ID') %>%
                ungroup()

BVAB_hits <- blast_BVAB %>%
              select(qseqid, sseqid) %>%
              left_join(select(seqtab.taxa.all, Seq_ID, Reads_per_taxa),
                          by=c('qseqid'='Seq_ID')) %>%
              separate(sseqid, c('Genus', 'Species')) %>%
              dplyr::rename(Seq_ID=qseqid) %>%
              select(Seq_ID, Genus, Reads_per_taxa) %>%
              unique()

all_hits <- bind_rows(BVAB_hits, OptiVag_hits) %>%
              unique()
```

## BLAST result info

```{r Information about BLAST result}
print(paste0("Number of sequences annotated by OptiVag:  ", length(unique(OptiVag_hits$Seq_ID))))

print(paste0("Number of sequences annotated by BVAB:     ", length(unique(BVAB_hits$Seq_ID))))

overlap <- intersect(BVAB_hits$Seq_ID, OptiVag_hits$Seq_ID)
ifelse(length(unique(all_hits$Seq_ID)) == length(all_hits$Seq_ID), "There is no duplicate species annotation", paste0(message("Resolving overlapping annotation between BVAB and OptiVag:"), paste(overlap,  collapse=', ')))

doNextChunk <- isFALSE(length(unique(all_hits$Seq_ID)) == length(all_hits$Seq_ID))

```

# Resolve Taxonomy
## Resolve overlapping annotation between BVAB and OptiVag

```{r Resolve Overlap, eval = doNextChunk}
# Identify which annotation that overlap for BVAB sequences 
duplicate <- all_hits %>%
  group_by(Seq_ID) %>% 
  filter(n()>1) %>%
  filter((grepl('BVAB', Genus))) %>%
  ungroup()

# Filter out duplicated annotation for BVAB
filtered_hits <- all_hits %>%
  filter(!(Seq_ID %in% duplicate$Seq_ID & !(grepl('BVAB', Genus))))

# is there duplicates other than BVAB?
d <- filtered_hits %>% group_by(Seq_ID) %>% summarize(n(), .groups='drop_last') %>% filter(`n()`>1)  %>% .$Seq_ID

# prints message about duplicate status
ifelse(length(unique(filtered_hits$Seq_ID)) == length(filtered_hits$Seq_ID), "Overlaping annotation resolved", paste0(message("Resolving duplicate taxonomy asignment:"), paste(d,  collapse=', ')))

# If duplicates are identified, run next chunk
doNextChunk <- isTRUE(length(d)>=1)

```

## Resolve duplicate taxonomy

```{r Resolve duplicate taxonomy, eval = doNextChunk}
# Collapse species annotation for sequences with several hits
filtered_2 <- aggregate(. ~Seq_ID, data = filtered_hits, na.action=NULL, FUN=function(x) { paste(unique(x), collapse = '/') })

# Sequence ids that are duplicate:
dd <- filtered_2 %>% group_by(Seq_ID) %>% summarize(n(), .groups='drop_last') %>% filter(`n()`>1)  %>% .$Seq_ID

# Message to terminal
ifelse(length(unique(filtered_hits$Seq_ID)) == length(filtered_2$Seq_ID), "Duplicate annotation resolved", paste0(message("WARNING! still duplicate annotation:"), paste(dd,  collapse=', ')))

filtered_hits <- filtered_2
```

## Merge New Annotation With Tax Table

```{r Save updated tax table}
#### Update taxa ####
update_taxa <- seqtab.taxa.all %>%
  left_join(., select(filtered_hits, Seq_ID, Genus_u="Genus",
                      Species_u="Species", Full_taxonomy), by='Seq_ID') %>%
  mutate(identical = ifelse(str_detect(.$Genus_u, "BVAB")|
                                     .$Genus_u == .$Genus|
                                           is.na(.$Genus), "yes", "no"))

new_taxtable <- update_taxa %>%
  mutate(Genus_f = ifelse(.$identical=="no"|is.na(.$identical), 
                          .$Genus, .$Genus_u)) %>%
  mutate(Species_f = ifelse(.$identical=="no"|is.na(.$identical), 
                            .$Species, .$Species_u)) %>%
  select(-Genus, -Species, -Genus_u, -Species_u, -identical, -Sequencing_run) %>%
  dplyr::rename(Genus="Genus_f", Species="Species_f") %>%
  select(Seq_ID, everything(), -Full_taxonomy, -Reads_per_taxa)

write.csv(new_taxtable, file=paste0("../../data/SeqID_to_taxa.csv"), quote=FALSE, row.names=FALSE)
```

## Report of unasigned sequences

```{r Saving unnasigned sequences}
#### Taxa that still have no Genus or Lacto spp. annotation ####
unnasigned_final <- new_taxtable %>% 
  left_join(select(seqtab.taxa.all, Sequence, Reads_per_taxa, Sequencing_run), by="Sequence") %>%
  filter(is.na(.$Genus)) %>%
  arrange(desc(Reads_per_taxa)) %>%
  arrange(desc(Family)) %>%
  filter(Reads_per_taxa >100)

write.csv(unnasigned_final, 
          file=paste0("../../results/Unnasigned_tax_report.csv"), quote=FALSE, row.names=FALSE)

unnasigned_final
```

# Save results

## merge seqtab with new taxonomy

```{r Create ASV tables}
# Add sample info and updated taxonomy 
ASV_wCtrl <- seqtab.nochim.reshaped %>% 
            map2(., sample_info, 
                 ~left_join(.x, select(.y, SampleID, ID, Sample_type, Visit, Place_sequenced,
                                            Description, Include, Include_final),
                            by=c("Sample_ID"="SampleID")) ) %>%
                  {. ->> ASV } %>%
            map(., ~ .x %>%
                  select(Sequence, Sample_ID, Reads) %>%
                  pivot_wider(., names_from=Sample_ID, values_from=Reads) %>%
                  left_join(., new_taxtable, by="Sequence")
                  )

#### ASV with all controls and replicate samples removed ####
ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3", "ASV_tissue_B1", 
               "ASV_CVL_V2_B2", "ASV_tissue_B2", "ASV_CVL_V2_C" )
ASV_list <- map(ASV, ~ .x %>%
                  filter(grepl("yes", .$Include)) %>%
                  group_split(Sample_type, Visit, .keep = T)
            ) %>%
            flatten() %>%
            set_names(ASV_names) %>%
            map(., ~ .x %>%
                select(Sequence, ID, Reads) %>%
                pivot_wider(., ,names_from=ID, values_from=Reads) %>%
                left_join(., new_taxtable, by="Sequence") 
                )

```

## Save csv

```{r Save ASV files}
imap(ASV_list, 
     ~write.csv(.x, file=paste0("../../data/",.y,".csv"), quote=FALSE, row.names=F) )

```

### 2021-03-18 reports/rmarkdown/manuscript/manuscript_template.Rmd

@@
 library(vegan)
-library(gridExtra)
+library(ggpubr)
 library(rafalib)
@@
 # separate tissue from CVLv2 samples in the second Boston run
-CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == 2) 
-tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == 3)
+CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v2")
+tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == "v3")
 
@@
 # Get rarefaction curves
-raremax <- map(seqtabs, ~min(rowSums(.x)))
-rare_runs <- map2(seqtabs, raremax, ~ .x %>%
-             rarecurve(., step = 20, sample = .y, 
-                       label = FALSE#,col = col, lty = lty,
-          )
+rare_runs <- map(seqtabs, ~ .x %>%
+             rarecurve(., step = 20, label = FALSE)
 )
@@
   rare_df_col <- rare_df %>%
-    rename(ASV = "value") %>%
+    dplyr::rename(ASV = "value") %>%
     group_by(Sample) %>%
@@
 # Create Plot
-rare_plot <- map2(rare_df_col, raremax, 
+rare_plot <- map(rare_df_col,
               ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
@@
               geom_line(aes(color = Counts, group = Sample)) +
-              theme(plot.title = element_text(hjust = 0.5))
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.margin=unit(c(0,0,0.5,0),"cm"))
             )
-
+raremax <- map(seqtabs, ~min(rowSums(.x)))
 r_plot <- pmap(list(rare_plot, raremax, limit, title),
                ~ ..1 + xlim(c(0,..3)) +
-               labs(title=paste0("Rarefaction Curve ", ..4)) +
+               labs(title=paste0("Sequencing ", ..4)) +
                geom_vline(xintercept = ..2,
@@
 
-grid.arrange(
-  r_plot[[1]],
-  r_plot[[3]],
-  r_plot[[4]],
+```
+
+```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
+ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),
+          r_plot[[3]] + theme(axis.title.x=element_blank()),
+          r_plot[[4]] + theme(axis.title.x=element_blank()),
           r_plot[[5]],
-  nrow = 4,
-  #height = c(1, 1, 1, 1),
-  clip = FALSE
+          ncol = 1,
+          labels = c("A","B","C","D")
           )
 ```
+
+**Supp.
+Figure 10.
+Rarefraction curves**.
+**A** Boston 1 sequencing run, **B** Sthlm sequencing run and **C** and **D** shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
 

### 2021-03-19 reports/rmarkdown/manuscript/manuscript_template.Rmd
add metadata to UMAP RNAseq

@@
        breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
-points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .01,length(sample_use)),
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .02,length(sample_use)),
         col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
@@
 points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
-        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
+        col= c("grey","red")[(!is.na(metadata[sample_use,"RepeatedPCR"]))+1][order(group_annotation[sample_use])],
         pch=16,xpd=T)
 points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.01,
-        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
+        col= c("grey","red")[as.numeric(metadata[sample_use,"HIVstatus"])][order(group_annotation[sample_use])],
         pch=16,xpd=T)
@@
 
-```{r, echo = FALSE, eval = TRUE,fig.asp=1.2}
+```{r, echo = FALSE, eval = TRUE,fig.asp=1}
 figlabels <- letters
@@
 ## Figure S13
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=.5}
+
+continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_ng_mL_v3","Plasma_S_Estradiol_pg_mL_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
+
+common <- colnames(top_TRX_counts)[colnames(top_TRX_counts) %in% rownames(metadata)]
+
+mypar(6,6,mar=c(1,1,1,1))
+
+plot(UMAP_TRX,
+       col=pal[metadata[common,"group_annotation"]],
+       main="",xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+empty_plot()
+legend("topleft",title.adj = 0,
+       legend = levels(metadata[common,"group_annotation"]),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+
+for(i in continuous_parameters){
+  xx <- metadata[common,i]
+  xx[is.na(xx)] <- 0
+  xx <- xx / max(xx)
+  o <- order(xx)
+  plot(UMAP_TRX[o,],
+       col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+
+}
+```
+
+\clearpage
+
+## Figure S14
+
+```{r, echo = FALSE, eval = TRUE,fig.asp=1}
+figlabels <- letters
+
+# g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
+# set.seed(1)
+# l <- layout_with_fr(g,niter=3000,start.temp=30)
+#
+
+categorical_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
+                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
+                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
+                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
+                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
+                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
+                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
+                         "blood_cervix_v2","blood_cervix_v3",
+                         "cervical_inflammation_v2","cervical_inflammation_v3",
+                         "HIVstatus","Contraception",
+                         "CT_v2","CT_v3","regular_partner",
+                         "vaginal_douching_v2","vaginal_douching_v3")
+
+common <- colnames(top_TRX_counts)[colnames(top_TRX_counts) %in% rownames(metadata)]
+
+mypar(6,5,mar=c(1,1,1,8))
+plot(UMAP_TRX,
+       col=pal[metadata[common,"group_annotation"]],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(metadata[common,"group_annotation"]),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+
+for(i in categorical_parameters){
+  xx <- factor(metadata[common,i])
+  plot(UMAP_TRX[,],
+       col=pal[xx],
+       main=i,xpd=T,
+       pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(xx),xjust = 0,yjust = 1,
+       bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
+}
+```
+
+\clearpage
+
+## Figure S15
 

### 2021-05-04 reports/rmarkdown/01_data_pre_processing.Rmd
Add Genus_lacto taxonomy

--- 2021-03-12 reports/rmarkdown/01_data_pre_processing.Rmd
@@
                       warning    = FALSE)
+
+setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
 ```
@@
 library(niceRplots)
+library(tidyverse)
 
@@
 ```{r, echo = FALSE}
-PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data
+PATH <- "../../" #Path to the data
 
@@
 
+```{r, Functions}
+# create "other" taxonomy
+other_taxa.fun <- function(df) {
+  other_taxa <- df %>%
+  unite("taxa_other", Kingdom:Species, sep = ";", remove = F, na.rm = T) %>%
+  mutate(taxa_other = ifelse(is.na(.$Species), paste0(.$taxa_other, ";other"), .$taxa_other)) %>%
+  #mutate(taxa_other = str_replace(.$taxa_other, ";NA.+|;NA", ";other")) %>%
+  mutate(taxa_other = str_extract(.$taxa_other, "([^;]+);([^;]+)$")) %>% # get the two last taxa levels
+  mutate(taxa_other = sub('^(.*);(other)', '\\2;\\1', .$taxa_other)) # place "other" first
+return(other_taxa)
+}
+
+# create genus lacto taxonomy
+genus_lacto.fun <- function(df){
+  genus_lacto <- df %>%
+    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% #ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species),
+    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste0(.$Genus_lacto, "crispatus/acidophilus"), 
+                                ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
+                                       as.character(.$Genus_lacto))) )
+  return(genus_lacto)
+}
+
+aggregate.fun <- function(df, level){
+  level <- enquo(level)
+  aggregated <- df %>%
+    dplyr::rename(Taxonomy = !!level) %>%
+    unite(., "Full_taxonomy", Kingdom:Species, sep = ";") %>%
+    dplyr::select(-any_of(c("Sequence", "Seq_ID", "Genus_lacto", "taxa_other", "Full_taxonomy"))) %>%
+    group_by(Taxonomy) %>%
+    summarise(across(where(is.numeric), sum)) %>%
+    ungroup()
+return(aggregated)
+}
+```
+
 # Loading data and metadata
@@
 # Load metadata
-metadata <- read.csv(paste0(PATH,"/Clinical_visit_2_3_updatesept28.csv"))
+dataset_names <- c("ASV_tissue_B2.csv",   # Tissue, Boston run 2 (95 samples)
+                     "ASV_tissue_B1.csv", # Tissue, Boston run 1 (1 sample)
+                     "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
+                     "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
+                     "ASV_CVL_V2_C.csv",  # CVL V2, CTMR         (62 samples)
+                     "ASV_CVL_V3.csv")
+
+datasets_ <- map(dataset_names, ~read.csv( paste0(PATH,"data/",.x))) %>% set_names(., dataset_names)
+metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
 rownames(metadata) <- as.character(metadata$ID)
+```
 
+# Taxonomy agglomeration
+
+```{r, echo = FALSE}
+temp <- map(datasets_, ~.x  %>%
+  select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
+  filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
+  other_taxa.fun(.) %>% # create "other" taxonomy 
+  genus_lacto.fun(.)  %>% # create genus lacto taxonomy 
+  {. ->> temp_ } %>% 
+  aggregate.fun(., "Genus_lacto") 
+)
+
+# Extract taxonomic information to file
+seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
+#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_taxa",".csv"),row.names = F )
+
+#arrange all datasets to have the same rownames with all bacteria
+common_microbes_fun <- function(x) {
+  t <- bind_rows(temp) %>%
+    select(Taxonomy) %>%
+    unique() %>%
+    left_join(x, by="Taxonomy") %>%
+    mutate(across(everything(), ~replace_na(.x, 0))) %>%
+    column_to_rownames(var = "Taxonomy")
+  return(t)
+}
+datasets_ <- map(temp, ~common_microbes_fun(.x))
+lapply(datasets_,dim)
+
+# datasets <- datasets[3] #in case you wanna test in 1 dataset
+
+# Transform phyloseq data into normalized matricies ( log2[CP1K] )
+# datasets <- lapply(datasets,function(x){
+#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
+# })
+# lapply(datasets,dim)
+
+```
+
+# Loading data and metadata Paulos code
+
+```{r, echo = FALSE}
+PATH <- "../../" #Path to the data
+# Load metadata
+metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
+rownames(metadata) <- as.character(metadata$ID)
 # Load datasets
@@
                      "ASV_CVL_V3.csv")
-
 #import datasets and return matrix with taxa-level resolution
 datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"/",x))
+  x <- read.csv( paste0(PATH,"data/",x))
   temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
@@
   use <- !is.na(taxa) & (taxa != "NA")
-
   temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
@@
 lapply(datasets,colnames)
-
 #arrange all datasets to have the same rownames with all bacteria
@@
 lapply(datasets,dim)
-
 # datasets <- datasets[3] #in case you wanna test in 1 dataset
-
 # Transform phyloseq data into normalized matricies ( log2[CP1K] )
@@
 # lapply(datasets,dim)
-
 # Extract taxonomic information to file
 taxa_info <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"/",x))
+  x <- read.csv( paste0(PATH,"data/",x))
   temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
@@
 ```{r, echo = FALSE}
+datasets <- datasets_
+names(datasets)
+
 #No batch correction needed in tissue_V3 dataset, only add it to the other samples
@@
 ```{r, echo = FALSE}
-mypar(3,3)
+#mypar(3,3)
+par(mar=c(3, 3, 0, 0))
 PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
@@
 
-if(file.exists(paste0(PATH,"../results/encoder.h5")) & 
-  file.exists(paste0(PATH,"../results/decoder.h5")) & 
-  file.exists(paste0(PATH,"../results/aen.h5")) ){
-  encoder <- load_model_hdf5(paste0(PATH,"../results/encoder.h5"))
-  decoder <- load_model_hdf5(paste0(PATH,"../results/decoder.h5"))
-  aen <- load_model_hdf5(paste0(PATH,"../results/aen.h5"))
+if(file.exists(paste0(PATH,"results/encoder.h5")) & 
+  file.exists(paste0(PATH,"results/decoder.h5")) & 
+  file.exists(paste0(PATH,"results/aen.h5")) ){
+  encoder <- load_model_hdf5(paste0(PATH,"results/encoder.h5"))
+  decoder <- load_model_hdf5(paste0(PATH,"results/decoder.h5"))
+  aen <- load_model_hdf5(paste0(PATH,"results/aen.h5"))
 
@@
 
-  save_model_hdf5(encoder, paste0(PATH,"../results/encoder.h5") )
-  save_model_hdf5(decoder, paste0(PATH,"../results/decoder.h5"))
-  save_model_hdf5(aen, paste0(PATH,"../results/aen.h5"))
+  save_model_hdf5(encoder, paste0(PATH,"results/encoder.h5") )
+  save_model_hdf5(decoder, paste0(PATH,"results/decoder.h5"))
+  save_model_hdf5(aen, paste0(PATH,"results/aen.h5"))
 }
@@
 samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
-datasets[["ASV_CVL_V2"]] <- cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] )
+datasets[["ASV_CVL_V2"]] <- as.data.frame.array(cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] ))
 datasets <- datasets[-c(1,2,3)]
@@
 
-datasets <- lapply(datasets, function(x) {
-  temp <- matrix(0,nrow = length(common_microbes),
-                 ncol = ncol(x),
-                 dimnames = list(common_microbes,colnames(x)))
-  temp[rownames(x),colnames(x)] <- x
-  return(temp)
-})
+common_microbes.fun <- function(x) {
+  xy <- rownames_to_column(x, var = "Taxa")
+  d <- data.frame(Taxa = common_microbes) %>%
+    left_join(., xy, by="Taxa") %>%
+    mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
+    column_to_rownames(var = "Taxa")
+  return(d)
+}
+datasets <- map(datasets, ~common_microbes.fun(.x))
 lapply(datasets,dim)
@@
 
-write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
 
@@
 batches_CVL2 <- t(t(setNames(batches_CVL2 , colnames(datasets[["ASV_CVL_V2_raw"]]))))
-write.csv( batches_CVL2 , paste0(PATH,"../results/batches_CVL2.csv"),row.names = T )
+write.csv( batches_CVL2 , paste0(PATH,"results/batches_CVL2.csv"),row.names = T )
 ```
@@
 ```{r, echo = FALSE}
-trx <- read.csv(paste0(PATH,"/TRX_rawcounts.csv"),row.names = 1)
+trx <- read.csv(paste0(PATH,"data/TRX_rawcounts.csv"),row.names = 1)
 trx <- rowsum(trx[,-c(1,2)],trx$symbol)
@@
 trx  <- trx[ rownames(trx) != "" , ]
+trx <- trx[ rownames(trx) != "NA" , ]
 colnames(trx) <- as.character(metadata$ID)[ match(colnames(trx) , as.character(metadata$PatID) ) ]
@@
 trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
-write.csv( trx , paste0(PATH,"../results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
+write.csv( trx , paste0(PATH,"results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
 ```

### 2021-05-06 reports/rmarkdown/01_data_pre_processing.Rmd
Genus_lacto correction

@@
   genus_lacto <- df %>%
-    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% #ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species),
-    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste0(.$Genus_lacto, "crispatus/acidophilus"), 
+    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
+    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste(.$Genus_lacto, "crispatus/acidophilus"), 
+                                ifelse(grepl("reuteri",.$Species), paste(.$Genus_lacto, "reuteri/oris/frumenti/antri"),
+                                  ifelse(grepl("gasseri",.$Species), paste(.$Genus_lacto,"gasseri/johnsonii/taiwanensis"),
+                                    ifelse(grepl("murinus",.$Species), paste(.$Genus_lacto,"murinus/animalis/apodemi/salivarius"),
+                                      ifelse(grepl("plantarum",.$Species), 
+                                           paste(.$Genus_lacto, "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
                                         ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
-                                       as.character(.$Genus_lacto))) )
+                                              as.character(.$Genus_lacto))))))) )
   return(genus_lacto)
@@
             other_taxa.fun(.) %>% # create "other" taxonomy 
-  genus_lacto.fun(.)  %>% # create genus lacto taxonomy 
+            genus_lacto.fun(.)  
+            ) %>% # create genus lacto taxonomy 
         {. ->> temp_ } %>% 
-  aggregate.fun(., "Genus_lacto") 
-)
+        map(., ~aggregate.fun(., "Genus_lacto"))
 
@@
 seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
-#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_taxa",".csv"),row.names = F )
+#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )
 
@@
 
-```{r, echo = FALSE}
+```{r, eval=F, echo = FALSE}
 PATH <- "../../" #Path to the data
@@
 taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
-write.csv2(taxa_info, paste0(PATH,"../results/taxonomy.csv"),row.names = T)
+write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
 ```

### 2021-05-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
pairwise group DGE with intercept

@@
 ## Figure 3
+## individual group comparison DGE
+```{r}
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+TRX_counts <- round(2^TRX - 1)
+TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
+sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
+TRX_counts <- TRX_counts[,sample_use]
+
+metadata <- metadata %>%
+  #select(Contraception, sexwork_months) %>%
+  mutate(sexwork_months=ifelse(is.na(.$sexwork_months),
+                               mean(metadata$sexwork_months,na.rm = TRUE),.$sexwork_months))%>%
+  mutate(Contraception=ifelse(.$Contraception=="no HC", "no_HC","DMPA")) %>%
+  select(ID, PatID, HIVstatus, Contraception, age, sexwork_months, everything())
+
+library(edgeR)
+
+#design <- model.matrix(~  group_annotation, data=metadata[sample_use,])
+#design <- model.matrix(~0 + group_annotation+Contraception+HIVstatus+age+sexwork_months, 
+                       #data=metadata[sample_use,])
+design <- model.matrix(~group_annotation+HIVstatus+Contraception+age+sexwork_months, 
+                       data=metadata[sample_use,])
+colnames(design) <- sub("group_annotation","",colnames(design))
+
+y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
+y <- calcNormFactors(y,method = "TMM")
+y <- estimateGLMCommonDisp(y,design)
+y <- estimateGLMTagwiseDisp(y,design)
+fit <- glmFit(y, design)
+
+#lrt <- glmLRT(fit,coef=2:5)
+lrt <- glmLRT(fit,coef=1:9)
+top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+colnames(top) <- sub("group_annotation","",colnames(top))
+#top <- cbind(logFC.Gard=0,top)
+
+saveRDS( y , "../../../results/TRX_EdgeR_estimations.rds" )
+write.csv2( top , "../../../results/TRX_EdgeR_top_dge.csv" )
+
+top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+top_dge <- rownames(top)[ top_dge ]
+
+top_TRX_counts <- cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
+top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
+top_TRX_counts[top_TRX_counts > 5] <- 5
+colnames(top_TRX_counts) <- colnames(TRX_counts)
+
+## individual group comparision ##
+# without intercept:
+n <- combn(sort(unique(group_annotation)),2) 
+g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
+s <- seq_along(1:ncol(n)) %>% set_names(g)
+#name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))
+
+# with intercept:
+n <- combn(sort(unique(group_annotation))[-1],2) 
+g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
+g <- c(sort(unique(group_annotation))[-1], g)
+
+CONTRASTS <- makeContrasts( contrasts=g,
+                            levels = design )
+lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
+top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]])
+
+#saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
+#saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
+## end of individual group comparision ##
+
+# save TopTables 
+library(openxlsx)
+l_top <- list(ModifiedM4 = top)
+imap(l_top, ~write.xlsx(.x, file=paste0("../../../results/","TopTable",.y,".xlsx"))) # Excel
+
+## individual GO comparison ##
+source("../enrichment_function.R")
+gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
+gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
+enrich_list <- function(x, gmt_list){
+  temp <- compute_enrichment(genes = x,
+                            gmt_list = gmt_list,
+                            min_terms_pathway = 10,
+                            max_terms_pathway = 300,
+                            min_overlap = 3,
+                            sort_by_pvalue = T)
+  temp <- temp[!grepl("REGULATION",rownames(temp)),]
+  return(temp)
+}
+
+info_logFC <- map(top, ~table(abs(.x$logFC) > 1))
+info_PValue <- map(top, ~table(.$PValue < 0.01))
+
+# up and down regulated genes seperatly
+t_split <- map(top, ~ .x %>%
+        dplyr::filter(., .$PValue < 0.01) %>%
+        #dplyr::filter(., abs(.$logFC) < 1) %>%
+        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
+        split(., .$reg)
+  )
+# up and down regulated genes toghether
+t <- map(top, ~ .x %>%
+        dplyr::filter(., .$PValue < 0.01) #%>%
+        #dplyr::filter(., abs(.$logFC) < 1) 
+  )
+
+## GO enrichment
+enrichment_groups <- map_depth(t, 1, ~enrich_list(rownames(.x), gmt_list_GO))
+enrichment_groups_split <- map_depth(t_split, 2, ~enrich_list(rownames(.x), gmt_list_GO)) %>%
+  map(., ~bind_rows(.x, .id = "regulation"))
+
+## save files
+l_enrich <- list(ModifiedM4_split = enrichment_groups_split, ModifiedM4 = enrichment_groups)
+imap(l_enrich, ~write.xlsx(.x, file=paste0("../../../results/","GO_enrichment_",.y,".xlsx"))) # Excel
+## end of individual GO comparison ##
+```
+
+\clearpage
 

### 2021-06-17 reports/rmarkdown/02_individual_reports.Rmd
update

@@
 params:
-  dataset: 1
+  dataset: 2
 ---
@@
                       warning    = FALSE)
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
 ```
@@
 ```{r, echo = FALSE}
-PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
-PATH <- "../.." #Path to the data
+PATH <- "../../" #Path to the data
 
@@
 # Load metadata
-metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
+metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
 rownames(metadata) <- as.character(metadata$ID)

### 2021-06-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
update

--- 2021-03-19 reports/rmarkdown/manuscript/manuscript_template.Rmd
@@
 editor_options:
-  chunk_output_type: console
+  chunk_output_type: console #inline
+#params:
+  #taxa_level: "Genus_lacto"
 ---
@@
 ```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
+params <- "Genus_lacto"
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
 options(tinytex.verbose = TRUE)
@@
 library(edgeR)
+library(openxlsx)
 remotes::install_github("czarnewski/niceRplots",force=T)
@@
 
-#color palettes
-pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
-         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
-heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
-cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
 ```
@@
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
-patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
+TissueGroupM4 <- read.csv("../../../data/210411_TissueGroupM4.csv", row.names = 1)
+#patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
+patient_groups <- factor(setNames(metadata$ModifiedM4,metadata$ID))
 bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
@@
 
+```{r}
+#color palettes
+Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
+pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
+         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
+heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
+cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
+```
+
 ```{r, echo = FALSE, eval = TRUE}
 #Assign names to patient groups
+# big_groups <- c(
+#   g1_L.crispatus= "1 NonIners",
+#   g2_L.iners=    "2 Iners",
+#   g3_Gard=   "3 Gard",
+#   g4_Prev=        "4 Prev",
+#   g5_Oth=   "5 Other"
+# )
 big_groups <- c(
-  g2_L.iners=    "5",
-  g3_L.iners2=   "4",
-  g1_L.crispatus= "2",
-  g5_Gard=        "3",
-  g6_Gard.Prev=   "6",
-  g7_Gard.Prev2=  "8",
-  g4_Gard.Lacto=  "7",
-  g8_Patho=       "1"
+  g1_L.crispatus= "NonIners",
+  g2_L.iners=     "Iners",
+  g3_Gard=        "Gard",
+  g4_Prev=        "Prev",
+  g5_Oth=         "Oth"
 )
+# big_groups <- c(
+#   g2_L.iners=    "5",
+#   g3_L.iners2=   "4",
+#   g1_L.crispatus= "2",
+#   g5_Gard=        "3",
+#   g6_Gard.Prev=   "6",
+#   g7_Gard.Prev2=  "8",
+#   g4_Gard.Lacto=  "7",
+#   g8_Patho=       "1"
+# )
 
@@
 group_annotation <- names(big_groups)[match(group_annotation,big_groups)]
-group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))
+group_annotation <- setNames(group_annotation,names(patient_groups))
+#group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))
 
 metadata$group_annotation <- group_annotation
-write.csv2(metadata,"../../../results/metadata_integration.csv",row.names = T)
+#write.csv2(metadata,"../../../results/metadata_integration_.csv",row.names = T)
+
+# modifications nessecary for dge analyser:
+metadata <- metadata %>%
+  mutate(ModifiedM4 = as.character(ModifiedM4_)) %>%
+  mutate(sexwork_months=ifelse(is.na(.$sexwork_months),
+                               mean(metadata$sexwork_months,na.rm = TRUE),.$sexwork_months))%>%
+  mutate(Contraception=ifelse(.$Contraception=="no HC", "no_HC","DMPA")) %>%
+  cbind(TissueGroupM4) %>%
+  #as_data_frame() %>%
+  #left_join(TissueGroupM4, by="ID") %>%
+  #mutate(TissueGroupM4= factor(.$TissueGroupM4)) %>%
+  select(ID, PatID, HIVstatus, Contraception, age, sexwork_months, everything())
+
+#write.csv2(metadata,"../../../results/metadata_integration_.csv",row.names = T)
 
@@
 BC_groups <- c(
-  BC01= "BC01_Bifidobacterium",
-  BC02= "BC02_Mycoplasma",
-  BC03= "BC03_Lactobacillus.spp",
-  BC04= "BC04_Escherichia.sp",
-  BC05= "BC05_Other1",
-  BC06= "BC06_Other2",
-  BC07= "BC07_Corynebacterium.spp",
-  BC08= "BC08_Other3",
-  BC09= "BC09_Ureaplasma.sp",
-  BC10= "BC10_Campylobacter",
-  BC11= "BC11_Mobiluncus_BVAB",
-  BC12= "BC12_Faecalibacterium",
-  BC13= "BC13_Dialister",
-  BC14= "BC14_P.timonesis",
-  BC15= "BC15_T.scotoductus"
+  BC01= "BC01_Peptococcus",
+  BC02= "BC02_Lactobacillus.spp",
+  BC03= "BC03_Micrococcus",
+  BC04= "BC04_Gardnerella",
+  BC05= "BC05_Paracoccus",
+  BC06= "BC06_Pseudomonas",
+  BC07= "BC07_Prvotella",
+  BC08= "BC08_Faecalibacterium",
+  BC09= "BC09_Escherichia/Shigella"
 )
+# BC_groups <- c(
+#   BC01= "BC01_Bifidobacterium",
+#   BC02= "BC02_Mycoplasma",
+#   BC03= "BC03_Lactobacillus.spp",
+#   BC04= "BC04_Escherichia.sp",
+#   BC05= "BC05_Other1",
+#   BC06= "BC06_Other2",
+#   BC07= "BC07_Corynebacterium.spp",
+#   BC08= "BC08_Other3",
+#   BC09= "BC09_Ureaplasma.sp",
+#   BC10= "BC10_Campylobacter",
+#   BC11= "BC11_Mobiluncus_BVAB",
+#   BC12= "BC12_Faecalibacterium",
+#   BC13= "BC13_Dialister",
+#   BC14= "BC14_P.timonesis",
+#   BC15= "BC15_T.scotoductus"
+# )
 
@@
 legend(par("usr")[2],par("usr")[4],title.adj = 0,
-       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
+       legend = levels(factor(group_annotation)),
+       xjust = 0,yjust = 1,
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
@@
 #######################
+# for Genus_lacto taxonomy:
+taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
+  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
+  unique() %>%
+  filter(!(is.na(.$Genus_lacto))) %>%
+  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
+  column_to_rownames(., var = "Genus_lacto")
+# For other_taxa taxonomy:
+#taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
 par(mar=c(4,5,1,12))
 temp <- t(t(2^x-1)/colSums(2^x-1))
-taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+
 taxa[is.na(taxa)] <- "NA"
@@
 temp <- temp[o,sample_o]*100
-temp <- temp[-grep("NA",rownames(temp)),]
+#temp <- temp[-grep("NA",rownames(temp)),]
 temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
@@
   points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
-        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
+        col= c("red","orange","grey")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
         pch=15,cex=.8,xpd=T)
-  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)
+  text(par("usr")[1],par("usr")[3]-2,labels = "BV Diagnosis",pos=2,cex=.8,xpd=T)
 
@@
 # Define Layout and set labels
-layout(matrix(c(2 ,3,4,5,      6,7,17,17,          17,17,17,17,
-                1,1,1,1,      1,8,17,17,          17,17,17,17,
-                1,1,1,1,      1,9,18,18,         18,18,18,18,
-                1,1,1,1,      1,10,19,19,         19,19,19,19,
-                16,15,14,13,  12,11,20,20,         20,20,20,20 ) , ncol= 12,byrow = T),widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
-# layout.show(12)
+layout(matrix(c(2,3,4,5,      6,7,11,11,          11,11,11,11,
+                1,1,1,1,      1,8,11,11,          11,11,11,11,
+                1,1,1,1,      1,9,12,12,         12,12,12,12,
+                1,1,1,1,      1,10,13,13,         13,13,13,13,
+                0,0,0,0,      0,0,14,14,         14,14,14,14 
+                ) , 
+              ncol= 12,byrow = T),widths = 1,heights = c(1.5,1.2,1.2,1.2,1.2))
+# layout.show(3)
 figlabels <- letters
@@
 ###################################
-taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
-taxa[is.na(taxa)] <- "NA"
-
+# for Genus_lacto taxonomy:
+taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
+  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
+  unique() %>%
+  filter(!(is.na(.$Genus_lacto))) %>%
+  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
+  column_to_rownames(., var = "Genus_lacto")
+# for other taxa
+# taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+# taxa[is.na(taxa)] <- "NA"
+
+#par(mar=c(1,10,1,15))
 par(mar=c(1,10,1,15))
@@
 y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
-# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
-#        plot_labels = T, plot_weights = F,order_1_by = "total",
-#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-#        pal=hue_pal()(length(unique(x))))
-
-# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
-#      plot_labels = T, plot_weights = F,order_1_by = "total",
-#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-#      pal=hue_pal()(length(unique(y))))
 
@@
 
-design <- model.matrix(~group_annotation, data=metadata[sample_use,])
+design <- model.matrix(~ group_annotation+Contraception+HIVstatus+age+sexwork_months, 
+                       data=metadata[sample_use,])
 
@@
 
-lrt <- glmLRT(fit,coef=2:8)
+lrt <- glmLRT(fit,coef=2:9)
 top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
 colnames(top) <- sub("group_annotation","",colnames(top))
-top <- cbind(logFC.Gard=0,top)
+top <- cbind(LogFC.intercept=0,top) # with intercept
 
@@
 points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
-        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
+        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]],
         pch=16)
 points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.005,
-        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
+        col= c("red","orange","grey")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
         pch=16,xpd=T)
@@
 
-mypar(2,3,mar=c(2,2,2,6))
+mypar(2.5,2.5,mar=c(2.5,2.5,2,6))
 set.seed(1)
@@
                  repulsion_strength = .4,negative_sample_rate = 3)
-plot(UMAP_picrust,bg=pal[group_annotation[sample_use]],main="UMAP on top functional profile",
+plot(UMAP_picrust,bg=pal[factor(group_annotation[sample_use])],main="UMAP on top functional profile",
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
@@
 ```{r, echo = FALSE, eval = TRUE}
-taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+genus_lacto.fun <- function(df){
+  genus_lacto <- df %>%
+    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
+    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste(.$Genus_lacto, "crispatus/acidophilus"), 
+                                ifelse(grepl("reuteri",.$Species), paste(.$Genus_lacto, "reuteri/oris/frumenti/antri"),
+                                  ifelse(grepl("gasseri",.$Species), paste(.$Genus_lacto,"gasseri/johnsonii/taiwanensis"),
+                                    ifelse(grepl("murinus",.$Species), paste(.$Genus_lacto,"murinus/animalis/apodemi/salivarius"),
+                                      ifelse(grepl("plantarum",.$Species), 
+                                           paste(.$Genus_lacto, "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
+                                        ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
+                                              as.character(.$Genus_lacto))))))) )
+  return(genus_lacto)
+}
+
+picrust_taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
+  column_to_rownames(var = "Seq_ID") %>%
+  mutate(Genus_lacto = ifelse(is.na(.$Genus_lacto), "NA", .$Genus_lacto))
+  #mutate(Genus_lacto = ifelse(is.na(.$Genus_lacto), "NA_", .$Genus_lacto))
+
+# for Genus_lacto taxonomy:
+taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
+  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
+  unique() %>%
+  filter(!(is.na(.$Genus_lacto))) %>%
+  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
+  column_to_rownames(., var = "Genus_lacto")
+# for other taxa
+#taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
 dim(picrust_taxa)
@@
 
-picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , as.character(picrust_taxa$Seq_ID)  ),"final_name"])
+picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , rownames(picrust_taxa)  ), params])
 picrust_KO_seq <- as.matrix(picrust_KO_seq)
@@
 sub(" [(].*","",names(pvalues))[1]
+
+#p <- sub(" [(].*","",names(pvalues))
+#x <- p
 
@@
 ## Figure 3
+## individual group comparison DGE
+```{r}
+TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+TRX_counts <- round(2^TRX - 1)
+TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
+sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
+TRX_counts <- TRX_counts[,sample_use]
+
+library(edgeR)
+#group <- "TissueGroupM4"
+#group <- "group_annotation"
+edgR_dge.fun <- function(group, confound) {
+  groups <- metadata[sample_use, group]
+
+  if(confound==FALSE) {
+    c <- "_no_conf"
+    design <- model.matrix(~  groups, data=metadata[sample_use,])} # without confounders
+  else{
+    c <- "_w_conf"
+    # design <- model.matrix(~groups+HIVstatus+Contraception+age+sexwork_months,
+    #                      data=metadata[sample_use,])
+    design <- model.matrix(~groups+HIVstatus+Contraception,
+                         data=metadata[sample_use,])
+  }
+  #design <- model.matrix(~0 + groups+Contraception+HIVstatus+age+sexwork_months, 
+                         #data=metadata[sample_use,]) # without intercept
+
+  colnames(design) <- sub("groups","",colnames(design))
+  colnames(design) <- sub("(Intercept)","Intercept",colnames(design))
+
+  y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
+  y <- calcNormFactors(y,method = "TMM")
+  y <- estimateGLMCommonDisp(y,design)
+  y <- estimateGLMTagwiseDisp(y,design)
+  fit <- glmFit(y, design)
+
+  #lrt <- glmLRT(fit,coef=1:ncol(design)) # without intercept
+  lrt <- glmLRT(fit,coef=2:5) # with intercept
+  top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
+  colnames(top) <- sub(group,"",colnames(top))
+  top <- cbind(LogFC.intercept=0,top) # with intercept
+  top <-  rownames_to_column(top, var = "Genes")
+
+  saveRDS( y , paste0("../../../results/TRX_EdgeR_estimations_",group,c,".rds" ))
+  #write.xlsx(top, file=paste0("../../../results/DGE/","dge_heatmap_",group,c,".xlsx"))
+  write.xlsx(top, file=paste0("../../../results/DGE/","dge_heatmap_",group,"_2:5_conf_2.xlsx"))
+  #write.csv2( top , paste0("../../../results/DGE/TRX_EdgeR_top_dge_",group,".csv" ))
+
+return(list(design=design, y=y, fit=fit, lrt=lrt, top=top, c=c))
+}
+
+dge_M4 <- edgR_dge.fun("ModifiedM4", confound=T)
+#dge_M4_nonconfound <- edgR_dge.fun("ModifiedM4", confound=F)
+#dge_TissueM4 <- edgR_dge.fun("TissueGroupM4", confound=T)
+
+```
 
@@
 
-TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-TRX_counts <- round(2^TRX - 1)
-TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
-sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
-TRX_counts <- TRX_counts[,sample_use]
-
-library(edgeR)
-
-design <- model.matrix(~  group_annotation, data=metadata[sample_use,])
-
-y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
-y <- calcNormFactors(y,method = "TMM")
-y <- estimateGLMCommonDisp(y,design)
-y <- estimateGLMTagwiseDisp(y,design)
-fit <- glmFit(y, design)
-
-lrt <- glmLRT(fit,coef=2:8)
-top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
-colnames(top) <- sub("group_annotation","",colnames(top))
-top <- cbind(logFC.Gard=0,top)
-
-saveRDS( y , "../../../results/TRX_EdgeR_estimations.rds" )
-write.csv2( top , "../../../results/TRX_EdgeR_top_dge.csv" )
-
-top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
-top_dge <- rownames(top)[ top_dge ]
+# group_name <- "ModifiedM4"
+# top <- dge_M4$top
+# y <- dge_M4$y
+# c <- dge_M4$c
+gene_mod_enrich.fun <- function(top, y, c, db, group_name, cuttoff){
+  p_val <- pull(top, "PValue")
+  top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+  top_dge <- top$Genes[ top_dge ]
 
@@
   h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
-cuttoff <- 2
+  cuttoff <- cuttoff[names(cuttoff) == db]
   gene_module <- cutree(h, h=cuttoff)
 
-layout(matrix(c(1,2,3,4,
-              1,2,3,4),
-              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
-par(mar=c(4,.1,2,.1))
-plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
-# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
-abline(v=cuttoff,xpd=F,col="red",lty=2)
-points( rep(-.1,length(gene_module)),
-        seq(length(gene_module),1,length.out = length(gene_module)),
-        col=taxa_pal[factor(gene_module[h$order])],
-        pch=16,cex=.5,xpd=T)
-
-image( t(top_TRX_counts[h$order,order(group_annotation[sample_use])][nrow(top_TRX_counts):1,]),
-       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
-       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
-points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .02,length(sample_use)),
-        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
-        pch=16,xpd=T)
-
-points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
-        col= c("grey","red")[(!is.na(metadata[sample_use,"RepeatedPCR"]))+1][order(group_annotation[sample_use])],
-        pch=16,xpd=T)
-points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.01,
-        col= c("grey","red")[as.numeric(metadata[sample_use,"HIVstatus"])][order(group_annotation[sample_use])],
-        pch=16,xpd=T)
-
   source("../enrichment_function.R")
-gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
+  if(db == "GO"){
+  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")}else{
+  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")}
   res_list <- lapply( unique( gene_module[h$order] ),function(x){
@@
   } )
+  res <- map(res_list, ~rownames_to_column(.x, var = "Terms")) %>% 
+    set_names(., paste0("gene_module_",  seq_along(res_list)))
+  #write.xlsx(res, file=paste0("../../../results/ENRICHMENT/",db,"_enrichment_heatmap_",group_name,c,".xlsx"))
+  write.xlsx(res, file=paste0("../../../results/ENRICHMENT/",db,"_enrichment_heatmap_",group_name,"_2:5_conf_2.xlsx"))
+
+  return(list(res_list=res_list, top_TRX_counts=top_TRX_counts, 
+              gene_module=gene_module, cuttoff=cuttoff, h=h, c=c))
+}
+
+db <- c("GO", "KEGG")
+heatmap_cvlM4 <- map(db, ~gene_mod_enrich.fun(dge_M4$top,dge_M4$y,dge_M4$c,
+                                              .x,"ModifiedM4", 
+                                              #cuttoff= c(GO=4, KEGG=3.5)
+                                              cuttoff= c(GO=2.5, KEGG=3)
+                                              )) %>% set_names(db)
+
+heatmap_cvlM4_nonconfound <- map(db,~gene_mod_enrich.fun(dge_M4_nonconfound$top,
+                                                      dge_M4_nonconfound$y,
+                                                      dge_M4_nonconfound$c,
+                                                      .x,"ModifiedM4",
+                                                      cuttoff= c(GO=2.5, KEGG=2)
+                                                      )) %>% set_names(db)
+
+# heatmap_tissueM4 <- map(db, ~gene_mod_enrich.fun(dge_TissueM4$top,
+#                                                  dge_TissueM4$y,
+#                                                  .x,"TissueM4", 
+#                                                  cuttoff=)) %>% set_names(db)
+
+###########
+#   GO    #
+###########
+gr <- heatmap_cvlM4_nonconfound
+gr <- heatmap_cvlM4
+
+name <- "ModifiedM4"
+groups <- metadata[sample_use, name]
+conf <- pluck(gr, "GO", "c")
+h <- pluck(gr, "GO", "h")
+cuttoff <- pluck(gr, "GO", "cuttoff")
+res_list <- pluck(gr, "GO", "res_list")
+top_TRX_counts <- pluck(gr, "GO", "top_TRX_counts")
+gene_module <- pluck(gr, "GO", "gene_module")
+
+layout(matrix(c(1,2,3,4,
+              1,2,3,4),
+              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.4,6)) #1,2,.3,8
+## Heatmap
+par(mar=c(4,.1,2,.1))
+plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
+# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
+abline(v=cuttoff,xpd=F,col="red",lty=2)
+points( rep(-.1,length(gene_module)),
+        seq(length(gene_module),1,length.out = length(gene_module)),
+        col=taxa_pal[factor(gene_module[h$order])],
+        pch=16,cex=.5,xpd=T)
+
+image( t(top_TRX_counts[h$order,order(groups)][nrow(top_TRX_counts):1,]),
+       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
+       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
+
+## Meta info bars
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
+        col=pal[factor(groups)[ order(groups)]],
+        pch=16)
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
+        col= c("red","orange","grey")[as.numeric(metadata[sample_use,"BV_Diagnosis_v3"])][order( groups )],
+        pch=16,xpd=T) 
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.015,
+        col= c("grey","red")[as.numeric(metadata[sample_use,"HIVstatus"])][order(groups)],
+        pch=16,xpd=T)
+
+## Gene module enrichment
 names(res_list) <- unique( gene_module[h$order] )
@@
 terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
+#terms <- unlist(lapply(res_list,function(x){ x$Terms [1:5] }))
+terms[is.na(terms)] <- ""
+genes <- unlist(lapply(res_list,function(x){ x$genes [1:5] }))
+genes[is.na(genes)] <- ""
+
+pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
+module_color <- unlist(lapply(names(res_list),function(x){ rep(x,5) }))
+
+temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
+plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))] , use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)
+
+par(mar=c(4,.1,2,20))
+
+barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
+         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",xlab="-log10(pvalue)",las=1,names.arg = "",xpd = FALSE)
+points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
+text( rev(pvalues)+1.5 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,xpd=T,
+      labels = rev(names(pvalues)))
+text( rev(pvalues)+1.5 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,
+      labels = rev(sub(" [(].*","",names(pvalues)) ),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
+# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
+#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
+
+#dev.copy2pdf(file=paste0("../../../results/Plots/","2:5_DGE_heatmap_GO_",name,conf,".pdf"), 
+dev.copy2pdf(file=paste0("../../../results/Plots/","2:5_DGE_heatmap_GO_","M4_conf_2",".pdf"),
+    width = 6.5,
+    height = 6, paper = "a4" #units = "cm", res = 300 
+    )
+
+###########
+#  KEGG   #
+###########
+gr <- heatmap_cvlM4_nonconfound
+gr <- heatmap_cvlM4
+
+name <- "ModifiedM4"
+groups <- metadata[sample_use, name]
+conf <- pluck(gr, "KEGG", "c")
+cuttoff <- pluck(gr, "KEGG", "cuttoff")
+h <- pluck(gr, "KEGG", "h")
+res_list <- pluck(gr, "KEGG", "res_list")
+top_TRX_counts <- pluck(gr, "KEGG", "top_TRX_counts")
+gene_module <- pluck(gr, "KEGG", "gene_module")
+
+layout(matrix(c(1,2,3,4,
+              1,2,3,4),
+              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.4,6)) #1,2,.3,8
+par(mar=c(4,.1,2,.1))
+
+## Heatmap
+plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
+# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
+abline(v=cuttoff,xpd=F,col="red",lty=2)
+points( rep(-.1,length(gene_module)),
+        seq(length(gene_module),1,length.out = length(gene_module)),
+        col=taxa_pal[factor(gene_module[h$order])],
+        pch=16,cex=.5,xpd=T)
+
+image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(groups)]),
+       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
+       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
+
+## Meta info bars
+points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
+        col=pal[factor(groups)[ order(groups)]],
+        pch=16)
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.005,
+        col= c("red","orange","grey")[as.numeric(metadata[sample_use,"BV_Diagnosis_v3"])][order( groups )],
+        pch=16,xpd=T)
+points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.015,
+        col= c("grey","red")[as.numeric(metadata[sample_use,"HIVstatus"])][order(groups)],
+        pch=16,xpd=T)
+
+## Gene module enrichment
+names(res_list) <- unique( gene_module[h$order] )
+
+pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:5] }))
+pvalues[is.na(pvalues)] <- 0
+terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
+#terms <- unlist(lapply(res_list,function(x){ x$Terms [1:5] }))
 terms[is.na(terms)] <- ""
@@
 
-layout(matrix(c(1,2,3,4,
-              1,2,3,4),
-              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
-par(mar=c(4,.1,2,.1))
-
-plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
-# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
-abline(v=cuttoff,xpd=F,col="red",lty=2)
-points( rep(-.1,length(gene_module)),
-        seq(length(gene_module),1,length.out = length(gene_module)),
-        col=taxa_pal[factor(gene_module[h$order])],
-        pch=16,cex=.5,xpd=T)
-
-image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(group_annotation[sample_use])]),
-       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
-       breaks = seq(-5,5,length.out = 100),axes=F , xlab="patient groups")
-points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[3] - .01,length(sample_use)),
-        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
-        pch=16,xpd=T)
-
-source("../enrichment_function.R")
-gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
-res_list <- lapply( unique( gene_module[h$order] ),function(x){
-  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
-                            gmt_list = gmt_list,
-                            min_terms_pathway = 10,
-                            max_terms_pathway = 300,
-                            min_overlap = 3,
-                            sort_by_pvalue = T)
-  temp <- temp[!grepl("REGULATION",rownames(temp)),]
-  return(temp)
-} )
-names(res_list) <- unique( gene_module[h$order] )
-
-pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:5] }))
-pvalues[is.na(pvalues)] <- 0
-terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:5] }))
-terms[is.na(terms)] <- ""
-genes <- unlist(lapply(res_list,function(x){ x$genes [1:5] }))
-genes[is.na(genes)] <- ""
-
-pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
-module_color <- unlist(lapply(names(res_list),function(x){ rep(x,5) }))
-
-temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
-plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))] , use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)
-
-par(mar=c(4,.1,2,30))
-
-barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
-         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",xlab="-log10(pvalue)",las=1,names.arg = "")
-points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
-text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,xpd=T,
-      labels = rev(names(pvalues)))
-text( rev(pvalues)+1 , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=.8,
-      labels = rev(sub(" [(].*","",names(pvalues)) ),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
-# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
-#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
-
-mypar(2,3,mar=c(2,2,2,6))
+#dev.copy2pdf(file=paste0("../../../results/Plots/","2:5_DGE_heatmap_KEGG_",name,conf,".pdf"), 
+dev.copy2pdf(file=paste0("../../../results/Plots/","2:5_DGE_heatmap_KEGG_","M4_conf_2",".pdf"), 
+    width = 6.5,
+    height = 6, paper = "a4" #units = "cm", res = 300
+)
+
+###########
+#  UMAP   #
+###########
+mypar(3,3,mar=c(3,3,2,6))
 set.seed(1)
@@
                  negative_sample_rate = 10)
-plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="UMAP on top DGE RNAseq",
+plot(UMAP_TRX,bg=pal[factor(groups)],main="UMAP on top DGE RNAseq",
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
 legend(par("usr")[2],par("usr")[4],title.adj = 0,
-       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
+       #legend = levels(factor(groups)),
+       legend = sort(unique(groups)),
+       xjust = 0,yjust = 1,
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+
+###############
+#  HALLMARK   #
+###############
 
@@
 
+```{r}
+## individual group comparision ##
+# for information about how to get pairwise comparison se:
+# edgeR User’s Guide section 3.4.2 Blocking and 4.2.8  Differential expression
+# https://www.bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf
+pairwise_dge.fun <- function(group, fit, design) {
+  groups <- metadata[sample_use, group]
+  levels <- as.character(sort(unique(groups)))
+
+  # without intercept:
+  n <- combn(levels,2) 
+  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
+  s <- seq_along(1:ncol(n)) %>% set_names(g)
+  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))
+
+  # with intercept:
+  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)
+
+  CONTRASTS <- makeContrasts( contrasts=g,
+                              levels = design )
+
+  lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
+  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% map(., ~rownames_to_column(.x, var = "Genes"))
+
+  #saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
+  #saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
+return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS))
+}
+pairwise_dge_M4 <- pairwise_dge.fun("ModifiedM4", 
+                                    fit=dge_M4$fit, 
+                                    design=dge_M4$design )
+# pairwise_dge_TissueM4 <- pairwise_dge.fun("TissueGroupM4", 
+#                                           fit=dge_TissueM4$fit, 
+#                                           design=dge_TissueM4$design)
+
+# save TopTables 
+l_top <- list(ModifiedM4 = pairwise_dge_M4$top) # TissueM4 = pairwise_dge_TissueM4$top
+imap(l_top, ~write.xlsx(.x, #file=paste0("../../../results/DGE","dge_pairwise_",.y,".xlsx"))) # Excel
+file=paste0("../../../results/DGE/","dge_pairwise_",.y,"_2:5_conf_2",".xlsx"))) # Excel
+## end of individual group comparision ##
+
+## individual GO/KEGG comparison ##
+source("../enrichment_function.R")
+gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
+gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
+enrich_list <- function(x, gmt_list){
+  temp <- compute_enrichment(genes = x,
+                            gmt_list = gmt_list,
+                            min_terms_pathway = 10,
+                            max_terms_pathway = 300,
+                            min_overlap = 3,
+                            sort_by_pvalue = T)
+  temp <- temp[!grepl("REGULATION",rownames(temp)),]
+  temp <- rownames_to_column(temp, var = "Terms")
+  return(temp)
+}
+
+#info_logFC <- map(top, ~table(abs(.x$logFC) > 1))
+#info_PValue <- map(top, ~table(.$PValue < 0.01))
+
+# up and down regulated genes seperatly
+t_split <- modify_depth(l_top, 2, ~ .x %>%
+        dplyr::filter(., .$PValue < 0.01) %>%
+        #dplyr::filter(., abs(.$logFC) < 1) %>%
+        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
+        split(., .$reg)
+  )
+# up and down regulated genes toghether
+t <- modify_depth(l_top, 2,~ .x %>%
+        dplyr::filter(., .$PValue < 0.01) #%>%
+        #dplyr::filter(., abs(.$logFC) < 1) 
+  )
+
+## GO enrichment rownames(.x)
+GO_enrichment <- map_depth(t, 2, ~enrich_list(.x$Genes, gmt_list_GO))
+GO_enrichment_split <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_GO)) %>%
+  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) %>% 
+  set_names(., paste0(names(.), "_split"))
+
+## KEGG enrichment
+KEGG_enrichment <- map_depth(t, 2, ~enrich_list(.x$Genes, gmt_list_KEGG))
+KEGG_enrichment_split <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_KEGG)) %>%
+  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) %>% 
+  set_names(., paste0(names(.), "_split"))
+
+## save files
+enrich_KEGG <- list(KEGG_enrichment, KEGG_enrichment_split) %>% flatten()
+enrich_GO <- list(GO_enrichment, GO_enrichment_split) %>% flatten()
+imap(enrich_KEGG, ~write.xlsx(.x, #file=paste0("../../../results/ENRICHMEN/T","GO_enrichment_pairwise_",.y,".xlsx"))) # Excel
+file=paste0("../../../results/ENRICHMENT/","GO_enrichment_pairwise_",.y,"_2:5_conf_2",".xlsx"))) # Excel
+imap(enrich_GO, ~write.xlsx(.x, #file=paste0("../../../results/ENRICHMENT/","KEGG_enrichment_pairwise_",.y,".xlsx"))) # Excel
+file=paste0("../../../results/ENRICHMENT/","KEGG_enrichment_pairwise_",.y, "_2:5_conf_2",".xlsx"))) # Excel
+
+## end of individual GO comparison ##
+```
+
 \clearpage
@@
   points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
-        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
+        col= c("grey","red", "black")[(!is.na(metadata$BV_Diagnosis_v3))+1][order( group_annotation )],
         pch=15,cex=1,xpd=T)
@@
          y = par("usr")[c(4)],title.adj = 0,
-         legend = levels(group_annotation),xjust = 0,yjust = 1,
+         legend = sort(unique(group_annotation)),
+         #legend = levels(group_annotation),
+         xjust = 0,yjust = 1,
          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
@@
 
-group_means <- sapply(levels(group_annotation), function(x){
+group_means <- sapply(levels(factor(group_annotation)), function(x){
   rowMeans( datasets[ , group_annotation == x ] )
@@
 ```{r, echo = FALSE, eval = TRUE,fig.asp=1.3}
-layout( matrix(1:(16*8), 16,8,byrow = F) )
+#layout( matrix(1:(16*8), 16,8,byrow = F) )
+layout( matrix(1:(10*8), 10,8,byrow = F) )
 par(mar=c(.5,.5,1,.5))
@@
 
-  plot(UMAP_picrust,bg=pal[group_annotation],main=gsub("ASV_|_norm.*","",j),
+  plot(UMAP_picrust,bg=pal[factor(group_annotation[sample_use])],
+       main=gsub("ASV_|_norm.*","",j),
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
@@
 
-  plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main=gsub("ASV_|_norm.*","",j),
+  plot(UMAP_TRX,bg=pal[factor(group_annotation[sample_use])],main=gsub("ASV_|_norm.*","",j),
      pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
@@
 
-violist(data=as.matrix(t(metadata[,continuous_parameters])),
-          srt=45,
-          genes=continuous_parameters,
-          col = pal, transparency = 50,
-          clustering=metadata$group_annotation,smooth = 2 )
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
+# violist(data=as.matrix(t(metadata[,continuous_parameters])),
+#           srt=45,
+#           genes=continuous_parameters,
+#           col = pal, transparency = 50,
+#           clustering=metadata$group_annotation,smooth = 2 )
+# add_letter(figlabels[1]); figlabels <- figlabels[-1]
+# 
+# m <-  na.omit(as.matrix(t(metadata[,continuous_parameters])))
+# m <- metadata[,continuous_parameters] %>% as.data.frame() %>%
+#   rownames_to_column(var = "ID") %>% pivot_longer(-ID, names_to= "Param", values_to= "Values") %>% pivot_wider(names_from = "ID", values_from="Values") %>%
+#   column_to_rownames(var = "Param")
 ```
@@
 plot(l,
-       col=pal[metadata$group_annotation],
-       main=i,xpd=T,
+       col=pal[factor(group_annotation[sample_use])],
+       main="SNN Louvain clustering",xpd=T,
        pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
@@
 legend("topleft",title.adj = 0,
-       legend = levels(metadata$group_annotation),xjust = 0,yjust = 1,
+       legend = sort(unique(metadata$group_annotation)),xjust = 0,yjust = 1,
        bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
@@
 plot(l,
-       col=pal[metadata$group_annotation],
-       main=i,xpd=T,
+       col=pal[factor(group_annotation[sample_use])],
+       main="SNN Louvain clustering",xpd=T,
        pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
 legend(par("usr")[2],par("usr")[4],title.adj = 0,
-       legend = levels(metadata$group_annotation),xjust = 0,yjust = 1,
+       legend = levels(factor(group_annotation[sample_use])),xjust = 0,yjust = 1,
        bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
@@
 plot(UMAP_TRX,
-       col=pal[metadata[common,"group_annotation"]],
-       main="",xpd=T,
+       col=pal[factor(group_annotation[sample_use])],
+       main=""UMAP trx"",xpd=T,
        pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
@@
 legend("topleft",title.adj = 0,
-       legend = levels(metadata[common,"group_annotation"]),xjust = 0,yjust = 1,
+       legend = levels(factor(group_annotation[sample_use])),xjust = 0,yjust = 1,
        bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
@@
 plot(UMAP_TRX,
-       col=pal[metadata[common,"group_annotation"]],
-       main=i,xpd=T,
+       col=pal[factor(group_annotation[sample_use])],
+       main="UMAP trx",xpd=T,
        pch=16,xlab="",ylab="",frame=F,axes=F,font.main=1,cex.main=1)
 legend(par("usr")[2],par("usr")[4],title.adj = 0,
-       legend = levels(metadata[common,"group_annotation"]),xjust = 0,yjust = 1,
+       legend = levels(factor(group_annotation[sample_use])),xjust = 0,yjust = 1,
        bty = "n",pch = 16,col = pal,pt.cex = 1,xpd=T)
@@
 
-```{r, echo = FALSE, eval = TRUE, fig.width=6, fig.height=10}
+```{r, echo = FALSE, eval = F, fig.width=6, fig.height=10}
 paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3
@@
 
-```{r Print rare curves, fig.width=6, fig.height=10, fig.align="center"}
+```{r Print rare curves, eval=F, fig.width=6, fig.height=10, fig.align="center"}
 ggarrange(r_plot[[1]] + theme(axis.title.x=element_blank()),

### 2021-08-16 reports/rmarkdown/01_data_pre_processing.Rmd

@@
     mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
-    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste(.$Genus_lacto, "crispatus/acidophilus"), 
-                                ifelse(grepl("reuteri",.$Species), paste(.$Genus_lacto, "reuteri/oris/frumenti/antri"),
-                                  ifelse(grepl("gasseri",.$Species), paste(.$Genus_lacto,"gasseri/johnsonii/taiwanensis"),
-                                    ifelse(grepl("murinus",.$Species), paste(.$Genus_lacto,"murinus/animalis/apodemi/salivarius"),
+    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste("L.", "crispatus/acidophilus"), 
+                                ifelse(grepl("reuteri",.$Species), paste("L.", "reuteri/oris/frumenti/antri"),
+                                  ifelse(grepl("gasseri",.$Species), paste("L.","gasseri/johnsonii/taiwanensis"),
+                                    ifelse(grepl("murinus",.$Species), paste("L.","murinus/animalis/apodemi/salivarius"),
                                       ifelse(grepl("plantarum",.$Species), 
-                                           paste(.$Genus_lacto, "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
-                                        ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
+                                           paste("L.", "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
+                                        ifelse(grepl("Lactobacillus",.$Genus), paste("L.", .$Species), 
                                               as.character(.$Genus_lacto))))))) )
@@
     left_join(x, by="Taxonomy") %>%
-    mutate(across(everything(), ~replace_na(.x, 0))) %>%
+    mutate(across(Taxonomy, ~replace_na(.x, "Not assigned"))) %>%
+    mutate(across(-Taxonomy, ~replace_na(.x, 0))) %>%
     column_to_rownames(var = "Taxonomy")
@@
 
+# raw files
+n <- c("ASV_CVL_V3", "ASV_tissue_V3", "ASV_CVL_V2")
+map2(datasets[1:3],n,  ~write.csv(.x, file=paste0(PATH,"results/",.y,"_raw_counts.csv"),row.names = T) )
+
 datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)

### 2021-08-16 reports/rmarkdown/03_integrated_datasets.Rmd

--- 2021-03-12 reports/rmarkdown/03_integrated_datasets.Rmd
@@
 library(niceRplots)
+library(tidyverse)
 ```
@@
 PATH <- "../../" #Path to the data
+Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
 
 #color palettes
-pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
+pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
          RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
@@
 ```{r, echo = FALSE}
-# Load metadata
-metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
+# Load meta data
+meta <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatefeb22.csv"))
+rownames(meta) <- as.character(meta$ID)
 
@@
 
-#import datasets and return matrix with taxa-level resolution
+#import datasets and return matrix with genus_lacto level resolution
+# datasets <- map(dataset_names, 
+#                 ~read.csv(paste0(PATH,"results/",.x, ".csv"),row.names = 1)) %>%
+#   set_names(., dataset_names)
+
+#fill all samples with 0s to make plotting easier
+#x <- datasets[[2]]
+# all_samples <- unique(unlist(lapply(datasets,colnames)))
+# datasets_all_samples <- lapply(datasets, function(x) {
+#   col <- setdiff(all_samples, colnames(x))
+#   temp <- tibble(t = rownames(x))  %>% 
+#      tibble::add_column(!!!set_names(as.list(rep(0, length(col))),nm=col)) %>% 
+#     column_to_rownames(., var = "t") %>% bind_cols(x) 
+#   return(temp)
+# })
+
+#import datasets and return matrix with genus_lacto level resolution
 datasets <- lapply(dataset_names, function(x){
@@
 
+# Define what samples to use
+samples_to_use <- intersect(colnames(datasets[["ASV_CVL_V3_normalized_batch_corrected"]]),
+                            colnames(datasets[["Tissue_RNAseq_V3_normalized"]]))
+
 #fill all samples with 0s to make plotting easier
+x <- datasets[[1]]
 datasets_all_samples <- lapply(datasets, function(x) {
   temp <- matrix(0,nrow = nrow(x),
-                 ncol = nrow(metadata),
-                 dimnames = list(rownames(x),rownames(metadata)))
+                 ncol = nrow(meta),
+                 dimnames = list(rownames(x),rownames(meta))
+                 )
   temp[rownames(x),colnames(x)] <- x
@@
 lapply(datasets_all_samples,dim)
+
+# Clean up metadata
+tissue_gr <- read.csv(paste0(PATH,"/data/210609_TissuegroupsJune.csv"))
+metadata <- meta %>%
+  left_join(tissue_gr, by="ID") %>%
+  mutate(Luminal_names = ifelse(.$ModifiedM4 == "NonIners", "L. crispatus/jensenii",
+                        ifelse(.$ModifiedM4 == "Iners", "L. iners",
+                          ifelse(.$ModifiedM4 == "Gard", "Gardnerella",
+                            ifelse(.$ModifiedM4 == "Prev", "High diverse", 
+                                    "Other")))) ) %>%
+  mutate(Luminal_gr = ifelse(.$ModifiedM4 == "NonIners", "L1",
+                        ifelse(.$ModifiedM4 == "Iners", "L2", 
+                          ifelse(.$ModifiedM4 == "Gard", "L3",
+                            ifelse(.$ModifiedM4 == "Prev", "L4", 
+                                    "L5")))) ) %>%
+  mutate(Tissue_names = ifelse(.$TissuegroupsJune == "Crisp", "L. crispatus/jensenii",
+                        ifelse(.$TissuegroupsJune == "Iners", "L. iners", 
+                          ifelse(.$TissuegroupsJune == "Gard", "Gardnerella",
+                            ifelse(.$TissuegroupsJune == "Prev", "High diverse", 
+                                    "Other")))) ) %>%
+  mutate(Tissue_gr = ifelse(.$TissuegroupsJune == "Crisp", "T1",
+                        ifelse(.$TissuegroupsJune == "Iners", "T2", 
+                          ifelse(.$TissuegroupsJune == "Gard", "T3",
+                            ifelse(.$TissuegroupsJune == "Prev", "T4", 
+                                    "T5")))) ) %>%
+  select(ID:Yeast_v3, group_M9_CVLv3, Luminal_names, Tissue_names, Luminal_gr,Tissue_gr)
+
+rownames(metadata) <- as.character(metadata$ID)
 ```
@@
 
-samples_to_use <- colnames(counts)[counts[c(3),] != 0]
 datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
@@
 #Graph construction  
-k <- 10 # k for KNN
-p <- 0.1 # prunning threshold of SNN
+k <- 12 # k for KNN
+p <- 0.1 # prunning threshold of SNN 0.14
 
@@
 
-metadata$joint_clustering <- cl$membership
+metadata$Louvain_clusters <- cl$membership
+metadata$joint_clustering <- metadata$Luminal_gr
+metadata$joint_clustering <- factor(metadata$joint_clustering) 
 set.seed(1)
@@
 
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
@@
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
@@
 empty_plot()
-legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
-       legend = levels(factor(metadata$group_M4_CVLv3)),
+legend("topleft",title = "Patient groups\nLuminal_gr",
+       legend = levels(factor(metadata$Luminal_gr)),
        bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
@@
 ```{r}
-mypar(mar=c(2,15,2,2))
+mypar(mar=c(3.2,15,2,2))
 
@@
          clustering = metadata$joint_clustering, 
-         col=pal,draw_mean_lines=F)
+         col=pal[factor(metadata$joint_clustering)],draw_mean_lines=F)
 add_letter("b")
@@
 barlist( data = all_microbiome, main = "detected features",
-         genes = c("Gardnerella vaginalis",
+         genes = c("Gardnerella",
                    "Lactobacillus iners",
@@
                    "Lactobacillus crispatus/acidophilus",
-                   "Prevotella amnii",
-                   "Prevotella timonensis",
-                   "Atopobium vaginae"),
+                   "Prevotella",
+                   "Atopobium"),
          clustering = metadata$joint_clustering,
-         col=pal,draw_mean_lines=F)
-
+         col=pal[factor(metadata$joint_clustering)], draw_mean_lines=F) # ,srt = 65
+#legend(0,legend = levels(factor(metadata$Luminal_gr)))
+
+# barlist( data = all_microbiome, main = "detected features",
+#          genes = c("Gardnerella vaginalis",
+#                    "Lactobacillus iners",
+#                    "Lactobacillus jensenii",
+#                    "Lactobacillus crispatus/acidophilus",
+#                    "Prevotella amnii",
+#                    "Prevotella timonensis",
+#                    "Atopobium vaginae"),
+#          clustering = metadata$joint_clustering,
+#          col=pal,draw_mean_lines=F)
 ```
@@
 x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
+xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
 xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
@@
          genes = c("BV","counts_16S"),
-         clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
+         clustering = metadata$joint_clustering, #srt = 60,
+         draw_mean_lines=F,col = pal[factor(metadata$joint_clustering)])
 
@@
          main = "Monbilicus",
-         genes = c("BV","counts_16S"),
+         genes = c("BV","counts_16S"), #srt = 60,
          clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
+         draw_mean_lines=F,col = pal[factor(metadata$joint_clustering)])
 
@@
                         counts_16S=xG),
-         main = "Gardanerella",
-         genes = c("BV","counts_16S"),
+         main = "Gardnerella",
+         genes = c("BV","counts_16S"), #srt = 60,
          clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
+         draw_mean_lines=F,col = pal[factor(metadata$joint_clustering)])
 ```
@@
 
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=4}
+```{r, eval=T,echo = FALSE,fig.width=10,fig.height=4}
 myletters <- letters
@@
 
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
+```{r, eval=T,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
@@
 
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
+```{r, eval=T,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
@@
 
-```{r, echo = FALSE}
-groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
-  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
-names(groupings) <- names(datasets_all_samples)[-1]
-all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))
-
-groupings_all_samples <- lapply(groupings,function(x){
-  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
-  df[rownames(x),1] <- as.character(x[,1])
-  return(df)
-})
-names(groupings_all_samples) <- names(groupings)
-
-df <- as.data.frame(groupings_all_samples)
-colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
-metadata <- cbind(metadata,df[rownames(metadata),])
-```
-
 # Computing differential expression across microbiome datasets
@@
 The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
+
+```{r, echo = FALSE}
+# groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
+#   x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
+# names(groupings) <- names(datasets_all_samples)[-1]
+# all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))
+# 
+# groupings_all_samples <- lapply(groupings,function(x){
+#   df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
+#   df[rownames(x),1] <- as.character(x[,1])
+#   return(df)
+# })
+# names(groupings_all_samples) <- names(groupings)
+# 
+# df <- as.data.frame(groupings_all_samples)
+# colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
+# metadata <- cbind(metadata,df[rownames(metadata),])
+
+Louvain_clusters <- bind_cols(ID = cl$names, louvain_CVL_V3 = cl$membership)
+
+df <- metadata %>%
+  left_join(., Louvain_clusters, by="ID")
+
+```
 
@@
                          "cervical_inflammation_v2","cervical_inflammation_v3",
+                         "BV_WBC_v2", "BV_WBC_v3","marital_status",
                          "HIVstatus","Contraception",
                          "CT_v2","CT_v3","regular_partner",
-                         "vaginal_douching_v2","vaginal_douching_v3","joint_clustering")
+                         "vaginal_douching_v2","vaginal_douching_v3","Luminal_gr", "Tissue_gr", 
+                         "Louvain_clusters")
 
@@
   bac_dataset <- datasets_all_samples[[i]]
-  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
+  dge <- read.csv2(paste0(PATH,"results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
   samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
@@
 ```{r, echo = FALSE,fig.width  = 10, fig.height=5}
-taxa <- read.csv2("../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
-taxa[is.na(taxa)] <- "NA"
-rownames(cors) %in% rownames(taxa)
+# for Genus_lacto taxonomy:
+taxa <- read_csv("../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
+  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
+  unique() %>%
+  filter(!(is.na(.$Genus_lacto))) %>%
+  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
+  column_to_rownames(., var = "Genus_lacto")
+
+# For other_taxa taxonomy:
+# taxa <- read.csv2("../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
+# taxa[is.na(taxa)] <- "NA"
+# rownames(cors) %in% rownames(taxa)
 
@@
 
-for( i in c("group_M4_CVLv3","group_M9_CVLv3","joint_clustering")){
+for( i in c("Luminal_gr","group_M9_CVLv3","Louvain_clusters")){
   for( j in c("ASV_tissue_V3_normalized_batch_corrected",
@@
     barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),col=pal,
+            genes=rownames(module_means_per_dataset[[j]]),col=pal[factor(temp_group)],
             labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
@@
 
-\clearpage
-
 ```{r, echo = FALSE,fig.width=10,fig.height=8}
+metadata <- select(metadata, -joint_clustering, -group_M9_CVLv3)
 write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)

### 2021-08-16 reports/rmarkdown/manuscript/Figure1.Rmd

---
title: "Figure 1. Microbiome abundances"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
#group_annotation <- as.character(metadata$Luminal_gr) %>% set_names()

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

#color palettes
#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")

pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
#pal <- c("#377EB8", "#4DAF4A", "#FF7F00", "#F781BF", "#984EA3", "#FFFF33", "#A65628",  "#999999")
#pal <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
#show_col(pal)
#show_col(taxa_pal)
```

```{r Figure 1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=6 ,fig.asp=1, fig.pos="t"}
### A + B
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################

figlabels <- letters
temp_legend <- NULL

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                4,4,4,4,4,4,
                5,5,8,9,6,6,
                5,5,8,7,6,6,
                5,5,8,7,7,7) , ncol= 6,byrow = T),
                               widths = c(1.2,1,1.85,1.35,1,1.4),
                               heights = c(.7,1.5,.7,1.5,1,.8,.2))
# layout(matrix(c(1,1,1,1,1,5,
#                 2,2,2,2,2,5,
#                 3,3,3,3,3,6,
#                 4,4,4,4,4,6,
#                 7,7,8,8,8,6) , ncol= 6,byrow = T),
#                                widths = c(.5,1,1,1,1,1.1),
#                                heights = c(.5,1,.5,1,1.5))

#i <- ASV[2]
ASV <- list(c("Luminal_gr","ASV_CVL_V3_normalized_batch_corrected"),
         c("Tissue_gr","ASV_tissue_V3_normalized_batch_corrected")
          #"3"="ASV_CVL_V2_normalized_NOT_batch_corrected"
         )
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))
  #x <- x[,sample_use]
  #x <- datasets_all_samples[[2]]

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="", #bg="white", #cex=.6, 
         genes = c("Shannon","simpson","Inv Simpson"),
         # clustering = group_annotation,
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  if(i[2] == "ASV_CVL_V3_normalized_batch_corrected"){
    end <- table(pal[factor(group_annotation[order(group_annotation)])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
    #print(par("usr"))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1,5,1,1.5)) #b,l,t,r
  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
  # taxa[is.na(taxa)] <- "NA"
  # taxa <- taxa[rownames(temp),]
  # temp <- rowsum(temp , taxa$Class)
  temp[is.nan(temp)] <- 0

  if(is.null(temp_legend)){
    temp_legend <- temp   #Use the same legend from the 1st dataset
  }

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  # temp <- temp[-grep("NA",rownames(temp)),]
  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=gsub("_gr","",i[1]),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = lvl)] ,names.arg = rep("",ncol(temp)))

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  #text(x = c(2), y = 3, ']', srt = 90, cex = 3, xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T) # pos=4,

}
### C
############################
# TOTAL RELATIVE ABUNDANCE #
############################
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]
raw <- c("ASV_CVL_V3", "ASV_tissue_V3") %>% set_names()

raw_list <- raw %>%
  map(., ~read_csv(paste0("../../../results/",.x,"_raw_counts.csv")))  %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.5514392, "Other")) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_CVL_V3) %>%
    dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
    `row.names<-`(., NULL) %>% 
    column_to_rownames(var = "name") %>%
    as.matrix()

#colSums(r)

par(mar=c(1,5,2,0)) #b,l,t,r
barplot( r, border = NA ,ylab="abundance", xlab ="",
         col = taxa_pal[factor(rownames(r),levels = rownames(temp))])
title(main = "Total relative abundance", line = 0.5, cex.main = .9, font.main= 1)
mtext("Luminal        Tissue ", line = 0, cex=.6, side=1)
#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]
# legend(x = par("usr")[c(2)],
#          y = par("usr")[c(4)],title.adj = 0,
#          legend = rownames(r),xjust = 0,yjust = 1,
#          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
#           title="top abundant bacteria",
#          pt.bg = taxa_pal)

### D
#############################
# LUMINAL-TISSUE COMPARISON #
#############################
par(mar=c(0,4,3,2)) #b,l,t,r
x <- as.character(metadata[,"Luminal_gr"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Tissue_gr"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
df <- data.frame(x1,y1)
plot_sankey( df,color_by = 1, 
             plot_labels = T, plot_weights = F,order_1_by = "NULL",
             order_2_by = "disentangled",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal= pal)
text(c(0,1),c(1,1),labels = c( "Luminal" , "Tissue" ),pos=3, xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###########
# LEGENDS #
###########
par(mar=c(0,.5,0,0)) #b,l,t,r
empty_plot()
legend(x = -0.04, #par("usr")[c(1)],
       y = 1.04, #par("usr")[c(4)],title.adj = 0,
       #legend = sort(unique(group_annotation)),
       legend = c("L. crispatus/jensenii", "L. iners", "Gardnerella", "High diverse", "Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1, #xjust=1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
       pt.bg = pal)

par(mar=c(0,.4,1,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],#title.adj = 0,
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          #title="",
         pt.bg = taxa_pal)

par(mar=c(0,.4,1,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],#title.adj = 0,
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          #title="",
         pt.bg = taxa_pal[13:17])

```

```{r include=FALSE}
dev.copy2pdf(file=paste0("./images/","Figure1",".pdf"), 
    width = 6.5,
    height = 7 , paper = "a4" #units = "cm", res = 300 # a4 = 8.3 x 11.7 inches
)
```

```{r eval=FALSE, include=FALSE}
##########################
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
y <- datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
z <- datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]

 # par(mar=c(1,5,1,1))
tempx <- t(t(2^x-1)/colSums(2^x-1))
tempy <- t(t(2^y-1)/colSums(2^y-1))
tempz <- t(t(2^z-1)/colSums(2^z-1))

tempx[is.nan(tempx)] <- 0
tempy[is.nan(tempy)] <- 0
tempz[is.nan(tempz)] <- 0
 # if(is.null(temp_legend)){
 #   temp_legend <- temp   #Use the same legend from the 1st dataset
#  }

temp1 <- rowSums(tempx)/100
temp2 <- rowSums(tempy)/108
temp3 <- rowSums(tempz)/93

Totrelab <- cbind(temp1, temp2, temp3)
Totrelab <- cbind(temp1, temp2)
Totrelab["Lactobacillus_crispatus/acidophilus", ] <- Totrelab["Lactobacillus_crispatus/acidophilus", ] + Totrelab["Lactobacillus_crispatus", ]
Totrelab[rownames(Totrelab) != "Lactobacillus_crispatus", ]
  o <- order( rowSums(Totrelab) , decreasing = T)
  #sample_o <- names(group_annotation)[ order( group_annotation )]
  temp_a <- Totrelab[o,]*100
  write.csv(temp_a, "totalRelAbundance.csv")

  # temp <- temp[-grep("NA",rownames(temp)),]
  temp_b <- rbind(temp_a[1:19,], Other = colSums(temp_a[20:nrow(temp_a),]) )

layout(matrix(c(2) , ncol= 2,byrow = T),heights = c(1))
#par(mar=c(2,7))
layout(matrix(c(2) , ncol= 2,byrow = T),
       heights = c(1))
  par(mar=c(2,7))
  barplot( temp_a, las=2 , border = NA ,ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp_a),levels = rownames(temp_a))] ,names.arg = rep("",ncol(temp_a)))
  legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],title.adj = 0,
         legend = rownames(temp_a),xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          title="top abundant bacteria",
         pt.bg = taxa_pal)
```

```{r fig1, eval=FALSE, fig.cap="Figure 1. Box plot of the total lengths of live and dead sparrows following a snowstorm in Providence", fig.height=5, fig.width=12, message=FALSE, warning=FALSE, include=FALSE}

#############
# LODA DATA #
#############
#ALL_DATA <- readRDS("~/Box/Projects/J_Mjosberg/NBIS_housekeeping_normalization/analysis_all_tissues/all_ilc_dataset.rds")
#pal <- c(scales::hue_pal()(8),RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2") )

########################
# DEFINE FIGURE LAYOUT #
########################
layout(matrix(c(1,2,5,
                1,2,6,
                1,2,7,
                1,2,8,
                1,2,9,
                1,3,10,
                1,3,11,
                1,3,12,
                1,3,13,
                1,3,14,
                1,4,15,
                1,4,16,
                1,4,16,
                17,4,16,
                17,4,16),
              nrow = 15, byrow = T), 
   widths=c(3,1.5,1.5), heights=c(1,1))
#layout.show(16)

################
# PLOT FIGURES #
################
figlabels <- letters

### A
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "RNA_snn_res.3", label = T, frame=F, main="",cex = .7)
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = "", col = "white", pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Clustering")

### B
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Tissue", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = levels(ALL_DATA$Tissue), col = pal[1:3], pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Tissue")

### C
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Celltype", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = levels(ALL_DATA$Celltype), col = pal[1:4], pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Celltype")

### D
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Donor", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = "", col = "white", pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Donor")

### E
################
my_gene_list <- c("PTGDR2","TNFSF10","TMEM273","KRT1","HPGD","FAM110A","ZP1","PKIB","TESPA1","CSGALNACT1")
par(mar=c(0.5,6,0,.5))
for(i in my_gene_list){ 
try({violins(data = ALL_DATA,gene = i,clustering = "Celltype",assay = "RNA",plot_y_axis = F,plot_x_axis = F,ylab="",main="",bw = 1)
mtext(side = 2, text = i, line = 1,las=1,cex = .8)})
}
mtext(side = 1, at=1:5, text = sort(unique(ALL_DATA@meta.data[,"Celltype"])), line = 1, las=1)
text(1:5, par("usr")[3], labels = sort(unique(ALL_DATA@meta.data[,"Celltype"])), srt = 45, adj = c(1,0.5), xpd = T, cex=1)

### F
################
my_image <- image_read("images/sparrow.jpg")
par(mar = c(1, 1, .5, .5), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot(my_image)
fig_label(figlabels[1],cex = 3,font = 2,region = "figure") ; figlabels <- figlabels[-1]

### G
################

```

### 2021-08-16 reports/rmarkdown/manuscript/manuscript_template.tex

@@
 Author Three\textsuperscript{1,2}}
-\date{March 18, 2021}
+\date{March 19, 2021}
 
@@
 \includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-1}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-2}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-3}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-4}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-5}
-\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-24-6}
+
+\clearpage
+
+\hypertarget{figure-s14}{%
+\subsection{Figure S14}\label{figure-s14}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-25-1}
+
+\clearpage
+
+\hypertarget{figure-s15}{%
+\subsection{Figure S15}\label{figure-s15}}
+
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-26-1}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-26-2}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-26-3}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-26-4}
+\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-26-5}
+
+\begin{center}\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/Print rare curves-1} \end{center}
+
+\textbf{Supp.
+Figure 10.
+Rarefraction curves}.
+\textbf{A} Boston 1 sequencing run, \textbf{B} Sthlm sequencing run and \textbf{C} and \textbf{D} shows Boston sequencing run 2 split in two plots showing the CVL visit 2 samples and the the TISSUE visit 3 samples respectively
 

### 2021-10-13 reports/rmarkdown/manuscript/Figure1.Rmd

@@
   library(niceRplots)
+  library(openxlsx)
 })  })  })
@@
          )
+div_list <- list()
 for(i in ASV){
@@
   divers[is.infinite(divers)] <- 0
-  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
+  rownames(divers) <- c("Shannon","Simpson","Inv Simpson")
+
+  div <- tibble(shann, simp, invsimp ) %>% rename_with(., ~paste0(c("Shannon","Simpson","Inv Simpson")))
+  div_list[[length(div_list) + 1]] <- div
+
   barlist( data = divers[,order(group_annotation)],
          main = "", xlab="", #bg="white", #cex=.6, 
-         genes = c("Shannon","simpson","Inv Simpson"),
+         genes = c("Shannon","Simpson","Inv Simpson"),
          # clustering = group_annotation,
@@
 }
+
+# save diversity index stats
+div_stat <- div_list %>% 
+  set_names(map(ASV,1)) %>%
+  map(., ~summarise_all(.x, list(median=median, IQR=IQR), na.rm = TRUE)) %>%
+  map(., ~pivot_longer(.x, everything(), names_sep = "_", names_to = c(".value", "Stats"), )) %>%
+  bind_rows(.id = "Sample type") %>%
+  mutate("Sample type" = gsub("_gr", "", .$"Sample type"))
+
+write.xlsx(div_stat, file=paste0("../../../results/","Suppl.Tbl.1 Diversity statistics",".xlsx"))
+
 ### C

### 2021-10-27 reports/rmarkdown/manuscript/Figure8.Rmd
Figure8.

---
title: "Differential Protein expression"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
})  })  })

#############
# LODA DATA #
#############
Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv") 
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)

#color palettes
#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

```{r DPE Functions, message=FALSE, warning=FALSE, include=FALSE}
################################
# DEG's FUNCTION ACROSS GROUPS #
################################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
limma_dpe.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~  groups, data=metadata[sample_use,]) } # without confounders
  else{
    c <- "_conf"
    # design <- model.matrix(~groups+HIVstatus+Contraception+age+sexwork_months,
    #                      data=metadata[sample_use,])
    design <- model.matrix(~ groups+Contraception,
                         data=metadata[sample_use,])
  }
  #design <- model.matrix(~0 + groups+Contraception+HIVstatus+age+sexwork_months, 
                         #data=metadata[sample_use,]) # without intercept

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- Protein[,sample_use]
  fit <- lmFit(y, design)
  fit2 <- contrasts.fit(fit, coefficients=2:n_gr)
  efit <- eBayes(fit2)

  # DPE across all groups
  top <- topTable(efit, adjust.method = "BH",number = "all",sort.by = "F") # ANOVA
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Antibody")

return(list(design=design, y=y, fit=fit, efit=efit, top=top, c=c))
}

###########################
# DEG's FUNCTION PAIRWISE #
###########################
pairwise_dpe.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups)))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  # DPE pairwise comparison
  fit2 <- contrasts.fit(fit, CONTRASTS)
  efit <- eBayes(fit2)
  top <- map(s, ~ topTable(efit, coef=.x, adjust.method = "BH",number = "all",sort.by = "P")) %>%
    map(., ~rownames_to_column(.x, var = "Antibody"))

  results <- decideTests(fit2, method="separate", p.value = 0.05)
  summary <- rownames_to_column(as.data.frame.array(summary(results)), var="FDR<0.05")
  top <- c(summary = list(summary), top)

  #saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
  #saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
return(list(efit=efit, top=top, CONTRASTS=CONTRASTS))
}

```

```{r format data, echo=FALSE, warning=FALSE, message=FALSE}
to_remove <- c("CSTB_HPA017380","SPINK5_HPA009067","SPRR3_HPA024330","FGA_HPA064755","S100A12_HPA002881","KRT1_HPA019797","KRT1_HPA062908","LGMN_HPA001426","TACSTD2_HPA043104","FLNA_HPA001115","GPX3_HPA071520","CSTA_HPA001031","SERPINB5_HPA019132","ITIH2_HPA059150","NCF2_HPA002327","ANXA3_HPA013431")

Protein <- Protein_V3 %>%
  left_join(select(metadata, PatID, ID), by="PatID") %>%
  select(ID, contains("HPA")) %>%
  filter(!is.na(.$ID)) %>%
  pivot_longer(., -ID, names_to = "Antibody") %>%
  pivot_wider(names_from = "ID", values_from = "value" ) %>%
  filter(!(Antibody %in% to_remove)) %>%
  column_to_rownames(var = "Antibody")

# Replace missing values with average in sexwork_months column
m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
metadata <- metadata %>%
  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))

```

```{r Across and Pairwise tables, include=FALSE}
# Samples to use
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- intersect(metadata$ID, colnames(TRX)[colSums(TRX)!=0])
sample_use <- intersect(colnames(Protein), sample_use)

# Across groups DPE
dpe_L <- limma_dpe.fun("Luminal_gr", confound=T)
dpe_T <- limma_dpe.fun("Tissue_gr", confound=T)

# save TopTables
write.xlsx(dpe_L$top, file=paste0("../../../results/DE/","dpe_across_","Luminal",dpe_L$c,".xlsx"))
write.xlsx(dpe_T$top, file=paste0("../../../results/DE/","dpe_across_","Tissue",dpe_T$c,".xlsx"))

# Pairwise DPE's
pairwise_dpe_L <- pairwise_dpe.fun("Luminal_gr", 
                                    fit=dpe_L$fit, 
                                    design=dpe_L$design )
pairwise_dge_T <- pairwise_dpe.fun("Tissue_gr",
                                   fit=dpe_T$fit,
                                   design=dpe_T$design)

# save TopTables 
l_top <- list(Luminal = pairwise_dpe_L$top, Tissue = pairwise_dge_T$top)
imap(l_top, ~write.xlsx(.x,file=paste0("../../../results/DE/","dpe_pairwise_",.y,"_conf",".xlsx"))) # Excel

```

```{r heatmap functions, echo=FALSE}

top_dpe.fun <- function(top){
  p_val <- pull(top, "P.Value")
  top_dpe <- (p_val < 0.05) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dpe <- top$Antibody[ top_dpe ]

  top_PROT_mfi <- Protein[top_dpe, sample_use]

  return(top_PROT_mfi)
}

# top_PROT_mfi <- t
# gr <- "Luminal_gr"
# plot_heatmap.fun(t, "Luminal_gr", scale = T)
plot_heatmap.fun <- function(top_PROT_mfi, gr, scale = FALSE){
  df <- na.omit(metadata[sample_use, c(gr, "ID")])
  df <- arrange(df, df[1])
  samples <- df[, "ID"]
  names(samples) <- df[, gr]

  if(scale == TRUE) {
    n <- colnames(top_PROT_mfi)
    top_PROT_mfi <- t(apply(top_PROT_mfi,1,function(x){scale(x,T,T)}))
    colnames(top_PROT_mfi) <- n
    breaks = seq(-5,5,length.out = 100)
  }else{breaks = seq(0,5,length.out = 100)}

  h <- hclust( as.dist( (1- cor(t(top_PROT_mfi)))/2 ), method = "ward.D2")

  layout(matrix(c(1,2,3,4,
                  1,2,3,4),
                  nrow = 1,ncol = 2,byrow = T),widths = c(1.,2,.4,6) )
    ## Heatmap
    par(mgp=c(0,0.2,0), mar=c(4,.4,1,0)) #(bottom, left, top, right) 
    plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = 0.7 )
    #title(xlab="height", line=3)
    mtext( c("Height"), side=1, las = 1, cex = 0.7, line =1,xpd=T)

    if(gr == "Tissue_gr") {
      height <-1.5 }else{height <- round(max(h$height))}
    lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob

    # abline(v=cuttoff,xpd=F,col="red",lty=2)
    # points( rep(-.1,length(gene_module)),
    #         seq(length(gene_module),1,length.out = length(gene_module)),
    #         col=taxa_pal[factor(gene_module[h$order])],
    #         pch=15,cex=.5,xpd=F)

    par(mar=c(4,.2,1,2)) #(bottom, left, top, right) 
    image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
           col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
           breaks = breaks,
           axes=F)

    mtext( c("patient groups"), side=1, las = 1, cex = 0.7, line =1,xpd=T)

    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,0,0), y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39), y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653), y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653), y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

    ## Meta info bars
    points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
            col=pal[factor(names(samples))[ order(df[1])]],
            pch=15,xpd=T,cex=.7)

    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.008,
            col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
            pch=15,xpd=T,cex=.5) 
    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
            col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
            pch=15,xpd=T,cex=.5)

    mtext( paste0(" ", gsub("_.+","",rownames(top_PROT_mfi)[rev(h$order)])), side=4, las = 1, cex = .45,
           at=seq(0,1,length.out = length(rownames(top_PROT_mfi))), xpd=T)

    mtext( c(" BV"," HIV"), side=4, las = 1, cex = .45,
           at=c(par("usr")[3]-.008,par("usr")[3]-.022), xpd=T)
}

```

### top dpe with p-value < 0.05 un-scaled and scaled respectively
```{r single plots, fig.height=5, fig.width=3, include=FALSE}
# top differentially expressed proteins Luminal groups
t <- top_dpe.fun(dpe_L$top)

library(cowplot)
plot_heatmap.fun(t, "Luminal_gr")
p.plot <- recordPlot()
plot_heatmap.fun(t, "Luminal_gr", scale = T)
p.plot_s <- recordPlot()

# top differentially expressed proteins Tissue groups
t <- top_dpe.fun(dpe_T$top)

library(cowplot)
plot_heatmap.fun(t, "Tissue_gr")
p.plot_T <- recordPlot()
plot_heatmap.fun(t, "Tissue_gr", scale = T)
p.plot_s_T <- recordPlot()

```

```{r top dpe Luminal plot, echo=FALSE, fig.height=5, fig.width=6}
# side by side
# plot_grid(ggdraw(p.plot), ggdraw(p.plot_s), 
#           labels = c('a', 'b'),rel_widths=c(1,1), greedy = F)

plot_grid(ggdraw(p.plot_s),ggdraw(p.plot_s_T), 
          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F) # , 'c', 'd'

```

```{r top dpe Tissue plot, eval=FALSE, fig.height=10, fig.width=6, include=FALSE}
# side by side Luminal & Tissue
# plot_grid(ggdraw(p.plot), ggdraw(p.plot_s),
#           ggdraw(p.plot_T), ggdraw(p.plot_s_T), 
#           labels = c('a', 'b', 'c', 'd'),rel_widths=c(1,1), greedy = F)

plot_grid(ggdraw(p.plot_s),ggdraw(p.plot_s_T), 
          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F) # , 'c', 'd'

```

**Figure 3. Protein Heatmap**. **a** and **c** Without scaling (normalized MFI), for Luminal and Tissue respectively **b** and **d** scaled by row for Luminal and Tissue respectively

## all dpe un-scaled and scaled respectively
### Luminal
```{r all dpe plot, eval=FALSE, fig.height=6, fig.width=3, include=FALSE}
# all proteins 
plot_heatmap.fun(Protein[,sample_use], "Luminal_gr")
plot_heatmap.fun(Protein[,sample_use], "Luminal_gr", scale = T)
```

### Tissue
```{r all dpe Tissue plot, eval=FALSE, fig.height=6, fig.width=3, include=FALSE}
# all proteins 
plot_heatmap.fun(Protein[,sample_use], "Tissue_gr")
plot_heatmap.fun(Protein[,sample_use], "Tissue_gr", scale = T)
```

### 2021-10-28 reports/rmarkdown/manuscript/Figure8.Rmd
Figure8. with cytokines

@@
   library(openxlsx)
+  library(rstatix)
+  library(broom)
+  library(niceRplots)
 })  })  })
@@
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
+cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
 
@@
 # plot_heatmap.fun(t, "Luminal_gr", scale = T)
-plot_heatmap.fun <- function(top_PROT_mfi, gr, scale = FALSE){
+plot_heatmap.fun <- function(top_PROT_mfi, gr, scale = FALSE, layout = "none"){
   df <- na.omit(metadata[sample_use, c(gr, "ID")])
@@
 
-  layout(matrix(c(1,2,3,4,
-                  1,2,3,4),
-                  nrow = 1,ncol = 2,byrow = T),widths = c(1.,2,.4,6) )
-    ## Heatmap
-    par(mgp=c(0,0.2,0), mar=c(4,.4,1,0)) #(bottom, left, top, right) 
+  if(layout == "multi") {
+    layout(matrix(c(1,2,5,6,
+                    7,7,8,8,
+                    3,3,4,4),
+           nrow = 3, ncol = 4,byrow = T),
+           widths = c(.9,2,.9,2), heights = c(1,1,.01) ) }
+  if(layout == "single") {
+    layout(matrix(c(1,2),
+                    nrow = 1, ncol = 2,byrow = T),widths = c(1,2) ) }
+
+    ## Dendogram
+    par(mgp=c(0,0.2,0), mar=c(4,.6,1,.2)) #(bottom, left, top, right) 
     plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
-          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = 0.7 )
+          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .7 )
     #title(xlab="height", line=3)
-    mtext( c("Height"), side=1, las = 1, cex = 0.7, line =1,xpd=T)
+    mtext( c("Height"), side=1, las = 1, cex = .7, line =1,xpd=T)
 
     if(gr == "Tissue_gr") {
-      height <-1.5 }else{height <- round(max(h$height))}
-    lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob
+      height <-1.5 
+      add_letter("b")
+    }else{height <- round(max(h$height))
+      add_letter("a")}
+    #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob
 
-    # abline(v=cuttoff,xpd=F,col="red",lty=2)
-    # points( rep(-.1,length(gene_module)),
-    #         seq(length(gene_module),1,length.out = length(gene_module)),
-    #         col=taxa_pal[factor(gene_module[h$order])],
-    #         pch=15,cex=.5,xpd=F)
-
-    par(mar=c(4,.2,1,2)) #(bottom, left, top, right) 
+    ## Heatmap
+    par(mar=c(4,.2,1,2.5)) #(bottom, left, top, right) 
     image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
@@
 
-    mtext( c("patient groups"), side=1, las = 1, cex = 0.7, line =1,xpd=T)
+    mtext( c("patient groups"), side=1, las = 1, cex = .7, line =1,xpd=T)
 
     if(gr == "Luminal_gr") {
-      lines(x=c(0,.603,.603,0,0), y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
-      lines(x=c(.39,.94,.94,.39,.39), y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
+      lines(x=c(0,.603,.603,0,0),
+            y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
+      lines(x=c(.39,.94,.94,.39,.39),
+            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
     if(gr == "Tissue_gr") {
-      lines(x=c(.653,.95,.95,.653,.653), y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
-      lines(x=c(.653,.95,.95,.653,.653), y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
+      lines(x=c(.653,.95,.95,.653,.653),
+            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
+      lines(x=c(.653,.95,.95,.653,.653),
+            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
     }
@@
 
-### top dpe with p-value < 0.05 un-scaled and scaled respectively
-```{r single plots, fig.height=5, fig.width=3, include=FALSE}
+```{r sig. cytokines function, include=FALSE}
+sig_stars.fun <- function(df_krusk, df_dunn, group, n = 6, order = NULL){
+  samples_use <- metadata[,c(group, "ID")] %>% na.omit() %>% filter(ID %in% colnames(cytokine_data))
+
+  # select only sig. cytokines 
+  sig <- df_krusk %>% filter(p.adj<0.05) %>% pull(Cytokines)
+  sig_df <- cytokine_data_log[sig,samples_use$ID] %>%
+    cbind(Max = c(rep(max(.)+n, nrow(.))))
+
+  groupings <- samples_use[[group]] [ match( colnames(sig_df[,1:ncol(sig_df)-1]) , as.character(samples_use$ID)  ) ]
+  groupings <- c(groupings, "Max")
+
+   if(is.null(order)){
+    g <- rownames(sig_df)
+  }else{g <- order}
+
+  stars_df <- df_dunn %>%
+    mutate(x = (as.integer(str_extract(.$group1, "\\d")) + 
+                  as.integer(str_extract(.$group2, "\\d")))/2) %>%
+    mutate(line_x0 = as.integer(str_extract(.$group1, "\\d"))) %>%
+    mutate(line_x1 = as.integer(str_extract(.$group2, "\\d"))) %>%
+    mutate(y = (as.numeric(factor(Cytokines, levels = rev(g))))-.3) %>%
+    arrange(group1, group2) %>%
+    dplyr::group_by(Cytokines) %>%
+    nest() %>%
+    mutate(y_ = map(data, ~accumulate(.$y, ~.x + .1))) %>%
+    unnest(c(data, y_))
+
+  return(list(sig_df=sig_df, stars_df=stars_df, groupings=groupings, samples_use=samples_use))
+}
+```
+
+```{r Cytokine stats tests, include=FALSE}
+cytokine_data$ID <- metadata [ match( cytokine_data$PatID , as.character(metadata$PatID)  ) ,"ID" ]
+cytokine_data <- cytokine_data[ !is.na(cytokine_data$ID) , ]
+rownames(cytokine_data) <- cytokine_data$ID
+cytokine_data <- cytokine_data[,grepl("CVL",colnames(cytokine_data))]
+colnames(cytokine_data) <- sub("CVL_","",colnames(cytokine_data))
+cytokine_data <- t(cytokine_data)
+cytokine_data[is.na(cytokine_data)] <- 0.01
+cytokine_data <- log2(cytokine_data+1)
+cytokine_data_log <- cytokine_data
+
+cyto2 <- as.data.frame(t(cytokine_data))
+cyto2$ID<-row.names(cyto2)
+
+#select data and re-format to long format
+cytogr.long <- metadata %>%
+  select(ID, Luminal_gr, Tissue_gr) %>%
+  left_join(.,cyto2, by="ID") %>%
+  pivot_longer(c(-Luminal_gr, -Tissue_gr, -ID), names_to = "Cytokines", values_to = "value")
+
+test.krusk_dunn.fun <- function(df, var, group) {
+  var <- enquo(var)
+
+  stat.test.krusk <- df %>%
+  group_by(!!(var)) %>%
+  kruskal_test(., as.formula(paste("value", "~", group))) %>%
+  adjust_pvalue(method = "BH") %>%
+  select(-.y., -n, -df) %>%
+  arrange(p) %>%
+  rename(p.value = "p") %>%
+  add_significance()
+
+# select the significant proteins from kruskal
+sig <- stat.test.krusk %>% filter(p.adj<0.05) %>% pull(!!var)
+
+### Post hoc test
+stat.test.dunn <- df %>%
+  filter(!!var %in% sig) %>%
+  group_by(!!var) %>%
+  dunn_test(., as.formula(paste("value", "~", group))) %>%
+  #dunn_test(value ~ !!group) %>%
+  select(-.y.) %>%
+  arrange(p) %>%
+  rename(p.value = "p") %>%
+  filter(p.adj<0.05)
+
+  sig2 <- stat.test.dunn[stat.test.dunn$p.adj<0.05,]
+
+ l <- list(stat.test.krusk, stat.test.dunn)
+ n <-  c("Kruskal", "Dunn's") %>% paste0(., "_", group, sep = "")
+ l <- set_names(l, n)
+
+return(l)
+}
+
+stat <- imap(c("Luminal_gr", "Tissue_gr"), ~test.krusk_dunn.fun(cytogr.long, Cytokines, .x))
+s <- flatten(stat)
+write.xlsx(s, file=paste0("../../../results/","Cytokine_stats",".xlsx"))
+
+```
+
+```{r Proteins plot, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
+figlabels <- letters
+
 # top differentially expressed proteins Luminal groups
 t <- top_dpe.fun(dpe_L$top)
-
-library(cowplot)
-plot_heatmap.fun(t, "Luminal_gr")
-p.plot <- recordPlot()
-plot_heatmap.fun(t, "Luminal_gr", scale = T)
-p.plot_s <- recordPlot()
+plot_heatmap.fun(t, "Luminal_gr", scale = T, layout = "multi")
 
@@
 t <- top_dpe.fun(dpe_T$top)
+par(mfg = c(2,1)) # adds the remaning plots to the previous layout
+plot_heatmap.fun(t, "Tissue_gr", scale = T)
 
-library(cowplot)
-plot_heatmap.fun(t, "Tissue_gr")
-p.plot_T <- recordPlot()
-plot_heatmap.fun(t, "Tissue_gr", scale = T)
-p.plot_s_T <- recordPlot()
+# Luminal Cytokine plot 
+g <- c("IL1a", "MIG", "IP10", "IL1b", "IL12_p70", "MCP1")
+Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g)
+temp <- as.factor(c(Luminal$groupings, "Max"))
+stars <- Luminal$stars_df
+
+par( mar=c(1.4,4,1.5,1)) #b,l,t,r 
+violist(data=Luminal$sig_df,
+          srt=45,
+          genes= g,
+          col = pal, transparency = 50,
+          clustering=c(Luminal$groupings),smooth = 2 )
+segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
+text( stars$x , stars$y_+.05,
+        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
+title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = .9)
+
+add_letter("c")
+
+# Tissue Cytokine plot
+Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
+temp <- factor(Tissue$groupings, levels=c("T1","T2","T3","T4","Max"))
+stars <- Tissue$stars_df
+
+par( mar=c(24.5,5,1.5,1)) #b,l,t,r
+violist(data=Tissue$sig_df,
+          srt=45, 
+          genes= rownames(Tissue$sig_df),
+          col = pal, transparency = 50,
+          clustering=Tissue$groupings, smooth = 2 )
+segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
+text( stars$x , stars$y_+.05,
+        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
+title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = .9)
+
+add_letter("d")
 
@@
 
-```{r top dpe Luminal plot, echo=FALSE, fig.height=5, fig.width=6}
-# side by side
-# plot_grid(ggdraw(p.plot), ggdraw(p.plot_s), 
-#           labels = c('a', 'b'),rel_widths=c(1,1), greedy = F)
+**Figure 3.
+Protein Heatmap**.
+**a** and **c** Without scaling (normalized MFI), for Luminal and Tissue respectively **b** and **d** scaled by row for Luminal and Tissue respectively.
+top dpe with p-value < 0.05
 
-plot_grid(ggdraw(p.plot_s),ggdraw(p.plot_s_T), 
+```{r All dpe plot, eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
+library(cowplot)
+
+plot_heatmap.fun(Protein[,sample_use], "Luminal_gr")
+p.plot <- recordPlot()
+
+plot_heatmap.fun(Protein[,sample_use], "Tissue_gr")
+p.plot_T <- recordPlot()
+
+# side by side unscaled
+plot_grid(ggdraw(p.plot),ggdraw(p.plot_T), 
           labels = c('a', 'b'),rel_widths=c(1,1), greedy = F) # , 'c', 'd'
@@
 ```
-
-```{r top dpe Tissue plot, eval=FALSE, fig.height=10, fig.width=6, include=FALSE}
-# side by side Luminal & Tissue
-# plot_grid(ggdraw(p.plot), ggdraw(p.plot_s),
-#           ggdraw(p.plot_T), ggdraw(p.plot_s_T), 
-#           labels = c('a', 'b', 'c', 'd'),rel_widths=c(1,1), greedy = F)
-
-plot_grid(ggdraw(p.plot_s),ggdraw(p.plot_s_T), 
-          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F) # , 'c', 'd'
-
-```
-
-**Figure 3. Protein Heatmap**. **a** and **c** Without scaling (normalized MFI), for Luminal and Tissue respectively **b** and **d** scaled by row for Luminal and Tissue respectively
-
-## all dpe un-scaled and scaled respectively
-### Luminal
-```{r all dpe plot, eval=FALSE, fig.height=6, fig.width=3, include=FALSE}
-# all proteins 
-plot_heatmap.fun(Protein[,sample_use], "Luminal_gr")
-plot_heatmap.fun(Protein[,sample_use], "Luminal_gr", scale = T)
-```
-
-### Tissue
-```{r all dpe Tissue plot, eval=FALSE, fig.height=6, fig.width=3, include=FALSE}
-# all proteins 
-plot_heatmap.fun(Protein[,sample_use], "Tissue_gr")
-plot_heatmap.fun(Protein[,sample_use], "Tissue_gr", scale = T)
-```

### 2021-10-29 reports/rmarkdown/manuscript/Figure1.Rmd
Fig.1 New Layout

@@
 ---
-title: "Figure 1. Microbiome abundances"
+title: "Main Figure 1.Luminal Patient Grouping"
 output: 
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data., message=FALSE, warning=FALSE, include=FALSE}
 ##################
@@
   library(scales)
+  library(igraph)
   library(RColorBrewer)
@@
 group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
-#group_annotation <- as.character(metadata$Luminal_gr) %>% set_names()
+
+SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
+bac_communities <- read.csv2("../../../results/bacterial_communities.csv",row.names = 1)
 
@@
 
-#color palettes
 #################
@@
 pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
-#pal <- c("#377EB8", "#4DAF4A", "#FF7F00", "#F781BF", "#984EA3", "#FFFF33", "#A65628",  "#999999")
-#pal <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
 taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
 #show_col(pal)
-#show_col(taxa_pal)
 ```
 
-```{r Figure 1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=6 ,fig.asp=1, fig.pos="t"}
-### A + B
-#############################
-# DIVERSITY + TAX ABUNDANCE #
-#############################
-
+```{r Figure 1., echo=FALSE, fig.asp=1, fig.height=7, fig.pos="t", fig.width=6, message=FALSE, warning=FALSE}
 figlabels <- letters
@@
                 3,3,3,3,3,3,
-                4,4,4,4,4,4,
-                5,5,8,9,6,6,
-                5,5,8,7,6,6,
-                5,5,8,7,7,7) , ncol= 6,byrow = T),
-                               widths = c(1.2,1,1.85,1.35,1,1.4),
-                               heights = c(.7,1.5,.7,1.5,1,.8,.2))
-# layout(matrix(c(1,1,1,1,1,5,
-#                 2,2,2,2,2,5,
-#                 3,3,3,3,3,6,
-#                 4,4,4,4,4,6,
-#                 7,7,8,8,8,6) , ncol= 6,byrow = T),
-#                                widths = c(.5,1,1,1,1,1.1),
-#                                heights = c(.5,1,.5,1,1.5))
+                5,6,7,7,8,8,
+                5,4,7,7,8,8,
+                5,4,7,7,8,8) , ncol= 6,byrow = T),
+                               widths = c(1.9,1.4,.7,.8,.8,.7,.8),
+                               heights = c(1.6,.65,1.5,.95,.9,.2))
+### A
+#################################
+# COHORT SCHEMATIC ILLUSTRATION #
+#################################
+par(mar=c(0,3.3,0,1))
+fig1a <- magick::image_read("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/supplementary_files/Scematic figure.pdf", density = 900 )
+plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
+rasterImage(fig1a, 0, 0, 1, 1)
 
-#i <- ASV[2]
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+### B
+#############################
+# DIVERSITY + TAX ABUNDANCE #
+#############################
+
 ASV <- list(c("Luminal_gr","ASV_CVL_V3_normalized_batch_corrected"),
@@
   gr <- factor(setNames(metadata[[i[1]]],metadata$ID))
-  #x <- x[,sample_use]
-  #x <- datasets_all_samples[[2]]
 
-  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
   temp <- t(t(2^x-1)/colSums(2^x-1))
@@
 
+  if(i[2] == "ASV_tissue_V3_normalized_batch_corrected"){break}
+
+  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
   barlist( data = divers[,order(group_annotation)],
@@
     map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
-    #print(par("usr"))
   }
@@
   par(mar=c(1,5,1,1.5)) #b,l,t,r
-  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
-  # taxa[is.na(taxa)] <- "NA"
-  # taxa <- taxa[rownames(temp),]
-  # temp <- rowsum(temp , taxa$Class)
   temp[is.nan(temp)] <- 0
@@
 
-# save diversity index stats
-div_stat <- div_list %>% 
-  set_names(map(ASV,1)) %>%
-  map(., ~summarise_all(.x, list(median=median, IQR=IQR), na.rm = TRUE)) %>%
-  map(., ~pivot_longer(.x, everything(), names_sep = "_", names_to = c(".value", "Stats"), )) %>%
-  bind_rows(.id = "Sample type") %>%
-  mutate("Sample type" = gsub("_gr", "", .$"Sample type"))
+###########
+# LEGENDS #
+###########
+par(mar=c(1,.3,0,.5)) #b,l,t,r
+empty_plot()
+legend(x = -0.04, y = 1.04, 
+       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other") ,
+       xjust = 0,yjust = 1,title.adj = 0,  ncol =1, #xjust=1,
+       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
+       pt.bg = pal)
 
-write.xlsx(div_stat, file=paste0("../../../results/","Suppl.Tbl.1 Diversity statistics",".xlsx"))
+par(mar=c(0,3.3,1.5,.5)) #b,l,t,r
+empty_plot()
+legend(x = par("usr")[c(1)],
+         y = par("usr")[c(4)],#title.adj = 0,
+         legend = lvl[1:12],xjust = 0,yjust = 1,
+         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+          #title="",
+         pt.bg = taxa_pal)
 
-### C
-############################
-# TOTAL RELATIVE ABUNDANCE #
-############################
-TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-sample_use <- colnames(TRX)[colSums(TRX)!=0]
+par(mar=c(0,.3,1.5,1)) #b,l,t,r
+empty_plot()
+legend(x = par("usr")[c(1)],
+         y = par("usr")[c(4)],#title.adj = 0,
+         legend = lvl[13:17],xjust = 0,yjust = 1,
+         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
+          #title="",
+         pt.bg = taxa_pal[13:17])
+
+### C + D
+####################
+# PATIENT GROUPING #
+####################
+
+### Louvain ###
+g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
+
+par(mar=c(0,0,1,1))
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
+      vertex.size=10, layout=l,
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+title(main = "Louvain", line = -2, cex.main = 1)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+### UMAP ###
+all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
+cors <- cor(all_microbiome[,], method = "pearson")
+adj <- (1-cors)/2
+U <- uwot::umap(adj,n_neighbors = 30)
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,#main="UMAP",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+title(main = "UMAP", line = -2, cex.main = 1)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+```
+
+```{r Suppl. Table 1., include=FALSE}
+# Total Relative Abundance 
 raw <- c("ASV_CVL_V3", "ASV_tissue_V3") %>% set_names()
-
 raw_list <- raw %>%
@@
     dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
-    `row.names<-`(., NULL) %>% 
-    column_to_rownames(var = "name") %>%
-    as.matrix()
+    rename(Taxa="name")
 
-#colSums(r)
+# Alfa Diversity Stats
+div_stat <- div_list %>%
+  set_names(map(ASV,1)) %>%
+  map(., ~summarise_all(.x, list(median=median, IQR=IQR), na.rm = TRUE)) %>%
+  map(., ~pivot_longer(.x, everything(), names_sep = "_", names_to = c(".value", "Stats"), )) %>%
+  bind_rows(.id = "Sample type") %>%
+  mutate("Sample type" = gsub("_gr", "", .$"Sample type"))
 
-par(mar=c(1,5,2,0)) #b,l,t,r
-barplot( r, border = NA ,ylab="abundance", xlab ="",
-         col = taxa_pal[factor(rownames(r),levels = rownames(temp))])
-title(main = "Total relative abundance", line = 0.5, cex.main = .9, font.main= 1)
-mtext("Luminal        Tissue ", line = 0, cex=.6, side=1)
-#add label
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-# legend(x = par("usr")[c(2)],
-#          y = par("usr")[c(4)],title.adj = 0,
-#          legend = rownames(r),xjust = 0,yjust = 1,
-#          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-#           title="top abundant bacteria",
-#          pt.bg = taxa_pal)
+# Bacterial Communities
+bac_communities <- bac_communities %>% 
+  rownames_to_column(var = "Taxa") %>%
+  mutate("bacterial communities"= factor(paste0("BC",sprintf("%02d",bac_communities$x)))) %>%
+  select(-x)
 
-### D
-#############################
-# LUMINAL-TISSUE COMPARISON #
-#############################
-par(mar=c(0,4,3,2)) #b,l,t,r
-x <- as.character(metadata[,"Luminal_gr"])
-x[is.na(x)] <- "NA"
-y <- as.character(metadata[,"Tissue_gr"])
-y[is.na(y)] <- "NA"
-x1 <- x [ (x!="NA") & (y!="NA") ]
-y1 <- y [ (x!="NA") & (y!="NA") ]
-df <- data.frame(x1,y1)
-plot_sankey( df,color_by = 1, 
-             plot_labels = T, plot_weights = F,order_1_by = "NULL",
-             order_2_by = "disentangled",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal= pal)
-text(c(0,1),c(1,1),labels = c( "Luminal" , "Tissue" ),pos=3, xpd=T)
-
-#add label
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
-###########
-# LEGENDS #
-###########
-par(mar=c(0,.5,0,0)) #b,l,t,r
-empty_plot()
-legend(x = -0.04, #par("usr")[c(1)],
-       y = 1.04, #par("usr")[c(4)],title.adj = 0,
-       #legend = sort(unique(group_annotation)),
-       legend = c("L. crispatus/jensenii", "L. iners", "Gardnerella", "High diverse", "Other") ,
-       xjust = 0,yjust = 1,title.adj = 0,  ncol =1, #xjust=1,
-       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
-       pt.bg = pal)
-
-par(mar=c(0,.4,1,.5)) #b,l,t,r
-empty_plot()
-legend(x = par("usr")[c(1)],
-         y = par("usr")[c(4)],#title.adj = 0,
-         legend = lvl[1:12],xjust = 0,yjust = 1,
-         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-          #title="",
-         pt.bg = taxa_pal)
-
-par(mar=c(0,.4,1,1)) #b,l,t,r
-empty_plot()
-legend(x = par("usr")[c(1)],
-         y = par("usr")[c(4)],#title.adj = 0,
-         legend = lvl[13:17],xjust = 0,yjust = 1,
-         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-          #title="",
-         pt.bg = taxa_pal[13:17])
+pages <- list("Total Relative Abundance"=r ,"Alfa Diversity"=div_stat, "Bacterial Communities"=bac_communities)
+write.xlsx(pages, file=paste0("./Suppl.Tbl/","Suppl.Tbl.01",".xlsx"))
 
 ```
-
-```{r include=FALSE}
-dev.copy2pdf(file=paste0("./images/","Figure1",".pdf"), 
-    width = 6.5,
-    height = 7 , paper = "a4" #units = "cm", res = 300 # a4 = 8.3 x 11.7 inches
-)
-```
-
-```{r eval=FALSE, include=FALSE}
-##########################
-x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-y <- datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
-z <- datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
-
- # par(mar=c(1,5,1,1))
-tempx <- t(t(2^x-1)/colSums(2^x-1))
-tempy <- t(t(2^y-1)/colSums(2^y-1))
-tempz <- t(t(2^z-1)/colSums(2^z-1))
-
-tempx[is.nan(tempx)] <- 0
-tempy[is.nan(tempy)] <- 0
-tempz[is.nan(tempz)] <- 0
- # if(is.null(temp_legend)){
- #   temp_legend <- temp   #Use the same legend from the 1st dataset
-#  }
-
-temp1 <- rowSums(tempx)/100
-temp2 <- rowSums(tempy)/108
-temp3 <- rowSums(tempz)/93
-
-Totrelab <- cbind(temp1, temp2, temp3)
-Totrelab <- cbind(temp1, temp2)
-Totrelab["Lactobacillus_crispatus/acidophilus", ] <- Totrelab["Lactobacillus_crispatus/acidophilus", ] + Totrelab["Lactobacillus_crispatus", ]
-Totrelab[rownames(Totrelab) != "Lactobacillus_crispatus", ]
-  o <- order( rowSums(Totrelab) , decreasing = T)
-  #sample_o <- names(group_annotation)[ order( group_annotation )]
-  temp_a <- Totrelab[o,]*100
-  write.csv(temp_a, "totalRelAbundance.csv")
-
-  # temp <- temp[-grep("NA",rownames(temp)),]
-  temp_b <- rbind(temp_a[1:19,], Other = colSums(temp_a[20:nrow(temp_a),]) )
-
-layout(matrix(c(2) , ncol= 2,byrow = T),heights = c(1))
-#par(mar=c(2,7))
-layout(matrix(c(2) , ncol= 2,byrow = T),
-       heights = c(1))
-  par(mar=c(2,7))
-  barplot( temp_a, las=2 , border = NA ,ylab="abundance",main=sub("_nor.*","",i),font.main=1,
-           col = taxa_pal[factor(rownames(temp_a),levels = rownames(temp_a))] ,names.arg = rep("",ncol(temp_a)))
-  legend(x = par("usr")[c(1)],
-         y = par("usr")[c(4)],title.adj = 0,
-         legend = rownames(temp_a),xjust = 0,yjust = 1,
-         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-          title="top abundant bacteria",
-         pt.bg = taxa_pal)
-```
-
-```{r fig1, eval=FALSE, fig.cap="Figure 1.
-Box plot of the total lengths of live and dead sparrows following a snowstorm in Providence", fig.height=5, fig.width=12, message=FALSE, warning=FALSE, include=FALSE}
-
-#############
-# LODA DATA #
-#############
-#ALL_DATA <- readRDS("~/Box/Projects/J_Mjosberg/NBIS_housekeeping_normalization/analysis_all_tissues/all_ilc_dataset.rds")
-#pal <- c(scales::hue_pal()(8),RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2") )
-
-########################
-# DEFINE FIGURE LAYOUT #
-########################
-layout(matrix(c(1,2,5,
-                1,2,6,
-                1,2,7,
-                1,2,8,
-                1,2,9,
-                1,3,10,
-                1,3,11,
-                1,3,12,
-                1,3,13,
-                1,3,14,
-                1,4,15,
-                1,4,16,
-                1,4,16,
-                17,4,16,
-                17,4,16),
-              nrow = 15, byrow = T), 
-   widths=c(3,1.5,1.5), heights=c(1,1))
-#layout.show(16)
-
-################
-# PLOT FIGURES #
-################
-figlabels <- letters
-
-### A
-################
-par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
-plot_meta(ALL_DATA,red = "umap",feat = "RNA_snn_res.3", label = T, frame=F, main="",cex = .7)
-fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
-legend("topright", legend = "", col = "white", pch = 16, bty="n",
-       pt.cex = 1,cex = 1,title = "Clustering")
-
-### B
-################
-par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
-plot_meta(ALL_DATA,red = "umap",feat = "Tissue", label = T, frame=F, main="")
-fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
-legend("topright", legend = levels(ALL_DATA$Tissue), col = pal[1:3], pch = 16, bty="n",
-       pt.cex = 1,cex = 1,title = "Tissue")
-
-### C
-################
-par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
-plot_meta(ALL_DATA,red = "umap",feat = "Celltype", label = T, frame=F, main="")
-fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
-legend("topright", legend = levels(ALL_DATA$Celltype), col = pal[1:4], pch = 16, bty="n",
-       pt.cex = 1,cex = 1,title = "Celltype")
-
-### D
-################
-par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
-plot_meta(ALL_DATA,red = "umap",feat = "Donor", label = T, frame=F, main="")
-fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
-legend("topright", legend = "", col = "white", pch = 16, bty="n",
-       pt.cex = 1,cex = 1,title = "Donor")
-
-### E
-################
-my_gene_list <- c("PTGDR2","TNFSF10","TMEM273","KRT1","HPGD","FAM110A","ZP1","PKIB","TESPA1","CSGALNACT1")
-par(mar=c(0.5,6,0,.5))
-for(i in my_gene_list){ 
-try({violins(data = ALL_DATA,gene = i,clustering = "Celltype",assay = "RNA",plot_y_axis = F,plot_x_axis = F,ylab="",main="",bw = 1)
-mtext(side = 2, text = i, line = 1,las=1,cex = .8)})
-}
-mtext(side = 1, at=1:5, text = sort(unique(ALL_DATA@meta.data[,"Celltype"])), line = 1, las=1)
-text(1:5, par("usr")[3], labels = sort(unique(ALL_DATA@meta.data[,"Celltype"])), srt = 45, adj = c(1,0.5), xpd = T, cex=1)
-
-### F
-################
-my_image <- image_read("images/sparrow.jpg")
-par(mar = c(1, 1, .5, .5), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
-plot(my_image)
-fig_label(figlabels[1],cex = 3,font = 2,region = "figure") ; figlabels <- figlabels[-1]
-
-### G
-################
-
-```

### 2021-10-29 reports/rmarkdown/manuscript/Figure2.Rmd
Fig.2 version 1

---
title: "Main Figure 2. Bacterial Communities and Function"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
sample_use <- metadata$ID

SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]

group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
gr <- "Luminal"

picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
picrust_KO_seq <- read.table("../../../data/KO_predicted.tsv",header = T,row.names = 1)
picrust_taxa <- read.table("../../../data/SeqID_to_taxa.csv",header = T,row.names = 1,sep = ",")

AA <- c("(?<!Phenyl)Alanine " = "Ala", "Arginine" = "Arg", "Asparagine" = "Asn", "Aspartic acid" = "Asp", "Cysteine" = "Cys", "Glutamic acid" = "Glu", "Glutamine" = "Gln", "Glycine" = "Gly", "Histidine" = "His", "Hydroxyproline" = "Hyp", "Isoleucine" = "Ile", "Leucine" = "Leu", "Lysine" = "Lys", "Methionine" = "Met", "Phenylalanine" = "Phe", "Proline" = "Pro", "Pyroglutamatic" = "Glp", "Serine" = "Ser", "Threonine" = "Thr", "Tryptophan" = "Trp", "Tyrosine" = "Tyr", "Valine" = "Val")
AA_ <- set_names(AA, map_chr(names(AA), ~ paste0("(?i)", .))) # make case insensitive

#################
# COLOR PALETTS #
#################
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
# fig.asp=.6
# Define Layout and set labels
# layout(matrix(c(2,3,4,5,      6,7,  11,12,13,14,
#                 1,1,1,1,      1,8,  11,12,13,14,
#                 1,1,1,1,      1,9,  11,12,13,14,
#                 1,1,1,1,      1,10, 11,12,13,14,
#                 15,15,15,15,  0,0,  11,12,13,14
#                 ) , 
#               ncol= 10,byrow = T),widths = c(1,1,1,1,1,1,1,2,.5,6),heights = c(1.5,1.2,1.2,1.2,1.2))
# layout.show(3)

layout(matrix(c(1,2,3,4,5,
                6,2,3,4,5,
                7,7,7,10,10,
                8,8,8,10,10,
                9,9,9,10,10),
              nrow = 5,ncol = 5,byrow = T), widths = c(1.9,.6,1,.2,2),
       heights = c(1.5,1.5,.7,.7,.7))
figlabels <- letters

#########################
# BACTERIAL COMMUNITIES #
#########################
bac_communities <- factor(paste0("BC",sprintf("%02d",bac_communities)))
gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
 set.seed(1)
lB <- layout_nicely(gB,niter=3000,start.temp=30)

par(mar=c(0,0,3.5,5))
plot( gB , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(bac_communities)] ,
      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
      vertex.size=10, #main="Bacterial Communities",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
title(main = "           Bacterial Communities- Louvain", line = 0.5, cex.main = 1)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = bact_pal,pt.cex = 1,xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

# par(mar=c(1,1,1,1))
# for(j in sort(unique(bac_communities)) ){
#   empty_plot()
#   text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
#   title(main=j,cex.main=.5)
# }

#####################
# BACTERIAL PICRUST #
#####################
picrust_data <- round(picrust_data[ rowSums(picrust_data>5)>=10, ])
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

ko_names <- readLines('../../../data/ko_info.tsv')
ko_names <- strsplit(ko_names,"\t")
ko_names <- setNames(unlist(lapply(ko_names,function(x){x[2]})), unlist(lapply(ko_names,function(x){x[1]})))
ko_names <- sub(" [[].*","",ko_names)
ko_names <- sub(".*; ","",ko_names)

named_picrust_data <- rowsum( picrust_data , group = ko_names[ rownames(picrust_data) ] )
#named_picrust_data[1:20,1:3]
#dim(named_picrust_data)
named_picrust_data <- round(named_picrust_data[ rowSums(named_picrust_data>5)>=5, ])

design <- model.matrix(~ group_annotation+Contraception+HIVstatus, 
                       data=metadata[sample_use,])

y <- DGEList(counts=named_picrust_data[,sample_use])
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y)
y <- estimateGLMTagwiseDisp(y)
fit <- glmFit(y, design)

lrt <- glmLRT(fit,coef=2:5)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(LogFC.intercept=0,top) # with intercept

#saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
#write.csv2( top , "../../../results/Picrust_EdgeR_top_dge.csv" )
write.xlsx(top, file=paste0("../../../results/DE/","dge_picrust_",gr,"_conf.xlsx"))

top_dge <- (top$FDR < 1e-5) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

top_picrust_data <- edgeR::cpm(y)[top_dge, ]
n <- colnames(top_picrust_data)
top_picrust_data <- t(apply(top_picrust_data,1,function(x){scale(x,T,T)}))
top_picrust_data[top_picrust_data > 5] <- 5
colnames(top_picrust_data) <- n

par(mar=c(4,.1,.5,.2)) #b,l,t,r

h <- hclust( as.dist( (1- cor(t(top_picrust_data)))/2 ), method = "ward.D2")
plot( rev(as.dendrogram(h)) ,xlim=c(max(h$height),-1), horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
cutoff <- 4
abline(v=cutoff,xpd=F,col="red",lty=2)
gene_module <- cutree(h, h = cutoff)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

points( rep(-.7,length(gene_module)),
        seq(length(gene_module),1,length.out = length(gene_module)),
        col=taxa_pal[factor(gene_module[h$order])],
        pch=15,cex=1,xpd=F)

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),
        rep(par("usr")[4],length(sample_use))+.005, pch=15, cex=.7,xpd=T,
        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]])

points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.008, pch=15, xpd=T,cex=.5,
        col= c("tomato","orange","#d7d7d7")[as.factor(metadata$BV_Diagnosis_v3)][order( group_annotation[sample_use])] )

points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.024, pch=15, xpd=T,cex=.5,
        col= c("#d7d7d7","tomato")[as.factor(metadata$HIVstatus)][order(group_annotation[sample_use])])

source("../enrichment_function.R")
gmt_list <- fgsea::gmtPathways("../../../data/KEGG_pathways_to_KO.tsv")
kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
names(gmt_list) <- kegg_names [ match ( names(gmt_list), as.character(kegg_names[,1]) ) , 2 ]
gmt_list <- lapply(gmt_list , function(x){
  na.omit(ko_names[x])
})
res_list <- lapply( unique( gene_module[h$order] ),function(x){
  temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  return(temp)
} )
res <- map(res_list, ~rownames_to_column(.x, var = "Terms")) %>% 
    set_names(., paste0("module_",  seq_along(res_list)))
write.xlsx(res, file=paste0("../../../results/ENRICHMENT/Heatmap/","KEGG","_enrichment_picrust_",gr,".xlsx"))

names(res_list) <- unique( gene_module[h$order] )

pvalues <- unlist(lapply(res_list,function(x){ -log10(as.numeric(x$pvalue)) [1:3] }))
pvalues[is.na(pvalues)] <- 0
terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:3] }))
terms[is.na(terms)] <- ""
terms <- stringr::str_replace_all(string = terms,pattern= AA_)
genes <- unlist(lapply(res_list,function(x){ x$genes [1:3] }))
genes[is.na(genes)] <- ""

#pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
pvalues <- setNames(pvalues,paste0(terms,""))
module_color <- unlist(lapply(names(res_list),function(x){ rep(x,3) }))

temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
df <- data.frame(temp,temp)
quietly(plot_sankey)( df, pal = taxa_pal[as.numeric(levels(temp))], use_w2 = F, plot_labels = T,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)$output

# par("usr")(xmin, ymin, xmax, ymax)
text(par("usr")[1],par("usr")[3]-.008,labels = "     BV",cex=.6, xpd=T)
text(par("usr")[1],par("usr")[3]-.024,labels = "      HIV",cex=.6, xpd=T)

par(mar=c(4,.1,.5,7))
barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
         xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,names.arg = "")
title(xlab="-log10(pvalue)", line=2)
abline(v=2,xpd=F,col="red",lty=2)
points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,bg=taxa_pal[factor(rev(module_color))] )
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,xpd=T,
      labels = rev(gsub("\\(\\)","",names(pvalues)) ),xpd=T)
text( rev(pvalues)+strwidth("M") , seq(1,length(pvalues) )*1.2-.5, adj=0,cex=1,
      labels = rev(names(pvalues)),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )

################
# PICRUST UMAP #
################

# based on 108 samples used in manuscript:
par(mar=c(2.5,2.5,3,4)) #b,l,t,r
set.seed(1)
UMAP_picrust <- uwot::umap(t(named_picrust_data[,sample_use]),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[factor(metadata$Luminal_gr)],
     #main="Top Functional Profile- UMAP",
     pch=21,frame=F,axes=F)
title(main = "Top Functional Profile- UMAP", line = 0.5, cex.main = 1)
#mtext("UMAP1", side=1, line=0.7, cex=0.8)
#mtext("UMAP2", side=2, line=0.7, cex=0.8)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(metadata$Luminal_gr)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T, cex=1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,1,2,1)) #b,l,t,r
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0

barlist( data = rbind(WS=mL ,
                      counts_16S=xL),
         main = "Lactobacillus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

barlist( data = rbind(WS=mM ,
                      counts_16S=xM),
         main = "Mobiluncus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(WS=mG ,
                      counts_16S=xG),
         main = "Gardnerella",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

empty_plot()
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

layout(matrix(c(1,
                2,
                3),
              nrow = 3,ncol = 1,byrow = T), widths = c(1,1,1))

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(3,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
xM <- rowsum(x , grepl("Monbilicus",rownames(x)) )[2,] / sum(grepl("Monbilicus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Mobiluncus",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardnerella",
         genes = c("WS","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal,srt=20)

```

### 2021-10-29 reports/rmarkdown/manuscript/Figure2.Rmd
Fig.2 Version 2

@@
   library(RColorBrewer)
-  #remotes::install_github("czarnewski/niceRplots",force=T)
+  remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
@@
 #################
-Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
-pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
-         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
 pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
@@
 bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')
-
 ```
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
-# fig.asp=.6
-# Define Layout and set labels
-# layout(matrix(c(2,3,4,5,      6,7,  11,12,13,14,
-#                 1,1,1,1,      1,8,  11,12,13,14,
-#                 1,1,1,1,      1,9,  11,12,13,14,
-#                 1,1,1,1,      1,10, 11,12,13,14,
-#                 15,15,15,15,  0,0,  11,12,13,14
-#                 ) , 
-#               ncol= 10,byrow = T),widths = c(1,1,1,1,1,1,1,2,.5,6),heights = c(1.5,1.2,1.2,1.2,1.2))
-# layout.show(3)
 
-layout(matrix(c(1,2,3,4,5,
-                6,2,3,4,5,
-                7,7,7,10,10,
-                8,8,8,10,10,
-                9,9,9,10,10),
-              nrow = 5,ncol = 5,byrow = T), widths = c(1.9,.6,1,.2,2),
-       heights = c(1.5,1.5,.7,.7,.7))
+layout(matrix(c(1,1,2,2,2,
+                1,1,3,3,3,
+                1,1,4,4,4,
+                5,6,7,8,9,
+                5,6,7,8,10),
+              nrow = 5,ncol = 5,byrow = T), widths = c(.6,1.5,.2,1.9,2),
+       heights = c(.6,.6,.6,1,1.5))
 figlabels <- letters
 
+## A
 #########################
@@
       edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
-      vertex.size=10, #main="Bacterial Communities",
+      vertex.size=10,
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
@@
 
-# par(mar=c(1,1,1,1))
-# for(j in sort(unique(bac_communities)) ){
-#   empty_plot()
-#   text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
-#   title(main=j,cex.main=.5)
-# }
+## B
+###################################
+# BACTERIAL VAGINOSIS COMPARISSON #
+###################################
+par(mar=c(1,5,2,1)) #b,l,t,r
+x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
+xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
+xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
 
+mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
+mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
+mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
+
+barlist( data = rbind(WS=mL ,
+                      "16S counts"=xL),
+         main = "Lactobacillus",
+         genes = c("WS","16S counts"),
+         clustering = group_annotation,
+         draw_mean_lines=F,col = pal,srt=20)
+
+#add label
+add_letter(figlabels[1]); figlabels <- figlabels[-1]
+
+barlist( data = rbind(WS=mM ,
+                      "16S counts"=xM),
+         main = "Mobiluncus",
+         genes = c("WS","16S counts"),
+         clustering = group_annotation,
+         draw_mean_lines=F,col = pal,srt=20)
+
+barlist( data = rbind(WS=mG ,
+                      "16S counts"=xG),
+         main = "Gardnerella",
+         genes = c("WS","16S counts"),
+         clustering = group_annotation,
+         draw_mean_lines=F,col = pal,srt=20)
+
+## C
 #####################
@@
 top <- cbind(LogFC.intercept=0,top) # with intercept
-
-#saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
-#write.csv2( top , "../../../results/Picrust_EdgeR_top_dge.csv" )
-write.xlsx(top, file=paste0("../../../results/DE/","dge_picrust_",gr,"_conf.xlsx"))
 
@@
     set_names(., paste0("module_",  seq_along(res_list)))
-write.xlsx(res, file=paste0("../../../results/ENRICHMENT/Heatmap/","KEGG","_enrichment_picrust_",gr,".xlsx"))
 
@@
       labels = rev(names(pvalues)),col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
-# text( rep(max(pvalues)+.5,length(pvalues)) , seq(length(pvalues),1 )*1.2-.5,
-#       adj=0,cex=.8,labels = names(pvalues), col = taxa_pal[factor(module_color)] )
 
+## D
 ################
@@
 plot(UMAP_picrust,bg=pal[factor(metadata$Luminal_gr)],
-     #main="Top Functional Profile- UMAP",
      pch=21,frame=F,axes=F)
@@
 
-###################################
-# BACTERIAL VAGINOSIS COMPARISSON #
-###################################
-par(mar=c(1,1,2,1)) #b,l,t,r
-x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
-xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
-xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
-
-mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
-mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
-mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
-
-barlist( data = rbind(WS=mL ,
-                      counts_16S=xL),
-         main = "Lactobacillus",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
-#add label
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
-barlist( data = rbind(WS=mM ,
-                      counts_16S=xM),
-         main = "Mobiluncus",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
-barlist( data = rbind(WS=mG ,
-                      counts_16S=xG),
-         main = "Gardnerella",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
 empty_plot()
@@
 
-```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r Suppl. Table 2., include=FALSE}
+#saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )
+top <- top %>% as_tibble(rownames="KO Terms")
 
-layout(matrix(c(1,
-                2,
-                3),
-              nrow = 3,ncol = 1,byrow = T), widths = c(1,1,1))
-
-###################################
-# BACTERIAL VAGINOSIS COMPARISSON #
-###################################
-par(mar=c(3,5,2,1))
-x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
-xM <- rowsum(x , grepl("Monbilicus",rownames(x)) )[2,] / sum(grepl("Monbilicus",rownames(x)))
-xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
-
-mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
-mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
-mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
-
-barlist( data = rbind(  BV=mL ,
-                        counts_16S=xL),
-         main = "Lactobacillus",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
-#add label
-add_letter(figlabels[1]); figlabels <- figlabels[-1]
-
-barlist( data = rbind(  BV=mM ,
-                        counts_16S=xM),
-         main = "Mobiluncus",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
-barlist( data = rbind(  BV=mG ,
-                        counts_16S=xG),
-         main = "Gardnerella",
-         genes = c("WS","counts_16S"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
-
+pages <- list(list("Across"=top), res) %>% flatten()
+write.xlsx(pages, file=paste0("./Suppl.Tbl/","Suppl.Tbl.02",".xlsx"))
 ```

### 2021-11-01 reports/rmarkdown/manuscript/Figure4.Rmd
Fig.4 Version 1

---
title: "DGE's and Enrichment (GO and KEGG)"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
sample_use <- metadata$ID

#color palettes
#################
# COLOR PALETTS #
#################
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

## Heatmap DGEs and enrichment
```{r Heatmap functions, echo=FALSE, warning=FALSE, message=FALSE}
##################
# DEG's FUNCTION #
##################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
edgR_dge.fun <- function(group, confound) {

  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~groups, data=metadata[sample_use,]) # without confounders
  }else{
    c <- "_conf"
    design <- model.matrix(~groups+HIVstatus+Contraception,
                         data=metadata[sample_use,])
  }

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
  y <- calcNormFactors(y,method = "TMM")
  y <- estimateGLMCommonDisp(y,design)
  y <- estimateGLMTagwiseDisp(y,design)
  fit <- glmFit(y, design)

  lrt <- glmLRT(fit,coef=2:n_gr) # with intercept
  top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
  colnames(top) <- sub(group,"",colnames(top))
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Genes")

  write.xlsx(top, file=paste0("../../../results/DE/","dge_heatmap_",group,c,".xlsx"))

return(list(design=design, y=y, fit=fit, lrt=lrt, top=top, c=c))
}

#######################
# ENRICHMENT FUNCTION #
#######################
# group_name <- "Luminal_gr"
# top <- dge_L$top
# y <- dge_L$y
# c <- dge_L$c
gene_mod_enrich.fun <- function(top, y, c, db, group_name, cuttoff){
  p_val <- pull(top, "PValue")
  top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dge <- top$Genes[ top_dge ]

  top_TRX_counts <- edgeR::cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
  top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
  top_TRX_counts[top_TRX_counts > 5] <- 5
  colnames(top_TRX_counts) <- colnames(TRX_counts)

  h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
  cuttoff <- cuttoff[names(cuttoff) == db]
  gene_module <- cutree(h, h=cuttoff)

  source("../enrichment_function.R")
  if(db == "GO"){
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")}else{
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")}
  res_list <- lapply( unique( gene_module[h$order] ),function(x){
    temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                              gmt_list = gmt_list,
                              min_terms_pathway = 10,
                              max_terms_pathway = 300,
                              min_overlap = 3,
                              sort_by_pvalue = T)
    temp <- temp[!grepl("REGULATION",rownames(temp)),]
    return(temp)
  } )
  res <- map(res_list, ~rownames_to_column(.x, var = "Terms")) %>% 
    set_names(., paste0("gene_module_",  seq_along(res_list)))

  write.xlsx(res, file=paste0("../../../results/ENRICHMENT/Heatmap/",db,"_enrichment_heatmap_",group_name,c,".xlsx"))

  return(list(res_list=res_list, top_TRX_counts=top_TRX_counts, 
              gene_module=gene_module, cuttoff=cuttoff, h=h, c=c))
}

```

```{r Heatmap plots and tables, fig.height=4, fig.width=7, echo = FALSE, eval = TRUE, warning=F}

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
TRX_counts <- TRX_counts[,sample_use]

#########
# DEG's #
#########
# Replace missing values with average in sexwork_months column
m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
metadata <- metadata %>%
  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))

dge_L <- edgR_dge.fun("Luminal_gr", confound=T)
dge_T <- edgR_dge.fun("Tissue_gr", confound=T)

###############
# ENRICHMENT #
##############
db <- c("GO", "KEGG") %>% set_names()
enrich_L <- db %>%
  map(., ~gene_mod_enrich.fun(dge_L$top,dge_L$y,dge_L$c,
                              db = .x,
                              group_name = "Luminal_gr", 
                              cuttoff = c(GO=2.5, KEGG=2.5)#3
                              )) 
enrich_T <- db %>%
  map(., ~gene_mod_enrich.fun(dge_T$top,dge_T$y,dge_T$c,
                              db = .x,
                              group_name = "Tissue_gr", 
                              cuttoff = c(GO=1.6, KEGG=1.6)
                              ))

########################
# PLOT GO/KEGG HEATMAP #
########################

sample_use <- metadata$ID
gr = "Luminal_gr"
name = "Luminal"
enrich = enrich_L
db <- "GO"
db <- "KEGG"
# plot_heatmap.fun(enrich_T, db, "Tissue_gr", "Tissue")
plot_heatmap.fun <- function(enrich, db, gr, name, N = 3){
df <- na.omit(metadata[sample_use, c(gr, "ID")])
df <- arrange(df, df[1])
samples <- df[, "ID"]
names(samples) <- df[, gr]

conf <- pluck(enrich, db, "c")
h <- pluck(enrich, db, "h")
cuttoff <- pluck(enrich, db, "cuttoff")
res_list <- pluck(enrich, db, "res_list")
top_TRX_counts <- pluck(enrich, db, "top_TRX_counts")[,samples]
gene_module <- pluck(enrich, db, "gene_module")

layout(matrix(c(1,2,3,4,5,
                1,2,3,4, 5),
                nrow = 2,ncol = 5,byrow = T),widths = c(1,3.5,.36,8,8)) #1,2,.3,8
  ## Heatmap
  par(mar=c(4,.1,2,.1))
  plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
  title(xlab="height", line=2)
  abline(v=cuttoff,xpd=F,col="red",lty=2)
  points( rep(-.1,length(gene_module)),
          seq(length(gene_module),1,length.out = length(gene_module)),
          col=taxa_pal[factor(gene_module[h$order])],
          pch=15,cex=.5,xpd=F)

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  title(xlab="patient groups", line=2)

  end <- table(pal[factor(names(samples))[ order(df[1])]])
  end <- end[order(factor(names(end), levels = pal))]
  end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.012) )
  start <- c(0, map_dbl(end[1:4], ~.x+0.012))

  axis(1, at = c(start, end), label = F, pos =1.02 , xpd=T, col="white",col.ticks="black")
  map2(start, end, ~lines(x=c(.y, .x), y= c(1.02, 1.02), xpd=T, cex=.8))
  # text labels
  t <- map2(start, end, ~(.x+.y)/2)
  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.03, .y,xpd=T, cex=.8))

  ## Meta info bars
  points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
          col=pal[factor(names(samples))[ order(df[1])]],
          pch=15,xpd=T,cex=.7)
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
          col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45) 
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
          col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45)

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )

  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))
  # pvalues <- unlist(lapply(res_list,function(x){ x$jaccard_index [1:N] }))
  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""

  #pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
  plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
               use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",
               yaxs="i",plot_weights = F)

  text(par("usr")[1],par("usr")[3]-.008,labels = "     BV",cex=.6, xpd=T)
  text(par("usr")[1],par("usr")[3]-.022,labels = "      HIV",cex=.6, xpd=T)

  par(mar=c(4,.1,2,5))

  # KEGG
  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(pvalue)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)

  # GO
  res_list <- pluck(enrich, "GO", "res_list")
  #top_TRX_counts <- pluck(enrich, "GO", "top_TRX_counts")[,samples]

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )

  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))
  # pvalues <- unlist(lapply(res_list,function(x){ x$jaccard_index [1:N] }))

  # pvalues <- unlist(lapply(res_list,function(x){ 
  #   xxx <- -log10(x$pvalue) [1:N]
  #   return( xxx / xxx[1] )
  # }))

  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""

  #pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(pvalue)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)

  # dev.copy2pdf(file=paste0("../../../results/ENRICHMENT/","DGE_heatmap_",db,"_",name,".pdf"),
  #     width = 6.5,
  #     height = 6, paper = "a4"
  #     )
}
#max(strwidth(pvalues))
#strwidth("M")

map(db, ~plot_heatmap.fun(enrich_L, .x, "Luminal_gr", "Luminal"))
map(db, ~plot_heatmap.fun(enrich_T, .x, "Tissue_gr", "Tissue"))

pdf("./test.pdf", width=7, height = 5, useDingbats = FALSE)
plot_heatmap.fun(enrich_L, "KEGG", "Luminal_gr", "Luminal")
dev.off()

```

```{r UMAP, echo=FALSE, fig.height=2.5, fig.width=3}
# , fig.height=3, fig.width=3.5
###############
#  TRX UMAP   #
###############
plot_UMAP.fun <- function(top_TRX_counts, group){
  #par(mar=c(4,4,2,6)) #b,l,t,r
  par(mar=c(2.5,2.5,2,3)) #b,l,t,r
  #par(mar=c(10,12,2,12)) #b,l,t,r
  set.seed(1)
  UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
                   metric = "correlation",min_dist = 0.1,spread = 5,
                   negative_sample_rate = 10)
  plot(UMAP_TRX,bg=pal[factor(group)],main="UMAP on top DGE RNAseq",
       pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  mtext("UMAP1", side=1, line=0.7, cex=0.8)
  mtext("UMAP2", side=2, line=0.7, cex=0.8)
  legend(par("usr")[2],par("usr")[4],title.adj = 0,
         #legend = levels(factor(metadata$Luminal_gr)),
         legend = sort(unique(group)),
         xjust = 0,yjust = 1, cex=0.7,
         bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
}

plot_UMAP.fun(enrich_L$GO$top_TRX_counts, metadata$Luminal_gr)
m <- filter(metadata, !is.na(Tissue_gr)) 
plot_UMAP.fun(enrich_T$GO$top_TRX_counts[,m$ID], m$Tissue_gr)

#plot_UMAP.fun(enrich_T$GO$top_TRX_counts, metadata$Tissue_gr)
dev.copy2pdf(file=paste0("./UMAP",".pdf"),
    width = 4,
    height = 3.5, paper = "a4" #units = "cm", res = 300 
    )

```

<!-- ## Pairwise DGEs and enrichment 
```{r Pairwise functions, message=FALSE, warning=FALSE, include=FALSE}
## individual group comparision ##
# for information about how to get pairwise comparison se:
# edgeR User’s Guide section 3.4.2 Blocking and 4.2.8  Differential expression
# https://www.bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf
###########################
# PAIRWISE DEG's FUNCTION #
###########################
pairwise_dge.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups),na.last = NA))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  #fit2 <- contrasts.fit(fit, CONTRASTS)
  #lrt_ <- glmLRT(fit2, coef = 2:5)
  lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% map(., ~rownames_to_column(.x, var = "Genes"))

  results <- map_df(lrt, ~decideTests(.x, method="separate", p.value = 0.05))
  summary <- bind_cols( "FDR<0.05"= c("Down","NotSig","Up"), summary(results))
  colnames(summary) <- c("FDR<0.05", g)
  top <- c(summary = list(summary), top)

return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS))
}

################################
# PAIRWISE ENRICHMENT FUNCTION #
################################
enrich_list <- function(x, gmt_list){
  temp <- compute_enrichment(genes = x,
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  temp <- rownames_to_column(temp, var = "Terms")
  return(temp)
} 
```

```{r Pairwise tables, include=FALSE}
##################
# PAIRWISE DEG's #
##################
pairwise_dge_L <- pairwise_dge.fun(group = "Luminal_gr", 
                                   fit=dge_L$fit, 
                                   design=dge_L$design )
pairwise_dge_T <- pairwise_dge.fun(group = "Tissue_gr",
                                   fit=dge_T$fit,
                                   design=dge_T$design)

# save TopTags 
top_dfs <- list(Luminal = pairwise_dge_L$top, Tissue = pairwise_dge_T$top) 
imap(top_dfs, ~write.xlsx(.x,file=paste0("../../../results/DE/","dge_pairwise_",.y,"_conf",".xlsx")))

###############################
# PAIRWISE ENRICHMENT GO/KEGG #
###############################
source("../enrichment_function.R")
gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")

#info_logFC <- map(top_dfs[[1]], ~table(abs(.x$logFC) > 1))
#info_PValue <- map(top_dfs[[1]], ~table(.$PValue < 0.01))

l_top <- list(Luminal = pairwise_dge_L$top[-1], Tissue = pairwise_dge_T$top[-1])
# up and down regulated genes seperatly
t_split <- modify_depth(l_top, 2, ~ .x %>%
        dplyr::filter(., .$PValue < 0.01) %>%
        #dplyr::filter(., abs(.$logFC) < 1) %>%
        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
        split(., .$reg)
  )
# up and down regulated genes together
t <- modify_depth(l_top, 2,~ .x %>%
        dplyr::filter(., .$PValue < 0.01) #%>%
        #dplyr::filter(., abs(.$logFC) < 1) 
  )

## GO enrichment 
GO_enrichment <- map_depth(t, 2, ~enrich_list(.x$Genes, gmt_list_GO))
GO_enrichment_split <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_GO)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) %>% 
  set_names(., paste0(names(.), "_split"))

## KEGG enrichment
KEGG_enrichment <- map_depth(t, 2, ~enrich_list(.x$Genes, gmt_list_KEGG))
KEGG_enrichment_split <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_KEGG)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) %>% 
  set_names(., paste0(names(.), "_split"))

## save files
enrich_KEGG <- list(KEGG_enrichment, KEGG_enrichment_split) %>% flatten()
enrich_GO <- list(GO_enrichment, GO_enrichment_split) %>% flatten()

imap(enrich_GO, ~write.xlsx(.x, 
file=paste0("../../../results/ENRICHMENT/Pairwise/","GO_enrichment_pairwise_",.y,"_conf",".xlsx"))) 
imap(enrich_KEGG, ~write.xlsx(.x, 
file=paste0("../../../results/ENRICHMENT/Pairwise/","KEGG_enrichment_pairwise_",.y, "_conf",".xlsx")))

```
 -->

### 2021-11-11 reports/rmarkdown/manuscript/Figure1.Rmd
Fig.1 minor changes

@@
 ---
-title: "Main Figure 1.Luminal Patient Grouping"
+title: "Figure 1.Luminal Patient Grouping"
 output: 
@@
 
-```{r Load data., message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
 ##################
@@
 figlabels <- letters
-temp_legend <- NULL
 
@@
                 5,4,7,7,8,8) , ncol= 6,byrow = T),
-                               widths = c(1.9,1.4,.7,.8,.8,.7,.8),
+                               widths = c(1.87,1.42,.7,.8,.8,.7,.8),
                                heights = c(1.6,.65,1.5,.95,.9,.2))
@@
 par(mar=c(0,3.3,0,1))
-fig1a <- magick::image_read("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/supplementary_files/Scematic figure.pdf", density = 900 )
+fig1a <- magick::image_read("../../../supplementary_files/Scematic figure.pdf", density = 900 )
 plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
@@
           c("Tissue_gr","ASV_tissue_V3_normalized_batch_corrected")
-          #"3"="ASV_CVL_V2_normalized_NOT_batch_corrected"
          )
@@
 
-  div <- tibble(shann, simp, invsimp ) %>% rename_with(., ~paste0(c("Shannon","Simpson","Inv Simpson")))
+  div <- tibble(ID=names(shann),shann, simp, invsimp ) %>% rename_with(., ~paste0(c("ID","Shannon","Simpson","Inv Simpson")))
   div_list[[length(div_list) + 1]] <- div
@@
          genes = c("Shannon","Simpson","Inv Simpson"),
-         # clustering = group_annotation,
          draw_mean_lines=F,col = pal[gr[order(group_annotation)]])
@@
 
-  if(is.null(temp_legend)){
-    temp_legend <- temp   #Use the same legend from the 1st dataset
-  }
-
   o <- order( rowSums(temp_legend) , decreasing = T)
@@
   temp <- temp[o,sample_o]*100
-  # temp <- temp[-grep("NA",rownames(temp)),]
   temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
@@
   text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
-  #text(x = c(2), y = 3, ']', srt = 90, cex = 3, xpd=T)
 
@@
         pch=15,cex=.7,xpd=T)
-  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T) # pos=4,
+  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T)
 
@@
 ###########
-par(mar=c(1,.3,0,.5)) #b,l,t,r
+par(mar=c(1,0,0,.5)) #b,l,t,r
 empty_plot()
@@
        legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other") ,
-       xjust = 0,yjust = 1,title.adj = 0,  ncol =1, #xjust=1,
-       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
+       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
+       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
        pt.bg = pal)
@@
 legend(x = par("usr")[c(1)],
-         y = par("usr")[c(4)],#title.adj = 0,
+         y = par("usr")[c(4)],
          legend = lvl[1:12],xjust = 0,yjust = 1,
          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-          #title="",
          pt.bg = taxa_pal)
 
-par(mar=c(0,.3,1.5,1)) #b,l,t,r
+par(mar=c(0,0,1.5,1)) #b,l,t,r
 empty_plot()
 legend(x = par("usr")[c(1)],
-         y = par("usr")[c(4)],#title.adj = 0,
+         y = par("usr")[c(4)],
          legend = lvl[13:17],xjust = 0,yjust = 1,
          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
-          #title="",
          pt.bg = taxa_pal[13:17])
@@
       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
-      vertex.size=10,#main="UMAP",
+      vertex.size=10,
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
@@
     dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
-    rename(Taxa="name")
+    rename(Taxa="name") %>%
+  mutate(across(where(is.numeric), ~round(.x, digits = 1)))
 
@@
   bind_rows(.id = "Sample type") %>%
-  mutate("Sample type" = gsub("_gr", "", .$"Sample type"))
+  select(-ID) %>%
+  mutate("Sample type" = gsub("_gr", "", .$"Sample type")) %>%
+  mutate(across(where(is.numeric), ~round(.x, digits = 2)))
 

### 2021-11-11 reports/rmarkdown/manuscript/Figure3.Rmd
Fig.3 Initial commit

---
title: "Figure 3. Tissue Microbiome Abundances"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

m <- metadata %>%
  select(ID, Luminal_gr, Tissue_gr) %>%
  arrange(Luminal_gr) %>%
  group_by(Luminal_gr) %>%
  nest() %>%
  mutate(L = map(data, ~arrange(.x, Tissue_gr))) %>%
  unnest(L)

# m <- metadata %>%
#   select(ID, Luminal_gr, Tissue_gr) %>%
#   left_join(div, by="ID") %>%
#   arrange(Tissue_gr) %>%
#   group_by(Tissue_gr) %>%
#   nest() %>%
#   mutate(L = map(data, ~arrange(.x, Luminal_gr, Shannon))) %>%
#   unnest(L)

# metadata <- metadata %>%
#   arrange(factor(ID, levels = m$ID))

group_annotation <- factor(setNames(m$Luminal_gr,m$ID))
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
group_annotation <- factor(setNames(metadata$Tissue_gr,metadata$ID))

#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")

```

```{r Figure 3., echo=FALSE, warning=FALSE, message=FALSE, fig.width=6 ,fig.asp=1, fig.pos="t"}
### A + B
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################
figlabels <- letters

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,6,7,4,4,
                3,3,6,5,4,4,
                8,8,8,8,8,8) , ncol= 6,byrow = T),
                               widths = c(1.2,1,1.83,1.7,.85,1.4),
                               heights = c(.7,1.5,.8,1.5,1,.8,.2))

ASV <- list(#c("Luminal_gr","ASV_CVL_V3_normalized_batch_corrected"),
             c("Tissue_gr","ASV_tissue_V3_normalized_batch_corrected")
         )
div_list <- list()
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]

  df <- na.omit(metadata[sample_use, c(i[1], "ID")])
  metadata <- metadata[df[,"ID"],]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))
  group_annotation <- na.omit(group_annotation)

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  temp <- t(t(2^x-1)/colSums(2^x-1))
  temp <- temp[, df[,"ID"]]

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","Simpson","Inv Simpson")

  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="",
         genes = c("Shannon","Simpson","Inv Simpson"),
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  div <- tibble(ID=names(shann),shann, simp, invsimp ) %>% rename_with(., ~paste0(c("ID","Shannon","Simpson","Inv Simpson")))
  div_list[[length(div_list) + 1]] <- div

  if(i[2] == "ASV_tissue_V3_normalized_batch_corrected"){
    end <- table(pal[factor(group_annotation[order(group_annotation)])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  temp <- rbind(temp[c(lvl[1:16]),], Other = colSums(temp[!(row.names(temp) %in% c(lvl[1:16])), ]))

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
           main=gsub("_gr","",i[1]),font.main=1,names.arg = rep("",ncol(temp)),
           col = taxa_pal[factor(rownames(temp),levels = lvl)])
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation)],
        pch=15,cex=.7,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T) # pos=4,
}

### C
############################
# TOTAL RELATIVE ABUNDANCE #
############################
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]
raw <- c("ASV_CVL_V3", "ASV_tissue_V3") %>% set_names()
raw_list <- raw %>%
  map(., ~read_csv(paste0("../../../results/",.x,"_raw_counts.csv")))  %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.5514392, "Other")) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_CVL_V3) %>%
    dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
    `row.names<-`(., NULL) %>% 
    column_to_rownames(var = "name") %>%
    as.matrix()

par(mar=c(6,5,2,.3)) #b,l,t,r
barplot( r, border = NA ,ylab="abundance", xlab ="", axisnames = F,
         col = taxa_pal[factor(rownames(r),levels = rownames(r))])
title(main = "Total relative abundance ", line = 0.5, cex.main = .9, font.main= 1)
mtext("Luminal        Tissue ", line = 0, cex=.6, side=1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### D
#############################
# LUMINAL-TISSUE COMPARISON #
#############################
par(mar=c(5,1.1,2,2)) #b,l,t,r
x <- as.character(metadata[,"Luminal_gr"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Tissue_gr"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
df <- data.frame(x1,y1)
plot_sankey( df,color_by = 1, 
             plot_labels = T, plot_weights = F,order_1_by = "NULL",
             order_2_by = "disentangled",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal= pal)
text(c(0,1),c(1,1),labels = c( "Luminal" , "Tissue" ),pos=3, xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###########
# LEGENDS #
###########
par(mar=c(0,0,0,.5)) #b,l,t,r # par(mar=c(1,.3,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, 
       y = 1.04, 
       legend = c("L. crispatus/jensenii", "L. iners", "Gardnerella", "High diverse", "Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,.2,1,.4)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,1,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

empty_plot()
```

### 2021-11-12 reports/rmarkdown/manuscript/Figure2.Rmd

@@
 ---
-title: "Main Figure 2. Bacterial Communities and Function"
+title: "Figure 2. Bacterial Communities and Function"
 output: 
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
 ##################
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
+```{r Figure 2.,echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
 
@@
               nrow = 5,ncol = 5,byrow = T), widths = c(.6,1.5,.2,1.9,2),
-       heights = c(.6,.6,.6,1,1.5))
+       heights = c(.68,.56,.56,1,1.5))
 figlabels <- letters
 
-## A
+### A
 #########################
@@
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] ,layout=lB)
-title(main = "           Bacterial Communities- Louvain", line = 0.5, cex.main = 1)
+title(main = "           Bacterial Communities- Louvain", line = 1, cex.main = 1)
 legend(par("usr")[2],par("usr")[4],title.adj = 0,
@@
 
-## B
+### B
 ###################################
-# BACTERIAL VAGINOSIS COMPARISSON #
+# WETT SMEAR BACTERIAL VALIDATION #
 ###################################
-par(mar=c(1,5,2,1)) #b,l,t,r
+par(mar=c(1,5,4,1)) #b,l,t,r
 x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
@@
                       "16S counts"=xL),
-         main = "Lactobacillus",
          genes = c("WS","16S counts"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
+         clustering = group_annotation,srt=20,
+         draw_mean_lines=F,col = pal)
+mtext("Lactobacillus", side=3, las = 1, cex = .6, line = .5,xpd=T)
+mtext("Wett Smear Bacteria Validation", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
 
@@
 
+par(mar=c(1,5,2,1)) #b,l,t,r
 barlist( data = rbind(WS=mM ,
                       "16S counts"=xM),
-         main = "Mobiluncus",
          genes = c("WS","16S counts"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
+         clustering = group_annotation,srt=20,
+         draw_mean_lines=F,col = pal)
+mtext("Mobiluncus", side=3, las = 1, cex = .6, line = .5,xpd=T)
 
@@
                       "16S counts"=xG),
-         main = "Gardnerella",
          genes = c("WS","16S counts"),
-         clustering = group_annotation,
-         draw_mean_lines=F,col = pal,srt=20)
+         clustering = group_annotation,srt=20,
+         draw_mean_lines=F,col = pal)
+mtext("Gardnerella", side=3, las = 1, cex = .6, line = .5,xpd=T)
 
-## C
+### C
 #####################
@@
 TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-sample_use <- colnames(TRX)[colSums(TRX)!=0]
+sample_use <- intersect(colnames(TRX)[colSums(TRX)!=0], colnames(picrust_data))
 
@@
 named_picrust_data <- rowsum( picrust_data , group = ko_names[ rownames(picrust_data) ] )
-#named_picrust_data[1:20,1:3]
-#dim(named_picrust_data)
 named_picrust_data <- round(named_picrust_data[ rowSums(named_picrust_data>5)>=5, ])
@@
 
-par(mar=c(4,.1,.5,.2)) #b,l,t,r
-
+par(mar=c(4,.1,4,.2)) #b,l,t,r
 h <- hclust( as.dist( (1- cor(t(top_picrust_data)))/2 ), method = "ward.D2")
 plot( rev(as.dendrogram(h)) ,xlim=c(max(h$height),-1), horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
+title(xlab="Height", line=2)
 cutoff <- 4
@@
        breaks = seq(-5,5,length.out = 100),axes=F)
+mtext("Differentialy Expressed KEGG Ontology Terms", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
+title(xlab="Study Groups", line=1)
+
+end <- table(pal[factor(group_annotation)[ order(group_annotation[sample_use])]] )
+end <- end[order(factor(names(end), levels = pal))]
+end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/ncol(picrust_data[sample_use])-.015) )
+start <- c(0, map_dbl(end[1:4], ~.x+0.015))
+
+axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
+map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
+# text labels
+t <- map2(start, end, ~(.x+.y)/2)
+map2(t,levels(factor(group_annotation)), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))
+
 points( seq(0,1,length.out = length(sample_use)),
@@
         col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]])
-
 points( seq(0,1,length.out = ncol(top_picrust_data)), 
@@
         col= c("tomato","orange","#d7d7d7")[as.factor(metadata$BV_Diagnosis_v3)][order( group_annotation[sample_use])] )
-
 points( seq(0,1,length.out = ncol(top_picrust_data)), 
@@
 
-par(mar=c(4,.1,.5,7))
+par(mar=c(4,.1,4,7))
 barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
          xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,names.arg = "")
-title(xlab="-log10(pvalue)", line=2)
+title(xlab="-log10(p-value)", line=2)
 abline(v=2,xpd=F,col="red",lty=2)
@@
 
-## D
+### D
 ################

### 2021-11-12 reports/rmarkdown/manuscript/Figure4-5.Rmd

---
title: "DGE's and Enrichment"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(readxl)
  library(enrichR)
  library(igraph)
  library(rafalib)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
sample_use <- metadata$ID

gr_name <- c("Luminal", "Tissue")
out <- paste0("./Suppl.Tbl/")

DEG_files <- c("Suppl.Tbl.04 DEGs_Luminal.xlsx",
               "Suppl.Tbl.08 DEGs_Tissue.xlsx")
enrichR_files <- c("Suppl.Tbl.07 TF_PPI_Luminal.xlsx",
                   "Suppl.Tbl.11 TF_PPI_Tissue.xlsx")

##################
# DEFINE CUTOFFS #
##################
dge_cutoff <- 0.01 # p-value
enrich_cutoff <- 0.1 # p-value
no_genes_cutoff <- 3
col_name <- paste0("genes_in(pValue<",dge_cutoff,")")

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

## Heatmap DGEs and enrichment
```{r Across functions, message=FALSE, warning=FALSE, include=FALSE}
##################
# DEG's FUNCTION #
##################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
edgR_dge.fun <- function(group, confound) {

  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~groups, data=metadata[sample_use,]) # without confounders
  }else{
    c <- "_conf"
    design <- model.matrix(~groups+HIVstatus+Contraception,
                         data=metadata[sample_use,])
  }

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
  y <- calcNormFactors(y,method = "TMM")
  y <- estimateGLMCommonDisp(y,design)
  y <- estimateGLMTagwiseDisp(y,design)
  fit <- glmFit(y, design)

  lrt <- glmLRT(fit,coef=2:n_gr) # with intercept
  top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
  colnames(top) <- sub(group,"",colnames(top))
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Genes")

return(list(design=design, y=y, fit=fit, lrt=lrt, top=top, c=c))
}

#######################
# ENRICHMENT FUNCTION #
#######################
# group_name <- "Luminal_gr"
# top <- dge_L$top
# y <- dge_L$y
# c <- dge_L$c
gene_mod_enrich.fun <- function(top, y, c, db, group_name, cuttoff){
  p_val <- pull(top, "PValue")
  top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dge <- top$Genes[ top_dge ]

  top_TRX_counts <- edgeR::cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
  top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
  top_TRX_counts[top_TRX_counts > 5] <- 5
  colnames(top_TRX_counts) <- colnames(TRX_counts)

  h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
  cuttoff <- cuttoff[names(cuttoff) == db]
  gene_module <- cutree(h, h=cuttoff)

  source("../enrichment_function.R")
  if(db == "GO"){
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")}else{
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")}
  res_list <- lapply( unique( gene_module[h$order] ),function(x){
    temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                              gmt_list = gmt_list,
                              min_terms_pathway = 10,
                              max_terms_pathway = 300,
                              min_overlap = 3,
                              sort_by_pvalue = T)
    temp <- temp[!grepl("REGULATION",rownames(temp)),]
    return(temp)
  } )

  return(list(res_list=res_list, top_TRX_counts=top_TRX_counts, 
              gene_module=gene_module, cuttoff=cuttoff, h=h, c=c))
}

```

```{r Pairwise functions, message=FALSE, warning=FALSE, include=FALSE}
###########################
# PAIRWISE DEG's FUNCTION #
###########################
pairwise_dge.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups),na.last = NA))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  #fit2 <- contrasts.fit(fit, CONTRASTS)
  #lrt_ <- glmLRT(fit2, coef = 2:5)
  lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% map(., ~rownames_to_column(.x, var = "Genes"))

  results <- map_df(lrt, ~decideTests(.x, method="separate", p.value = 0.05))
  summary <- bind_cols( "FDR<0.05"= c("Down","NotSig","Up"), summary(results))
  colnames(summary) <- c("FDR<0.05", g)
  top <- c(summary = list(summary), top)

return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS))
}
################################
# PAIRWISE ENRICHMENT FUNCTION #
################################
enrich_list <- function(x, gmt_list){
  temp <- compute_enrichment(genes = x,
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  temp <- rownames_to_column(temp, var = "Terms")
  return(temp)
} 
```

```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
TRX_counts <- TRX_counts[,sample_use]

#########
# DEG's #
#########
# Replace missing values with average in sexwork_months column
m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
metadata <- metadata %>%
  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))

dge_L <- edgR_dge.fun("Luminal_gr", confound=T)
dge_T <- edgR_dge.fun("Tissue_gr", confound=T)

###############
# ENRICHMENT #
##############
db <- c("GO", "KEGG") %>% set_names()
enrich_L <- db %>%
  map(., ~gene_mod_enrich.fun(dge_L$top,dge_L$y,dge_L$c,
                              db = .x,
                              group_name = "Luminal_gr", 
                              cuttoff = c(GO=2.5, KEGG=2.5)#3
                              )) 
enrich_T <- db %>%
  map(., ~gene_mod_enrich.fun(dge_T$top,dge_T$y,dge_T$c,
                              db = .x,
                              group_name = "Tissue_gr",
                              cuttoff = c(GO=1.6, KEGG=1.6)
                              ))

```

```{r Pairwise tables, include=FALSE}
##################
# PAIRWISE DEG's #
##################
pairwise_dge_L <- pairwise_dge.fun(group = "Luminal_gr", 
                                   fit=dge_L$fit, 
                                   design=dge_L$design )
pairwise_dge_T <- pairwise_dge.fun(group = "Tissue_gr",
                                   fit=dge_T$fit,
                                   design=dge_T$design)

###############################
# PAIRWISE ENRICHMENT GO/KEGG #
###############################
source("../enrichment_function.R")
gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")

#info_logFC <- map(top_dfs[[1]], ~table(abs(.x$logFC) > 1))
#info_PValue <- map(top_dfs[[1]], ~table(.$PValue < 0.01))

l_top <- list(Luminal = pairwise_dge_L$top[-1], Tissue = pairwise_dge_T$top[-1])
# up and down regulated genes seperatly
t_split <- modify_depth(l_top, 2, ~ .x %>%
        dplyr::filter(., .$PValue < 0.01) %>%
        #dplyr::filter(., abs(.$logFC) < 1) %>%
        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
        split(., .$reg)
  )

## GO enrichment 
pairwise_GO <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_GO)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) 
## KEGG enrichment
pairwise_KEGG <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_KEGG)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation"))

```

```{r Suppl. Tbl. 4 & 8, include=FALSE}
# DEG's across and pairwise
pages <- list(list(Across=dge_L$top), pairwise=pairwise_dge_L$top) %>% flatten()
write.xlsx(pages, file=paste0(out,DEG_files[1]))

pages <- list(list(Across=dge_T$top), pairwise=pairwise_dge_T$top) %>% flatten()
write.xlsx(pages, file=paste0(out,DEG_files[2]))
```

```{r Suppl. Tbl. 5,6 & 9,10, eval=FALSE, include=FALSE}
# Enrichment Across
df_across <- enframe(list(Luminal=enrich_L, Tissue=enrich_T)) %>%  
  unnest_longer(value) %>%
  mutate(res = map(.$value, "res_list")) %>%
  unnest_longer(res) %>%
  mutate(res = map(res, ~rownames_to_column(.x, var = "Terms"))) %>%
  mutate(res = map(res, ~filter(.x, pvalue <= 0.01))) %>%
  group_by(name, value_id) %>%
  mutate(mod = row_number()) %>% ungroup() %>%
  mutate(res = set_names(.$res, paste0(.$value_id,"_gene_module_",.$mod))) %>%
  group_split(name) %>% set_names(., map(., ~unique(.x$name)))

write.xlsx(df_across[[1]]$res, file=paste0(out,"Suppl.Tbl.05 Enrichment_Across_Luminal",".xlsx"))
write.xlsx(df_across[[2]]$res, file=paste0(out,"Suppl.Tbl.09 Enrichment_Across_Tissue",".xlsx"))

# Enrichment pairwise
df_pairwise <- enframe(list(GO=pairwise_GO, KEGG=pairwise_KEGG)) %>%
  unnest_longer(value) %>%
  unnest_longer(value, names_repair ="unique") %>%
  mutate(value = map(value, ~filter(.x, FDR <= 0.05))) %>%
  mutate(value = set_names(.$value, paste0(.$name,"_",.$value_id...3))) %>%
  group_split(value_id...4) %>% set_names(., map(., ~unique(.x$value_id...4)))

write.xlsx(df_pairwise[[1]]$value, file=paste0(out,"Suppl.Tbl.06 Enrichment_Pairwise_Luminal",".xlsx"))
write.xlsx(df_pairwise[[2]]$value, file=paste0(out,"Suppl.Tbl.10 Enrichment_Pairwise_Tissue",".xlsx"))
```

```{r TF PPI Analysis function, include=FALSE}
TF_PPI_analysis.fun <- function(out, filename, gr_name, outfile) {
  # Get DEG's files and filter p-value < 0.01
  sheets <- excel_sheets(paste0(out, filename)) %>% set_names(.) %>% .[-c(1:2)]
  detables <- map(sheets, ~read_xlsx(paste0(out, filename), sheet = .x, col_names = TRUE)) %>%
    map(.,~mutate(.x, DE = ifelse(.x$PValue > dge_cutoff , 0, ifelse(.x$logFC < 0, -1, 1 )))) %>% 
    map(.,~filter(., .$PValue <= 0.01))

    # Running enrichments for up- or down-regulated genes
    save_l <- list()
    for(i in sheets ){
      message(paste0("Processing comparisson   ", i))
      temp <- detables[[i]]
      temp <- temp[ !duplicated(temp$Genes) & !is.na(temp$Genes) , ]
      temp <- mutate(temp, rownames = Genes)
      temp <- column_to_rownames(temp, var = "rownames")

      save_f <- list()
      for(j in c(1,-1)){
        gene_list <- temp$Genes[ temp$DE == j ]

        #Transcription_Factor_PPIs
        f <- enrichr(genes = gene_list, databases = "Transcription_Factor_PPIs")[[1]]
        if (is.data.frame(f) && nrow(f)>0){
          f <- f[order(f$P.value,decreasing = F), -c(5:6)]
          f <- cbind("Transcription Factor"=f[["Term"]], setNames(list(length(gene_list)), col_name), f[,-c(1)])
          f <- f[f$P.value < enrich_cutoff & as.numeric(sapply(strsplit(f[,"Overlap"],"/"),"[[", 1)) >= no_genes_cutoff,]
          f_n <- set_names(list(f), paste0(ifelse(j==1,"up","down"))) 
          save_f <- append(save_f, f_n)

        }else{if(length(save_f) == 0){save_f <- append(save_f, f_n)} }
      }
      l <- bind_rows(save_f, .id = "Regulation") %>%
        filter(Adjusted.P.value <= 0.05)
      l <- set_names(list(l), paste0(i))
      save_l <- append(save_l, l ) 
    } 
    write.xlsx(save_l, file=paste0(out, outfile), row.names=FALSE)
  }
```

```{r TF PPI Analysis, eval=FALSE, include=FALSE}
TF_PPI_analysis.fun(out, DEG_files[1], gr_name[1], enrichR_files[1]) # writes Suppl.Tbl.07
TF_PPI_analysis.fun(out, DEG_files[2], gr_name[2], enrichR_files[2]) # writes Suppl.Tbl.11
```

```{r plot functions, warning=FALSE, include=FALSE}
############################
# GO/KEGG HEATMAP FUNCTION #
############################
# sample_use <- metadata$ID
# gr = "Tissue_gr"
# name = "Tissue"
# enrich = enrich_T
# db <- "GO"
# db <- "KEGG"
# plot_heatmap.fun(enrich_T, db, "Tissue_gr", "Tissue")
plot_heatmap.fun <- function(enrich, db, gr, name, N=3, Layout){
  figlabels <- letters
  df <- na.omit(metadata[sample_use, c(gr, "ID")])
  df <- arrange(df, df[1])
  samples <- df[, "ID"]
  names(samples) <- df[, gr]

  conf <- pluck(enrich, db, "c")
  h <- pluck(enrich, db, "h")
  cuttoff <- pluck(enrich, db, "cuttoff")
  res_list <- pluck(enrich, db, "res_list")
  top_TRX_counts <- pluck(enrich, db, "top_TRX_counts")[,samples]
  gene_module <- pluck(enrich, db, "gene_module")

  if(missing(Layout)){
    layout(matrix(c(1,2,3,4,5,
                  1,2,3,4, 5),
                  nrow = 2,ncol = 5,byrow = T),widths = c(1,3.5,.36,8,8))
  }else{layout(Layout[[1]],Layout[[2]],Layout[[3]])}

  ## Heatmap
  par(mar=c(4,.1,2.5,.1))
  plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
  title(xlab="Height", line=2)
  abline(v=cuttoff,xpd=F,col="red",lty=2)
  points( rep(-.1,length(gene_module)),
          seq(length(gene_module),1,length.out = length(gene_module)),
          col=taxa_pal[factor(gene_module[h$order])],
          pch=15,cex=.5,xpd=F)

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("           Differentialy Expressed Genes", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

  end <- table(pal[factor(names(samples))[ order(df[1])]])
  end <- end[order(factor(names(end), levels = pal))]
  end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
  start <- c(0, map_dbl(end[1:4], ~.x+0.015))

  axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
  map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
  # text labels
  t <- map2(start, end, ~(.x+.y)/2)
  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

  ## Meta info bars
  points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
          col=pal[factor(names(samples))[ order(df[1])]],
          pch=15,xpd=T,cex=.7)
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
          col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45) 
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
          col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45)

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )

  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))
  # pvalues <- unlist(lapply(res_list,function(x){ x$jaccard_index [1:N] }))
  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  terms <- gsub("endoplasmic reticulum", "ER", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""

  #pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
  plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
               use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",
               yaxs="i",plot_weights = F)

  text(par("usr")[1],par("usr")[3]-.008,labels = "    BV",cex=.6, xpd=T)
  text(par("usr")[1],par("usr")[3]-.028,labels = "    HIV",cex=.6, xpd=T)

  # KEGG
  if(name == "Tissue")
  {par(mar=c(4,.1,2.5,13))}else #bottom, left, top, right  
      {par(mar=c(4,.1,2.5,8))}
  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(p-value)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)

  # GO
  res_list <- pluck(enrich, "GO", "res_list")
  #top_TRX_counts <- pluck(enrich, "GO", "top_TRX_counts")[,samples]

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )
  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))

  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  terms <- gsub("endoplasmic reticulum", "ER", terms, perl = TRUE)
  terms <- gsub("establishment of ", "", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(p-value)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
}

###########################
# ASSEMBLE PLOTS FUNCTION #
###########################
arrange_plots.fun <- function(degs, group, gr_name, comparison, enrichment_db, file, Layout){

  ### A
  ###################
  # GO/KEGG HEATMAP #
  ###################
  plot_heatmap.fun(degs, "KEGG", group, gr_name, Layout=Layout)

  ### B
  #####################
  # TF PPI ENRICHMENT #
  #####################
  # Create plot for each enrichment
  n <- 10
  for( x in comparison ){
  for( i in enrichment_db ){
    DD <- Inf
    UU <- 0
    names_DD <- ""
    names_UU <- ""

    f_DOWN <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "down")
    f_UP <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "up")

    if (is.data.frame(f_DOWN) && nrow(f_DOWN)>0){
      tempD_e <- f_DOWN
      DD <- log10(tempD_e$P.value)[1:min(n,length(tempD_e$P.value))]
      names_DD <- (as.character(tempD_e$"Transcription Factor"))[1:min(n,length(tempD_e$P.value))]
    }
    if (is.data.frame(f_DOWN) && nrow(f_DOWN)>0){
      tempU_e <- f_UP
      UU <- rev(-log10(tempU_e$P.value)[1:min(n,length(tempU_e$P.value))])
      names_UU <- rev(as.character(tempU_e$"Transcription Factor")[1:min(n,length(tempU_e$P.value))])
    }

    comb <- c( DD/abs(max(abs(DD))) , UU/abs(max(abs(UU))) )
    comb <- replace_na(comb, 0)

    if (identical(comb, c(0,0))){next}

    par(mar = c(1.2, 4.7, 2.2, 1))  #bottom, left, top, right
    barplot( c(rep(NA, each=20-length(comb)), comb),  
             xlim=c(-1.1,1.1),
             # xlim=c(min(c(comb,-2/abs(max(abs(DD))) ))*1.1, 
             #                          max(c(comb,2/abs(max(abs(DD)))) )*1.1), 
                        cex.names = .9, #mar=c(1,25,2,2), #main=paste(x, i, sep="\n"),
             names.arg = c(rep("", each=20-length(comb)),names_DD, names_UU ), 
             horiz = T,las= 1, border=NA,xaxt='n',line=-.2, #xlab="-log10 p-value",
             col = c(rep(NA, each=20-length(comb)), 
                     rev(colorRampPalette( c("firebrick3","gray","navy") )(19))[(comb+1)*9+1])
    )
    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=.9)

    if(x == "L1-L2" | x == "T1-T2"){add_letter("b")}
    points_y <- rev(seq(par("usr")[4]-1.45, by=-1.2, length=length(c(UU,DD))))
             points( comb , points_y,
                     #mar=c(1,20,2,2),
                     bg=colorRampPalette( c("navy","gray","firebrick3") )(19)[(comb+1)*9+1],
                     pch=21, col="black",cex=1.5)
             lines(x=c(0, 0), y= c(points_y[1]-.6, par("usr")[4]-1)) 
             lines(x=c(-2/abs(max(abs(DD))),-2/abs(max(abs(DD)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3)) # pval=0.01
             lines(x=c(2/abs(max(abs(UU))),2/abs(max(abs(UU)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3))  # pval=0.01
             text(0,points_y[1]-1.5, "-log10 p-value", xpd=T)
    }
  }
}

dev.copy2pdf(file=paste0("./test.pdf"),
    width = 6.6,
    height = 6, paper = "a4"
    )
```

```{r Figure 4., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
# Layout
l <- list(mat = matrix(c(1,2,3,4,4,4,5,5,5,
                         1,2,3,4,4,4,5,5,5,
                         6,6,6,7,7,8,8,9,9,
                         10,10,10,11,11,12,12,13,13),
                nrow = 3,ncol = 9,byrow = T),
          widths = c(1.14,3.5,.36,2.5,2.5,2.5,2.5,2.5,2.5),
          heigth = c(1,1,1.5)
          )
#layout.show(layout(l[[1]],l[[2]],l[[3]]))

sheets <- excel_sheets(paste0(out, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)
```

```{r Figure 5., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
# comparison <- sheets
# enrichment_db <- "TF_PPI_enrichment"
# degs <- enrich_T
# group <- "Tissue_gr" 
# gr_name <- "Tissue"

# Layout
l <- list(mat = matrix(c(1,2,3,4,4,4,5,5,5,
                         1,2,3,4,4,4,5,5,5,
                         6,6,6,7,7,8,8,9,9,
                         10,10,10,11,11,12,12,13,13),
                nrow = 3,ncol = 9,byrow = T),
          widths = c(1.14,3.5,.36,2.5,2.5,2.5,2.5,2.5,2.5),
          heigth = c(1,1,1.5)
          )

sheets <- excel_sheets(paste0(out, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)
```

```{r UMAP, eval=FALSE, fig.height=2.5, fig.width=3, include=FALSE}
# , fig.height=3, fig.width=3.5
###############
#  TRX UMAP   #
###############
plot_UMAP.fun <- function(top_TRX_counts, group){
  #par(mar=c(4,4,2,6)) #b,l,t,r
  par(mar=c(2.5,2.5,2,3)) #b,l,t,r
  #par(mar=c(10,12,2,12)) #b,l,t,r
  set.seed(1)
  UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
                   metric = "correlation",min_dist = 0.1,spread = 5,
                   negative_sample_rate = 10)
  plot(UMAP_TRX,bg=pal[factor(group)],main="UMAP on top DGE RNAseq",
       pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  mtext("UMAP1", side=1, line=0.7, cex=0.8)
  mtext("UMAP2", side=2, line=0.7, cex=0.8)
  legend(par("usr")[2],par("usr")[4],title.adj = 0,
         #legend = levels(factor(metadata$Luminal_gr)),
         legend = sort(unique(group)),
         xjust = 0,yjust = 1, cex=0.7,
         bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
}

plot_UMAP.fun(enrich_L$GO$top_TRX_counts, metadata$Luminal_gr)
m <- filter(metadata, !is.na(Tissue_gr)) 
plot_UMAP.fun(enrich_T$GO$top_TRX_counts[,m$ID], m$Tissue_gr)

#plot_UMAP.fun(enrich_T$GO$top_TRX_counts, metadata$Tissue_gr)
dev.copy2pdf(file=paste0("./UMAP",".pdf"),
    width = 4,
    height = 3.5, paper = "a4" #units = "cm", res = 300 
    )

```

 <!-- -->

### 2021-11-12 reports/rmarkdown/manuscript/Figure8.Rmd

@@
 
-# save TopTables
-write.xlsx(dpe_L$top, file=paste0("../../../results/DE/","dpe_across_","Luminal",dpe_L$c,".xlsx"))
-write.xlsx(dpe_T$top, file=paste0("../../../results/DE/","dpe_across_","Tissue",dpe_T$c,".xlsx"))
-
 # Pairwise DPE's
@@
 
-# save TopTables 
-l_top <- list(Luminal = pairwise_dpe_L$top, Tissue = pairwise_dge_T$top)
-imap(l_top, ~write.xlsx(.x,file=paste0("../../../results/DE/","dpe_pairwise_",.y,"_conf",".xlsx"))) # Excel
-
 ```
@@
 ```{r heatmap functions, echo=FALSE}
-
 top_dpe.fun <- function(top){
@@
 
-stat <- imap(c("Luminal_gr", "Tissue_gr"), ~test.krusk_dunn.fun(cytogr.long, Cytokines, .x))
+n <- c("Luminal_gr", "Tissue_gr") %>% set_names(c("Luminal", "Tissue"))
+stat <- imap(n, ~test.krusk_dunn.fun(cytogr.long, Cytokines, .x))
 s <- flatten(stat)
-write.xlsx(s, file=paste0("../../../results/","Cytokine_stats",".xlsx"))
+```
+
+```{r Save stats to excel, include=FALSE}
+# save TopTables pairwise
+l_top_p <- list(Luminal = pairwise_dpe_L$top, Tissue = pairwise_dge_T$top)
+imap(l_top_p, ~write.xlsx(.x,file=paste0("../../../results/DE/","dpe_pairwise_",.y,".xlsx"))) # Excel
+
+# save TopTables across + cytokines
+l_top_a <- list(Luminal = list(Across = dpe_L$top), Tissue = list(Across = dpe_T$top))
+l <- list(l_top_a, stat, l_top_p)
+names <- c("Luminal", "Tissue") %>% set_names()
+p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))
+
+imap(p_list, ~write.xlsx(.x,file=paste0("../../../results/DE/","Proteins_",.y,".xlsx"))) # Excel
 
@@
 Protein Heatmap**.
-**a** and **c** Without scaling (normalized MFI), for Luminal and Tissue respectively **b** and **d** scaled by row for Luminal and Tissue respectively.
-top dpe with p-value < 0.05
+**a** and **b** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively.
+(p-value < 0.05) **b** and **d** sig.
+cytokines for Luminal and Tissue respectively.
 

### 2021-11-15 reports/rmarkdown/manuscript/Figure4-5.Rmd
Add Supp. Tbl 12

@@
 
+```{r Suppl. Tbl. 12, include=FALSE}
+metadata <- metadata %>% 
+  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
+  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))
+
+dge_LT <- edgR_dge.fun("L_T_groups", confound=T)
+enrich_LT <- db %>%
+  map(., ~gene_mod_enrich.fun(dge_LT$top,dge_LT$y,dge_LT$c,
+                              db = .x,
+                              group_name = "L_T_groups",
+                              cuttoff = c(GO=2, KEGG=2) # c(GO=2, KEGG=2.7)
+                              ))
+
+#pal <- c("#54ba63","#ad58c6","#d03e7a","#4cb6a9","#bb6fac","#cca442","#8fb53c","#6a7ecd","#63813e","#c46570","#b3743e","#d35238")
+#plot_heatmap.fun(enrich_LT, "KEGG", "L_T_groups", "L_T_groups")
+
+t <- dge_LT$top %>% filter(PValue <= 0.01)
+
+df_across <- enframe(list(LtoT=enrich_LT)) %>%  
+  unnest_longer(value) %>%
+  mutate(res = map(.$value, "res_list")) %>%
+  unnest_longer(res) %>%
+  mutate(res = map(res, ~rownames_to_column(.x, var = "Terms"))) %>%
+  mutate(res = map(res, ~filter(.x, pvalue <= 0.01))) %>%
+  group_by(name, value_id) %>%
+  mutate(mod = row_number()) %>% ungroup() %>%
+  mutate(res = set_names(.$res, paste0(.$value_id,"_gene_module_",.$mod))) %>%
+  group_split(name) %>% set_names(., map(., ~unique(.x$name)))
+
+pages <- list(list("DEGs L2T2 vs L2T3"=t), df_across[[1]]$res) %>% flatten()
+write.xlsx(pages, file=paste0(out,"Suppl.Tbl.12 L2T2_L2T3_DEGs_and_enrichment.xlsx")) 
+```
+
 ```{r UMAP, eval=FALSE, fig.height=2.5, fig.width=3, include=FALSE}
@@
  <!-- -->
+
+## igraph network plot
+```{r igraph network, echo=FALSE, warning=FALSE} 
+# 1. Change the groups of interest by assigning condition_use
+# 2. Change the number of genes to be included in the network by changing n       # , results="hold"
+#    look at tempD_i and tempU_i to see how many genes was significant
+# 3. Adjust the size of the gene nodes and text in the graph by changing; square, circle and txt
+
+# combinations
+g <- sheets[-1]
+
+# Set enrichment conditions
+n <- 10  # maximum number of TF's to plot
+# sizes:
+square <- 10
+cricle <- 8
+txt <- .7
+
+# Create a graph from the list of UP and DOWN regulated genes
+i_graph <- function(temp_i, n, square, cricle){
+  temp <- lapply( 1:n, function(x){ strsplit(as.character(temp_i$Genes[x]),split = ";")[[1]]  } )
+  temp <- map(temp, purrr::discard, is.na) %>% compact() # removes NA
+  names(temp) <- temp_i$Term[1:length(temp)]
+  temp2 <- lapply( names(temp) , function(x){ rep(x,length(temp[[x]])) })
+  temp <- cbind(unlist(temp2),unlist(temp))
+  g <- graph_from_data_frame(temp,directed = F)
+  v_shape <- ifelse( names(V(g)) %in% unique(temp[,1]) , "square" , "circle" )
+  v_color <- ifelse( names(V(g)) %in% unique(temp[,1]) , "honeydew" , "cornsilk" )
+  v_size <- ifelse( names(V(g)) %in% unique(temp[,1]) , square , cricle )
+
+  return(list(g=g, v_shape=v_shape, v_color=v_color, v_size=v_size))
+  }
+
+# Test parameters
+# i <- "TF_PPI_enrichment"
+# i <- "TF_TRRUST_enrichment"
+# n <- 4
+# g <- sheets[-1]
+# condition_use <- g[1]
+# g <-  g[1]
+# temp_i_names <- paste0(strsplit(i, "enrichment"),condition_use,"_UP_", gr_name,".csv") %>% set_names()
+
+# Save all combination plots in one file
+save_pdf.fun <- function(i){
+
+  temp_i_names <- grep(gr_name,list.files(paste0(out,i,"/",gr_name),full.names=F),value=T) %>%
+    set_names()
+  temp_i <- imap(temp_i_names, ~read.csv2(paste0(out,i,"/",gr_name,"/",.x)))
+  obj_igraph <- imap(temp_i, ~i_graph(.x, n, square, cricle))
+  pattern <-paste(g , collapse="|")
+
+  par(mar = c(0, 0, .7, 0))  #bottom, left, top, right
+  p <- imap(obj_igraph,
+       ~with(.x, 
+               plot.igraph(g, label=V(g), 
+               main = paste0(str_extract(.y,pattern)," ",str_extract(.y, "UP|DOWN")," regulated genes"),
+               layout=layout_with_kk(g), vertex.label.color="black" ,
+               vertex.size=v_size, vertex.label.cex=txt , vertex.label.font=2, 
+               vertex.shape=v_shape, vertex.color=v_color, marigin = c(-1,-4,0,0),
+               vertex.frame.color="grey60", curved=F) 
+  ))
+  #text(0,1.15, paste0(str_extract(.y,pattern)," ",str_extract(.y, "UP|DOWN")," regulated genes"), cex=1.2, font=2)
+  dev.copy2pdf(file = paste0(out, "Plots/", gr_name, "_network_", i,".pdf", sep=""))
+}
+
+```

### 2021-11-18 reports/rmarkdown/01_data_pre_processing.Rmd
update

@@
 # Load metadata
-dataset_names <- c("ASV_tissue_B2.csv",   # Tissue, Boston run 2 (95 samples)
-                     "ASV_tissue_B1.csv", # Tissue, Boston run 1 (1 sample)
+dataset_names <- c("ASV_tissue_V3_B2.csv",   # Tissue, Boston run 2 (95 samples)
+                     "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
                      "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
                      "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
-                     "ASV_CVL_V2_C.csv",  # CVL V2, CTMR         (62 samples)
-                     "ASV_CVL_V3.csv")
+                     "ASV_CVL_V2_S.csv",  # CVL V2, CTMR         (62 samples)
+                     "ASV_CVL_V3_B1.csv") # CVL V3, Boston run 1 (111 samples)
 
@@
             other_taxa.fun(.) %>% # create "other" taxonomy 
-            genus_lacto.fun(.)  
-            ) %>% # create genus lacto taxonomy 
+            genus_lacto.fun(.)  # create genus lacto taxonomy
+            ) %>%  
         {. ->> temp_ } %>% 
-        map(., ~aggregate.fun(., "Genus_lacto"))
+        map(., ~aggregate.fun(., "Genus_lacto")) # agglomerate on genus/lacto spp. lvl.
 
 # Extract taxonomic information to file
-seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
+seq_taxa <- bind_rows(temp_, .id = "dataset") %>% select(1:12) %>% unique() 
 #write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )
+
+# make long format raw non agglom ASV files
+# remove_from_CVL_V2 <- intersect(colnames(temp_[[5]]), colnames(temp_[[3]]))[-c(1:11)]
+# names <- c("ASV_tissue_V3_Boston run2","ASV_tissue_V3_Boston run1", "ASV_CVL_V2_Boston run1", 
+#                      "ASV_CVL_V2_Boston run2", "ASV_CVL_V2_Stockholm", "ASV_CVL_V3_Boston run 1") 
+# ASV_long <- temp_ %>%
+#   modify_at(., c("ASV_CVL_V2_C.csv"), ~select(.x, -any_of(remove_from_CVL_V2))) %>% 
+#   map(., ~pivot_longer(.x, matches("^P\\d"), names_to = "ID", values_to = "Reads")) %>%
+#   set_names(names ) %>%
+#   bind_rows(., .id = "dataset") %>%
+#   separate(., dataset, into=c(NA, "Sample_type", "Visit","Place_sequenced"), sep = "_", remove = F) %>% 
+#   mutate(dataset = str_extract(.$dataset, "CVL_V\\d|tissue_V3")) %>% 
+#   group_split(dataset, .keep=F) %>% set_names(c("tissue", "CVL_V2", "CVL_V3"))
+# 
+# saveRDS(ASV_long, paste0(PATH,"results/","ASV_all_data_long_format.RDS"))
 
@@
 lapply(datasets,colnames)
+
 #arrange all datasets to have the same rownames with all bacteria
@@
 lapply(datasets,dim)
+
 # datasets <- datasets[3] #in case you wanna test in 1 dataset
@@
 # lapply(datasets,dim)
+
 # Extract taxonomic information to file
@@
 write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
+
 ```
@@
 
+# normalized files
 datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)

### 2021-11-18 reports/rmarkdown/01_data_pre_processing.Rmd
Remove CVL V2 and batch testing

@@
 
-```{r, echo = FALSE}
+```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
+##################
+# LOAD LIBRARIES #
+##################
+suppressWarnings({suppressMessages({suppressPackageStartupMessages({
 library(igraph)
@@
 library(tidyverse)
+})  })  })
 
-```
-
-#Defining some variables for the analysis
-
-```{r, echo = FALSE}
+#############
+# LODA DATA #
+#############
 PATH <- "../../" #Path to the data
 
-#color palettes
+# Load metadata
+dataset_names <- c("ASV_tissue_V3_B2.csv",   # Tissue, Boston run 2 (95 samples)
+                     "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
+                     #"ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
+                     #"ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
+                     #"ASV_CVL_V2_S.csv",  # CVL V2, CTMR         (62 samples)
+                     "ASV_CVL_V3_B1.csv") # CVL V3, Boston run 1 (111 samples)
+
+datasets_ <- map(dataset_names, ~read.csv( paste0(PATH,"data/",.x))) %>% set_names(., dataset_names)
+metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
+rownames(metadata) <- as.character(metadata$ID)
+
+#################
+# COLOR PALETTS #
+#################
 pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
@@
 
-```{r, Functions}
+```{r Functions, include=FALSE}
 # create "other" taxonomy
@@
 
-# Loading data and metadata
-
-```{r, echo = FALSE}
-# Load metadata
-dataset_names <- c("ASV_tissue_V3_B2.csv",   # Tissue, Boston run 2 (95 samples)
-                     "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
-                     "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
-                     "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
-                     "ASV_CVL_V2_S.csv",  # CVL V2, CTMR         (62 samples)
-                     "ASV_CVL_V3_B1.csv") # CVL V3, Boston run 1 (111 samples)
-
-datasets_ <- map(dataset_names, ~read.csv( paste0(PATH,"data/",.x))) %>% set_names(., dataset_names)
-metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
-```
-
 # Taxonomy agglomeration
 
-```{r, echo = FALSE}
+```{r Agglomerate Taxa, echo = FALSE}
 temp <- map(datasets_, ~.x  %>%
@@
 
-# Loading data and metadata Paulos code
-
-```{r, eval=F, echo = FALSE}
-PATH <- "../../" #Path to the data
-# Load metadata
-metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
-# Load datasets
-dataset_names <- c("ASV_tissue_B2.csv",
-                     "ASV_tissue_B1.csv",
-                     "ASV_CVL_V2_B1.csv",
-                     "ASV_CVL_V2_B2.csv",
-                     "ASV_CVL_V2_C.csv",
-                     "ASV_CVL_V3.csv")
-#import datasets and return matrix with taxa-level resolution
-datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"data/",x))
-  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
-  temp[is.na(temp)] <- "NA"
-  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
-  not_complete <- grepl("_NA",sps)
-
-  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
-  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
-  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
-  taxa[!not_complete] <- sps[!not_complete]
-  taxa <- sub("_"," ",taxa)
-  use <- !is.na(taxa) & (taxa != "NA")
-  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
-  temp <- rowsum(temp[use,],taxa[use])
-  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
-  return(temp)
-})
-names(datasets) <- dataset_names
-lapply(datasets,dim)
-lapply(datasets,colnames)
-
-#arrange all datasets to have the same rownames with all bacteria
-common_microbes <- unique(unlist(lapply(datasets,rownames)))
-datasets <- lapply(datasets, function(x) {
-  temp <- matrix(0,nrow = length(common_microbes),
-                 ncol = ncol(x),
-                 dimnames = list(common_microbes,colnames(x)))
-  temp[rownames(x),colnames(x)] <- x
-  return(temp)
-})
-lapply(datasets,dim)
-
-# datasets <- datasets[3] #in case you wanna test in 1 dataset
-# Transform phyloseq data into normalized matricies ( log2[CP1K] )
-# datasets <- lapply(datasets,function(x){
-#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
-# })
-# lapply(datasets,dim)
-
-# Extract taxonomic information to file
-taxa_info <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"data/",x))
-  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
-  temp[is.na(temp)] <- "NA"
-  return(t(temp))
-})
-taxa_info <- t(as.data.frame(taxa_info))
-rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
-rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
-sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
-not_complete <- grepl("_NA",sps)
-taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
-taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
-taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
-taxa[!not_complete] <- sps[!not_complete]
-taxa <- sub("_"," ",taxa)
-use <- !is.na(taxa) & (taxa != "NA")
-taxa[taxa == "NA"] <- "other"
-rownames(taxa_info) <- taxa
-taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
-write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
-
-```
-
 # Merging, renaming and correcting batch effects on datasets
 
-```{r, echo = FALSE}
+```{r merge sequencing runs, echo = FALSE}
 datasets <- datasets_
@@
 #No batch correction needed in tissue_V3 dataset, only add it to the other samples
-datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_B2.csv"]],datasets[["ASV_tissue_B1.csv"]] )
+datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_V3_B2.csv"]],datasets[["ASV_tissue_V3_B1.csv"]] )
 
@@
 
-#Batch correction for the sample CVL_V2
-batch1 <- cbind( datasets[["ASV_CVL_V2_B1.csv"]],datasets[["ASV_CVL_V2_B2.csv"]] )
-batch1 <- t( log2( t(batch1) / colSums(batch1) * 1000 + 1 ))
-batch2 <- datasets[["ASV_CVL_V2_C.csv"]]
-batch2 <- t( log2( t(batch2) / colSums(batch2) * 1000 + 1 ))
-
-common_samples <- sort(colnames(batch1) [ colnames(batch1) %in% colnames(batch2) ])
-colnames(batch1) <- paste0("boston_",colnames(batch1))
-colnames(batch2) <- paste0("CTMR_",colnames(batch2))
-
-merged <- cbind(batch1,batch2)
-batch <- factor(sub("_.*","",colnames(merged)))
-common_samples2 <- sort(colnames(merged) [ sub(".*_","",colnames(merged)) %in% common_samples])
-
-nonzeros <- rowSums(merged)>0
-merged <- merged[nonzeros,]
-
-PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
 ```
 
-# Batch correction with different approaches
-
-```{r, echo = FALSE}
-#mypar(3,3)
-par(mar=c(3, 3, 0, 0))
-PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
-
-#COMBAT
-mod0 <- model.matrix(~1,data = data.frame(colnames(merged),row.names = colnames(merged)))
-corrected_combat <- ComBat(dat=merged, batch=batch, mod=mod0)
-
-#mnnCorrect
-corrected_mnnCorrect <- batchelor::mnnCorrect( batch2[nonzeros,],
-                                               batch1[nonzeros,] ,
-                                               cos.norm.out = F,cos.norm.in = F,
-                                         restrict = list((1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ],
-                                                         (1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ]))
-corrected_mnnCorrect <- corrected_mnnCorrect@assays@data$corrected
-
-#regressBatches
-corrected_regressBatches <- batchelor::regressBatches( batch1[nonzeros,] ,
-                                         batch2[nonzeros,],
-                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
-                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
-corrected_regressBatches <- corrected_regressBatches@assays@data$corrected
-
-#rescaleBatches
-corrected_rescaleBatches <- batchelor::rescaleBatches( batch1[nonzeros,] ,
-                                         batch2[nonzeros,],
-                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
-                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
-corrected_rescaleBatches <- corrected_rescaleBatches@assays@data$corrected
-```
-
-# Batch correction with Autoencoder
-
-```{r, echo = FALSE}
-library(keras)
-library(caret)
-
-xtrain <- t(batch1[nonzeros,paste0("boston_",sort(common_samples))])
-xtrain <- xtrain/max(xtrain)
-ytrain <- t(batch2[nonzeros,paste0("CTMR_",sort(common_samples))])
-ytrain <- ytrain/max(ytrain)
-dim(xtrain)
-
-input_size = dim(xtrain)[2]
-latent_size = 10
-print(input_size)
-
-x_train = array_reshape(xtrain, dim=c(dim(xtrain)[1], input_size))
-y_train = array_reshape(ytrain, dim=c(dim(ytrain)[1], input_size))
-
-print(dim(x_train))
-print(dim(y_train))
-
-if(file.exists(paste0(PATH,"results/encoder.h5")) & 
-  file.exists(paste0(PATH,"results/decoder.h5")) & 
-  file.exists(paste0(PATH,"results/aen.h5")) ){
-  encoder <- load_model_hdf5(paste0(PATH,"results/encoder.h5"))
-  decoder <- load_model_hdf5(paste0(PATH,"results/decoder.h5"))
-  aen <- load_model_hdf5(paste0(PATH,"results/aen.h5"))
-
-} else {
-  enc_input = layer_input(shape = input_size)
-  enc_output = enc_input %>% 
-    layer_dense(units=200, activation = "relu") %>% 
-    layer_activation_leaky_relu() %>% 
-    layer_dense(units=latent_size,activation = "sigmoid") %>% 
-    layer_activation_leaky_relu()
-
-  encoder = keras_model(enc_input, enc_output)
-  summary(encoder)
-
-  dec_input = layer_input(shape = latent_size)
-  dec_output = dec_input %>% 
-    layer_dense(units=200, activation = "relu") %>% 
-    layer_activation_leaky_relu() %>% 
-    layer_dense(units = input_size, activation = "sigmoid") %>% 
-    layer_activation_leaky_relu()
-
-  decoder = keras_model(dec_input, dec_output)
-  summary(decoder)
-
-  aen_input = layer_input(shape = input_size)
-  aen_output = aen_input %>% 
-    encoder() %>% 
-    decoder()
-
-  aen = keras_model(aen_input, aen_output)
-  summary(aen)
-
-  aen %>% compile(optimizer="rmsprop", loss="binary_crossentropy")
-
-  aen %>% fit( y_train , x_train , epochs=2000, batch_size=round(dim(y_train)[1]/2) )
-
-  save_model_hdf5(encoder, paste0(PATH,"results/encoder.h5") )
-  save_model_hdf5(decoder, paste0(PATH,"results/decoder.h5"))
-  save_model_hdf5(aen, paste0(PATH,"results/aen.h5"))
-}
-
-encoded_imgs = encoder %>% predict(y_train)
-decoded_imgs = decoder %>% predict(encoded_imgs)
-
-# plot( x_train , decoded_imgs , cex=.3)
-# abline(0,1,col="red")
-# plot( x_train, y_train , cex=.3 )
-# abline(0,1,col="red")
-```
-
-# Plotting the batch-corrected data for comparisson
-
-```{r, echo = FALSE,fig.height=7*2,fig.width=7}
-layout(matrix(c(1,1,1,1,     
-                2,2,2,3,
-                4,4,4,5,
-                6,6,6,7,
-                8,8,8,9,
-                10,10,10,11,
-                12,12,12,13
-                ) , ncol= 4,byrow = T),heights = c(1))
-
-fig_letters <- letters
-par(mar=c(3,3,3,1))
-pc_var <- PC_raw$sdev^2 / sum(PC_raw$sdev^2) * 100
-barplot(pc_var,ylim=c(0,60),las=2,names.arg = paste0("PC",1:length(pc_var)),
-        border = NA,col="grey20",ylab="% variance explained")
-abline(h=0,lwd=2)
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="uncorrected") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(batch2[nonzeros, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="ComBat") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_combat[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="mnnCorrect") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_mnnCorrect[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="regressBatches") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_regressBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="rescaleBatches") ; axis(2)
-abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(corrected_rescaleBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="AutoEncoder") ; axis(2)
-# abline(h=0,col="grey90",lty=1)
-lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
-b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:20){
-  abline(v=i,col="grey90",lty=1)
-  x <- range(PC_raw$x[,i])
-  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
-  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
-  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
-  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
-  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
-  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
-}
-add_letter(fig_letters); fig_letters <- fig_letters[-1]
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
-points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")
-
-```
-
-As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits and learns the structure of the data.
-
-```{r, echo = FALSE,fig.height=5,fig.width=5}
-datasets[["ASV_CVL_V2_raw"]] <- cbind( datasets[["ASV_CVL_V2_B2.csv"]] , datasets[["ASV_CVL_V2_B1.csv"]], datasets[["ASV_CVL_V2_C.csv"]][,!
-colnames(datasets[["ASV_CVL_V2_C.csv"]]) %in% common_samples] )
-
-batch2_all <- t(batch2[nonzeros,])
-batch2_all <- batch2_all/max(batch2_all)
-
-batch2_encoded_data <- encoder %>% predict(batch2_all)
-batch2_decoded_data <- decoder %>% predict(batch2_encoded_data)
-batch2_decoded_data <- t(round(t(t(batch2_decoded_data*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4))
-rownames(batch2_decoded_data) <- colnames(batch2_all) ; colnames(batch2_decoded_data) <- rownames(batch2_all)
-
-mypar()
-plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", main="AutoEncoder",xlim=c(-20,15),ylim=c(-15,15),type="n")
-b2_batch_projection <- scale(t(batch2_decoded_data), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-b1_projection <- scale(round(t(batch1[nonzeros,]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-
-b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
-for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey20") }
-
-points(b2_batch_projection[,1:2], cex=.8, pch=15, col= "navy")
-points(b1_projection[,1:2], cex=.8, pch=16, col= "firebrick")
-legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")
-
-samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
-datasets[["ASV_CVL_V2"]] <- as.data.frame.array(cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] ))
-datasets <- datasets[-c(1,2,3)]
-colnames(datasets[["ASV_CVL_V2"]]) <- sub(".*_","",colnames(datasets[["ASV_CVL_V2"]]))
-lapply(datasets,dim)
-lapply(datasets,colnames)
-
-common_microbes <- unique(unlist(lapply(datasets,rownames)))
-all_samples <- rownames(metadata)
-
-common_microbes.fun <- function(x) {
-  xy <- rownames_to_column(x, var = "Taxa")
-  d <- data.frame(Taxa = common_microbes) %>%
-    left_join(., xy, by="Taxa") %>%
-    mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
-    column_to_rownames(var = "Taxa")
-  return(d)
-}
-datasets <- map(datasets, ~common_microbes.fun(.x))
-lapply(datasets,dim)
+```{r save raw and norm microbiome files, echo = FALSE,fig.height=5,fig.width=5}
 
@@
 # raw files
-n <- c("ASV_CVL_V3", "ASV_tissue_V3", "ASV_CVL_V2")
-map2(datasets[1:3],n,  ~write.csv(.x, file=paste0(PATH,"results/",.y,"_raw_counts.csv"),row.names = T) )
+n <- c("ASV_CVL_V3", "ASV_tissue_V3")
+map2(datasets, n,  ~write.csv(.x, file=paste0(PATH,"results/",.y,"_raw_counts.csv"),row.names = T) )
 
 # normalized files
-datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
+datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3_B1.csv"]]) / colSums(datasets[["ASV_CVL_V3_B1.csv"]]) * 1000 + 1 )),4)
 datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
-datasets[["ASV_CVL_V2_raw"]] <- round(t( log2( t(datasets[["ASV_CVL_V2_raw"]]) / colSums(datasets[["ASV_CVL_V2_raw"]]) * 1000 + 1 )),4)
-datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)
 
-write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
 write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
 write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
-
-batches_CVL2 <- ifelse( colnames(datasets[["ASV_CVL_V2_raw"]]) %in% sub(".*_","",colnames(batch1)),"boston","CTMR" )
-batches_CVL2 <- t(t(setNames(batches_CVL2 , colnames(datasets[["ASV_CVL_V2_raw"]]))))
-write.csv( batches_CVL2 , paste0(PATH,"results/batches_CVL2.csv"),row.names = T )
 ```
 
-```{r, echo = FALSE}
+```{r save raw and norm trx files, echo = FALSE}
 trx <- read.csv(paste0(PATH,"data/TRX_rawcounts.csv"),row.names = 1)

### 2022-01-06 reports/rmarkdown/manuscript/Figure6.Rmd
Figure 6. Differential Protein expression

--- 2021-11-12 reports/rmarkdown/manuscript/Figure8.Rmd
@@
 ---
-title: "Differential Protein expression"
+title: "Figure 6. Differential Protein expression"
 output: 
@@
   library(rstatix)
-  library(broom)
   library(niceRplots)
@@
 #group <- "Luminal_gr"
+#group <- "L_T_groups"
 limma_dpe.fun <- function(group, confound) {
@@
   # DPE across all groups
-  top <- topTable(efit, adjust.method = "BH",number = "all",sort.by = "F") # ANOVA
+  top <- topTable(efit, adjust.method = "BH",number = "all") # ANOVA
   top <- cbind(LogFC.intercept=0,top) # with intercept
@@
 
+# L2 to T2/T3
+metadata <- metadata %>% 
+  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
+  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))
+
+dpe_LT <- limma_dpe.fun("L_T_groups", confound=T)
+
 # Pairwise DPE's
@@
            nrow = 3, ncol = 4,byrow = T),
-           widths = c(.9,2,.9,2), heights = c(1,1,.01) ) }
+           widths = c(1,2,.8,2), heights = c(1,1,.01) ) }
   if(layout == "single") {
@@
     ## Dendogram
-    par(mgp=c(0,0.2,0), mar=c(4,.6,1,.2)) #(bottom, left, top, right) 
+    par(mgp=c(0,0.2,0), mar=c(4,.6,2.5,.2)) #(bottom, left, top, right) 
     plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
@@
       height <-1.5 
-      add_letter("b")
+      add_letter("c")
     }else{height <- round(max(h$height))
@@
     ## Heatmap
-    par(mar=c(4,.2,1,2.5)) #(bottom, left, top, right) 
+    par(mar=c(4,.2,2.5,2.5)) #(bottom, left, top, right) 
     image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
@@
            axes=F)
+    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, line =1.5,xpd=T)
+    mtext( c("Study Groups"), side=1, las = 1, cex = .7, line = 1,xpd=T)
 
-    mtext( c("patient groups"), side=1, las = 1, cex = .7, line =1,xpd=T)
+    # Study group annotation
+    end <- table(pal[factor(names(samples))[ order(df[1])]])
+    end <- end[order(factor(names(end), levels = pal))]
+    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
+    start <- c(0, map_dbl(end[1:4], ~.x+0.015))
 
+    axis(1, at = c(start, end), label = F, pos =1.03 , xpd=T, col="white",col.ticks="black")
+    map2(start, end, ~lines(x=c(.y, .x), y= c(1.03, 1.03), xpd=T, cex=.8))
+    # text labels
+    t <- map2(start, end, ~(.x+.y)/2)
+    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.04, .y,xpd=T, cex=.8))
+
+    # Boxes 
     if(gr == "Luminal_gr") {
@@
 
-```{r Save stats to excel, include=FALSE}
+```{r Suppl. Tbl. 15 16 & 17, include=FALSE}
+names <- c("15 Luminal", "16 Tissue") %>% set_names()
+
 # save TopTables pairwise
-l_top_p <- list(Luminal = pairwise_dpe_L$top, Tissue = pairwise_dge_T$top)
-imap(l_top_p, ~write.xlsx(.x,file=paste0("../../../results/DE/","dpe_pairwise_",.y,".xlsx"))) # Excel
+l_top_p <- list(pairwise_dpe_L$top, pairwise_dge_T$top)  %>% set_names(names)
 
 # save TopTables across + cytokines
-l_top_a <- list(Luminal = list(Across = dpe_L$top), Tissue = list(Across = dpe_T$top))
+l_top_a <- list( list(Across = dpe_L$top), list(Across = dpe_T$top))  %>% set_names(names)
 l <- list(l_top_a, stat, l_top_p)
-names <- c("Luminal", "Tissue") %>% set_names()
 p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))
 
-imap(p_list, ~write.xlsx(.x,file=paste0("../../../results/DE/","Proteins_",.y,".xlsx"))) # Excel
-
+imap(p_list, ~write.xlsx(.x,file=paste0("./Suppl.Tbl/","Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
+write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0("./Suppl.Tbl/","Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
 ```
 
-```{r Proteins plot, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
+```{r Figure 6., echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
 figlabels <- letters
@@
 
-par( mar=c(1.4,4,1.5,1)) #b,l,t,r 
+par( mar=c(1.4,5,1.5,0)) #b,l,t,r 
 violist(data=Luminal$sig_df,
@@
 
-add_letter("c")
+rect(7, -0.5, 5.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
+     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
+     )
+
+add_letter("b")
 
@@
 
-par( mar=c(24.5,5,1.5,1)) #b,l,t,r
+par( mar=c(24.4,6,1.5,0)) #b,l,t,r
 violist(data=Tissue$sig_df,
@@
 
+rect(6, -0.5, 4.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
+     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
+     )
 add_letter("d")

### 2022-01-06 reports/rmarkdown/manuscript/Figure1.Rmd
Adding BV and HIV colour coding

@@
 ---
-title: "Figure 1.Luminal Patient Grouping"
+title: "Figure 1.Luminal Study Groups"
 output: 
@@
 
-  o <- order( rowSums(temp_legend) , decreasing = T)
+  o <- order( rowSums(temp) , decreasing = T)
   sample_o <- names(group_annotation)[ order( group_annotation )]
@@
 
+  l <- c("BV:","Normal","Intermediate","BV")
+  legend(x = 10, y = 111, #"bottom", inset = c(0, -0.5),
+         bty = "n", horiz=TRUE, cex=0.7, xpd = T, 
+         border = "white", 
+         legend = l, x.intersp=0.5, xjust=0,
+         text.width = c(0, map_dbl(l, ~strwidth(.x)/2)) ,
+         fill=c(NA, "#d7d7d7","orange","tomato"))
+  l2 <- c("HIV:","HIV-","HIV+")
+  legend(x = 80, y = 111, # bottom position: par("usr")[3]-5
+         bty = "n", horiz=TRUE, cex=0.7, xpd = T,
+         border = "white",
+         legend = l2, x.intersp=0.5, 
+         text.width =c(0, map_dbl(l2, ~strwidth(.x)/2)),
+         fill=c(NA, "#d7d7d7", "tomato"))
 }
@@
   mutate("bacterial communities"= factor(paste0("BC",sprintf("%02d",bac_communities$x)))) %>%
-  select(-x)
+  select(-x) %>%
+  arrange(`bacterial communities`)
 

### 2022-01-06 reports/rmarkdown/manuscript/Figure2.Rmd
Correcting spelling

@@
 mtext("Lactobacillus", side=3, las = 1, cex = .6, line = .5,xpd=T)
-mtext("Wett Smear Bacteria Validation", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
+mtext("Wet Smear Bacteria Validation", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
 

### 2022-01-06 reports/rmarkdown/manuscript/Figure3.Rmd
correcting total abundance and adding colour coding

@@
 
-  par(mar=c(1,5,1,1.5)) #b,l,t,r
+  par(mar=c(1.5,5,1,1.5)) #b,l,t,r
   temp[is.nan(temp)] <- 0
 
-  o <- order( rowSums(temp_legend) , decreasing = T)
+  o <- order( rowSums(temp) , decreasing = T)
   sample_o <- names(group_annotation)[ order( group_annotation )]
@@
            col = taxa_pal[factor(rownames(temp),levels = lvl)])
-  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
-        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation)],
-        pch=15,cex=.7,xpd=T)
-  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
+
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, 
+          bg="white", pch=15,cex=.7,xpd=T, 
+          col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation)])
+  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-7,bg="white",
         col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
@@
   text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
-  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T) # pos=4,
+  text(par("usr")[2],par("usr")[3]-7,labels = "       HIV",cex=.5,xpd=T) # pos=4,
+
+ l <- c("BV:","Normal","Intermediate","BV")
+  legend(x = 10, y = 111, #"bottom", inset = c(0, -0.5),
+         bty = "n", horiz=TRUE, cex=0.7, xpd = T, 
+         border = "white", 
+         legend = l, x.intersp=0.5, xjust=0,
+         text.width = c(0, map_dbl(l, ~strwidth(.x)/2)) ,
+         fill=c(NA, "#d7d7d7","orange","tomato"))
+  l2 <- c("HIV:","HIV-","HIV+")
+  legend(x = 70, y = 111, # bottom position: par("usr")[3]-5
+         bty = "n", horiz=TRUE, cex=0.7, xpd = T,
+         border = "white",
+         legend = l2, x.intersp=0.5, 
+         text.width =c(0, map_dbl(l2, ~strwidth(.x)/2)),
+         fill=c(NA, "#d7d7d7", "tomato"))
 }
@@
     pivot_wider(names_from = ID, values_from = value) %>%
-    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.5514392, "Other")) %>%
+    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.55, "Other")) %>%
     group_by(name) %>%
@@
 barplot( r, border = NA ,ylab="abundance", xlab ="", axisnames = F,
-         col = taxa_pal[factor(rownames(r),levels = rownames(r))])
+         col = taxa_pal[factor(rownames(r),levels = lvl)])
 title(main = "Total relative abundance ", line = 0.5, cex.main = .9, font.main= 1)

### 2022-01-06 reports/rmarkdown/manuscript/Suppl.Figure1.Rmd
Suppl. Figure 1. initial commit

---
title: "Bacterial abundance across datasets"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")

#################
# COLOR PALETTS #
#################
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
```

```{r Bacterial abundance, echo = FALSE, eval = TRUE, fig.height=3, fig.width=7}
par(mar=c(6,2.5,1,1)) #b,l,t,r

top_all <- lapply( datasets_all_samples[2:3] , function(x){
   sort(apply(x,1,var),decreasing = T)[1:20]} )
top <- unique( sub( ".*[.]" , "" , names(unlist(top_all)) ) )

x <- t(sapply(datasets_all_samples[2:3],function(x){
  c(t(x))
}))
colnames(x) <- c(sapply(rownames(datasets_all_samples[[2]]),
                        function(i){rep(i,ncol(datasets_all_samples[[2]]))}))
#rownames(x) <- gsub("ASV_|_norm.*", "",rownames(x))
rownames(x) <- c("Tissue", "Luminal")

top <- sort(sapply( unique(colnames(x)),
               function(i) { mean( x[,colnames(x)==i ] ) } ),decreasing = T)
top <- names(top)[1:30]

temp <- factor(top, levels=top)
violist(data = x[,colnames(x)%in%top],
        genes = rownames(x),
        clustering = factor( colnames(x)[colnames(x)%in%top],levels = top) ,
        srt=45, cex.axis=0.7, pt.cex=0.2,
        transparency = 50,col = pal)
```

### 2022-01-07 reports/rmarkdown/manuscript/Suppl.Figure2.Rmd
Suppl. Figure 2

---
title: "Suppl. Figure 2. Differential bacterial abundance across tissue and CVL microbiome"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))

#Graph construction  
fct <- .25 # FC threshold for differential expression
pvt <- 0.01 # Pvalue threshold for differential expression
min_pct <- .05 # minimun level of detected bacteria in each sample group

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
```

```{r Suppl. Figure 9, eval=T,warning=FALSE, message=FALSE,echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
datasets <- factor(c( rep("Tissue",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("Luminal",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    # calc % of samples that have a presence of bacteria j 
    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i) 
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
# dim(res)

write.xlsx(res, file=paste0("../../../results/","DBA_across_Luminal_vs_Tissue",".xlsx"))

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

figlabels <- letters
par(mfrow=c(1,2),mar=c(2,12,1,3.1))
#par(mar=c(2,17,1,2)) #b,l,t,r
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,
           show_grid = T,cex.main=1,font.main=1,
           cex.row = .8,cex.col = .8,srt = 0)
add_letter("a")
barlist( all_microbiome , names(sort(ord)) , clustering = datasets,
         show_grid = T,cex.main=1,font.main=1,
         cex.axis = .8,srt = 0)
add_letter("b")
```

**Suppl. Figure 2.** Differential bacterial abundance across tissue and CVL microbiome week3 datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` (from the wilcoxon test) were considered significant and were sorted by the highest expression. The colour scale indicates the difference in total abundance between the datsets as a proportion, where Max is the higest abundace of the two datasets and the other becomes a proportion of this value.The size of the dots indicates the average abundance of the given bacteria in the given dataset.

### 2022-01-07 reports/rmarkdown/manuscript/Suppl.Figure3-4.Rmd
Suppl. Figure 3 & 4

--- 2021-03-16 reports/rmarkdown/03_integrated_datasets.Rmd
@@
 ---
-title: "`r params$project_title`"
+title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
+output: 
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+header-includes: 
+- \usepackage{float}
 editor_options: 
   chunk_output_type: console
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  pdf_document:
-    toc: yes
-  html_document:
-    code_folding: hide
-    toc: yes
-    toc_float:
-      collapsed: no
-params:
-  organisation: Karolinska Institutet
-  pi_mail: prin.cipal@ki.se
-  pi_name: Prin Cipal
-  project_title: Broliden_5325
-  redmine_issue: '#5325'
-  requester_mail: kristina.broliden@ki.se
-  requester_name: Kristina Broliden
-  staff_mail: supp.ortstaff@nbis.se
-  staff_name: Supp Ortstaff
-  staff_web: https://nbis.se/about/staff/supp-ortstaff/
 ---
 
-```{r Setup, echo = FALSE}
-knitr::opts_chunk$set(fig.width  = 10,
-                      results    = "hide",
-                      message    = FALSE,
-                      warning    = FALSE)
-```
+```{r message=FALSE, warning=FALSE, include=FALSE}
+##################
+# LOAD LIBRARIES #
+##################
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+suppressWarnings({suppressMessages({suppressPackageStartupMessages({
+  library(tidyverse)
+  library(openxlsx)
+  library(fgsea)
+  library(rafalib)
+  library(Rcpp)
+  #remotes::install_github("czarnewski/niceRplots",force=T)
+  library(niceRplots)
+})  })  })
 
-#Load libraries and other scripts
-
-```{r, echo = FALSE}
-library(igraph)
-library(rafalib)
-library(fgsea)
-library(scales)
-remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
-library(niceRplots)
-```
-
-#Defining some variables for the analysis
-
-```{r, echo = FALSE}
-PATH <- "../../" #Path to the data
-
-#color palettes
-pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
-         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
-heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
-cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
-
-#Graph construction  
-fct <- .25 # FC threshold for differential expression
-pvt <- 0.01 # Pvalue threshold for differential expression
-min_pct <- .05 # minimun level of detected bacteria in each sample group
+#############
+# LODA DATA #
+#############
+#Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv") 
+datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
+PATH <- "../../.."
+SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
+bac_communities <- read.xlsx("./Suppl.Tbl/Suppl.Tbl.01.xlsx", sheet = "Bacterial Communities")
 
@@
 
-# Loading data and metadata
-
-```{r, echo = FALSE}
-# Load metadata
-metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
-
-# Load datasets
-dataset_names <- c("Tissue_RNAseq_V3_normalized",
-                     "ASV_tissue_V3_normalized_batch_corrected",
-                     "ASV_CVL_V3_normalized_batch_corrected",
-                     "ASV_CVL_V2_normalized_NOT_batch_corrected")
-
-#import datasets and return matrix with taxa-level resolution
-datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"results/",x,".csv"),row.names = 1)
-  return(as.matrix(x)) })
-names(datasets) <- dataset_names
-
-#fill all samples with 0s to make plotting easier
-datasets_all_samples <- lapply(datasets, function(x) {
-  temp <- matrix(0,nrow = nrow(x),
-                 ncol = nrow(metadata),
-                 dimnames = list(rownames(x),rownames(metadata)))
-  temp[rownames(x),colnames(x)] <- x
-  return(temp)
-})
-lapply(datasets_all_samples,dim)
+```{r echo=FALSE}
+counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
+counts[is.na(counts)] <- 0
 ```
 
-# Calculate QC metrics
+```{r, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
+myletter <- letters
+#reprod. <- FALSE
+reprod. <- TRUE
 
-\clearpage
+ASV <- list(c("Luminal","ASV_CVL_V3_normalized_batch_corrected"),
+            c("Tissue","ASV_tissue_V3_normalized_batch_corrected")
+               )
+# i <- ASV[[1]]
+for(i in ASV){
+  cat(paste0("Correlation between RNAseq and ",i[1], " 16S","\n"))
 
-```{r, echo = FALSE,fig.width=10,fig.height=5}
-counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
-counts[is.na(counts)] <- 0
-detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
-detected_counts[is.na(detected_counts)] <- 0
-rownames(detected_counts) <- paste0(rownames(detected_counts),">0")
-
-mypar(2,1,mar=c(2,10,2,1))
-barlist( data = rbind( counts ),main = "total counts",
-         genes = c(rownames(counts)), labels = sub("_norm.*","",rownames(counts)),
-         clustering = rep(1,nrow(metadata)),
-         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
-add_letter("a")
-
-barlist( data = rbind( detected_counts),main = "detected features",
-         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
-         clustering = rep(1,nrow(metadata)),
-         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
-add_letter("b")
-
-sum((colSums(counts == 0)>=3)*1)
-colnames(counts)[(colSums(counts == 0)==3)]
-paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")
-
-metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
-
-samples_to_use <- colnames(counts)[counts[c(3),] != 0]
-datasets_all_samples <- lapply( datasets_all_samples , function(x) { x [ , samples_to_use] } )
-metadata <- metadata[ samples_to_use , ]
-
-detected_counts <- detected_counts[,samples_to_use]
-barlist( data = rbind( detected_counts),main = "detected features",
-         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
-         clustering = rep(1,nrow(metadata)),
-         col="grey",draw_mean_lines=F)
-add_letter("b")
-
-saveRDS(datasets_all_samples,"../../results/datasets_all_samples.RDS")
-```
-
-**Figure 1.** Comparative barplot for the **a)** total counts and **b)** number of non-zero detected features (genes / bacteria) for each of the sequencing datasets.
-Samples are ordered alfabetically according to the patient ID.
-Samples that are present in exactly two datasets are shown in orange (`r paste0(colnames(counts)[(colSums(counts == 0)==2)],collapse = ",")`).
-Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`).
-
-```{r, echo = FALSE,fig.width=10,fig.height=9}
-# all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-#                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-# all_microbiome <- t( t(all_microbiome) / colSums(all_microbiome) )
-all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
-# all_microbiome <- all_microbiome + rnorm( length(all_microbiome) , mean = 0, sd = .00000001)
-dim(all_microbiome)
-
-zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
-zscore[ zscore > 4 ] <- 4
-zscore[ zscore < -4 ] <- -4
-rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
-mypar(mar=c(2,15,2,2))
-top_var <- apply(all_microbiome,1,var)
-top_var_t <- sort(top_var,decreasing = T)[100]
-is_top_var <- top_var > top_var_t 
-barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
-
-#Compute sample-wise correlations
-cors <- cor(all_microbiome[,], method = "pearson")
-adj <- (1-cors)/2
-# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
-
-# cors <- cor(all_microbiome, method = "pearson")
-# adj <- (1-cors)/2
-# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
-
-#Graph construction  
-k <- 10 # k for KNN
-p <- 0.1 # prunning threshold of SNN
-
-#Compute KNN graph from correlation matrix
-knn <- RANN::nn2(adj,k = k)
-nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
-for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
-
-#Computing SNN matrix
-N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
-P <- t(nn) %*% nn
-B <- N - nn
-C <- t(B) %*% B
-Q <- nrow(nn)*N - C
-J <- P/Q
-J[is.nan(J)] <- 0
-J[J < p] <- 0
-colnames(J) <- colnames(all_microbiome)
-rownames(J) <- colnames(all_microbiome)
-
-#Transforming SNN into an igraph object
-g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
-write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)
-
-mypar(3,3,mar=c(0,0,2,0))
-
-set.seed(1)
-set.seed(1)
-l <- layout_with_fr(g,niter=3000,start.temp=30)
-
-cl <- igraph::cluster_louvain(g)
-# cl <- igraph::walktrap.community(g)
-
-metadata$joint_clustering <- cl$membership
-set.seed(1)
-U <- uwot::umap(adj,n_neighbors = 30)
-h <- hclust(as.dist(adj),method = "ward.D2")
-# metadata$joint_clustering <- cutree(h,k = 5)
-write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,main="FR embedding\nlouvain",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
-      vertex.size=10,main="umap",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-empty_plot()
-legend("topleft",title = "Patient groups",
-       legend = levels(factor(cl$membership)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,main="FR embedding",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M9_CVLv3)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
-      vertex.size=10,main="umap",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-empty_plot()
-legend("topleft",title = "Patient groups\ngroup_M9_CVLv3",
-       legend = levels(factor(metadata$group_M9_CVLv3)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
-# title("Louvain clusters",cex.main=1,font.main=1)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,main="FR embedding",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$group_M4_CVLv3)] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
-      vertex.size=10,main="umap",
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-empty_plot()
-legend("topleft",title = "Patient groups\ngroup_M4_CVLv3",
-       legend = levels(factor(metadata$group_M4_CVLv3)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
-# title("Louvain clusters",cex.main=1,font.main=1)
-
-```
-
-**Figure 1b.** Sample embedding of `r k`-SNN graph clustered using Louvain based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).
-
-```{r}
-mypar(mar=c(2,15,2,2))
-
-barlist( data = rbind( detected_counts),main = "detected features",
-         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
-         clustering = metadata$joint_clustering,
-         col=pal,draw_mean_lines=F)
-add_letter("b")
-
-barlist( data = all_microbiome, main = "detected features",
-         genes = c("Gardnerella vaginalis",
-                   "Lactobacillus iners",
-                   "Lactobacillus jensenii",
-                   "Lactobacillus crispatus/acidophilus",
-                   "Prevotella amnii",
-                   "Prevotella timonensis",
-                   "Atopobium vaginae"),
-         clustering = metadata$joint_clustering,
-         col=pal,draw_mean_lines=F)
-
-```
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=5,fig.height=5}
-x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
-xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
-xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))
-
-mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
-mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
-mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0
-
-mypar(3,1,mar=c(2,6,2,2))
-barlist( data = rbind(  BV=mL ,
-                        counts_16S=xL),
-         main = "Lactobacillus",
-         genes = c("BV","counts_16S"),
-         clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
-
-barlist( data = rbind(  BV=mM ,
-                        counts_16S=xM),
-         main = "Monbilicus",
-         genes = c("BV","counts_16S"),
-         clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
-
-barlist( data = rbind(  BV=mG ,
-                        counts_16S=xG),
-         main = "Gardanerella",
-         genes = c("BV","counts_16S"),
-         clustering = metadata$joint_clustering,
-         draw_mean_lines=F,col = pal)
-```
-
-\clearpage
-
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=4}
-myletters <- letters
-
-mypar(1,2,mar=c(2,15,1,1))
-
-for (m in c("ASV_tissue_V3_normalized_batch_corrected",
-            "ASV_CVL_V3_normalized_batch_corrected",
-            "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-
-  all_microbiome <- datasets_all_samples[[m]]
-
-  datasets <- factor(metadata$joint_clustering)
-  # datasets <- datasets[ colSums(all_microbiome)>0 ]
-  # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-  NN <- min(table(datasets))
-
-  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-  for(i in levels(datasets)){
-    for(j in rownames(all_microbiome) ){
-
-      set.seed(1)
-      a <- c(sample(all_microbiome[j,datasets == i],NN))
-      set.seed(1)
-      b <- sample(all_microbiome[j,datasets != i],NN)
-
-      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
-      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
-
-      temp <- wilcox.test(x=a, y=b)
-
-      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                                 c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-    }
-  }
-  res <- res[-1,]
-  res$pvalue <- as.numeric(res$pvalue)
-  res$perc.1 <- as.numeric(res$perc.1)
-  res$perc.2 <- as.numeric(res$perc.2)
-  res$perc.diff <- res$perc.1 - res$perc.2
-
-  res$fc <- as.numeric(res$fc)
-  res$FDR <- p.adjust(res$pvalue)
-  # res <- res[order(res$pvalue),]
-  res <- res[abs(res$fc) > fct,]
-  res <- res[abs(res$pvalue) < 0.05,]
-  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-  dim(res)
-
-  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
-
-  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-            cex.row = .8,cex.col = .8,srt = 0,main = m)
-  add_letter(myletters); myletters <- myletters[-1]
-
-  barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-            cex.axis = .8,srt = 0,main = m)
-  add_letter(myletters); myletters <- myletters[-1]
-
-}
-```
-
-**Figure 2.** Differential bacterial abundance across joint clustered samples and compared across microbiome datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
-
-\clearpage
-
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
-datasets <- datasets[ colSums(all_microbiome)>0 ]
-all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
-NN <- min(table(datasets))
-
-res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-for(i in levels(datasets)){
-  for(j in rownames(all_microbiome) ){
-
-    a <- all_microbiome[j,datasets == i]
-    b <- all_microbiome[j,datasets != i]
-
-    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
-    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
-
-    temp <- wilcox.test(x=a, y=b)
-
-    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-  }
-}
-res <- res[-1,]
-res$pvalue <- as.numeric(res$pvalue)
-res$perc.1 <- as.numeric(res$perc.1)
-res$perc.2 <- as.numeric(res$perc.2)
-res$perc.diff <- res$perc.1 - res$perc.2
-
-res$fc <- as.numeric(res$fc)
-res$FDR <- p.adjust(res$pvalue)
-res <- res[order(res$pvalue),]
-res <- res[abs(res$fc) > fct,]
-res <- res[abs(res$pvalue) < pvt,]
-res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-dim(res)
-
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
-
-mypar(1,2,mar=c(2,15,1,1))
-plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.row = .8,cex.col = .8,srt = 0)
-add_letter("a")
-barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.axis = .8,srt = 0)
-add_letter("b")
-```
-
-**Figure 2.** Differential bacterial abundance across microbiome datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
-
-\clearpage
-
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- factor(c( rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
-datasets <- datasets[ colSums(all_microbiome)>0 ]
-all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
-NN <- min(table(datasets))
-
-res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-for(i in levels(datasets)){
-  for(j in rownames(all_microbiome) ){
-
-    a <- all_microbiome[j,datasets == i]
-    b <- all_microbiome[j,datasets != i]
-
-    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
-    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
-
-    temp <- wilcox.test(x=a, y=b)
-
-    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-  }
-}
-res <- res[-1,]
-res$pvalue <- as.numeric(res$pvalue)
-res$perc.1 <- as.numeric(res$perc.1)
-res$perc.2 <- as.numeric(res$perc.2)
-res$perc.diff <- res$perc.1 - res$perc.2
-
-res$fc <- as.numeric(res$fc)
-res$FDR <- p.adjust(res$pvalue)
-res <- res[order(res$pvalue),]
-res <- res[abs(res$fc) > fct,]
-res <- res[abs(res$pvalue) < pvt,]
-res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-dim(res)
-
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
-
-mypar(1,2,mar=c(2,15,1,1))
-plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.row = .8,cex.col = .8,srt = 0)
-add_letter("a")
-barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.axis = .8,srt = 0)
-add_letter("b")
-```
-
-**Figure 2.** Differential bacterial abundance across CVL2 and CVL3 microbiome datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
-
-\clearpage
-
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
-datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
-datasets <- datasets[ colSums(all_microbiome)>0 ]
-all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
-NN <- min(table(datasets))
-
-res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-for(i in levels(datasets)){
-  for(j in rownames(all_microbiome) ){
-
-    a <- all_microbiome[j,datasets == i]
-    b <- all_microbiome[j,datasets != i]
-
-    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
-    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
-
-    temp <- wilcox.test(x=a, y=b)
-
-    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-  }
-}
-res <- res[-1,]
-res$pvalue <- as.numeric(res$pvalue)
-res$perc.1 <- as.numeric(res$perc.1)
-res$perc.2 <- as.numeric(res$perc.2)
-res$perc.diff <- res$perc.1 - res$perc.2
-
-res$fc <- as.numeric(res$fc)
-res$FDR <- p.adjust(res$pvalue)
-res <- res[order(res$pvalue),]
-res <- res[abs(res$fc) > fct,]
-res <- res[abs(res$pvalue) < pvt,]
-res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-dim(res)
-
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
-
-mypar(1,2,mar=c(2,15,1,1))
-plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.row = .8,cex.col = .8,srt = 0)
-add_letter("a")
-barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.axis = .8,srt = 0)
-add_letter("b")
-```
-
-**Figure 2b.** Differential bacterial abundance across tissue and CVL microbiome week3 datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
-
-\clearpage
-
-```{r eval=FALSE, include=FALSE}
-groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
-  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
-names(groupings) <- names(datasets_all_samples)[-1]
-all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))
-
-groupings_all_samples <- lapply(groupings,function(x){
-  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
-  df[rownames(x),1] <- as.character(x[,1])
-  return(df)
-})
-names(groupings_all_samples) <- names(groupings)
-
-df <- as.data.frame(groupings_all_samples)
-colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
-metadata <- cbind(metadata,df[rownames(metadata),])
-```
-
-# Computing differential expression across microbiome datasets
-
-```{r, echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
-datasets <- factor(c( rep("tissue_v3",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("CVL_v3",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) , rep("CVL_v2",ncol(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])) ))
-
-datasets <- datasets[ colSums(all_microbiome)>0 ]
-all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
-NN <- min(table(datasets))
-```
-
-```{r, eval=F,echo = FALSE,fig.width=10,fig.height=12}
-res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-for(i in levels(datasets)){
-  for(j in rownames(all_microbiome) ){
-
-    a <- all_microbiome[j,datasets == i]
-    b <- all_microbiome[j,datasets != i]
-
-    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
-    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)
-
-    temp <- wilcox.test(x=a, y=b)
-
-    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-  }
-}
-res <- res[-1,]
-res$pvalue <- as.numeric(res$pvalue)
-res$perc.1 <- as.numeric(res$perc.1)
-res$perc.2 <- as.numeric(res$perc.2)
-res$perc.diff <- res$perc.1 - res$perc.2
-
-res$fc <- as.numeric(res$fc)
-res$FDR <- p.adjust(res$pvalue)
-res <- res[order(res$pvalue),]
-res <- res[abs(res$fc) > fct,]
-res <- res[abs(res$pvalue) < pvt,]
-res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-dim(res)
-
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
-
-mypar(1,1,mar=c(5,15,1,1))
-plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
-          cex.row = .7,cex.col = .8,srt = 30)
-add_letter("a")
-```
-
-**Figure 4.** Differential bacterial abundance across all groups and all microbiome datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.
-
-```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}
-sample_order <- order(df$louvain_CVL_V3)
-
-# zscore_datasets_all_samples <- lapply(datasets_all_samples,function(x){
-#   zscore <- t(apply(x,1,function(i){scale(i,T,T)}))
-#   zscore[ zscore > 6 ] <- 6
-#   zscore[ zscore < -6 ] <- -6
-#   rownames(zscore) <- rownames(x); colnames(zscore) <- colnames(x)
-#   zscore[is.nan(zscore)] <- -6
-#   return(zscore)
-# })
-
-mypar(3,1,mar=c(1,2,1,0.5))
-
-image(t(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
-        [ names(sort(ord)),sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_tissue_V3)), pch=16,
-        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_tissue_V3)), pch=16,
-        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-
-image(t(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
-        [ names(sort(ord)),sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V3)), pch=16,
-        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V3)), pch=16,
-        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-
-image(t(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
-        [ names(sort(ord)),sample_order]),
-      col=heat_pal,axes=F,border=NA,
-      main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V2)), pch=16,
-        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-            length(df$louvain_CVL_V2)), pch=16,
-        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-
-# **Figure 5.** Comparisson of microbiome datasets, showing only the significant bacteria.
-Samples are ordered by the CVL3 groupings.
-The colors represent their respective bacterial groupings for each dataset.
-
-```
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=8,fig.height=8}
-metadata_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
-                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
-                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
-                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
-                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
-                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
-                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
-                         "blood_cervix_v2","blood_cervix_v3",
-                         "cervical_inflammation_v2","cervical_inflammation_v3",
-                         "HIVstatus","Contraception",
-                         "CT_v2","CT_v3","regular_partner",
-                         "vaginal_douching_v2","vaginal_douching_v3","joint_clustering")
-
-Ftest <- sapply(metadata_parameters,function(x){
-  cat(paste0("\n",x,"\n"))
-  temp <- sapply(metadata_parameters,x=x,function(x,y){
-    if (x==y){
-      return( 1 )
-    } else {
-      # if( !is.na(sum(as.numeric(metadata[,x]))) ){ f1 <- cut(as.numeric(metadata[,x]),breaks = 5 ) }
-      if( is.numeric(metadata[,x]) ){ f1 <- cut(metadata[,x],breaks = 5 ) }
-    f1 <- as.character(metadata[,x])
-    f1[is.na(f1)] <- "NA"
-    f1[f1==""] <- "NA"
-
-    # if( !is.na(sum(as.numeric(metadata[,y]))) ){ f1 <- cut(as.numeric(metadata[,y]),breaks = 5 ) }
-    if( is.numeric(metadata[,y]) ){ f2 <- cut(metadata[,y],breaks = 5 ) }
-    f2 <- as.character(metadata[,y])
-    f2[is.na(f2)] <- "NA"
-    f2[f2==""] <- "NA"
-
-    f11 <- f1 [ (f1!="NA") & (f2!="NA") ]
-    f22 <- f2 [ (f1!="NA") & (f2!="NA") ]
-    contT1  <- table(list(factor(f11,levels=unique(f1)),factor(f22,levels=unique(f2))))
-    cat(".")
-    f <- NULL
-    try(f <- fisher.test(contT1,simulate.p.value = T),silent = T)
-    if( is.null(f) ){ 
-      return( 1  ) 
-      } else {
-      return( c(f$p.value)  ) 
-    }
-    # return( chisq.test(contT1,simulate.p.value = T)$p.value )
-    }
-  })
-
-  return( temp )
-})
-
-clean_Ftest <- Ftest
-# clean_Ftest <- ( clean_Ftest + t(clean_Ftest) ) / 2
-clean_Ftest <- -log10(clean_Ftest)
-clean_Ftest[clean_Ftest > 6] <- 6
-# clean_Ftest[ clean_Ftest > 0.05 ] <- 1
-
-write.csv2(Ftest , "../../results/metadata_association_results_complete.csv",row.names = T)
-write.csv2(clean_Ftest , "../../results/metadata_association_results_filtered.csv",row.names = T)
-
-mypar(1,1,mar=c(12,1,1,12))
-image( t(clean_Ftest [nrow(Ftest):1,]),axes=F,
-       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
-text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(Ftest)) , 
-       labels = rownames(Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-text( seq(0,1,length.out = nrow(Ftest)) , par("usr")[3]-.02 , 
-      labels = rownames(Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-```
-
-**Figure 6.** Association analysis across several patient categorical metadata parameters, including patient groupings annotations from microbiome.
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=10}
-mypar(6,4,mar=c(1,4,1,4) )
-x <- as.character(metadata[,"Contraception"])
-x[is.na(x)] <- "NA"
-y <- as.character(metadata[,"Antibiotics"])
-y[is.na(y)] <- "NA"
-x1 <- x [ (x!="NA") & (y!="NA") ]
-y1 <- y [ (x!="NA") & (y!="NA") ]
-plot_sankey( data.frame(x1,y1),color_by = 2,
-             plot_labels = T, plot_weights = F,order_1_by = "total",
-             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(10))
-text(c(0,1),c(1,1),labels = c( "Contraception" , "Antibiotics" ),pos=3,xpd=T)
-
-x <- as.character(metadata[,"BV_Monbilicus_v2"])
-x[is.na(x)] <- "NA"
-y <- as.character(metadata[,"Antibiotics"])
-y[is.na(y)] <- "NA"
-x1 <- x [ (x!="NA") & (y!="NA") ]
-y1 <- y [ (x!="NA") & (y!="NA") ]
-plot_sankey( data.frame(x1,y1),color_by = 2,
-             plot_labels = T, plot_weights = F,order_1_by = "total",
-             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(10))
-text(c(0,1),c(1,1),labels = c( "BV_Monbilicus_v2" , "Antibiotics" ),pos=3,xpd=T)
-
-x <- as.character(metadata[,"HIVstatus"])
-x[is.na(x)] <- "NA"
-y <- as.character(metadata[,"Antibiotics"])
-y[is.na(y)] <- "NA"
-x1 <- x [ (x!="NA") & (y!="NA") ]
-y1 <- y [ (x!="NA") & (y!="NA") ]
-plot_sankey( data.frame(x1,y1),color_by = 2,
-             plot_labels = T, plot_weights = F,order_1_by = "total",
-             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(10))
-text(c(0,1),c(1,1),labels = c( "HIVstatus" , "Antibiotics_v2" ),pos=3,xpd=T)
-
-x <- as.character(metadata[,"blood_cervix_v3"])
-x[is.na(x)] <- "NA"
-y <- as.character(metadata[,"cervical_inflammation_v3"])
-y[is.na(y)] <- "NA"
-x1 <- x [ (x!="NA") & (y!="NA") ]
-y1 <- y [ (x!="NA") & (y!="NA") ]
-plot_sankey( data.frame(x1,y1),
-             plot_labels = T, plot_weights = F,order_1_by = "total",
-             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(9))
-text(c(0,1),c(1,1),labels = c( "blood_cervix_v3" , "cervical_inflammation_v3" ),pos=3,xpd=T)
-
-```
-
-**Figure 7.** A few examples of significant association between metadata parameters shown as sankey plots.
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=10}
-mypar(min(length(metadata_parameters),6),4,mar=c(1,4,1,4))
-
-# metadata$Contraception
-#Antibiotics usage
-for(i in metadata_parameters){
-  empty_plot()
-  text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
-  for(j in c("joint_clustering")){
-    x <- as.character(metadata[,i])
-    x[is.na(x)] <- "NA"
-    y <- as.character(metadata[,j])
-    y[is.na(y)] <- "NA"
-    x1 <- x [ (x!="NA") & (y!="NA") ]
-    y1 <- y [ (x!="NA") & (y!="NA") ]
-    plot_sankey( data.frame(x1,y1),
-             plot_labels = T, plot_weights = F,order_1_by = "total",
-             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
-             pal=hue_pal()(13))
-    text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
-  }
-}
-```
-
-**Figure 8.** Sankey plots for all tested associations between the patient groups identified in in the microbiome datasets.
-
-\clearpage
-
-```{r, eval=F, echo = FALSE,fig.width=10,fig.height=13}
-myletter <- letters
-
-for(i in c("ASV_tissue_V3_normalized_batch_corrected",
-           "ASV_CVL_V3_normalized_batch_corrected",
-           "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-
-  cat(paste0("Correlation between RNAseq and ",sub("_norm.*","",i)))
+  # because hclust is non reproducible 
+  if(reprod. == TRUE){
+    NESes <- read.csv2(paste0("../../../results/",i[1],"_TRX_Bact_norm_enrichment_scores",".csv"), row.names = 1)
+  }else{
 
     TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-  bac_dataset <- datasets_all_samples[[i]]
-  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
-  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]
+    bac_dataset <- datasets_all_samples[[i[2]]]
+    samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i[2]),] > 0)==2 ]
 
-  # mypar(4,3,mar=c(4,2,2,2))
     #filter TRX dataset
@@
     TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
+    # only top 5000 most variabe genes are included in the correlation
     top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
-  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
     TRX <- TRX[ top_vars_TRX , ]
@@
     bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
+    # keeping taxa that is present in at least two samples :
     bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
@@
     cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
-  # hist(cors,xlim=c(-1,1),breaks = 100)
-  dim(cors)
 
     bacteria_use <- colnames(cors)
-  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
-  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
-  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
-  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T))
+    gmt <- gmtPathways(paste0(PATH, "/supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt"))
+
+    enrichments <- lapply(bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
+      res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
       return(res)
     })
-  parallel::stopCluster(cl)
     names(enrichments) <- bacteria_use
@@
     top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
-  # mypar(mar=c(4,15,2,2))
-  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )
 
-  # pvalues[ pvalues < -log10(0.05)] <- 0
-  # sort(rowSums(pvalues>0),T)
-  # sort(colSums(pvalues>0),T)
     NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
@@
     NESes[is.na(NESes)] <- 0
+    # replace all NESes values with 0 if they don't have a significant p-value < 0.05:
     NESes[ pvalues < -log10(0.05) ] <- 0 
+    # filter out bacterium and pathways with <10 significant NES scores 
     NESes <- NESes[rowSums(NESes!=0) >= 10 , colSums(NESes!=0) >= 10]
 
-  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
-  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
+    set.seed(1)
     o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
     o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
+    NESes <- NESes[o_bacs, o_kegg]
 
-  NESes <- NESes[o_bacs,o_kegg]
+    bact <- gsub("\\.{1}", "/", rownames(NESes))
+    bact <- gsub("\\/{2}", ". ", bact)
+    rownames(NESes) <- bact
 
-  mypar(1,1,mar=c(12,2,2,12))
+    #write.csv2(NESes, paste0("../../../results/",i[1],"_TRX_Bact_norm_enrichment_scores",".csv"))
+
+  }
+
+  terms <- gsub("_", " ", colnames(NESes))
+  terms <- gsub("(GO |KEGG )", "", terms, perl = TRUE)
+  colnames(NESes) <- terms
+
+  par(mfrow=c(1,1),mar=c(12,0,2,4)) #b,l,t,r
   image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
@@
         labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
-  add_letter(myletter); myletter <- myletter[-1]
-  # empty_plot()
-  cat("\\clearpage")
 
@@
 
-# **Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+```
+
+**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
 Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
@@
 The heatmap shows the normalized enrichment score (NES).
-Only enrichments with pvalue below 0.05 are shown.
-Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.
+Only enrichments with pvalue below 0.05 are shown.Only bacterium and pathways with at least 10 significant NES scores were included in the heatmap.
 
-```
+```{r eval=FALSE, include=FALSE}
+g <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
+cl <- igraph::cluster_louvain(g)
+cors <- cor(t(datasets_all_samples[[i]][rowSums(datasets_all_samples[[i]] > 0)>=10,]))
 
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=10}
+module_means <- bac_communities %>%
+  left_join(as_tibble(datasets_all_samples[[i]], rownames = "Taxa"), by="Taxa")  %>%
+  group_by(bacterial.communities) %>%
+  summarise(across(where(is.numeric), ~sum(.x))) %>%
+  mutate(across(where(is.numeric), ~.x/c(table(bac_communities$bacterial.communities))) ) %>%
+  select(-bacterial.communities) %>%
+  #column_to_rownames(var="bacterial.communities") %>%
+  as.matrix() %>%
+  `rownames<-`(unique( bac_communities$bacterial.communities)) 
+bact_df <- module_means
 
@@
                         datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
+#cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
 
-cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))
+module_means_per_dataset <- lapply(datasets_all_samples[[i]],function(x){
+  return( rowsum( x[bac_communities$Taxa,] , group = bac_communities$bacterial.communities ) / c(table(bac_communities$bacterial.communities)) )
+})
 
-#Graph construction 
-k_nearest_neighbours <- 5 # k for KNN
-snn_threshold <- 0.1 # prunning threshold of SNN
+module_means <- bac_communities %>%
+  left_join(as_tibble(datasets_all_samples[[i]], rownames = "Taxa"), by="Taxa")  %>%
+  group_by(bacterial.communities) %>%
+  summarise(across(where(is.numeric), ~sum(.x))) %>%
+  mutate(across(where(is.numeric), ~.x/c(table(bac_communities$bacterial.communities))) )
 
-#Compute kNN graph
-knn <- RANN::nn2( (1-cors)/2 ,k = k_nearest_neighbours)
-nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
-for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
-
-#Calculate SNN and prune based on Jaccard index
-N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
-P <- t(nn) %*% nn
-B <- N - nn
-C <- t(B) %*% B
-Q <- nrow(nn)*N - C
-J <- P/Q
-J[is.nan(J)] <- 0
-J[J < snn_threshold] <- 0
-rownames(J) <- rownames(cors)
-colnames(J) <- colnames(cors)
-
-#Build a graph object, perform clustering and define the force-directed layout
-g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
-write.csv(J,"../../results/bacteria_SNN_graph.csv",row.names = T)
-
-cl <- igraph::cluster_louvain(g)
-# cl <- igraph::walktrap.community(g)
-set.seed(1)
-l <- layout_with_fr(g,niter=3000,start.temp=1)
-
-#Define layout with UMAP
-mU <- uwot::umap((1-cors)/2, n_neighbors = 10)
-
-#Plot
-mypar(3,3,mar=c(0,0,2,0))
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
-    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-    vertex.size=10, main="graph layout FR",
-    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
-    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
-    vertex.size=10, main="UMAP",
-    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-par(mar=c(0,0,0,0))
-empty_plot()
-legend("topleft",title = "bact. communities",
-     legend = levels(factor(cl$membership)),
-     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
+b  <-  datasets_all_samples[[i]][bac_communities$Taxa,]
+rowsum(datasets_all_samples[[i]][bac_communities$Taxa,] , group = bac_communities$bacterial.communities )
 ```
 
-**Figure 9a.** Bacterial community embedding of `r k`-SNN graph clustered using walktrap community detection algorithm based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).
-
-```{r, echo = FALSE,fig.width  = 10, fig.height=12}
-mypar(3,6,mar=c(1,1,2,1))
-for(j in sort(unique(cl$membership)) ){
-  empty_plot()
-  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
-  title(main=paste0("community",j))
-}
-write.csv2(cl$membership,"../../results/bacterial_communities.csv",row.names = rownames(cors))
-```
-
-**Figure 9b.** List of bacterial belongin to each community.
-
-```{r, echo = FALSE,fig.width  = 10, fig.height=5}
-taxa <- read.csv2("../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
-taxa[is.na(taxa)] <- "NA"
-rownames(cors) %in% rownames(taxa)
-
-mypar(2,2,mar=c(1,10,2,10))
-x <- taxa[match(rownames(cors),rownames(taxa)),2]
-y <- taxa[match(rownames(cors),rownames(taxa)),3]
-plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
-       plot_labels = T, plot_weights = F,order_1_by = "total",
-       order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-       pal=hue_pal()(length(unique(x))))
-
-plot_sankey( data.frame( y[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
-     plot_labels = T, plot_weights = F,order_1_by = "total",
-     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-     pal=hue_pal()(length(unique(y))))
-
-plot_sankey( data.frame( x[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
-     plot_labels = T, plot_weights = F,order_1_by = "total",
-     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
-     pal=hue_pal()(length(unique(x))))
-
-```
-
-**Figure 9b.** Taxonomic annotation of bacterial communities.
-
-```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}
-
-module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
-  return( rowsum( x[rownames(cors),] , group = cl$membership ) / c(table(cl$membership)) )
-})
-
-mypar(4,3,mar=c(3,5,2,2))
-
-# mypar(3,1,mar=c(1,2,1,3))
-# sample_order <- order(df$louvain_tissue_V3)
-# image(t(module_means_per_dataset[["ASV_tissue_V3_normalized_batch_corrected"]]
-#         [ ,sample_order]),
-#       col=heat_pal,axes=F,border=NA,
-#       main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_tissue_V3)), pch=16,
-#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
-#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_tissue_V3)), pch=16,
-#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
-# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
-#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-# 
-# 
-# sample_order <- order(df$louvain_CVL_V3)
-# image(t(module_means_per_dataset[["ASV_CVL_V3_normalized_batch_corrected"]]
-#         [ ,sample_order]),
-#       col=heat_pal,axes=F,border=NA,
-#       main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_CVL_V3)), pch=16,
-#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
-#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_CVL_V3)), pch=16,
-#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
-# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
-#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-# 
-# 
-# sample_order <- order(df$louvain_CVL_V2)
-# image(t(module_means_per_dataset[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
-#         [ ,sample_order]),
-#       col=heat_pal,axes=F,border=NA,
-#       main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
-# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_CVL_V2)), pch=16,
-#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
-#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
-#             length(df$louvain_CVL_V2)), pch=16,
-#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
-# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
-#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-
-mypar(3,3,mar=c(3,5,2,2))
-
-for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
-  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-          "ASV_CVL_V3_normalized_batch_corrected",
-          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-    temp_group <- as.character(df[,i])
-    temp_group[ is.na(temp_group) ] <- "NA"
-    temp_group[ temp_group == "" ] <- "NA"
-    temp_group <- factor(temp_group)
-
-    barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),col = pal,
-            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
-            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
-  }
-}
-
-```
-
-```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}
-
-module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
-  return( rowsum( x[rownames(cors),] , group = cl$membership ) / c(table(cl$membership)) )
-})
-
-mypar(3,3,mar=c(3,5,2,2))
-
-for( i in c("group_M4_CVLv3","group_M9_CVLv3","joint_clustering")){
-  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
-          "ASV_CVL_V3_normalized_batch_corrected",
-          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
-
-    temp_group <- as.character(metadata[,i])
-    temp_group[ is.na(temp_group) ] <- "NA"
-    temp_group[ temp_group == "" ] <- "NA"
-    temp_group <- factor(temp_group)
-
-    barlist(data=module_means_per_dataset[[j]],
-            genes=rownames(module_means_per_dataset[[j]]),col=pal,
-            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
-            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
-  }
-}
-```
-
-**Figure 5.** Comparisson of microbiome communities across all datasets and all clustering methods.
-
-```{r, echo = FALSE,fig.width=10,fig.height=8}
+```{r eval=FALSE, fig.height=8, fig.width=10, include=FALSE}
 myletter <- letters
@@
   bacteria_use <- colnames(cors)
-  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
+  #gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
+  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
   cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
@@
 ```
-
-**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
-Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
-Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the Hallmark MSigDB gene annotation database.
-This, in turn, will result in a matrix associating every bacteria with every Hallmark MSigDB process in the tissue.
-The heatmap shows the normalized enrichment score (NES).
-Only enrichments with pvalue below 0.05 are shown.
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=8}
-myletter <- letters
-
-NESse_list <- list()
-pvalues_list <- list()
-
-for(i in c("ASV_tissue_V3_normalized_batch_corrected",
-           "ASV_CVL_V3_normalized_batch_corrected")){
-
-  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))
-
-  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
-  bac_dataset <- module_means_per_dataset[[i]]
-  samples_TRX_and_CVL3 <- colnames(bac_dataset)
-
-  # mypar(4,3,mar=c(4,2,2,2))
-  #filter TRX dataset
-  TRX <- TRX [ , samples_TRX_and_CVL3 ]
-  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]
-
-  #filter bac_dataset dataset
-  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
-  dim(bac_dataset)
-
-  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
-  dim(cors)
-
-  bacteria_use <- colnames(cors)
-  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
-  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
-  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
-  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
-    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
-    return(res)
-  })
-  parallel::stopCluster(cl)
-  names(enrichments) <- paste0( "bac.com.", bacteria_use )
-
-  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
-  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)
-
-  pvalues <- lapply(enrichments,function(x){
-    temp <- allpvals
-    temp[x$pathway] <- x$pval
-    return(temp)
-  })
-  pvalues <- t(as.data.frame(pvalues))
-  pvalues <- -log10( pvalues )
-  pvalues[is.na(pvalues)] <- 0
-
-  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
-  NESes <- lapply(enrichments,function(x) {
-    temp <- allNES
-    temp[x$pathway] <- x$NES
-    return(temp)
-  })
-  NESes <- t(as.data.frame(NESes))
-  NESes[is.na(NESes)] <- 0
-  # NESes[ pvalues < -log10(0.1) ] <- 0
-
-  NESse_list[[i]] <- NESes
-  pvalues_list[[i]] <- pvalues
-}
-
-write.csv2(NESse_list,"../../results/Kegg_NESse_list.csv")
-write.csv2(pvalues_list,"../../results/Kegg_pvalues_list.csv")
-
-common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[colSums(x!=0)>=1] ) ) )
-
-layout(matrix(c(1,2,3,4,4),nrow = 1))
-par(mar=c(6,1,2,0))
-for( i in names(NESse_list)){
-  x <- NESse_list[[i]][,common_pathways]
-  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
-         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-  # add_letter(myletter); myletter <- myletter[-1]
-  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-}
-empty_plot(xaxs="i",yaxs="i",
-           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
-text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
-      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-
-cat("\\clearpage")
-
-#Combined representation of the two datasets
-layout(matrix(c(1,2,2,2,2),nrow = 1))
-par(mar=c(6,1,3,0))
-x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
-x[NESse_list[[1]][,common_pathways] == 0] <- 0
-x[NESse_list[[2]][,common_pathways] == 0] <- 0
-x <- x[, colSums(x!=0)>0]
-image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
-       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-# add_letter(myletter); myletter <- myletter[-1]
-text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-empty_plot(xaxs="i",yaxs="i",
-           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
-text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
-      labels = rev(sub("KEGG_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-
-#Combined representation of the two datasets
-layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
-par(mar=c(6,30,3,0))
-x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
-x[NESse_list[[1]][,common_pathways] == 0] <- 0
-x[NESse_list[[2]][,common_pathways] == 0] <- 0
-x <- x[, colSums(x!=0)>0]
-
-TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-non_zeros <- colSums(TRX)!=0
-TRX <- TRX [ , non_zeros ]
-pathway_means <- t(sapply( colnames(x) , function(i){
-  i <- gmt[[i]]
-  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
-  if(ncol(i)>1){
-    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
-    i <- colMeans(i)
-  } else {
-    i <- (i - min(i)) / (max(i) - min(i))
-  }
-  return( i - min(i) )
-} ))
-plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
-        genes=rownames(pathway_means),
-        clustering=factor(metadata$joint_clustering[non_zeros]),
-        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))
-
-par(mar=c(6,2,3,2))
-image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
-       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
-# add_letter(myletter); myletter <- myletter[-1]
-text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
-    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
-# empty_plot(xaxs="i",yaxs="i",
-#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
-# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
-      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
-```
-
-**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
-Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
-Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database.
-This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue.
-The heatmap shows the normalized enrichment score (NES).
-Only enrichments with pvalue below 0.05 are shown.
-Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.
-
-\clearpage
-
-```{r, echo = FALSE,fig.width=10,fig.height=8}
-write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
-```

### 2022-01-08 reports/rmarkdown/manuscript/Suppl.Figure5.Rmd
Suppl. Figure 5

---
title: "Suppl. Figure 5. Rarefaction curves"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(vegan)
  library(phyloseq)
  library(ggpubr)
})  })  })

#############
# LODA DATA #
#############
paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVL V2)
             '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
             '../../../data/phyloseq_Sthlm.RDS')       # CVL V2, (only v4 region)

n <- c("boston_r1", "boston_r2", "Sthlm")
ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)

```

```{r rarefaction curves, fig.height=6, fig.width=8, fig.asp=0.75,  include=FALSE}
##################
# FILTER SAMPLES #
##################
# remove controls and duplicates
# Include column keeps replicate samples
# Include_final column removes all replicates and women that don't have trx samples
ps_filt <- map(ps_list, ~ .x %>%
  subset_samples(., Include_final == "yes"))

# separate tissue from CVLv2 samples in the second Boston run
CVLv3 <- ps_filt[["boston_r1"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v3")
CVLv2 <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "CVL" & Visit == "v2")
tissue <- ps_filt[["boston_r2"]] %>% subset_samples(., Sample_type == "tissue" & Visit == "v3")

# merge the lists
ps_types <- c(ps_filt, "CVLv3"=CVLv3, "CVLv2"=CVLv2, "tissue"=tissue)
# Get ASV tables
seqtabs <- map(ps_types, ~data.frame(.x@otu_table@.Data))

#######################
# RAREFACTION CURVES #
######################

# Get rarefaction curves
rare_runs <- map(seqtabs, ~ .x %>%
             rarecurve(., step = 20, label = FALSE)
)

ggplot_df_fun <- function(rare_obj, count) {
  # Add sample names
  names(rare_obj) <- c(paste0("sample",sprintf("%03.0f", 1:length(rare_obj))))

  # Coerce data into "long" form.
  rare_df <- imap_dfr(rare_obj, ~ .x %>%
              as_tibble(., rownames = "Sample size") %>%
              mutate(Sample = .y) %>%
              mutate(`Sample size` = as.numeric(gsub("N","",`Sample size`)))
  )
  # Get nr. of counts in each sample
  c <- formatC(count, format="f", big.mark = " ", digits=0)
  rare_df_col <- rare_df %>%
    dplyr::rename(ASV = "value") %>%
    group_by(Sample) %>%
    nest() %>%
    mutate(Counts = map(data, ~max(.$`Sample size`))) %>%
    unnest(c(data, Counts)) %>%
    mutate(Counts_ = Counts) %>%
    mutate(Counts = ifelse(Counts > count, paste0("> ", c), paste0("< ", c)))

  return(rare_df_col)
}

# Define some plot variables
count <- c(33000, 10000, 33000, 40000, 20000, 2500)
limit <- c(65000, 65000, 65000, 65000, 65000, 10000)
title <- c("Run B1. CVLv3 + CVLv2 rep.", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B1. CVLv3", "Run B2. CVLv2", "Run B2. TISSUEv3")
title <- c("Run CVL", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B2. CVLv2", "Run Tissue")

# Extract data frame from the rarecurve object
rare_df_col<- map2(rare_runs, count, ~ggplot_df_fun(.x, .y))
# Create Plot
rare_plot <- map(rare_df_col,
              ~ggplot(.x, aes(x = `Sample size`, y = ASV)) +
              theme_bw() +
              scale_color_manual(values=c('tomato', alpha('black', 0.1))) +
              geom_line(aes(color = Counts, group = Sample)) +
              theme(plot.title = element_text(hjust = 0.5))
            )
raremax <- map(seqtabs, ~min(rowSums(.x)))

# sample information:
 read_counts <- map(rare_df_col, ~ .x %>%
            select(Sample, Counts, Counts_) %>%
            unique(.))

# Save png
#imap(r_plot, ~ggsave(filename = paste0('Rarefaction_plot_', .y, '.png'), plot = .x,
#        path = paste0("../../../results/"),
#        width = 10, height = 5,)
#)

```

```{r Print rare curves, warning=FALSE, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.asp=0.75,}
plot <- ggarrange(
          #rare_plot[[1]] + theme(axis.title.x=element_blank(),plot.margin=unit(c(.5,.5,.5,.5),"cm")),
          #rare_plot[[3]] + theme(axis.title.x=element_blank()),
          rare_plot[[4]] + theme(axis.title.x=element_blank(), plot.margin=unit(c(.5,.5,.5,.5),"cm")),
          rare_plot[[6]] + theme(plot.margin=unit(c(.5,.5,.5,.5),"cm")),
          ncol = 1,
          labels = c("a","b","c","d")
          )
print(plot)

ggsave(filename = paste0('Rarefaction_plot.pdf'), plot = plot,
       path = paste0("../../../results/Plots"),
       width = 8, height = 6
)
```

**Supp. Figure 1. Rarefraction curves**. Rarefaction curves showing number of unique ASVs detected in each sample when simulating increasing sequencing depth. Low abundant taxa may be undetected at low sequencing depth but are expected to be detected with an increased sequencing depth (x-axis). When the curve flattens out, all taxa in the sample are considered detected. **a)** The Luminal sequencing run and **b)** the Tissue sequencing run. The sequencing depth was > `r as.character(count[4])` reads in all but `r filter(read_counts$CVLv3, Counts == "< 40 000") %>% nrow()` samples for the Luminal dataset while `r filter(read_counts$tissue, Counts == "< 2 500") %>% nrow()` samples had fewer than `r as.character(count[6])` reads in the tissue dataset.

### 2022-01-30 reports/rmarkdown/manuscript/Figure2.Rmd
Fgure text update

@@
 ###################################
-# WETT SMEAR BACTERIAL VALIDATION #
+# WET SMEAR BACTERIAL VALIDATION #
 ###################################
@@
 mtext("Lactobacillus", side=3, las = 1, cex = .6, line = .5,xpd=T)
-mtext("Wet Smear Bacteria Validation", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
+mtext("Wet Smear Bacteria Validation", side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
 
@@
        breaks = seq(-5,5,length.out = 100),axes=F)
-mtext("Differentialy Expressed KEGG Ontology Terms", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
+mtext("                        Differentially expressed KEGG Ontology terms and pathways", 
+      side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
 title(xlab="Study Groups", line=1)

### 2022-01-31 reports/rmarkdown/manuscript/Figure4-5.Rmd
minor changes
change txt size and update igrap network

--- 2021-11-12 reports/rmarkdown/manuscript/Figure4-5.Rmd
@@
   lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
-  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% map(., ~rownames_to_column(.x, var = "Genes"))
+  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% 
+         map(., ~rownames_to_column(.x, var = "Genes"))
 
@@
 ```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}
-
 TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
@@
          breaks = seq(-5,5,length.out = 100),axes=F)
-  mtext("           Differentialy Expressed Genes", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
+  mtext("                                                          Differentially expressed genes and KEGG/GO pathways", side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
   title(xlab="Study Groups", line=1)
@@
   res_list <- pluck(enrich, "GO", "res_list")
-  #top_TRX_counts <- pluck(enrich, "GO", "top_TRX_counts")[,samples]
 
@@
     barplot( c(rep(NA, each=20-length(comb)), comb),  
-             xlim=c(-1.1,1.1),
-             # xlim=c(min(c(comb,-2/abs(max(abs(DD))) ))*1.1, 
-             #                          max(c(comb,2/abs(max(abs(DD)))) )*1.1), 
-                        cex.names = .9, #mar=c(1,25,2,2), #main=paste(x, i, sep="\n"),
+             xlim=c(-1.1,1.1), cex.names = .9,
              names.arg = c(rep("", each=20-length(comb)),names_DD, names_UU ), 
-             horiz = T,las= 1, border=NA,xaxt='n',line=-.2, #xlab="-log10 p-value",
+             horiz = T, las= 1, border=NA, xaxt='n', line=-.2,
              col = c(rep(NA, each=20-length(comb)), 
@@
     )
-    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=.9)
+    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=1)
 
@@
              points( comb , points_y,
-                     #mar=c(1,20,2,2),
                      bg=colorRampPalette( c("navy","gray","firebrick3") )(19)[(comb+1)*9+1],
@@
 }
-
-dev.copy2pdf(file=paste0("./test.pdf"),
-    width = 6.6,
-    height = 6, paper = "a4"
-    )
 ```
@@
 arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)
+
+# test figure
+# dev.copy2pdf(file=paste0("./test.pdf"),
+#     width = 6.6,
+#     height = 6, paper = "a4"
+#     )
 ```
@@
 
-```{r UMAP, eval=FALSE, fig.height=2.5, fig.width=3, include=FALSE}
-# , fig.height=3, fig.width=3.5
-###############
-#  TRX UMAP   #
-###############
-plot_UMAP.fun <- function(top_TRX_counts, group){
-  #par(mar=c(4,4,2,6)) #b,l,t,r
-  par(mar=c(2.5,2.5,2,3)) #b,l,t,r
-  #par(mar=c(10,12,2,12)) #b,l,t,r
-  set.seed(1)
-  UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
-                   metric = "correlation",min_dist = 0.1,spread = 5,
-                   negative_sample_rate = 10)
-  plot(UMAP_TRX,bg=pal[factor(group)],main="UMAP on top DGE RNAseq",
-       pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
-  mtext("UMAP1", side=1, line=0.7, cex=0.8)
-  mtext("UMAP2", side=2, line=0.7, cex=0.8)
-  legend(par("usr")[2],par("usr")[4],title.adj = 0,
-         #legend = levels(factor(metadata$Luminal_gr)),
-         legend = sort(unique(group)),
-         xjust = 0,yjust = 1, cex=0.7,
-         bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
+```{r Suppl. Tbl. 12, include=FALSE}
+metadata <- metadata %>% 
+  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
+  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))
+
+dge_LT <- edgR_dge.fun("L_T_groups", confound=T)
+enrich_LT <- db %>%
+  map(., ~gene_mod_enrich.fun(dge_LT$top,dge_LT$y,dge_LT$c,
+                              db = .x,
+                              group_name = "L_T_groups",
+                              cuttoff = c(GO=2, KEGG=2) # c(GO=2, KEGG=2.7)
+                              ))
+
+# Plot the L2T2 vs L2T3 heatmap:
+#pal <- c("#54ba63","#ad58c6","#d03e7a","#4cb6a9","#bb6fac","#cca442","#8fb53c","#6a7ecd","#63813e","#c46570","#b3743e","#d35238")
+#plot_heatmap.fun(enrich_LT, "KEGG", "L_T_groups", "L_T_groups")
+
+t <- dge_LT$top %>% filter(PValue <= 0.01)
+
+df_across <- enframe(list(LtoT=enrich_LT)) %>%  
+  unnest_longer(value) %>%
+  mutate(res = map(.$value, "res_list")) %>%
+  unnest_longer(res) %>%
+  mutate(res = map(res, ~rownames_to_column(.x, var = "Terms"))) %>%
+  mutate(res = map(res, ~filter(.x, pvalue <= 0.01))) %>%
+  group_by(name, value_id) %>%
+  mutate(mod = row_number()) %>% ungroup() %>%
+  mutate(res = set_names(.$res, paste0(.$value_id,"_gene_module_",.$mod))) %>%
+  group_split(name) %>% set_names(., map(., ~unique(.x$name)))
+
+pages <- list(list("DEGs L2T2 vs L2T3"=t), df_across[[1]]$res) %>% flatten()
+write.xlsx(pages, file=paste0(out,"Suppl.Tbl.12 L2T2_L2T3_DEGs_and_enrichment.xlsx")) 
+```
+
+## igraph network plot
+```{r igraph network, echo=FALSE, warning=FALSE} 
+# 1. Change the groups of interest by assigning condition_use
+# 2. Change the number of genes to be included in the network by changing n
+#    look at tempD_i and tempU_i to see how many genes was significant
+# 3. Adjust the size of the gene nodes and text in the graph by changing; square, circle and txt
+
+# combinations
+sheets <- map(DEG_files, ~excel_sheets(paste0(out, .x))[-c(1:2)] ) %>%
+  map(., ~set_names(.x))
+
+# Set enrichment conditions
+n <- 10  # maximum number of TF's to plot
+# sizes:
+square <- 10
+cricle <- 8
+txt <- .7
+
+# Create a graph from the list of UP and DOWN regulated genes
+i_graph <- function(temp_i, n, square, cricle){
+  temp <- lapply( 1:n, function(x){ strsplit(as.character(temp_i$Genes[x]),split = ";")[[1]]  } )
+  temp <- map(temp, purrr::discard, is.na) %>% compact() # removes NA
+  names(temp) <- temp_i$`Transcription Factor`[1:length(temp)]
+  temp2 <- lapply( names(temp) , function(x){ rep(x,length(temp[[x]])) })
+  temp <- cbind(unlist(temp2),unlist(temp))
+  g <- graph_from_data_frame(temp,directed = F)
+  v_shape <- ifelse( names(V(g)) %in% unique(temp[,1]) , "square" , "circle" )
+  v_color <- ifelse( names(V(g)) %in% unique(temp[,1]) , "honeydew" , "cornsilk" )
+  v_size <- ifelse( names(V(g)) %in% unique(temp[,1]) , square , cricle )
+
+  return(list(g=g, v_shape=v_shape, v_color=v_color, v_size=v_size))
   }
 
-plot_UMAP.fun(enrich_L$GO$top_TRX_counts, metadata$Luminal_gr)
-m <- filter(metadata, !is.na(Tissue_gr)) 
-plot_UMAP.fun(enrich_T$GO$top_TRX_counts[,m$ID], m$Tissue_gr)
+# Test parameters
+# i <- "TF_PPI_Tissue"
+# i <- "TF_TRRUST_enrichment"
+# gr_name <- "Tissue"
+# comp <- sheets
 
-#plot_UMAP.fun(enrich_T$GO$top_TRX_counts, metadata$Tissue_gr)
-dev.copy2pdf(file=paste0("./UMAP",".pdf"),
-    width = 4,
-    height = 3.5, paper = "a4" #units = "cm", res = 300 
-    )
+# Save all combination plots in one file
+save_pdf.fun <- function(i, gr_name, comp){
+  temp_i_name <- grep(i,list.files(paste0(out),full.names=F),value=T)
+  p <- map(comp, ~read_xlsx(paste0(out, temp_i_name), sheet = .x))
+
+  p_split <- p %>%
+    imap(., ~.x %>%
+          mutate(Regulation = paste0(.y, "_", .$Regulation)) %>%
+          split(., .$Regulation)
+          ) %>% compact() %>% flatten()
+
+  obj_igraph <- imap(p_split, ~i_graph(.x, n, square, cricle))
+
+  par(mar = c(0, 0, .7, 0))  #bottom, left, top, right
+  pdf(file = paste0(out, "Plots/", i, "_network_",".pdf", sep=""))
+    p <- imap(obj_igraph,
+         ~with(.x, 
+                 plot.igraph(g, label=V(g), page = T,
+                 main = paste0(str_replace(.y, "_", " ")," regulated genes"),
+                 layout=layout_with_kk(g), vertex.label.color="black" ,
+                 vertex.size=v_size, vertex.label.cex=txt, vertex.label.font=2, 
+                 vertex.shape=v_shape, vertex.color=v_color, marigin = c(-1,-4,0,0),
+                 vertex.frame.color="grey60", curved=F)
+
+    ))
+    dev.off()
+}
+
+# if(isFALSE(dir.exists(paste0(out,"Plots")))){dir.create(paste0("../",result.dir),recursive = TRUE)}
+# save_pdf.fun("TF_PPI_Luminal", "Luminal", comp = sheets[[1]])
+# save_pdf.fun("TF_PPI_Tissue", "Tissue", comp = sheets[[2]])
 
 ```
-
- <!-- -->

### 2022-01-31 reports/rmarkdown/manuscript/Figure6.Rmd
Updating to accomodate Paulos new code in niceRplots

@@
   library(rstatix)
+  #remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
@@
            axes=F)
-    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, line =1.5,xpd=T)
+    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, font = 2, line =1.5,xpd=T)
     mtext( c("Study Groups"), side=1, las = 1, cex = .7, line = 1,xpd=T)
@@
   sig <- df_krusk %>% filter(p.adj<0.05) %>% pull(Cytokines)
-  sig_df <- cytokine_data_log[sig,samples_use$ID] %>%
-    cbind(Max = c(rep(max(.)+n, nrow(.))))
+  sig_df <- cytokine_data_log[sig,samples_use$ID]
 
   groupings <- samples_use[[group]] [ match( colnames(sig_df[,1:ncol(sig_df)-1]) , as.character(samples_use$ID)  ) ]
-  groupings <- c(groupings, "Max")
+  groupings <- c(groupings)
 
@@
 Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g)
-temp <- as.factor(c(Luminal$groupings, "Max"))
 stars <- Luminal$stars_df
 
-par( mar=c(1.4,5,1.5,0)) #b,l,t,r 
+par( mar=c(1.4,5,1.5,2)) #b,l,t,r 
 violist(data=Luminal$sig_df,
-          srt=45,
-          genes= g,
+          srt=45, y_padding = 2, smooth = .2,
+          genes= g, #add_ylims = T, #gives starting value of y-axis and max value of all data points
           col = pal, transparency = 50,
-          clustering=c(Luminal$groupings),smooth = 2 )
+          clustering=c(Luminal$groupings))
 segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
@@
 
-rect(7, -0.5, 5.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
-     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
-     )
-
 add_letter("b")
@@
 Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
-temp <- factor(Tissue$groupings, levels=c("T1","T2","T3","T4","Max"))
 stars <- Tissue$stars_df
 
-par( mar=c(24.4,6,1.5,0)) #b,l,t,r
+par( mar=c(20.4,6,1.5,3)) #b,l,t,r
 violist(data=Tissue$sig_df,
-          srt=45, 
+          srt=45, y_padding = 2, smooth = .2,
           genes= rownames(Tissue$sig_df), 
-          col = pal, transparency = 50,
-          clustering=Tissue$groupings, smooth = 2 )
+          col = pal, transparency = 50, #add_ylims = T,
+          clustering=Tissue$groupings )
 segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
@@
 
-rect(6, -0.5, 4.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
-     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
-     )
 add_letter("d")
@@
 plot_grid(ggdraw(p.plot),ggdraw(p.plot_T), 
-          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F) # , 'c', 'd'
+          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F)
 

### 2022-01-31 reports/rmarkdown/manuscript/Suppl.Figure1.Rmd
minor changes

@@
 ---
-title: "Bacterial abundance across datasets"
+title: "Suppl. Figure 1. Bacterial abundance across datasets"
 output: 
@@
 suppressWarnings({suppressMessages({suppressPackageStartupMessages({
+  #remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
@@
 
-```{r Bacterial abundance, echo = FALSE, eval = TRUE, fig.height=3, fig.width=7}
-par(mar=c(6,2.5,1,1)) #b,l,t,r
+```{r Bacterial abundance, echo = FALSE, eval = TRUE,  fig.asp=.5}
+par(mar=c(6,2.5,0,0)) #b,l,t,r
 
@@
                         function(i){rep(i,ncol(datasets_all_samples[[2]]))}))
-#rownames(x) <- gsub("ASV_|_norm.*", "",rownames(x))
+
 rownames(x) <- c("Tissue", "Luminal")

### 2022-01-31 reports/rmarkdown/manuscript/Suppl.Figure2.Rmd
changes to accomoadate Paulos new niceRplot code

@@
 ---
-title: "Suppl. Figure 2. Differential bacterial abundance across tissue and CVL microbiome"
+title: "Suppl. Figure 2. Differential bacterial abundance across Tissue and Luminal microbiome"
 output: 
@@
 
-write.xlsx(res, file=paste0("../../../results/","DBA_across_Luminal_vs_Tissue",".xlsx"))
+#write.xlsx(res, file=paste0("../../../results/","DBA_across_Luminal_vs_Tissue",".xlsx"))
 
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))
+ord <- getcluster(data = all_microbiome, genes =  unique(as.character(res$bacteria)), clustering =  datasets)
+ord <- ord[unique(as.character(res$bacteria))]
 
@@
 par(mfrow=c(1,2),mar=c(2,12,1,3.1))
-#par(mar=c(2,17,1,2)) #b,l,t,r
-plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,
+plot_dots( all_microbiome, genes = names(sort(ord)) , clustering = datasets,
            show_grid = T,cex.main=1,font.main=1,
@@
 add_letter("a")
-barlist( all_microbiome , names(sort(ord)) , clustering = datasets,
+barlist( all_microbiome, names(sort(ord)) , clustering = datasets,
          show_grid = T,cex.main=1,font.main=1,
@@
 add_letter("b")
+
+dev.off()
 ```

### 2022-01-31 reports/rmarkdown/manuscript/Figure2.Rmd
update meta data points

@@
   library(RColorBrewer)
-  remotes::install_github("czarnewski/niceRplots",force=T)
+  #remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
@@
 
+# Dendogram
 par(mar=c(4,.1,4,.2)) #b,l,t,r
@@
 
+# Heatmap
 image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
@@
 
+# Study Group brackets
 end <- table(pal[factor(group_annotation)[ order(group_annotation[sample_use])]] )
@@
 end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/ncol(picrust_data[sample_use])-.015) )
-start <- c(0, map_dbl(end[1:4], ~.x+0.015))
+start <- c(0, map_dbl(end[1:4], ~.x+0.017))
+end <- c(end[1:length(end)-1], par("usr")[2]-.005)
 
-axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
-map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
+axis(1, at = c(start, end), label = F, pos =1.021 , xpd=T, col="white",col.ticks="black")
+map2(start, end, ~lines(x=c(.y, .x), y= c(1.021, 1.021), xpd=T, cex=.8))
 # text labels
 t <- map2(start, end, ~(.x+.y)/2)
-map2(t,levels(factor(group_annotation)), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))
+map2(t,levels(factor(group_annotation)), ~text(x=.x, y=1.037, .y,xpd=T, cex=.8))
 
-points( seq(0,1,length.out = length(sample_use)),
-        rep(par("usr")[4],length(sample_use))+.005, pch=15, cex=.7,xpd=T,
-        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]])
+# Meta bar info
+par(mar=c(3,.1,2,.1)) #bottom, left, top, right  
+map(c(.0021, .006), ~points( seq(0,1,length.out = length(sample_use)),
+        rep(par("usr")[4],length(sample_use))+.x, pch=15, cex=.3,xpd=F,
+        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]]) )
 points( seq(0,1,length.out = ncol(top_picrust_data)), 
-        rep(par("usr")[3],ncol(top_picrust_data))-.008, pch=15, xpd=T,cex=.5,
+        rep(par("usr")[3],ncol(top_picrust_data))-.008, pch=15, xpd=F,cex=.3,
         col= c("tomato","orange","#d7d7d7")[as.factor(metadata$BV_Diagnosis_v3)][order( group_annotation[sample_use])] )
 points( seq(0,1,length.out = ncol(top_picrust_data)), 
-        rep(par("usr")[3],ncol(top_picrust_data))-.024, pch=15, xpd=T,cex=.5,
+        rep(par("usr")[3],ncol(top_picrust_data))-.024, pch=15, xpd=F,cex=.3,
         col= c("#d7d7d7","tomato")[as.factor(metadata$HIVstatus)][order(group_annotation[sample_use])])
@@
 df <- data.frame(temp,temp)
+par(mar=c(4,0,4,.1)) #bottom, left, top, right  
 quietly(plot_sankey)( df, pal = taxa_pal[as.numeric(levels(temp))], use_w2 = F, plot_labels = T,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)$output

### 2022-01-31 reports/rmarkdown/manuscript/Figure4-5.Rmd
update meta data points

@@
 
-  ## Heatmap
-  par(mar=c(4,.1,2.5,.1))
+  # Dendogram
+  par(mar=c(4,.1,2.5,.1)) #bottom, left, top, right  
   plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
@@
 
+  # Heatmap
   image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
@@
 
+  # Study Group brackets
   end <- table(pal[factor(names(samples))[ order(df[1])]])
@@
   end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
-  start <- c(0, map_dbl(end[1:4], ~.x+0.015))
+  start <- c(0, map_dbl(end[1:4], ~.x+0.017))
+  end <- c(end[1:length(end)-1], par("usr")[2]-.005)
 
-  axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
-  map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
+  axis(1, at = c(start, end), label = F, pos =1.029 , xpd=T, col="white",col.ticks="black")
+  map2(start, end, ~lines(x=c(.y, .x), y= c(1.029, 1.029), xpd=T, cex=.8))
   # text labels
   t <- map2(start, end, ~(.x+.y)/2)
-  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))
+  par(mar=c(3,.2,2,.2)) #bottom, left, top, right  
+  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.045, .y,xpd=T, cex=.8))
 
   ## Meta info bars
-  points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
+  par(mar=c(3,.1,2,.1)) #bottom, left, top, right  
+  map(c(.0025, .007), ~points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.x,
           col=pal[factor(names(samples))[ order(df[1])]],
-          pch=15,xpd=T,cex=.7)
+          pch=15,xpd=F,cex=.3) )
   points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
           col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
-          pch=15,xpd=T,cex=.45) 
+          pch=15,xpd=F,cex=.3) 
   points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
           col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
-          pch=15,xpd=T,cex=.45)
+          pch=15,xpd=F,cex=.3)
 
@@
   temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
+  par(mar=c(4,0,2.5,.1)) #bottom, left, top, right  
   plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
@@
 
-  text(par("usr")[1],par("usr")[3]-.008,labels = "    BV",cex=.6, xpd=T)
-  text(par("usr")[1],par("usr")[3]-.028,labels = "    HIV",cex=.6, xpd=T)
+  text(par("usr")[1]+0.008,par("usr")[3]-.008,labels = "     BV",cex=.6, xpd=T)
+  text(par("usr")[1]+0.008,par("usr")[3]-.028,labels = "      HIV",cex=.6, xpd=T)
 
@@
 # test figure
-# dev.copy2pdf(file=paste0("./test.pdf"),
-#     width = 6.6,
-#     height = 6, paper = "a4"
-#     )
+dev.copy2pdf(file=paste0("./test.pdf"),
+    width = 6.6,
+    height = 6, paper = "a4"
+    )
 ```
@@
 arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)
+
+dev.copy2pdf(file=paste0("./test.pdf"),
+    width = 6.6,
+    height = 6, paper = "a4"
+    )
 ```

### 2022-02-02 reports/rmarkdown/manuscript/Figure3.Rmd
Remove white space from figure

@@
   library(RColorBrewer)
-  remotes::install_github("czarnewski/niceRplots",force=T)
+  # remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
@@
 
-```{r Figure 3., echo=FALSE, warning=FALSE, message=FALSE, fig.width=6 ,fig.asp=1, fig.pos="t"}
-### A + B
+```{r Figure 3., echo=FALSE, warning=FALSE, message=FALSE,  fig.pos="t"}
+#  fig.width=6 ,fig.asp=1, fig.height=7,
+### A
 #############################
@@
                 3,3,6,7,4,4,
-                3,3,6,5,4,4,
-                8,8,8,8,8,8) , ncol= 6,byrow = T),
+                3,3,6,5,4,4) , ncol= 6,byrow = T),
                                widths = c(1.2,1,1.83,1.7,.85,1.4),
-                               heights = c(.7,1.5,.8,1.5,1,.8,.2))
+                               heights = c(.7,1.6,1,1.1,1,.8,.2))
 
@@
 
-### C
+### B
 ############################
@@
 
-par(mar=c(6,5,2,.3)) #b,l,t,r
+par(mar=c(1.5,5,2,.3)) #b,l,t,r
 barplot( r, border = NA ,ylab="abundance", xlab ="", axisnames = F,
@@
 
-### D
+### C
 #############################
@@
 #############################
-par(mar=c(5,1.1,2,2)) #b,l,t,r
+par(mar=c(0,1.1,2,2)) #b,l,t,r
 x <- as.character(metadata[,"Luminal_gr"])
@@
 
-par(mar=c(0,.2,1,.4)) #b,l,t,r
+par(mar=c(0,.2,1.5,.4)) #b,l,t,r
 empty_plot()
@@
 
-par(mar=c(0,0,1,1)) #b,l,t,r
+par(mar=c(0,0,1.5,1)) #b,l,t,r
 empty_plot()
@@
 
-empty_plot()
 ```

### 2022-02-10 reports/rmarkdown/01_data_pre_processing.Rmd
removing code not in use

@@
 ---
-title: "`r params$project_title`"
+title: "Normalization"
 editor_options:
@@
       collapsed: no
-params:
-  organisation: Karolinska Institutet
-  pi_mail: prin.cipal@ki.se
-  pi_name: Prin Cipal
-  project_title: Broliden_5325
-  redmine_issue: '#5325'
-  requester_mail: kristina.broliden@ki.se
-  requester_name: Kristina Broliden
-  staff_mail: supp.ortstaff@nbis.se
-  staff_name: Supp Ortstaff
-  staff_web: https://nbis.se/about/staff/supp-ortstaff/
 ---
@@
 
-setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
 ```
 
-#Load libraries and other scripts
+# Load libraries and other scripts
 
@@
 suppressWarnings({suppressMessages({suppressPackageStartupMessages({
-library(igraph)
-library(rafalib)
-library(sva)
-library(batchelor)
-remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
+#remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
 library(niceRplots)
@@
 
-# Load metadata
 dataset_names <- c("ASV_tissue_V3_B2.csv", # Tissue, Boston run 2 (95 samples)
@@
 
-# make long format raw non agglom ASV files
-# remove_from_CVL_V2 <- intersect(colnames(temp_[[5]]), colnames(temp_[[3]]))[-c(1:11)]
-# names <- c("ASV_tissue_V3_Boston run2","ASV_tissue_V3_Boston run1", "ASV_CVL_V2_Boston run1", 
-#                      "ASV_CVL_V2_Boston run2", "ASV_CVL_V2_Stockholm", "ASV_CVL_V3_Boston run 1") 
-# ASV_long <- temp_ %>%
-#   modify_at(., c("ASV_CVL_V2_C.csv"), ~select(.x, -any_of(remove_from_CVL_V2))) %>% 
-#   map(., ~pivot_longer(.x, matches("^P\\d"), names_to = "ID", values_to = "Reads")) %>%
-#   set_names(names ) %>%
-#   bind_rows(., .id = "dataset") %>%
-#   separate(., dataset, into=c(NA, "Sample_type", "Visit","Place_sequenced"), sep = "_", remove = F) %>% 
-#   mutate(dataset = str_extract(.$dataset, "CVL_V\\d|tissue_V3")) %>% 
-#   group_split(dataset, .keep=F) %>% set_names(c("tissue", "CVL_V2", "CVL_V3"))
-# 
-# saveRDS(ASV_long, paste0(PATH,"results/","ASV_all_data_long_format.RDS"))
-
 #arrange all datasets to have the same rownames with all bacteria
@@
 
-# datasets <- datasets[3] #in case you wanna test in 1 dataset
-
-# Transform phyloseq data into normalized matricies ( log2[CP1K] )
-# datasets <- lapply(datasets,function(x){
-#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
-# })
-# lapply(datasets,dim)
-
 ```
 
-# Merging, renaming and correcting batch effects on datasets
+# Merging and renaming
 
@@
 ```
+
+# normalize and save microbiome datasets
 
@@
 
+# normalize and save trx datasets
+
 ```{r save raw and norm trx files, echo = FALSE}

### 2022-02-10 reports/rmarkdown/01_taxonomy_improved.Rmd
Clean up + minor changes

@@
 ---
-title: "`r params$project_title`"
+title: "Improve Taxonomy Assignment"
 editor_options:
@@
       collapsed: no
-params:
-  organisation: Karolinska Institutet
-  pi_mail: krisina.broliden@ki.se
-  pi_name: Kristina Broliden
-  project_title: Broliden_5325
 ---
@@
 
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
 ```
@@
 ```{r Get ASV and tax tables, message=FALSE}
-paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3
+paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
              '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
@@
 #### ASV with all controls and replicate samples removed ####
-ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3", "ASV_tissue_B1", 
-               "ASV_CVL_V2_B2", "ASV_tissue_B2", "ASV_CVL_V2_C" )
+ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3_B1", "ASV_tissue_V3_B1", 
+               "ASV_CVL_V2_B2", "ASV_tissue_V3_B2", "ASV_CVL_V2_S" )
 ASV_list <- map(ASV, ~ .x %>%
@@
 
+saveRDS(ASV, paste0(PATH,"results/","ASV_wCtrl_long_format.RDS"))
 ```

### 2022-02-10 reports/rmarkdown/02_individual_reports.Rmd
rm unused code reformat to clustering

--- 2021-03-16 reports/rmarkdown/02_individual_reports.Rmd
@@
 ---
-title: "`r params$project_title`"
+title: "Graph Based Clustering"
+output: 
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+header-includes: 
+- \usepackage{float}
 editor_options: 
   chunk_output_type: console
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  pdf_document:
-    toc: yes
-  html_document:
-    code_folding: hide
-    toc: yes
-    toc_float:
-      collapsed: no
-params:
-  dataset: 1
 ---
 
-```{r Setup, echo = FALSE}
-knitr::opts_chunk$set(fig.width  = 10,
-                      results    = "hold",
-                      message    = FALSE,
-                      warning    = FALSE)
+```{r message=FALSE, warning=FALSE, include=FALSE}
+##################
+# LOAD LIBRARIES #
+##################
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+suppressWarnings({suppressMessages({suppressPackageStartupMessages({
+  library(igraph)
+  library(rafalib)
+  library(fgsea)
+  library(scales)
+  #remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
+  library(niceRplots)
+  library(tidyverse)
+})  })  })
+
+#############
+# LODA DATA #
+#############
+datasets_all_samples <- readRDS("../../results/datasets_all_samples.RDS")
+metadata <- read.csv2("../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
+
+#color palettes
+#################
+# COLOR PALETTS #
+#################
+pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
+taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
+bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')
 ```
 
-#Load libraries and other scripts
+```{r, echo = FALSE}
 
-```{r, echo = FALSE}
-library(igraph)
-library(rafalib)
-library(sva)
-library(batchelor)
-library(scales)
-remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
-library(niceRplots)
+all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
+dim(all_microbiome)
+
+zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
+zscore[ zscore > 4 ] <- 4
+zscore[ zscore < -4 ] <- -4
+rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)
+mypar(mar=c(2,15,2,2))
+top_var <- apply(all_microbiome,1,var)
+top_var_t <- sort(top_var,decreasing = T)[100]
+is_top_var <- top_var > top_var_t 
+barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)
+
+#Compute sample-wise correlations
+cors <- cor(all_microbiome[,], method = "pearson")
+adj <- (1-cors)/2
+# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
+
+# cors <- cor(all_microbiome, method = "pearson")
+# adj <- (1-cors)/2
+# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )
+
+#Graph construction  
+k <- 12 # k for KNN
+p <- 0.1 # prunning threshold of SNN 0.14
+
+#Compute KNN graph from correlation matrix
+knn <- RANN::nn2(adj,k = k)
+nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
+for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }
+
+#Computing SNN matrix
+N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
+P <- t(nn) %*% nn
+B <- N - nn
+C <- t(B) %*% B
+Q <- nrow(nn)*N - C
+J <- P/Q
+J[is.nan(J)] <- 0
+J[J < p] <- 0
+colnames(J) <- colnames(all_microbiome)
+rownames(J) <- colnames(all_microbiome)
+
+#Transforming SNN into an igraph object
+g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
+write.csv(J,"../../results/patient_SNN_graph.csv",row.names = T)
 
@@
 
-#Defining some variables for the analysis
+```{r Plot Study Groups, echo=FALSE,fig.width=10,fig.height=9}
+mypar(3,3,mar=c(0,0,2,0))
 
-```{r, echo = FALSE}
-PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
-PATH <- "../.." #Path to the data
+set.seed(1)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=30)
 
-#color palettes
-pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
-         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
-heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
-cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
+cl <- igraph::cluster_louvain(g)
+# cl <- igraph::walktrap.community(g)
+
+metadata$Louvain_clusters <- cl$membership
+metadata$joint_clustering <- metadata$Luminal_gr
+metadata$joint_clustering <- factor(metadata$joint_clustering) 
+set.seed(1)
+U <- uwot::umap(adj,n_neighbors = 30)
+h <- hclust(as.dist(adj),method = "ward.D2")
+# metadata$joint_clustering <- cutree(h,k = 5)
+#write.csv2(metadata,"../../results/metadata_integration.csv",row.names = T)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,main="FR embedding\nlouvain",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups\nLouvain_gr",
+       legend = levels(factor(cl$membership)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,main="FR embedding",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups\nLuminal_gr",
+       legend = levels(factor(metadata$Luminal_gr)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
+# title("Louvain clusters",cex.main=1,font.main=1)
+
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Tissue_gr)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
+      vertex.size=10,main="FR embedding",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Tissue_gr)] , 
+      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
+      vertex.size=10,main="umap",
+      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+empty_plot()
+legend("topleft",title = "Patient groups\nTissue_gr",
+       legend = levels(factor(metadata$Tissue_gr)),
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
+# title("Louvain clusters",cex.main=1,font.main=1)
+
+```
+
+```{r Bacterial communities, echo = FALSE}
+
+all_microbiome <- cbind(#datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+                        )
+
+cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))
 
 #Graph construction 
-k <- 10 # k for KNN
-P <- 0.35 # prunning threshold of SNN
+k_nearest_neighbours <- 5 # k for KNN
+snn_threshold <- 0.1 # prunning threshold of SNN
 
-#Graph construction  
-fct <- .25 # FC threshold for differential expression
-pvt <- 0.05 # Pvalue threshold for differential expression
-min_pct <- .1 # minimun level of detected bacteria in each sample group
-```
-
-# Loading data and metadata
-
-```{r, echo = FALSE}
-# Load metadata
-metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
-rownames(metadata) <- as.character(metadata$ID)
-
-# Load datasets
-dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
-                   "ASV_CVL_V3_normalized_batch_corrected.csv",
-                   "ASV_CVL_V2_normalized_batch_corrected.csv",
-                   "ASV_CVL_V2_normalized_NOT_batch_corrected.csv")
-
-#import datasets and return matrix with taxa-level resolution
-datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"/results/",x),row.names = 1)
-  return(x)
-})
-names(datasets) <- dataset_names
-lapply(datasets,dim)
-```
-
-# Merging microbiome datasets
-
-```{r, echo = FALSE}
-# Define the common microbes
-common_microbes <- sort(unique(unlist(lapply(datasets,rownames))))
-sample_names <- sort(unique(unlist(lapply(datasets,colnames))))
-
-# Create empty matrix for the full dataset
-microbiome <- as.matrix(datasets[[ params$dataset ]])
-microbiome <- microbiome[ rowSums(microbiome>0)>=2 , colSums(microbiome>0)>=2 ]
-
-dataset_origin <- factor(rep(sub("_norma.*","",names(datasets)[2]),ncol(datasets[[2]])))
-```
-
-# Organise the datasets
-
-```{r, echo = FALSE}
-zscore <- t(apply(microbiome,1,function(i){scale(i,T,T)}))
-zscore[ zscore > 6 ] <- 6
-zscore[ zscore < -6 ] <- -6
-rownames(zscore) <- rownames(microbiome); colnames(zscore) <- colnames(microbiome)
-
-eucl <- as.matrix(dist(t(zscore),method = "euclidean"))
-h_bacs <- hclust(dist(zscore),method = "ward.D2")
-h_samples <- hclust(dist(t(zscore)),method = "ward.D2")
-
-microbiome <- microbiome[h_bacs$order,h_samples$order]
-zscore <- zscore[h_bacs$order,h_samples$order]
-```
-
-# Organise the datasets
-
-```{r, echo = FALSE,fig.width  = 10, fig.height=6}
-mypar(1,2,mar=c(2,2,2,5))
-image(t(microbiome[nrow(microbiome):1,]),
-      col=heat_pal,
-      axes=F,main="log2(CP1K+1)",
-      xlab="samples",ylab="bacteria",
-      line=.4,cex.main=1,font.main=1)
-text(  1+.02 ,seq(1,0,length.out = nrow(zscore)) , labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
-
-image(t(zscore[nrow(zscore):1,]),
-      col=heat_pal,
-      axes=F,main="Z-score log2(CP1K+1)",
-      xlab="samples",ylab="bacteria",
-      line=.4,cex.main=1,font.main=1)
-text(  1+.02 ,seq(1,0,length.out = nrow(zscore)) , labels = rownames(zscore), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
-```
-
-\clearpage
-
-# Computing a SNN graph from sample correlations
-
-```{r, echo = FALSE}
-mypar(2,2)
-
-#Compute sample-wise correlations
-cors <- cor(zscore,method = "pearson")
-eucl <- as.matrix(dist(t(zscore)))
-mean_local <- quantile(eucl,probs = .9)
-hist(eucl,breaks = seq(min(eucl),max(eucl),length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="euclidean distance",xlim=c(min(eucl),max(eucl)))
-abline(v=median(eucl),col="red",lty=1,xpd=F)
-abline(v=mean_local,col="red",lty=2,xpd=F)
-
-#Compute KNN graph from correlation matrix
-knn <- RANN::nn2(cors,k = k)
-# knn <- RANN::nn2(t(zscore),k = k)
+#Compute kNN graph
+knn <- RANN::nn2( (1-cors)/2 ,k = k_nearest_neighbours)
 nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
@@
 
-#Compute SNN graph from KNN
-jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
+#Calculate SNN and prune based on Jaccard index
+N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
+P <- t(nn) %*% nn
+B <- N - nn
+C <- t(B) %*% B
+Q <- nrow(nn)*N - C
+J <- P/Q
+J[is.nan(J)] <- 0
+J[J < snn_threshold] <- 0
+rownames(J) <- rownames(cors)
+colnames(J) <- colnames(cors)
 
-#Prune SNN graph based on the Jaccard index
-pSNN <- jSNN
-pSNN <- pSNN / max(pSNN)
+#Build a graph object, perform clustering and define the force-directed layout
+g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
+write.csv(J,"../../results/bacteria_SNN_graph.csv",row.names = T)
 
-hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="SNN jaccard",xlim=c(0,1))
-abline(v=P,col="red",lty=2,xpd=F)
+cl <- igraph::cluster_louvain(g)
+# cl <- igraph::walktrap.community(g)
+set.seed(1)
+l <- layout_with_fr(g,niter=3000,start.temp=1)
 
-hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
-     border=NA,col="grey60",las=1,
-     xaxs="i",yaxs="i",
-     main="correlation coeficient",xlim=c(-1,1))
-abline(v=0,col="red",lty=1,xpd=F)
-
-pSNN[ pSNN < P ] <- 0
-pSNN[ pSNN > P ] <- cors[pSNN > P]
-pSNN[ pSNN < 0 ] <- 0
-
-g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
+#Define layout with UMAP
+mU <- uwot::umap((1-cors)/2, n_neighbors = 10)
 
@@
 
-\clearpage
+```{r Plot BC, echo=FALSE,fig.width=10,fig.height=10}
 
-# Visualise the data
+mypar(3,3,mar=c(0,0,2,0))
 
-```{r, echo = FALSE,fig.width  = 10, fig.height=3}
-mypar(2,2,mar=c(0,0,2,10))
-
-set.seed(1)
-# l <- layout_with_graphopt(g,niter = 1000)
-l <- layout_with_fr(g,niter=3000,start.temp=30)
-
-cl <- igraph::cluster_louvain(g)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
+plot( g , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(cl$membership)] , 
     edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
+    vertex.size=10, main="graph layout FR",
     edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
+plot( g , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(cl$membership)] , 
+    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
+    vertex.size=10, main="UMAP",
+    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
+par(mar=c(0,0,0,0))
+empty_plot()
+legend("topleft",title = "bact. communities",
      legend = levels(factor(cl$membership)),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
-title("Louvain clusters",cex.main=1,font.main=1)
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[dataset_origin] , 
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-       y = par("usr")[c(4)],
-       legend = levels(dataset_origin),
-       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
-title("dataset",cex.main=1,font.main=1)
-
-# plot( g , vertex.label.cex=0.000001 , 
-#       vertex.color = hue_pal()(20)[factor(metadata[common_samples,"group_M9_CVLv3"])] ,
-#       edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-#       vertex.size=10,
-#       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-# legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
-#        y = par("usr")[c(4)],
-#        legend = levels(factor(metadata[common_samples,"group_M9_CVLv3"])),
-#        bty = "n",pch = 21,pt.bg = hue_pal()(20),pt.cex = 1,xpd=T)
-# title("group_M9_CVLv3",cex.main=1,font.main=1)
-
+     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
 ```
 
-\clearpage
-
-# Computing differential expression across clusters
-
-```{r, echo = FALSE}
-
-NN <- min(table(cl$membership))
-
-res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
-for(i in unique(cl$membership)){
-  for(j in rownames(microbiome) ){
-
-    set.seed(1)
-    a <- c(sample(microbiome[j,cl$membership == i],NN))
-    set.seed(1)
-    b <- sample(microbiome[j,cl$membership != i],NN)
-
-    perc1 <- sum(microbiome[j,cl$membership == i]>0) / sum(cl$membership == i)
-    perc2 <- sum(microbiome[j,cl$membership != i]>0) / sum(cl$membership != i)
-
-    temp <- wilcox.test(x=a, y=b)
-
-    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )
-
-    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
-                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
-    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
-  }
-}
-res <- res[-1,]
-res$pvalue <- as.numeric(res$pvalue)
-res$perc.1 <- as.numeric(res$perc.1)
-res$perc.2 <- as.numeric(res$perc.2)
-res$perc.diff <- res$perc.1 - res$perc.2
-
-res$fc <- as.numeric(res$fc)
-res$FDR <- p.adjust(res$pvalue)
-res <- res[order(res$pvalue),]
-res <- res[abs(res$fc) > fct,]
-res <- res[abs(res$pvalue) < pvt,]
-res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
-dim(res)
-```
-
-#Plotting the most significant bacteria across clusters
-
-```{r, echo = FALSE, fig.width  = 10, fig.height=14}
-library(dplyr)
-res %>% group_by(cluster)  %>% top_n(-20, pvalue) %>% top_n(10, fc) -> top7
-top7 <- top7[top7$fc > 0 , ]
-
-groupings <- factor(cl$membership)
-ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(microbiome, x, groupings)}))
-
-mypar(2,2,mar=c(2,15,1,1))
-
-plot_dots(microbiome, names(sort(ord)) ,
-          clustering = cl$membership,
-          show_grid = T,cex.main=1,font.main=1,
-          cex.row = .8,cex.col = .8,srt = 0)
-
-barlist( microbiome , names(sort(ord)), clustering = cl$membership, col=pal,cex.axis = .8)
-
-violist( microbiome , names(sort(ord)), clustering = cl$membership,cex.axis = .8,
-         col=pal,smooth = 2,transparency = 60,pt.col = "black")
-
-o <- order(cl$membership)
-image(t(zscore[rev(names(sort(ord))),o]),col=heat_pal,axes=F,border=NA,
-      main="",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
-text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)])) , seq(1,0,length.out = length(names(sort(ord)))),
-      labels = names(sort(ord)), srt = 0, adj = c(1,.5), xpd = TRUE, cex=.8)
-points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
-        col=pal[factor(cl$membership)][o] ,xpd=T)
-
-```
-
-#Plotting the most significant bacteria across PREVIOUS ANNOTATION
-
-```{r, echo = FALSE, fig.width  = 10, fig.height=11}
-mypar(2,2,mar=c(2,15,1,1))
-
-common_samples <- colnames(microbiome)[colnames(microbiome) %in% metadata$ID]
-
-plot_dots(microbiome[,common_samples], names(sort(ord)) ,cex.row = .8,cex.col = .8,
-          clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),
-          show_grid = T,cex.main=1,font.main=1,
-          srt = 90)
-
-barlist( microbiome[,common_samples] , names(sort(ord)),cex.axis = .8,
-         clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),
-         col=hue_pal()(20),srt = 90)
-
-violist( microbiome[,common_samples] , names(sort(ord)), 
-         clustering = factor(metadata[common_samples,"group_M9_CVLv3"]),cex.axis = .8,
-         col=hue_pal()(20),smooth = 2,transparency = 60,pt.col = "black",srt = 45)
-
-o <- order(factor(metadata[common_samples,"group_M9_CVLv3"]))
-image(t(zscore[rev(names(sort(ord))),o]),col=heat_pal,axes=F,border=NA,
-      main="",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
-text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)])) , seq(1,0,length.out = length(names(sort(ord)))),
-      labels = names(sort(ord)), srt = 0, adj = c(1,.5), xpd = TRUE, cex=.8)
-points( (0:(length(factor(metadata[common_samples,"group_M9_CVLv3"]))-1) )/(length(factor(metadata[common_samples,"group_M9_CVLv3"]))-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(factor(metadata[common_samples,"group_M9_CVLv3"]))), pch=16,
-        col=hue_pal()(20)[factor(metadata[common_samples,"group_M9_CVLv3"])][o] ,xpd=T)
-
-```
-
-\clearpage
-
-# Plotting the most significant bacteria across clusters
-
-```{r, echo = FALSE, fig.width  = 10, fig.height=11}
-mypar(6,6,mar=c(.5,.5,2,.5))
-
-plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] ,
-      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
-      vertex.size=10,
-      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-empty_plot()
-legend("topleft",legend = levels(factor(cl$membership)),bty = "n",pch = 21,pt.bg = pal,pt.cex = .8)
-
-# plot( l , bg=pal[factor(cl$membership)],
-#       pch=21,main="clusters",cex.main=1,font.main=1,
-#       frame=F,axes=F,cex=1,xlab="",ylab="")
-# empty_plot()
-# legend("topleft",legend = levels(factor(cl$membership)),bty = "n",pch = 21,pt.bg = pal)
-
-res2 <- res[order(res$cluster),]
-for( j in 1:nrow(res2) ){
-  i <- res2$bacteria[j]
-  temp <- microbiome[i,]
-  temp <- temp / max(temp)
-  o <- order(temp)
-  plot( l[o,] , col=c("grey85", colorRampPalette(c("grey80","grey60","navy"))(50)) [temp[o]*49+1],
-        pch=16,main=paste0(i,"\ncluster=",res2$cluster[j]," (",ifelse(res2$fc[j]>0,"up)","down)")),
-        cex.main=1,font.main=1,frame=F,axes=F,cex=1,xlab="",ylab="")
-}
-```
-
-\clearpage
-
-#Plotting bacteria across clusters
-
-```{r, echo = FALSE, fig.width  = 10, fig.height=6}
-mypar(2,2,mar=c(12,6,2,2))
-o <- order(cl$membership)
-image(zscore[rev(names(sort(ord))),o][,ncol(zscore):1],col=heat_pal,axes=F,border=NA,
-      main="bacteria",xlab="",ylab="samples",line=.4,cex.main=1,font.main=1)
-text( seq(1,0,length.out = length(names(sort(ord)))) ,
-      par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]) , 
-      labels = names(sort(ord)), srt = 60, adj = c(1,.5), xpd = TRUE, cex=.7)
-points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl$membership)), pch=16,
-    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
-    col=pal[factor(cl$membership)][o] ,xpd=T)
-
-par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)])
-
-image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
-      main="sample correlation matrix",xlab="samples",ylab="samples",line=.4,cex.main=1,font.main=1)
-points( (0:(length(cl$membership)-1) )/(length(cl$membership)-1),
-        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(cl$membership)), pch=16,
-        col=pal[factor(cl$membership)][o] ,xpd=T)
-points( rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(cl$membership)), pch=16,
-    rev((0:(length(cl$membership)-1) )/(length(cl$membership)-1)),
-    col=pal[factor(cl$membership)][o] ,xpd=T)
-```
-
-\clearpage
-
-#Plotting bacteria across clusters
-
-```{r, echo = FALSE, fig.width  = 10, fig.height=6}
-mypar(2,2)
-barlist( rbind( BacsDetected = colSums(microbiome > 0) ,
-                TotalCounts = colSums(microbiome),
-                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
-         , c("BacsDetected","TotalCounts","CountsPerBac"),
-         clustering = metadata[colnames(microbiome),"group_M4_CVLv3"],
-         col=pal)
-
-violist( rbind( BacsDetected = colSums(microbiome > 0) ,
-                TotalCounts = colSums(microbiome),
-                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
-         , c("BacsDetected","TotalCounts","CountsPerBac"),
-         clustering = metadata[colnames(microbiome),"group_M4_CVLv3"],
-         col=pal,transparency = 50,pt.col = "black",smooth = 2)
-
-barlist( rbind( BacsDetected = colSums(microbiome > 0) ,
-                TotalCounts = colSums(microbiome),
-                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
-         , c("BacsDetected","TotalCounts","CountsPerBac"),
-         clustering = cl$membership,
-         col=pal)
-
-violist( rbind( BacsDetected = colSums(microbiome > 0) ,
-                TotalCounts = colSums(microbiome),
-                CountsPerBac = colSums(microbiome)/colSums(microbiome > 0) )
-         , c("BacsDetected","TotalCounts","CountsPerBac"),
-         clustering = cl$membership,
-         col=pal,transparency = 50,pt.col = "black",smooth = 2)
-
-table(list(cl$membership,metadata[colnames(microbiome),"group_M4_CVLv3"]))
-```
-
-\clearpage
-
-# Saving clusters and differentially expressed bacteria
-
-```{r, eval=T,echo = FALSE}
-write.csv2( data.frame(cl$membership,row.names = colnames(microbiome)) ,row.names = T,
-            paste0(PATH,"/results/Sample_Louvain_clusters_",params$dataset) )
-
-write.csv2( data.frame(l,row.names = colnames(microbiome)) ,row.names = T,
-            paste0(PATH,"/results/Sample_graph_layout_",params$dataset) )
-
-write.csv2( res ,
-            paste0(PATH,"/results/DGE_",params$dataset),row.names = T )
-```
+**Figure 3.** Bacterial community embedding of `r k`-SNN graph clustered using Louvain community detection algorithm based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).

### 2022-02-10 reports/rmarkdown/02_individual_reports.Rmd
don't save tex file

@@
     fig_caption: yes
-    keep_tex: yes
+    toc: true
 header-includes: 

### 2022-02-10 reports/rmarkdown/manuscript/Figure3.Rmd
Save pdf image to common folder

@@
     fig_caption: yes
-    keep_tex: yes
 header-includes: 
@@
 
+```{r setup}
+knitr::opts_chunk$set(
+  fig.path="./Figures/",
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+```
+
 ```{r message=FALSE, warning=FALSE, include=FALSE}
+
 ##################
@@
 ##################
-#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
 suppressWarnings({suppressMessages({suppressPackageStartupMessages({
@@
 
-# m <- metadata %>%
-#   select(ID, Luminal_gr, Tissue_gr) %>%
-#   left_join(div, by="ID") %>%
-#   arrange(Tissue_gr) %>%
-#   group_by(Tissue_gr) %>%
-#   nest() %>%
-#   mutate(L = map(data, ~arrange(.x, Luminal_gr, Shannon))) %>%
-#   unnest(L)
-
-# metadata <- metadata %>%
-#   arrange(factor(ID, levels = m$ID))
-
 group_annotation <- factor(setNames(m$Luminal_gr,m$ID))
@@
 
-```{r Figure 3., echo=FALSE, warning=FALSE, message=FALSE,  fig.pos="t"}
-#  fig.width=6 ,fig.asp=1, fig.height=7,
+```{r Figure 3, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.pos="t"}
+# fig.asp=1, fig.height=7,
 ### A

### 2022-02-10 reports/rmarkdown/manuscript/Suppl.Figure1.Rmd
Save pdf image to common folder

@@
 
+```{r setup}
+knitr::opts_chunk$set(
+  fig.path="./Figures/",
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+```
+
 ```{r message=FALSE, warning=FALSE, include=FALSE}
@@
 
-```{r Bacterial abundance, echo = FALSE, eval = TRUE,  fig.asp=.5}
-par(mar=c(6,2.5,0,0)) #b,l,t,r
-
+```{r Suppl.Fig.1 Bacterial_abundance, echo = FALSE, eval = TRUE,  fig.asp=.5}
 top_all <- lapply( datasets_all_samples[2:3] , function(x){
@@
 
-temp <- factor(top, levels=top)
+###################################
+# BACTERIAL ABUNDANCE IN DATASETS #
+###################################
+par(mar=c(6,2.5,0,0)) #b,l,t,r
 violist(data = x[,colnames(x)%in%top],

### 2022-02-10 reports/rmarkdown/01_data_pre_processing.Rmd

--- 2021-11-18 reports/rmarkdown/01_data_pre_processing.Rmd
@@
                       warning    = FALSE)
-
-setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
 ```
@@
 library(niceRplots)
-library(tidyverse)
 
@@
 ```{r, echo = FALSE}
-PATH <- "../../" #Path to the data
+PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data
+PATH <- "../../data/" #Path to the data
 
@@
 
-```{r, Functions}
-# create "other" taxonomy
-other_taxa.fun <- function(df) {
-  other_taxa <- df %>%
-  unite("taxa_other", Kingdom:Species, sep = ";", remove = F, na.rm = T) %>%
-  mutate(taxa_other = ifelse(is.na(.$Species), paste0(.$taxa_other, ";other"), .$taxa_other)) %>%
-  #mutate(taxa_other = str_replace(.$taxa_other, ";NA.+|;NA", ";other")) %>%
-  mutate(taxa_other = str_extract(.$taxa_other, "([^;]+);([^;]+)$")) %>% # get the two last taxa levels
-  mutate(taxa_other = sub('^(.*);(other)', '\\2;\\1', .$taxa_other)) # place "other" first
-return(other_taxa)
-}
-
-# create genus lacto taxonomy
-genus_lacto.fun <- function(df){
-  genus_lacto <- df %>%
-    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
-    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste("L.", "crispatus/acidophilus"), 
-                                ifelse(grepl("reuteri",.$Species), paste("L.", "reuteri/oris/frumenti/antri"),
-                                  ifelse(grepl("gasseri",.$Species), paste("L.","gasseri/johnsonii/taiwanensis"),
-                                    ifelse(grepl("murinus",.$Species), paste("L.","murinus/animalis/apodemi/salivarius"),
-                                      ifelse(grepl("plantarum",.$Species), 
-                                           paste("L.", "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
-                                        ifelse(grepl("Lactobacillus",.$Genus), paste("L.", .$Species), 
-                                              as.character(.$Genus_lacto))))))) )
-  return(genus_lacto)
-}
-
-aggregate.fun <- function(df, level){
-  level <- enquo(level)
-  aggregated <- df %>%
-    dplyr::rename(Taxonomy = !!level) %>%
-    unite(., "Full_taxonomy", Kingdom:Species, sep = ";") %>%
-    dplyr::select(-any_of(c("Sequence", "Seq_ID", "Genus_lacto", "taxa_other", "Full_taxonomy"))) %>%
-    group_by(Taxonomy) %>%
-    summarise(across(where(is.numeric), sum)) %>%
-    ungroup()
-return(aggregated)
-}
-```
-
 # Loading data and metadata
@@
 # Load metadata
-dataset_names <- c("ASV_tissue_V3_B2.csv",   # Tissue, Boston run 2 (95 samples)
-                     "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
-                     "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
-                     "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
-                     "ASV_CVL_V2_S.csv",  # CVL V2, CTMR         (62 samples)
-                     "ASV_CVL_V3_B1.csv") # CVL V3, Boston run 1 (111 samples)
+metadata <- read.csv(paste0(PATH,"/Clinical_visit_2_3_updatesept28.csv"))
+rownames(metadata) <- as.character(metadata$ID)
 
-datasets_ <- map(dataset_names, ~read.csv( paste0(PATH,"data/",.x))) %>% set_names(., dataset_names)
-metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
-```
-
-# Taxonomy agglomeration
-
-```{r, echo = FALSE}
-temp <- map(datasets_, ~.x  %>%
-            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
-            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
-            other_taxa.fun(.) %>% # create "other" taxonomy 
-            genus_lacto.fun(.)  # create genus lacto taxonomy
-            ) %>%  
-        {. ->> temp_ } %>% 
-        map(., ~aggregate.fun(., "Genus_lacto")) # agglomerate on genus/lacto spp. lvl.
-
-# Extract taxonomic information to file
-seq_taxa <- bind_rows(temp_, .id = "dataset") %>% select(1:12) %>% unique() 
-#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )
-
-# make long format raw non agglom ASV files
-# remove_from_CVL_V2 <- intersect(colnames(temp_[[5]]), colnames(temp_[[3]]))[-c(1:11)]
-# names <- c("ASV_tissue_V3_Boston run2","ASV_tissue_V3_Boston run1", "ASV_CVL_V2_Boston run1", 
-#                      "ASV_CVL_V2_Boston run2", "ASV_CVL_V2_Stockholm", "ASV_CVL_V3_Boston run 1") 
-# ASV_long <- temp_ %>%
-#   modify_at(., c("ASV_CVL_V2_C.csv"), ~select(.x, -any_of(remove_from_CVL_V2))) %>% 
-#   map(., ~pivot_longer(.x, matches("^P\\d"), names_to = "ID", values_to = "Reads")) %>%
-#   set_names(names ) %>%
-#   bind_rows(., .id = "dataset") %>%
-#   separate(., dataset, into=c(NA, "Sample_type", "Visit","Place_sequenced"), sep = "_", remove = F) %>% 
-#   mutate(dataset = str_extract(.$dataset, "CVL_V\\d|tissue_V3")) %>% 
-#   group_split(dataset, .keep=F) %>% set_names(c("tissue", "CVL_V2", "CVL_V3"))
-# 
-# saveRDS(ASV_long, paste0(PATH,"results/","ASV_all_data_long_format.RDS"))
-
-#arrange all datasets to have the same rownames with all bacteria
-common_microbes_fun <- function(x) {
-  t <- bind_rows(temp) %>%
-    select(Taxonomy) %>%
-    unique() %>%
-    left_join(x, by="Taxonomy") %>%
-    mutate(across(Taxonomy, ~replace_na(.x, "Not assigned"))) %>%
-    mutate(across(-Taxonomy, ~replace_na(.x, 0))) %>%
-    column_to_rownames(var = "Taxonomy")
-  return(t)
-}
-datasets_ <- map(temp, ~common_microbes_fun(.x))
-lapply(datasets_,dim)
-
-# datasets <- datasets[3] #in case you wanna test in 1 dataset
-
-# Transform phyloseq data into normalized matricies ( log2[CP1K] )
-# datasets <- lapply(datasets,function(x){
-#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
-# })
-# lapply(datasets,dim)
-
-```
-
-# Loading data and metadata Paulos code
-
-```{r, eval=F, echo = FALSE}
-PATH <- "../../" #Path to the data
-# Load metadata
-metadata <- read.csv(paste0(PATH,"data/Clinical_visit_2_3_updatefeb22.csv"))
-rownames(metadata) <- as.character(metadata$ID)
 # Load datasets
@@
                      "ASV_CVL_V3.csv")
+
 #import datasets and return matrix with taxa-level resolution
 datasets <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"data/",x))
+  x <- read.csv( paste0(PATH,"/",x))
   temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
@@
   use <- !is.na(taxa) & (taxa != "NA")
+
   temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
@@
 # datasets <- datasets[3] #in case you wanna test in 1 dataset
+
 # Transform phyloseq data into normalized matricies ( log2[CP1K] )
@@
 taxa_info <- lapply(dataset_names, function(x){
-  x <- read.csv( paste0(PATH,"data/",x))
+  x <- read.csv( paste0(PATH,"/",x))
   temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
@@
 taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
-write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
-
+write.csv2(taxa_info, paste0(PATH,"../results/taxonomy.csv"),row.names = T)
 ```
@@
 ```{r, echo = FALSE}
-datasets <- datasets_
-names(datasets)
-
 #No batch correction needed in tissue_V3 dataset, only add it to the other samples
@@
 ```{r, echo = FALSE}
-#mypar(3,3)
-par(mar=c(3, 3, 0, 0))
+mypar(3,3)
 PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
@@
 
-if(file.exists(paste0(PATH,"results/encoder.h5")) & 
-  file.exists(paste0(PATH,"results/decoder.h5")) & 
-  file.exists(paste0(PATH,"results/aen.h5")) ){
-  encoder <- load_model_hdf5(paste0(PATH,"results/encoder.h5"))
-  decoder <- load_model_hdf5(paste0(PATH,"results/decoder.h5"))
-  aen <- load_model_hdf5(paste0(PATH,"results/aen.h5"))
+if(file.exists(paste0(PATH,"../results/encoder.h5")) & 
+  file.exists(paste0(PATH,"../results/decoder.h5")) & 
+  file.exists(paste0(PATH,"../results/aen.h5")) ){
+  encoder <- load_model_hdf5(paste0(PATH,"../results/encoder.h5"))
+  decoder <- load_model_hdf5(paste0(PATH,"../results/decoder.h5"))
+  aen <- load_model_hdf5(paste0(PATH,"../results/aen.h5"))
 
@@
 
-  save_model_hdf5(encoder, paste0(PATH,"results/encoder.h5") )
-  save_model_hdf5(decoder, paste0(PATH,"results/decoder.h5"))
-  save_model_hdf5(aen, paste0(PATH,"results/aen.h5"))
+  save_model_hdf5(encoder, paste0(PATH,"../results/encoder.h5") )
+  save_model_hdf5(decoder, paste0(PATH,"../results/decoder.h5"))
+  save_model_hdf5(aen, paste0(PATH,"../results/aen.h5"))
 }
@@
 samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
-datasets[["ASV_CVL_V2"]] <- as.data.frame.array(cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] ))
+datasets[["ASV_CVL_V2"]] <- cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] )
 datasets <- datasets[-c(1,2,3)]
@@
 
-common_microbes.fun <- function(x) {
-  xy <- rownames_to_column(x, var = "Taxa")
-  d <- data.frame(Taxa = common_microbes) %>%
-    left_join(., xy, by="Taxa") %>%
-    mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
-    column_to_rownames(var = "Taxa")
-  return(d)
-}
-datasets <- map(datasets, ~common_microbes.fun(.x))
+datasets <- lapply(datasets, function(x) {
+  temp <- matrix(0,nrow = length(common_microbes),
+                 ncol = ncol(x),
+                 dimnames = list(common_microbes,colnames(x)))
+  temp[rownames(x),colnames(x)] <- x
+  return(temp)
+})
 lapply(datasets,dim)
@@
 
-# raw files
-n <- c("ASV_CVL_V3", "ASV_tissue_V3", "ASV_CVL_V2")
-map2(datasets[1:3],n,  ~write.csv(.x, file=paste0(PATH,"results/",.y,"_raw_counts.csv"),row.names = T) )
-
-# normalized files
 datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
@@
 
-write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
-write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )
+write.csv( datasets[["ASV_CVL_V2_raw"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = T )
 
@@
 batches_CVL2 <- t(t(setNames(batches_CVL2 , colnames(datasets[["ASV_CVL_V2_raw"]]))))
-write.csv( batches_CVL2 , paste0(PATH,"results/batches_CVL2.csv"),row.names = T )
+write.csv( batches_CVL2 , paste0(PATH,"../results/batches_CVL2.csv"),row.names = T )
 ```
@@
 ```{r, echo = FALSE}
-trx <- read.csv(paste0(PATH,"data/TRX_rawcounts.csv"),row.names = 1)
+trx <- read.csv(paste0(PATH,"/TRX_rawcounts.csv"),row.names = 1)
 trx <- rowsum(trx[,-c(1,2)],trx$symbol)
@@
 trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
-write.csv( trx , paste0(PATH,"results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
+write.csv( trx , paste0(PATH,"../results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
 ```

### 2022-02-10 reports/rmarkdown/01_taxonomy_improved.Rmd

@@
 
-#Load libraries and functions
+# Load libraries, data and metadata
 
-```{r, echo = FALSE}
+```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
+##################
+# LOAD LIBRARIES #
+##################
+suppressWarnings({suppressMessages({suppressPackageStartupMessages({
 library(tidyverse)
@@
 library(Biostrings)
+})  })  })
+
+#############
+# LODA DATA #
+#############
+paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
+             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
+             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region
+
+n <- c("boston_r1", "boston_r2", "Sthlm")
+ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+
+seqtabs <- map(ps_list, ~data.frame(.x@otu_table@.Data))
+taxtabs <- map(ps_list, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
+sample_info <- map(ps_list, ~as_tibble(sample_data(.x), rownames = 'SampleID'))
+
+######################
+# DOWNLOAD DATABASES #
+######################
+db_dir <- "../../data/database/"
+if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }
+
+url <- c( "https://raw.githubusercontent.com/ctmrbio/BVAB-and-Lac-sequences/master/BVAB_rRNA_database.fa",
+          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_db.fasta.gz",
+          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_seqinfo.csv")
+
+walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
+system("gunzip ../../data/database/*.gz")
+
+OptiVag <- read_csv("../../data/database/optivag_seqinfo.csv", col_names = T)
 ```
@@
 ```{r function}
-######################## FASTA Function ########################
+################## 
+# FASTA FUNCTION #
+##################
 write.fasta <- function(df, file.out, open = "w"){
@@
 
-```
-
-# Loading data and metadata
-
-```{r Get ASV and tax tables, message=FALSE}
-paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
-             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
-             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region
-
-n <- c("boston_r1", "boston_r2", "Sthlm")
-ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
-
-seqtabs <- map(ps_list, ~data.frame(.x@otu_table@.Data))
-taxtabs <- map(ps_list, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
-sample_info <- map(ps_list, ~as_tibble(sample_data(.x), rownames = 'SampleID')) 
 ```
@@
 ```{r SeqID dict, include=FALSE, results='hide'}
-############### Create dictonary for Sequence id ####
+########################
+# CREATE SEQUENCE ID's #
+########################
 Seq_ID_key <-seqtab.taxa %>% 
@@
 # BLAST
-## Download databases
-
-```{r Download data base}
-db_dir <- "../../data/database/"
-if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }
-
-url <- c( "https://raw.githubusercontent.com/ctmrbio/BVAB-and-Lac-sequences/master/BVAB_rRNA_database.fa",
-          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_db.fasta.gz",
-          "https://github.com/ctmrbio/optivag/raw/master/database/db/16S/v0.1/optivag_seqinfo.csv")
-
-walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
-system("gunzip ../../data/database/*.gz")
-
-OptiVag <- read_csv("../../data/database/optivag_seqinfo.csv", col_names = T)
-```
 

### 2022-02-10 reports/rmarkdown/03_integrated_datasets.Rmd

@@
 \clearpage
+
+```{r eval=FALSE, include=FALSE}
+groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
+  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
+names(groupings) <- names(datasets_all_samples)[-1]
+all_names <- unique(unlist(lapply(groupings,function(x)rownames(x))))
+
+groupings_all_samples <- lapply(groupings,function(x){
+  df <- data.frame(group=rep("NA",length(all_names)), row.names = all_names,stringsAsFactors = F)
+  df[rownames(x),1] <- as.character(x[,1])
+  return(df)
+})
+names(groupings_all_samples) <- names(groupings)
+
+df <- as.data.frame(groupings_all_samples)
+colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
+metadata <- cbind(metadata,df[rownames(metadata),])
+```
 

### 2022-02-10 reports/rmarkdown/manuscript/Figure1.Rmd

@@
 
+```{r setup}
+knitr::opts_chunk$set(
+  fig.path="./Figures/",
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+```
+
 ```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
@@
 
-```{r Figure 1., echo=FALSE, fig.asp=1, fig.height=7, fig.pos="t", fig.width=6, message=FALSE, warning=FALSE}
+```{r Figure 1, echo=FALSE, fig.pos="t", fig.width=6, fig.height=5, fig.asp=1, message=FALSE, warning=FALSE}
 figlabels <- letters
+# fig.height=7, fig.asp=1,
 
@@
                                widths = c(1.87,1.42,.7,.8,.8,.7,.8),
-                               heights = c(1.6,.65,1.5,.95,.9,.2))
+                               heights = c(1.3,.55,1.33,.65,.9,0)) 
 ### A
@@
 #################################
-par(mar=c(0,3.3,0,1))
+par(mar=c(0,3.3,0,1)) #b,l,t,r
 fig1a <- magick::image_read("../../../supplementary_files/Scematic figure.pdf", density = 900 )
@@
 
-  par(mar=c(1,5,1,1.5)) #b,l,t,r
+  par(mar=c(1.5,5,1,1.5)) #b,l,t,r
   temp[is.nan(temp)] <- 0
@@
   text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
-
   points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
@@
 ###########
-par(mar=c(1,0,0,.5)) #b,l,t,r
+par(mar=c(0,0,0,.5)) #b,l,t,r
 empty_plot()
 legend(x = -0.04, y = 1.04, 
-       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other") ,
+       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other"),
        xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
@@
 
-par(mar=c(0,3.3,1.5,.5)) #b,l,t,r
+par(mar=c(0,3.3,0,.5)) #b,l,t,r
 empty_plot()
@@
 
-par(mar=c(0,0,1.5,1)) #b,l,t,r
+par(mar=c(0,0,0,1)) #b,l,t,r
 empty_plot()
@@
 
-par(mar=c(0,0,1,1))
+par(mar=c(0,0,0,1)) #b,l,t,r
 plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
@@
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-title(main = "Louvain", line = -2, cex.main = 1)
+title(main = "Louvain", line = -1, cex.main = 1)
 
@@
       edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
-title(main = "UMAP", line = -2, cex.main = 1)
+title(main = "UMAP", line = -1, cex.main = 1)
 

### 2022-02-10 reports/rmarkdown/manuscript/Suppl.Figure3-4.Rmd

@@
 
-```{r, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
+```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
 myletter <- letters
@@
 
+  ### Suppl. Figure 3-4
+  #######################################
+  # FUNCTIONAL ASSOCIATION BACT AND TRX #
+  #######################################
   par(mfrow=c(1,1),mar=c(12,0,2,4)) #b,l,t,r
@@
 
-**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
+**Suppl.
+Figure 3-4.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
 Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
@@
 
-#Combined representation of the two datasets
+# Combined representation of the two datasets
 layout(matrix(c(1,2,2,2,2),nrow = 1))
@@
 
-#Combined representation of the two datasets
+# Combined representation of the two datasets
 layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))

### 2022-02-10 test.Rmd

---
title: "test"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-02-10 reports/rmarkdown/manuscript/Figure2.Rmd

@@
 ---
+
+```{r setup}
+knitr::opts_chunk$set(
+  fig.path="./Figures/",
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+```
 
@@
 
-```{r Figure 2.,echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
+```{r Figure 2,echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
 

### 2022-02-10 reports/rmarkdown/manuscript/manuscript_template.Rmd
Check distribution

@@
 
+hist(x,breaks = 400)
+
 top <- sort(sapply( unique(colnames(x)),

### 2022-02-10 reports/02_Clustering.Rmd
move and rename file

--- 2022-02-10 reports/rmarkdown/02_individual_reports.Rmd
@@
   pdf_document:
-    fig_caption: yes
     toc: true

### 2022-02-10 reports/manuscript/Figure4-5.Rmd
Move files one lvl up

--- 2022-01-31 reports/rmarkdown/manuscript/Figure4-5.Rmd
@@
 ###########################
+# comparison <- sheets
+# enrichment_db <- "TF_PPI_enrichment"
+# degs <- enrich_T
+# group <- "Tissue_gr" 
+# gr_name <- "Tissue"
 arrange_plots.fun <- function(degs, group, gr_name, comparison, enrichment_db, file, Layout){
@@
 
-```{r Figure 4., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
+```{r Figure 4, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
 # Layout
@@
 # test figure
-dev.copy2pdf(file=paste0("./test.pdf"),
-    width = 6.6,
-    height = 6, paper = "a4"
-    )
+# dev.copy2pdf(file=paste0("./test.pdf"),
+#     width = 6.6,
+#     height = 6, paper = "a4"
+#     )
 ```
 
-```{r Figure 5., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
-# comparison <- sheets
-# enrichment_db <- "TF_PPI_enrichment"
-# degs <- enrich_T
-# group <- "Tissue_gr" 
-# gr_name <- "Tissue"
-
+```{r Figure 5, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
 # Layout
@@
 
-dev.copy2pdf(file=paste0("./test.pdf"),
-    width = 6.6,
-    height = 6, paper = "a4"
-    )
+# dev.copy2pdf(file=paste0("./test.pdf"),
+#     width = 6.6,
+#     height = 6, paper = "a4"
+#     )
 ```

### 2022-02-10 reports/manuscript/Figure6.Rmd
Move files one lvl up

--- 2022-01-31 reports/rmarkdown/manuscript/Figure6.Rmd
@@
 #############
+datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
 Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv")
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
+cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
 metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
-cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
 
-#color palettes
 #################
@@
 
-```{r Figure 6., echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
+```{r Figure 6, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
 figlabels <- letters
-
-# top differentially expressed proteins Luminal groups
+####################################
+# TOP DIFF. EXPRESSED PROT HEATMAP #
+####################################
+### A
+# Luminal groups
 t <- top_dpe.fun(dpe_L$top)
@@
 
-# top differentially expressed proteins Tissue groups
+### C
+# Tissue groups
 t <- top_dpe.fun(dpe_T$top)
@@
 
-# Luminal Cytokine plot 
+##############################
+# SIG. CYTOKINES VIOLIN PLOT #
+##############################
+### B
+# Luminal groups
 g <- c("IL1a", "MIG", "IP10", "IL1b", "IL12_p70", "MCP1")
@@
 
-# Tissue Cytokine plot
+### D
+# Tissue groups
 Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")

### 2022-02-10 reports/manuscript/Suppl.Figure2.Rmd
Move files one lvl up

--- 2022-01-31 reports/rmarkdown/manuscript/Suppl.Figure2.Rmd
@@
 
+```{r setup}
+knitr::opts_chunk$set(
+  fig.path="./Figures/",
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+```
+
 ```{r echo=FALSE}
@@
 
-```{r Suppl. Figure 9, eval=T,warning=FALSE, message=FALSE,echo = FALSE,fig.width=10,fig.height=9}
+```{r Suppl.Fig.2 Diff_bact_abundance, eval=T,warning=FALSE, message=FALSE,echo = FALSE,fig.width=10,fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
@@
 
+### A & B
+####################################
+# DIFFERENTIAL BACTERIAL ABUNDANCE #
+####################################
 figlabels <- letters

### 2022-02-10 reports/manuscript/Suppl.Figure5.Rmd
Move files one lvl up

--- 2022-01-08 reports/rmarkdown/manuscript/Suppl.Figure5.Rmd
@@
 
-```{r Print rare curves, warning=FALSE, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.asp=0.75,}
+```{r Suppl.Fig5 Rarefraction_plot, warning=FALSE, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.asp=0.75,}
+### A & B
+###########################
+# RAREFRACTION CURVE PLOT #
+###########################
 plot <- ggarrange(

### 2022-02-24 reports/01_taxonomy_improved.Rmd
initial commit

--- 2022-02-10 reports/rmarkdown/01_taxonomy_improved.Rmd
@@
 output:
-  pdf_document:
-    toc: yes
   html_document:
@@
 
-```{r Setup, echo = FALSE}
-knitr::opts_chunk$set(fig.width  = 10,
-                      results    = "hold",
+```{r Setup, include=FALSE}
+knitr::opts_chunk$set(results    = "hold",
                       message    = FALSE,
@@
 
-#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown")
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
 ```
@@
 
+#########
+# PATHS #
+#########
+result_dir <- "../results/01_taxonomy_improved_output/"
+db_dir <- "../resources/OptiVag_BVAB_database/"
+
 #############
@@
 #############
-paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
-             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
-             '../../data/phyloseq_Sthlm.RDS')       # CVL V2, only v4 region
+#### new ####
+paths_p <- c('../results/00_dada2_output/ps_objects/phyloseq_Boston_run1.RDS', # 108 Luminal samples, 1 Tissue sample
+             '../results/00_dada2_output/ps_objects/phyloseq_Boston_run2.RDS'  # 92 Tissue samples
+             )
+##### old ####
+paths_p <- c('phyloseq_boston_r1.RDS',   # 108 Luminal samples, 1 Tissue sample
+             'phyloseq_boston_r2.RDS')   # 92 Tissue samples
 
-n <- c("boston_r1", "boston_r2", "Sthlm")
+n <- c("boston_r1", "boston_r2")
 ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+
+#########################
+# SELECT SAMPLES TO USE #
+#########################
+mapping <- c("../data/Mapping/Mapping_info_Boston_run1.csv", "../data/Mapping/Mapping_info_Boston_run2.csv")
+m <- mapping %>% 
+  map(., ~read_csv(file = .x, col_names = TRUE)) %>% 
+  map(., ~pull(.x, "Sample_ID"))  %>% 
+  flatten()
+ps_list <- map(ps_list, ~subset_samples(.x, Sample_ID %in% m))
+#map2(ps_list, paths_p, ~saveRDS(.x, file = paste0("../data/", .y)))
 
@@
 ######################
-db_dir <- "../../data/database/"
 if( isFALSE(dir.exists(db_dir)) ) { dir.create(db_dir,recursive = TRUE) }
@@
 walk(url, ~download.file(.x, paste0(db_dir, basename(.x)), method="auto"))
-system("gunzip ../../data/database/*.gz")
+system("gunzip ../resources/OptiVag_BVAB_database/*.gz")
 
 OptiVag <- read_csv("../../data/database/optivag_seqinfo.csv", col_names = T)
+```
+
+```{r test identity, eval=FALSE, include=FALSE}
+paths_p <- c('../results/00_dada2_output/ps_objects/phyloseq_Boston_run1.RDS', # 108 Luminal samples, 1 Tissue sample
+             '../results/00_dada2_output/ps_objects/phyloseq_Boston_run2.RDS'  # 92 Tissue samples
+             )
+paths_p <- c('../../Gabriella_repo/data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVLv2)
+             '../../Gabriella_repo/data/phyloseq_boston_r2.RDS')       # CVL V2, only v4 region
+
+n <- c("boston_r1", "boston_r2")
+ps_list_new <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+
+seqtabs_ <- map(ps_list_new, ~data.frame(.x@otu_table@.Data))
+taxtabs_ <- map(ps_list_new, ~as_tibble(.x@tax_table@.Data, rownames = 'Sequence'))
+sample_info_ <- map(ps_list_new, ~as_tibble(sample_data(.x), rownames = 'SampleID'))
+
+int <- intersect(colnames(seqtabs[[1]]), colnames(seqtabs_[[1]]))
+setdiff(colnames(seqtabs[[1]]), colnames(seqtabs_[[1]]))
+setdiff(colnames(seqtabs_[[1]]), colnames(seqtabs[[1]]))
+
+old <- seqtabs_[[1]][row.names(seqtabs[[1]]), int]
+new <- seqtabs[[1]][, int]
+
+all_equal(old, new)
+
+old[1,1:5]
+new[1,1:5]
+
+head(seqtabs)
+
 ```
@@
 
-write.csv(Seq_ID_key, file=paste0("../../results/Seq_ID_key.csv"), quote=FALSE, row.names=FALSE)
+if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }
+write.csv(Seq_ID_key, file=paste0(result_dir,"Seq_ID_key.csv"), quote=FALSE, row.names=FALSE)
 
@@
 # Create Query files
-write.fasta(Seq_ID_key, paste0("../../results/BVAB_Query.fasta"))
-write.fasta(unnasigned_0, paste0("../../results/optivag_Query.fasta"))
+write.fasta(Seq_ID_key, paste0(result_dir,"BVAB_Query.fasta"))
+write.fasta(unnasigned_0, paste0(result_dir,"optivag_Query.fasta"))
 ```
@@
 names(fa) = paste(ID, names(fa))
-writeXStringSet(fa, "../../data/database/BVAB_db.fasta")
+writeXStringSet(fa, paste0(db_dir,"BVAB_db.fasta"))
 
 for(name in c("BVAB", "optivag")) {
-  db_dir     = "../../data/database/"
-  result_dir = "../../results/"
   if (isFALSE(is_empty(list.files(path = db_dir, pattern = paste0(name,".+\\.nhr"))))) {
@@
 
-  write_tsv(blast_out, paste0("../../results/", name, "_hits.tsv"))
+  write_tsv(blast_out, paste0(result_dir, name, "_hits.tsv"))
 }
@@
 ```{r Load BLAST hits, message=FALSE}
-blast_OptiVag <- read_table2(paste0("../../results/OptiVag_hits.tsv"))
-blast_BVAB <- read_table2(paste0("../../results/BVAB_hits.tsv"))
+blast_OptiVag <- read_table2(paste0(result_dir,"optivag_hits.tsv"))
+blast_BVAB <- read_table2(paste0(result_dir, "BVAB_hits.tsv"))
 
@@
 
-write.csv(new_taxtable, file=paste0("../../data/SeqID_to_taxa.csv"), quote=FALSE, row.names=FALSE)
+write.csv(new_taxtable, file=paste0(result_dir,"SeqID_to_taxa.csv"), quote=FALSE, row.names=FALSE)
 ```
@@
 write.csv(unnasigned_final, 
-          file=paste0("../../results/Unnasigned_tax_report.csv"), quote=FALSE, row.names=FALSE)
+          file=paste0(result_dir,"Unnasigned_tax_report.csv"), quote=FALSE, row.names=FALSE)
 
@@
 #### ASV with all controls and replicate samples removed ####
-ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3_B1", "ASV_tissue_V3_B1", 
-               "ASV_CVL_V2_B2", "ASV_tissue_V3_B2", "ASV_CVL_V2_S" )
+# ASV_names <- c("ASV_CVL_V2_B1", "ASV_CVL_V3_B1", "ASV_tissue_V3_B1", 
+#                "ASV_CVL_V2_B2", "ASV_tissue_V3_B2" )
+ASV_names <- c("ASV_CVL_V3_B1", "ASV_tissue_V3_B1", "ASV_tissue_V3_B2" )
 ASV_list <- map(ASV, ~ .x %>%
@@
 imap(ASV_list, 
-     ~write.csv(.x, file=paste0("../../data/",.y,".csv"), quote=FALSE, row.names=F) )
+     ~write.csv(.x, file=paste0(result_dir,.y,".csv"), quote=FALSE, row.names=F) )
 
-saveRDS(ASV, paste0(PATH,"results/","ASV_wCtrl_long_format.RDS"))
+saveRDS(ASV, paste0(result_dir,"ASV_wCtrl_long_format.RDS"))
 ```

### 2022-02-24 reports/02_data_preprocessing.Rmd
move normalization to seperate file

---
title: "Merge sequencing runs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)

#setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
```

# Load libraries and other scripts

```{r Load data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
library(tidyverse)
})  })  })

#########
# PATHS #
#########
#input_dir <- "../../Gabriella_repo/data/"
input_dir <- "../results/01_taxonomy_improved_output/"
result_dir <- "../results/02_data_preprocessing_output/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
dataset_names <- c("ASV_tissue_V3_B2.csv", # Tissue, Boston run 2 (95 samples)
                   "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
                   #"ASV_CVL_V2_B1.csv",   # CVL V2, Boston run 1 (27 samples)
                   #"ASV_CVL_V2_B2.csv",   # CVL V2, Boston run 2 (49 samples)
                   #"ASV_CVL_V2_S.csv",    # CVL V2, CTMR         (62 samples)
                    "ASV_CVL_V3_B1.csv")   # CVL V3, Boston run 1 (111 samples)

datasets <- map(dataset_names, ~read.csv(paste0(input_dir,.x))) %>% set_names(., dataset_names)
metadata <- read.csv(paste0("../data/", "metadata.csv"),row.names = 1)
trx <- read.csv(paste0("../data/","Raw_gene_counts_matrix.csv"),row.names = 1)

map(datasets, ~dim(.x))
```

```{r Functions}
# create "other" taxonomy
other_taxa.fun <- function(df) {
  other_taxa <- df %>%
  unite("taxa_other", Kingdom:Species, sep = ";", remove = F, na.rm = T) %>%
  mutate(taxa_other = ifelse(is.na(.$Species), paste0(.$taxa_other, ";other"), .$taxa_other)) %>%
  #mutate(taxa_other = str_replace(.$taxa_other, ";NA.+|;NA", ";other")) %>%
  mutate(taxa_other = str_extract(.$taxa_other, "([^;]+);([^;]+)$")) %>% # get the two last taxa levels
  mutate(taxa_other = sub('^(.*);(other)', '\\2;\\1', .$taxa_other)) # place "other" first
return(other_taxa)
}

# create genus lacto taxonomy
genus_lacto.fun <- function(df){
  genus_lacto <- df %>%
    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste("L.", "crispatus/acidophilus"), 
                                ifelse(grepl("reuteri",.$Species), paste("L.", "reuteri/oris/frumenti/antri"),
                                  ifelse(grepl("gasseri",.$Species), paste("L.","gasseri/johnsonii/taiwanensis"),
                                    ifelse(grepl("murinus",.$Species), paste("L.","murinus/animalis/apodemi/salivarius"),
                                      ifelse(grepl("plantarum",.$Species), 
                                           paste("L.", "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
                                        ifelse(grepl("Lactobacillus",.$Genus), paste("L.", .$Species), 
                                              as.character(.$Genus_lacto))))))) )
  return(genus_lacto)
}

aggregate.fun <- function(df, level){
  level <- enquo(level)
  aggregated <- df %>%
    dplyr::rename(Taxonomy = !!level) %>%
    unite(., "Full_taxonomy", Kingdom:Species, sep = ";") %>%
    dplyr::select(-any_of(c("Sequence", "Seq_ID", "Genus_lacto", "taxa_other", "Full_taxonomy"))) %>%
    group_by(Taxonomy) %>%
    summarise(across(where(is.numeric), sum)) %>%
    ungroup()
return(aggregated)
}
```

# Taxonomy agglomeration

```{r Agglomerate Taxa}
temp <- map(datasets, ~.x  %>%
            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
            other_taxa.fun(.) %>% # create "other" taxonomy 
            genus_lacto.fun(.)  # create genus lacto taxonomy
            ) %>%  
        {. ->> temp_ } %>% 
        map(., ~aggregate.fun(., "Genus_lacto")) # agglomerate on genus/lacto spp. lvl.

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_, .id = "dataset") %>% select(1:12) %>% unique() 
#write.csv(seq_taxa, paste0(result_dir,"SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )

#arrange all datasets to have the same rownames with all bacteria
common_microbes_fun <- function(x) {
  t <- bind_rows(temp) %>%
    select(Taxonomy) %>%
    unique() %>%
    left_join(x, by="Taxonomy") %>%
    mutate(across(Taxonomy, ~replace_na(.x, "Not assigned"))) %>%
    mutate(across(-Taxonomy, ~replace_na(.x, 0))) %>%
    column_to_rownames(var = "Taxonomy")
  return(t)
}
datasets <- map(temp, ~common_microbes_fun(.x))
lapply(datasets,dim)

```

# Merging and renaming

```{r merge sequencing runs}
names(datasets)

#No batch correction needed in tissue_V3 dataset, only add it to the other samples
datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_V3_B2.csv"]],datasets[["ASV_tissue_V3_B1.csv"]] )

datasets <- datasets[-c(1,2)]
lapply(datasets,dim)

```

```{r save raw microbiome files, fig.height=5, fig.width=5}

datasets <- lapply(datasets,function(x){
  x <- x[,sort(colnames(x))]
  return(x)
})

# raw files
n <- c("ASV_Luminal", "ASV_Tissue")
map2(datasets, n,  ~write.csv(.x, file=paste0(result_dir,.y,"_raw_counts.csv"),row.names = T) )

```

### 2022-02-24 reports/04_clustering.Rmd
improve readability

---
title: "Graph Based Clustering"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r save new metadata file, eval=FALSE, include=FALSE }
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")

# metadata <- read.csv2("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
# 
# meta <- metadata %>%
#   select(ID, PatID, Contraception, HIVstatus, BV_Diagnosis_v3, Luminal_names, Tissue_names, Luminal_gr, Tissue_gr,
#          BV_Lactobacillus_v3, BV_Monbilicus_v3, BV_Vaginal_Garda_v3)
# 
# write.csv(meta, "../data/metadata.csv")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6.6929133858, 
                      message    = FALSE,
                      warning    = FALSE)
#setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
```

```{r Load data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(igraph)
  library(rafalib)
  #remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
  library(niceRplots)
})  })  })

#########
# PATHS #
#########
input_dir <- "../results/03_normalize_data_output/"
result_dir <- "../results/04_clustering_output/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS(paste0(input_dir, "datasets_all_samples.RDS"))
# datasets_all_samples <- readRDS("../../Gabriella_repo/results/datasets_all_samples.RDS") %>%
#   set_names(., c("Tissue_RNAseq_V3_normalized", "ASV_Tissue_normalized", "ASV_Luminal_normalized", "ASV_CVL_V2_normalized_NOT_batch_corrected"))
metadata <- read.csv(paste0("../data/", "metadata.csv"),row.names = 1)

#########################
# SELECT SAMPLES TO USE #
#########################
# datasets_all_samples <- datasets_all_samples %>%
#   map(., ~as_tibble(.x, rownames = NA )) %>%
#   map(., ~select(.x, any_of(metadata$ID)))

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
bact_pal <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77', '#999933','#CC6677', '#882255', '#AA4499', '#DDDDDD')
```

```{r, fig.height=6}
###################################
# STUDY GROUPS GRAPH CONSTRUCTION #
####################################
all_microbiome <- datasets_all_samples[["ASV_Luminal_normalized"]]
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
dim(all_microbiome)

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 4 ] <- 4
zscore[ zscore < -4 ] <- -4
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)

mypar(mar=c(2,15,2,2))
top_var <- apply(all_microbiome,1,var)
top_var_t <- sort(top_var,decreasing = T)[100]
is_top_var <- top_var > top_var_t 
barplot(sort(top_var[is_top_var],decreasing = T)[30:1],horiz = T,las=1)

#Compute sample-wise correlations
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

# cors <- cor(all_microbiome, method = "pearson")
# adj <- (1-cors)/2
# pheatmap::pheatmap(adj,border_color = NA,color = colorRampPalette(c("grey90","grey70","grey70","grey70","orange3","firebrick"))(90) )

#Graph construction  
k <- 12 # k for KNN
p <- 0.1 # prunning threshold of SNN 0.14

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(adj,k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Computing SNN matrix
N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < p] <- 0
colnames(J) <- colnames(all_microbiome)
rownames(J) <- colnames(all_microbiome)

#Transforming SNN into an igraph object
g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,paste0(result_dir,"participant_SNN_graph.csv"),row.names = T)

```

```{r Plot Study Groups, fig.height=9, fig.width=10}
#####################
# PLOT STUDY GROUPS #
#####################
mypar(3,3,mar=c(0,0,2,0))

set.seed(1)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)

metadata$Louvain_clusters <- cl$membership
metadata$joint_clustering <- metadata$Luminal_gr
metadata$joint_clustering <- factor(metadata$joint_clustering) 
set.seed(1)
U <- uwot::umap(adj,n_neighbors = 30)
h <- hclust(as.dist(adj),method = "ward.D2")
# metadata$joint_clustering <- cutree(h,k = 5)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding\nlouvain",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Participant groups\nLouvain_gr",
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Luminal_gr)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Participant groups\nLuminal_gr",
       legend = levels(factor(metadata$Luminal_gr)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Tissue_gr)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,main="FR embedding",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(metadata$Tissue_gr)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,main="umap",
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
empty_plot()
legend("topleft",title = "Participant groups\nTissue_gr",
       legend = levels(factor(metadata$Tissue_gr)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1.2,xpd=T)
# title("Louvain clusters",cex.main=1,font.main=1)

```

```{r Bacterial communities}
#######################################
# BACTERIAL GROUPS GRAPH CONSTRUCTION #
#######################################
all_microbiome <- cbind(#datasets_all_samples[["ASV_Tissue_normalized"]],
                        datasets_all_samples[["ASV_Luminal_normalized"]]
                        )

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))

#Graph construction 
k_nearest_neighbours <- 5 # k for KNN
snn_threshold <- 0.1 # prunning threshold of SNN

#Compute kNN graph
knn <- RANN::nn2( (1-cors)/2 ,k = k_nearest_neighbours)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Calculate SNN and prune based on Jaccard index
N <- matrix(1,ncol = ncol(nn),nrow = nrow(nn))
P <- t(nn) %*% nn
B <- N - nn
C <- t(B) %*% B
Q <- nrow(nn)*N - C
J <- P/Q
J[is.nan(J)] <- 0
J[J < snn_threshold] <- 0
rownames(J) <- rownames(cors)
colnames(J) <- colnames(cors)

#Build a graph object, perform clustering and define the force-directed layout
g <- graph_from_adjacency_matrix(J,mode = "undirected",diag = F,weighted = T)
write.csv(J,paste0(result_dir,"bacteria_SNN_graph.csv"),row.names = T)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)
Bact.com <- tibble(Taxa=rownames(J), Bact_com=cl$membership)
write.csv(Bact.com,paste0(result_dir,"bacterial_communities.csv"),row.names = T)

set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=1)

#Define layout with UMAP
mU <- uwot::umap((1-cors)/2, n_neighbors = 10)

```

```{r Plot BC, fig.height=10, fig.width=10}
##############################
# PLOT BACTERIAL COMMUNITIES #
##############################
mypar(3,3,mar=c(0,0,2,0))
plot( g , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
    vertex.size=10, main="graph layout FR",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
plot( g , vertex.label.cex=0.000001 , vertex.color = bact_pal[factor(cl$membership)] , 
    edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=mU,
    vertex.size=10, main="UMAP",
    edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )

par(mar=c(0,0,0,0))
empty_plot()
legend("topleft",title = "bact. communities",
     legend = levels(factor(cl$membership)),
     bty = "n",pch = 21,pt.bg = pal,cex = .8,pt.cex = .9,xpd=T)
```

**Figure** Bacterial community embedding of `r k`-SNN graph clustered using Louvain community detection algorithm based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).

### 2022-02-24 reports/manuscript/Figure1.Rmd
Tweaking plot sizes and output

---
title: "Figure 1.Luminal Study Groups"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6.6929133858,
  fig.path="./Figures/",
  dev=c('pdf', 'png'), dpi=300,
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)

      # Create PNG
      # path2 <- xfun::with_ext(new_filename, "png")
      # img <- magick::image_read(new_filename)
      # magick::image_write(img, path2, format = "png")
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")

#### Figure requirements Microbiome: ####
  # - width of 85 mm for half page width figure
  # - width of 170 mm for full page width figure
  # - maximum height of 225 mm for figure and legend
  # - image resolution of approximately 300 dpi (dots per inch) at the final size
```

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

#########
# PATHS #
#########
result_dir <- "./Suppl.Tbl/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
fig1a <- magick::image_read("../../resources/Scematic figure.pdf", density = 900 )
datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
metadata <- read.csv("../../data/metadata.csv",row.names = 1)
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))

SNN_participant <- read.csv("../../results/04_clustering_output/participant_SNN_graph.csv",row.names = 1)
bac_communities <- read.csv("../../results/04_clustering_output/bacterial_communities.csv",row.names = 1)

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")

pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
# show_col(pal)
```

```{r Figure 1, echo=FALSE, fig.pos="t", fig.height=6.2, message=FALSE, warning=FALSE}
# , fig.height=6.2
figlabels <- letters
# fig.height=7, fig.asp=1,
print(knitr::is_latex_output())
layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                5,6,7,7,8,8,
                5,4,7,7,8,8,
                5,4,7,7,8,8) , ncol= 6,byrow = T),
                               widths = c(1.87,1.42,.7,.8,.8,.7,.8),
                               heights = lcm(c(4.7,2,4.5,2,.5)) # height in cm
                              # heights = c(1.4,.55,1.4,.61,.75,0)
       ) 
### A
#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0.1,3.7,0.3,1)) #b,l,t,r
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### B
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################

ASV <- list(c("Luminal_gr","ASV_Luminal_normalized"),
            c("Tissue_gr","ASV_Tissue_normalized")
         )
div_list <- list()
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))

  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","Simpson","Inv Simpson")

  div <- tibble(ID=names(shann),shann, simp, invsimp ) %>% rename_with(., ~paste0(c("ID","Shannon","Simpson","Inv Simpson")))
  div_list[[length(div_list) + 1]] <- div

  if(i[2] == "ASV_Tissue_normalized"){break}

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="", #bg="white", #cex=.6, 
         genes = c("Shannon","Simpson","Inv Simpson"),
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  if(i[2] == "ASV_Luminal_normalized"){
    end <- table(pal[factor(group_annotation[order(group_annotation)])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1.5,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

  o <- order( rowSums(temp) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=gsub("_gr","",i[1]),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = lvl)] ,names.arg = rep("",ncol(temp)))

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-7,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T)

  l <- c("BV:","Normal","Intermediate","BV")
  legend(x = 10, y = 111, #"bottom", inset = c(0, -0.5),
         bty = "n", horiz=TRUE, cex=0.7, xpd = T, 
         border = "white", 
         legend = l, x.intersp=0.5, xjust=0,
         text.width = c(0, map_dbl(l, ~strwidth(.x)/2)) ,
         fill=c(NA, "#d7d7d7","orange","tomato"))
  l2 <- c("HIV:","HIV-","HIV+")
  legend(x = 80, y = 111, # bottom position: par("usr")[3]-5
         bty = "n", horiz=TRUE, cex=0.7, xpd = T,
         border = "white",
         legend = l2, x.intersp=0.5, 
         text.width =c(0, map_dbl(l2, ~strwidth(.x)/2)),
         fill=c(NA, "#d7d7d7", "tomato"))
}

###########
# LEGENDS #
###########
par(mar=c(0,0,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, y = 1.04, 
       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other"),
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,4.5,0,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,0,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

### C + D
################
# STUDY GROUPS #
################

### Louvain ###
g <- graph_from_adjacency_matrix(as.matrix(SNN_participant), mode = "undirected",diag = F,weighted = T)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

par(mar=c(0,0,0,1)) #b,l,t,r
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10, layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "Louvain", line = -1, cex.main = 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### UMAP ###
all_microbiome <- datasets_all_samples[["ASV_Luminal_normalized"]]
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
U <- uwot::umap(adj,n_neighbors = 30)
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "UMAP", line = -1, cex.main = 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

```

```{r Suppl. Table 1., include=FALSE}
# Total Relative Abundance 
raw <- c("ASV_Luminal", "ASV_Tissue") %>% set_names()
raw_list <- raw %>%
  map(., ~read_csv(paste0("../../results/02_data_preprocessing_output/",.x,"_raw_counts.csv")))  %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_Luminal<0.5514392, "Other")) %>%
    mutate(name = ifelse(.$name=="Not assigned", "Other", .$name)) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_Luminal) %>%
    dplyr::rename(Luminal="ASV_Luminal",Tissue="ASV_Tissue") %>%
    rename(Taxa="name") %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 1)))

# Alfa Diversity Stats
div_stat <- div_list %>%
  set_names(map(ASV,1)) %>%
  map(., ~summarise_if(.x, is.numeric, list(median=median, IQR=IQR), na.rm = TRUE)) %>%
  map(., ~pivot_longer(.x, everything(), names_sep = "_", names_to = c(".value", "Stats"), )) %>%
  bind_rows(.id = "Sample type") %>%
  mutate("Sample type" = gsub("_gr", "", .$"Sample type")) %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 2)))

# Bacterial Communities
bac_communities <- bac_communities %>% 
  mutate("bacterial communities"= factor(paste0("BC",sprintf("%02d",bac_communities$Bact_com)))) %>%
  select(-Bact_com) %>%
  arrange(`bacterial communities`)

pages <- list("Total Relative Abundance"=r ,"Alfa Diversity"=div_stat, "Bacterial Communities"=bac_communities)
write.xlsx(pages, file=paste0(result_dir,"Suppl.Tbl.01",".xlsx"))

```

### 2022-02-24 reports/manuscript/Figure2.Rmd
Tweaking plot sizes and output

--- 2022-01-31 reports/rmarkdown/manuscript/Figure2.Rmd
@@
 title: "Figure 2. Bacterial Communities and Function"
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 output: 
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  fig.width = 6.6929133858,
+  fig.path="./Figures/",
+  dev=c('pdf', 'png'), dpi=300,
+  fig.process = function(filename){
+    new_filename <- stringr::str_remove(string = filename, 
+                                        pattern = "-1")
+    fs::file_move(path = filename, new_path = new_filename)
+    ifelse(fs::file_exists(new_filename), new_filename, filename)
+})
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
+```
+
 ```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
@@
 
+#########
+# PATHS #
+#########
+result_dir <- "./Suppl.Tbl/"
+if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }
+
 #############
@@
 #############
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
+datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)
 sample_use <- metadata$ID
-
-SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
-bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]
-
 group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
@@
 
-picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
-picrust_KO_seq <- read.table("../../../data/KO_predicted.tsv",header = T,row.names = 1)
-picrust_taxa <- read.table("../../../data/SeqID_to_taxa.csv",header = T,row.names = 1,sep = ",")
+SNN_bacteria <- read.csv("../../results/04_clustering_output/bacteria_SNN_graph.csv",row.names = 1)
+bac_communities <- as.matrix(read.csv("../../results/04_clustering_output/bacterial_communities.csv",row.names = 2))[,2]
+
+picrust_data <- read.table("../../results/05_picrust_output/out_2021-01-20/KO_metagenome_out/pred_metagenome_unstrat.tsv", header = T,row.names = 1)
+ko_names <- readLines('../../resources/picrust_data/ko_info.tsv')
+kegg_names <- read.delim('../../resources/picrust_data/KEGG_pathways_info.tsv',header = F)
+gmt_list <- fgsea::gmtPathways("../../resources/picrust_data/KEGG_pathways_to_KO.tsv")
+source("../../code/enrichment_function.R")
 
@@
 par(mar=c(1,5,4,1)) #b,l,t,r
-x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
+x <- datasets_all_samples[["ASV_Luminal_normalized"]]
 xL <- rowsum(x , grepl("L.",rownames(x)) )[2,] / sum(grepl("L.",rownames(x)))
@@
 
-ko_names <- readLines('../../../data/ko_info.tsv')
 ko_names <- strsplit(ko_names,"\t")
@@
 
-source("../enrichment_function.R")
-gmt_list <- fgsea::gmtPathways("../../../data/KEGG_pathways_to_KO.tsv")
-kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)
 names(gmt_list) <- kegg_names [ match ( names(gmt_list), as.character(kegg_names[,1]) ) , 2 ]
@@
 pages <- list(list("Across"=top), res) %>% flatten()
-write.xlsx(pages, file=paste0("./Suppl.Tbl/","Suppl.Tbl.02",".xlsx"))
+write.xlsx(pages, file=paste0(result_dir,"Suppl.Tbl.02",".xlsx"))
 ```

### 2022-02-24 reports/manuscript/Figure3.Rmd
Tweaking plot sizes and output

---
title: "Figure 3. Tissue Microbiome Abundances"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6.6929133858,
  fig.path="./Figures/",
  dev=c('pdf', 'png'), dpi=300,
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(RColorBrewer)
  # remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)
raw_data <- c("ASV_Luminal", "ASV_Tissue") %>% 
  set_names() %>%
  map(., ~read_csv(paste0("../../results/02_data_preprocessing_output/",.x,"_raw_counts.csv")))
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- intersect(colnames(raw_data$ASV_Tissue), colnames(TRX)[colSums(TRX)!=0])

m <- metadata %>%
  select(ID, Luminal_gr, Tissue_gr) %>%
  arrange(Luminal_gr) %>%
  group_by(Luminal_gr) %>%
  nest() %>%
  mutate(L = map(data, ~arrange(.x, Tissue_gr))) %>%
  unnest(L)

group_annotation <- factor(setNames(m$Luminal_gr,m$ID))
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
group_annotation <- factor(setNames(metadata$Tissue_gr,metadata$ID))

#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")

```

```{r Figure 3, echo=FALSE, warning=FALSE, message=FALSE, fig.pos="t"}
# fig.asp=1, fig.height=7,
### A
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################
figlabels <- letters

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,6,7,4,4,
                3,3,6,5,4,4) , ncol= 6,byrow = T),
                               widths = c(1.2,1,1.83,1.7,.85,1.4),
                               heights = lcm(c(2,4.5,2.5,2.3))
                               # heights = c(.7,1.6,1,1.1,1,.8,.2)
       )

ASV <- list(#c("Luminal_gr","ASV_Luminal_normalized"),
             c("Tissue_gr","ASV_Tissue_normalized")
         )
div_list <- list()
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]

  df <- na.omit(metadata[sample_use, c(i[1], "ID")])
  metadata <- metadata[df[,"ID"],]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))
  group_annotation <- na.omit(group_annotation)

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  temp <- t(t(2^x-1)/colSums(2^x-1))
  temp <- temp[, df[,"ID"]]

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","Simpson","Inv Simpson")

  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="",
         genes = c("Shannon","Simpson","Inv Simpson"),
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  div <- tibble(ID=names(shann),shann, simp, invsimp ) %>% 
    rename_with(., ~paste0(c("ID","Shannon","Simpson","Inv Simpson")))
  div_list[[length(div_list) + 1]] <- div

  if(i[2] == "ASV_Tissue_normalized"){
    end <- table(pal[factor(group_annotation[sample_use])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col=NA, col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1.5,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

  o <- order( rowSums(temp) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  temp <- rbind(temp[c(lvl[1:16]),], Other = colSums(temp[!(row.names(temp) %in% c(lvl[1:16])), ]))

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
           main=gsub("_gr","",i[1]),font.main=1,names.arg = rep("",ncol(temp)),
           col = taxa_pal[factor(rownames(temp),levels = lvl)])

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, 
          bg="white", pch=15,cex=.7,xpd=T, 
          col= c("tomato","orange","#d7d7d7")[factor(metadata$BV_Diagnosis_v3)][order( group_annotation)])
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-7,bg="white",
        col= c("#d7d7d7","tomato")[factor(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  text(par("usr")[2],par("usr")[3]-7,labels = "       HIV",cex=.5,xpd=T) # pos=4,

 l <- c("BV:","Normal","Intermediate","BV")
  legend(x = 10, y = 111, #"bottom", inset = c(0, -0.5),
         bty = "n", horiz=TRUE, cex=0.7, xpd = T, 
         border = NA, #pch=15,
         legend = l, x.intersp=0.5, xjust=0,
         text.width = c(0, map_dbl(l, ~strwidth(.x)/2)) ,
         fill=c(NA, "#d7d7d7","orange","tomato"))
  l2 <- c("HIV:","HIV-","HIV+")
  legend(x = 70, y = 111, # bottom position: par("usr")[3]-5
         bty = "n", horiz=TRUE, cex=0.7, xpd = T,
         border = NA,
         legend = l2, x.intersp=0.5, 
         text.width =c(0, map_dbl(l2, ~strwidth(.x)/2)),
         fill=c(NA, "#d7d7d7", "tomato"))
}

### B
############################
# TOTAL RELATIVE ABUNDANCE #
############################
sample_use <- colnames(TRX)[colSums(TRX)!=0]
raw_list <- raw_data %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_Luminal<0.55, "Other")) %>%
    mutate(name = ifelse(.$name=="Not assigned", "Other", .$name)) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_Luminal) %>%
    dplyr::rename(Luminal="ASV_Luminal",Tissue="ASV_Tissue") %>%
    `row.names<-`(., NULL) %>% 
    column_to_rownames(var = "name") %>%
    as.matrix()

par(mar=c(1.3,5,2,.3)) #b,l,t,r
barplot( r, border = NA ,ylab="abundance", xlab ="", axisnames = F,
         col = taxa_pal[factor(rownames(r),levels = lvl)])
title(main = "Total relative abundance ", line = 0.5, cex.main = .9, font.main= 1)
mtext("Luminal        Tissue ", line = 0, cex=.6, side=1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### C
#############################
# LUMINAL-TISSUE COMPARISON #
#############################
par(mar=c(0,1.1,2,2)) #b,l,t,r
x <- as.character(metadata[,"Luminal_gr"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Tissue_gr"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
df <- data.frame(x1,y1)
p <- plot_sankey(df, color_by = 1, 
             plot_labels = T, plot_weights = F, order_1_by = "NULL",
             order_2_by = "disentangled",use_w1 = T,use_w2 = T,
             gap2v = 0,smoothing = .1,pal= pal
             )
text(c(0,1),c(1,1),labels = c( "Luminal" , "Tissue" ),pos=3, xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###########
# LEGENDS #
###########
par(mar=c(0,0,0,.5)) #b,l,t,r # par(mar=c(1,.3,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, 
       y = 1.04, 
       legend = c("L. crispatus/jensenii", "L. iners", "Gardnerella", "High diverse", "Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,.2,1.5,.4)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,1.5,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

```

### 2022-02-24 reports/manuscript/Figure4-5.Rmd
Tweaking plot sizes and output

@@
 ---
-title: "DGE's and Enrichment"
+title: "Figure 4. & 5. DGE's and Enrichment"
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 output: 
@@
 
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(
+  fig.width = 6.6929133858,
   fig.path="./Figures/",
+  dev=c('pdf', 'png'), dpi=300,
   fig.process = function(filename){
@@
 })
-#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
 ```
@@
 
-#############
-# LODA DATA #
-#############
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
-sample_use <- metadata$ID
-
-gr_name <- c("Luminal", "Tissue")
-out <- paste0("./Suppl.Tbl/")
+#########
+# PATHS #
+#########
+result_dir <- "./Suppl.Tbl/"
+if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }
 
@@
                    "Suppl.Tbl.11 TF_PPI_Tissue.xlsx")
+
+#############
+# LODA DATA #
+#############
+datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)
+sample_use <- metadata$ID
+gr_name <- c("Luminal", "Tissue")
+
+source("../../code/enrichment_function.R")
+GO_database <- fgsea::gmtPathways("../../resources/KEGG_GO_database/c5.bp.v6.2.symbols.gmt.txt")
+KEGG_database <- fgsea::gmtPathways("../../resources/KEGG_GO_database/c2.cp.kegg.v6.2.symbols.gmt.txt")
 
@@
 
-  source("../enrichment_function.R")
   if(db == "GO"){
-  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")}else{
-  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")}
+  gmt_list <- GO_database}else{
+  gmt_list <- KEGG_database}
   res_list <- lapply( unique( gene_module[h$order] ),function(x){
@@
 #########
-# Replace missing values with average in sexwork_months column
-m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
-metadata <- metadata %>%
-  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))
-
 dge_L <- edgR_dge.fun("Luminal_gr", confound=T)
@@
 ###############################
-source("../enrichment_function.R")
-gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
-gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")
-
 #info_logFC <- map(top_dfs[[1]], ~table(abs(.x$logFC) > 1))
@@
 ## GO enrichment 
-pairwise_GO <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_GO)) %>%
+pairwise_GO <- map_depth(t_split, 3, ~enrich_list(.x$Genes, GO_database)) %>%
   map_depth(., 2, ~bind_rows(.x, .id = "regulation")) 
 ## KEGG enrichment
-pairwise_KEGG <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_KEGG)) %>%
+pairwise_KEGG <- map_depth(t_split, 3, ~enrich_list(.x$Genes, KEGG_database)) %>%
   map_depth(., 2, ~bind_rows(.x, .id = "regulation"))
@@
 pages <- list(list(Across=dge_L$top), pairwise=pairwise_dge_L$top) %>% flatten()
-write.xlsx(pages, file=paste0(out,DEG_files[1]))
+write.xlsx(pages, file=paste0(result_dir,DEG_files[1]))
 
 pages <- list(list(Across=dge_T$top), pairwise=pairwise_dge_T$top) %>% flatten()
-write.xlsx(pages, file=paste0(out,DEG_files[2]))
+write.xlsx(pages, file=paste0(result_dir,DEG_files[2]))
 ```
@@
 
-write.xlsx(df_across[[1]]$res, file=paste0(out,"Suppl.Tbl.05 Enrichment_Across_Luminal",".xlsx"))
-write.xlsx(df_across[[2]]$res, file=paste0(out,"Suppl.Tbl.09 Enrichment_Across_Tissue",".xlsx"))
+write.xlsx(df_across[[1]]$res, file=paste0(result_dir,"Suppl.Tbl.05 Enrichment_Across_Luminal",".xlsx"))
+write.xlsx(df_across[[2]]$res, file=paste0(result_dir,"Suppl.Tbl.09 Enrichment_Across_Tissue",".xlsx"))
 
@@
 
-write.xlsx(df_pairwise[[1]]$value, file=paste0(out,"Suppl.Tbl.06 Enrichment_Pairwise_Luminal",".xlsx"))
-write.xlsx(df_pairwise[[2]]$value, file=paste0(out,"Suppl.Tbl.10 Enrichment_Pairwise_Tissue",".xlsx"))
+write.xlsx(df_pairwise[[1]]$value, file=paste0(result_dir,"Suppl.Tbl.06 Enrichment_Pairwise_Luminal",".xlsx"))
+write.xlsx(df_pairwise[[2]]$value, file=paste0(result_dir,"Suppl.Tbl.10 Enrichment_Pairwise_Tissue",".xlsx"))
 ```
@@
 ```{r TF PPI Analysis function, include=FALSE}
-TF_PPI_analysis.fun <- function(out, filename, gr_name, outfile) {
+# filename <- DEG_files[1]
+# gr_name <- gr_name[1]
+# outfile <- enrichR_files[1]
+TF_PPI_analysis.fun <- function(result_dir, filename, gr_name, outfile) {
   # Get DEG's files and filter p-value < 0.01
-  sheets <- excel_sheets(paste0(out, filename)) %>% set_names(.) %>% .[-c(1:2)]
-  detables <- map(sheets, ~read_xlsx(paste0(out, filename), sheet = .x, col_names = TRUE)) %>%
+  sheets <- excel_sheets(paste0(result_dir, filename)) %>% set_names(.) %>% .[-c(1:2)]
+  detables <- map(sheets, ~read_xlsx(paste0(result_dir, filename), sheet = .x, col_names = TRUE)) %>%
     map(.,~mutate(.x, DE = ifelse(.x$PValue > dge_cutoff , 0, ifelse(.x$logFC < 0, -1, 1 )))) %>% 
@@
     } 
-    write.xlsx(save_l, file=paste0(out, outfile), row.names=FALSE)
+    write.xlsx(save_l, file=paste0(result_dir, outfile), row.names=FALSE)
   }
@@
 ```{r TF PPI Analysis, eval=FALSE, include=FALSE}
-TF_PPI_analysis.fun(out, DEG_files[1], gr_name[1], enrichR_files[1]) # writes Suppl.Tbl.07
-TF_PPI_analysis.fun(out, DEG_files[2], gr_name[2], enrichR_files[2]) # writes Suppl.Tbl.11
+TF_PPI_analysis.fun(result_dir, DEG_files[1], gr_name[1], enrichR_files[1]) # writes Suppl.Tbl.07
+TF_PPI_analysis.fun(result_dir, DEG_files[2], gr_name[2], enrichR_files[2]) # writes Suppl.Tbl.11
 ```
@@
 
-    f_DOWN <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "down")
-    f_UP <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "up")
+    f_DOWN <- read_xlsx(paste0(result_dir, file), sheet = x) %>% filter(Regulation == "down")
+    f_UP <- read_xlsx(paste0(result_dir, file), sheet = x) %>% filter(Regulation == "up")
 
@@
 
-sheets <- excel_sheets(paste0(out, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
+sheets <- excel_sheets(paste0(result_dir, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
 arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)
@@
 
-```{r Figure 5, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
+```{r Figure 5, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6, crop = TRUE}
+knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
 # Layout
@@
 
-sheets <- excel_sheets(paste0(out, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
+sheets <- excel_sheets(paste0(result_dir, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
 arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)
@@
 pages <- list(list("DEGs L2T2 vs L2T3"=t), df_across[[1]]$res) %>% flatten()
-write.xlsx(pages, file=paste0(out,"Suppl.Tbl.12 L2T2_L2T3_DEGs_and_enrichment.xlsx")) 
+write.xlsx(pages, file=paste0(result_dir,"Suppl.Tbl.12 L2T2_L2T3_DEGs_and_enrichment.xlsx")) 
 ```
@@
 ## igraph network plot
-```{r igraph network, echo=FALSE, warning=FALSE} 
+```{r igraph network, eval=FALSE, warning=FALSE, include=FALSE}
 # 1. Change the groups of interest by assigning condition_use
@@
 # combinations
-sheets <- map(DEG_files, ~excel_sheets(paste0(out, .x))[-c(1:2)] ) %>%
+sheets <- map(DEG_files, ~excel_sheets(paste0(result_dir, .x))[-c(1:2)] ) %>%
   map(., ~set_names(.x))
@@
 save_pdf.fun <- function(i, gr_name, comp){
-  temp_i_name <- grep(i,list.files(paste0(out),full.names=F),value=T)
-  p <- map(comp, ~read_xlsx(paste0(out, temp_i_name), sheet = .x))
+  temp_i_name <- grep(i,list.files(paste0(result_dir),full.names=F),value=T)
+  p <- map(comp, ~read_xlsx(paste0(result_dir, temp_i_name), sheet = .x))
 
@@
   par(mar = c(0, 0, .7, 0))  #bottom, left, top, right
-  pdf(file = paste0(out, "Plots/", i, "_network_",".pdf", sep=""))
+  pdf(file = paste0(result_dir, "Plots/", i, "_network_",".pdf", sep=""))
     p <- imap(obj_igraph,
@@
 
-# if(isFALSE(dir.exists(paste0(out,"Plots")))){dir.create(paste0("../",result.dir),recursive = TRUE)}
+# if(isFALSE(dir.exists(paste0(result_dir,"")))){dir.create(paste0("../",result.dir),recursive = TRUE)}
 # save_pdf.fun("TF_PPI_Luminal", "Luminal", comp = sheets[[1]])

### 2022-02-24 reports/manuscript/Figure6.Rmd
Tweaking plot sizes and output

@@
 title: "Figure 6. Differential Protein expression"
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 output: 
@@
   library(openxlsx)
+  library(readxl)
   library(rstatix)
@@
 #############
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv")
-cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
-metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
+datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI_Bradley_et.al.xlsx")
+cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
+metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)
 
@@
 Protein <- Protein_V3 %>%
-  left_join(select(metadata, PatID, ID), by="PatID") %>%
   select(ID, contains("HPA")) %>%
-  filter(!is.na(.$ID)) %>%
+  filter(ID %in% metadata$ID) %>%
   pivot_longer(., -ID, names_to = "Antibody") %>%
@@
   column_to_rownames(var = "Antibody")
-
-# Replace missing values with average in sexwork_months column
-m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
-metadata <- metadata %>%
-  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))
 
@@
     start <- c(0, map_dbl(end[1:4], ~.x+0.015))
+    end <- c(end[1:length(end)-1], par("usr")[2]-.005)
 
-    axis(1, at = c(start, end), label = F, pos =1.03 , xpd=T, col="white",col.ticks="black")
-    map2(start, end, ~lines(x=c(.y, .x), y= c(1.03, 1.03), xpd=T, cex=.8))
+    axis(1, at = c(start, end), label = F, pos =1.032 , xpd=T, col="white",col.ticks="black")
+    map2(start, end, ~lines(x=c(.y, .x), y= c(1.032, 1.032), xpd=T, cex=.8))
     # text labels
     t <- map2(start, end, ~(.x+.y)/2)
-    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.04, .y,xpd=T, cex=.8))
+    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))
 
@@
     if(gr == "Luminal_gr") {
-      lines(x=c(0,.603,.603,0,0),
-            y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
+      lines(x=c(0,.603,.603,-.002,-.002),
+            y=c(.488,.488,1.008, 1.008,.488), col="red", lty=1, lwd = 1.5)
       lines(x=c(.39,.94,.94,.39,.39),
@@
     ## Meta info bars
+    if(gr == "Luminal_gr"){size = .5}else{size=.6}
     points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
@@
             col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
-            pch=15,xpd=T,cex=.5) 
+            pch=15,xpd=T,cex=size) 
     points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
             col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
-            pch=15,xpd=T,cex=.5)
+            pch=15,xpd=T,cex=size)
 
@@
 
-imap(p_list, ~write.xlsx(.x,file=paste0("./Suppl.Tbl/","Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
-write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0("./Suppl.Tbl/","Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
+imap(p_list, ~write.xlsx(.x,file=paste0(result_dir,"Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
+write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0(result_dir,"Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
 ```
@@
 
-**Figure 3.
+**Figure 6.
 Protein Heatmap**.
-**a** and **b** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively.
+**a** and **c** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively.
 (p-value < 0.05) **b** and **d** sig.

### 2022-08-24 reports/manuscript/Figure1.Rmd
Change figure format to tiff, jpeg and pdf

@@
 geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
-output: 
-  pdf_document:
-    fig_caption: yes
 header-includes: 
@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/Figure1/", ...)
+    })
 ---
@@
 knitr::opts_chunk$set(
+  results = "hold",
+  message    = FALSE,
+  warning    = FALSE,
   fig.width = 6.6929133858,
   fig.path="./Figures/",
-  dev=c('pdf', 'png'), dpi=300,
   fig.process = function(filename){
@@
     ifelse(fs::file_exists(new_filename), new_filename, filename)
-
-      # Create PNG
-      # path2 <- xfun::with_ext(new_filename, "png")
-      # img <- magick::image_read(new_filename)
-      # magick::image_write(img, path2, format = "png")
 })
 # setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
+#.libPaths("/Users/vilkal/Applications/miniconda3/envs/Paulos_repo/lib/R/library")
 
@@
   # - image resolution of approximately 300 dpi (dots per inch) at the final size
+
 ```
 
-```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data}
 ##################
@@
 
-```{r Figure 1, echo=FALSE, fig.pos="t", fig.height=6.2, message=FALSE, warning=FALSE}
+```{r Figure 1, fig.align = "center", fig.pos="t", fig.height=6.2}
 # , fig.height=6.2
+# fig.height=7, fig.asp=1,
 figlabels <- letters
-# fig.height=7, fig.asp=1,
-print(knitr::is_latex_output())
+
 layout(matrix(c(1,1,1,1,1,1,
@@
 
-```{r Suppl. Table 1., include=FALSE}
+```{r Suppl. Table 1.}
 # Total Relative Abundance 
@@
 write.xlsx(pages, file=paste0(result_dir,"Suppl.Tbl.01",".xlsx"))
+alpa_div <- set_names(div_list, c("Luminal","Tissue"))
+write.xlsx(alpa_div, file=paste0(result_dir,"alpha_diversity",".xlsx"))
+```
 
+```{r eval=FALSE, include=FALSE}
+#############
+# TOP DGEs #
+#############
+p_val <- pull(top, "PValue")
+top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
+top_dge <- top$Genes[ top_dge ]
+
+top_TRX_counts <- edgeR::cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
+top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
+top_TRX_counts[top_TRX_counts > 5] <- 5
+colnames(top_TRX_counts) <- colnames(TRX_counts)
+
+##########################
+# UMAP ON TOP DGE RNAseq #
+##########################
+mypar(2,3,mar=c(2,2,2,6))
+set.seed(1)
+UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
+                 metric = "correlation",min_dist = 0.1,spread = 5,
+                 negative_sample_rate = 10)
+plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="UMAP on top DGE RNAseq",
+     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
+legend(par("usr")[2],par("usr")[4],title.adj = 0,
+       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
+       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
 ```

### 2022-08-24 reports/manuscript/Figure4-5.Rmd
Change figure format to tiff, jpeg and pdf

@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/Figure4-5/", ...)
+    })
 ---
@@
 knitr::opts_chunk$set(
+  results = "hold",
+  message    = FALSE,
+  warning    = FALSE,
   fig.width = 6.6929133858,
   fig.path="./Figures/",
-  dev=c('pdf', 'png'), dpi=300,
+  fig.align = "center",
   fig.process = function(filename){
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r message=FALSE, warning=FALSE}
 ##################
@@
 ## Heatmap DGEs and enrichment
-```{r Across functions, message=FALSE, warning=FALSE, include=FALSE}
+```{r Across functions, message=FALSE, warning=FALSE}
 ##################
@@
 
-```{r Pairwise functions, message=FALSE, warning=FALSE, include=FALSE}
+```{r Pairwise functions, message=FALSE, warning=FALSE}
 ###########################
@@
 
-```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}
+```{r Across tables, fig.height=4, fig.width=7, warning=F}
 TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
@@
 
-```{r Pairwise tables, include=FALSE}
+```{r Pairwise tables}
 ##################
@@
 
-```{r Suppl. Tbl. 4 & 8, include=FALSE}
+```{r Suppl. Tbl. 4 & 8}
 # DEG's across and pairwise
@@
 
-```{r Suppl. Tbl. 5,6 & 9,10, eval=FALSE, include=FALSE}
+```{r Suppl. Tbl. 5,6 & 9,10, eval=FALSE}
 # Enrichment Across
@@
 
-```{r Suppl. Tbl. 12, include=FALSE}
+```{r Suppl. Tbl. 12}
 metadata <- metadata %>% 
@@
 ## igraph network plot
-```{r igraph network, eval=FALSE, warning=FALSE, include=FALSE}
+```{r igraph network, eval=FALSE, warning=FALSE}
 # 1. Change the groups of interest by assigning condition_use

### 2022-08-24 reports/manuscript/Figure6.Rmd
Change figure format to tiff, jpeg and pdf

@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/Figure6/", ...)
+    })
 ---

### 2022-08-25 reports/manuscript/Suppl.Figure1.Rmd
cahange figure output format

---
title: "Suppl. Figure 1. Bacterial abundance across datasets"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, out_dir, ...) {
    source("../../code/knit_function.R");
    custom_knit(inputFile, "../../lab_book/SupplFigure1/", ...)
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path="./Suppl.Figures/",
  fig.align = "center",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r Load data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(tidyverse)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/results/datasets_all_samples.RDS")
datasets_all_samples_ <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")

#################
# COLOR PALETTS #
#################
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
```

```{r Suppl.Fig.1 Bacterial_abundance, fig.asp=.5}
# test to get Paulos code to work with the new files who has less taxa due to filtering after normalization with only selected samples
# it does not currently work
# top_all_ <- lapply( datasets_all_samples_[1:2] , function(x){
#    sort(apply(x,1,var),decreasing = T)[1:20]} )
# top_ <- unique( sub( ".*[.]" , "" , names(unlist(top_all_)) ) )
# 
# x_ <- t(sapply(datasets_all_samples_[1:2],function(x){
#   c(t(x))
# }))
# colnames(x_) <- c(sapply(rownames(datasets_all_samples_[[1]]),
#                         function(i){rep(i,ncol(datasets_all_samples_[[1]]))}))
# rownames(datasets_all_samples_[[1]]) == rownames(datasets_all_samples_[[2]])

top_all <- lapply( datasets_all_samples[2:3] , function(x){
   sort(apply(x,1,var),decreasing = T)[1:20]} )
top <- unique( sub( ".*[.]" , "" , names(unlist(top_all)) ) )

x <- t(sapply(datasets_all_samples[2:3],function(x){
  c(t(x))
}))
colnames(x) <- c(sapply(rownames(datasets_all_samples[[2]]),
                        function(i){rep(i,ncol(datasets_all_samples[[2]]))}))

rownames(x) <- c("Tissue", "Luminal")

top <- sort(sapply( unique(colnames(x)),
               function(i) { mean( x[,colnames(x)==i ] ) } ),decreasing = T)
top <- names(top)[1:30]

###################################
# BACTERIAL ABUNDANCE IN DATASETS #
###################################
par(mar=c(6.5,2.5,0,0)) #b,l,t,r
violist(data = x[,colnames(x)%in%top],
        genes = rownames(x),
        clustering = factor( colnames(x)[colnames(x)%in%top],levels = top) ,
        srt=45, cex.axis=0.7, pt.cex=0.2,  smooth = .1,
        transparency=50,col = pal)
```

### Paulos code with tidyverse
```{r New_files, fig.asp=.5}
Lum <- datasets_all_samples_[[1]][rownames(datasets_all_samples_[[1]])%in%top,]
Tis <- datasets_all_samples_[[2]][rownames(datasets_all_samples_[[1]])%in%top,]

x_Lum <- Lum %>% as_tibble(., rownames ="tax") %>%
  pivot_longer(cols = -tax, names_to = 'ID') #%>% 
  #pivot_wider(names_from = tax, values_from = value)

x_Tis <- Tis %>% as_tibble(., rownames ="tax") %>%
  pivot_longer(cols = -tax, names_to = 'ID') #%>% 
  #pivot_wider(names_from = tax, values_from = value)

xx <- x_Lum %>% left_join(x_Tis, by=c("tax","ID")) %>% select(-ID)
xx_ <- xx %>% filter(!(is.na(.$value.y))) # filter out NA from Tissue

Luminal <- set_names(xx$value.x, xx$tax)
Tissue <- set_names(xx_$value.y, xx_$tax)

t <- rbind(Tissue, Luminal)

###################################
# BACTERIAL ABUNDANCE IN DATASETS #
###################################
par(mar=c(6.5,2.5,0,0)) #b,l,t,r
violist(data = t, smooth = .1,
        genes = c("Tissue", "Luminal"),
        clustering = factor( colnames(t),levels = top),
        srt=45, cex.axis=0.7, pt.cex=0.2,
        transparency = 50,col = pal)
```

I belive that the reson for the diference seen in the tissue distribution is that some of the samples was removed later on.

### 2022-08-25 reports/manuscript/Suppl.Figure2.Rmd
cahange figure output format

@@
 title: "Suppl. Figure 2. Differential bacterial abundance across Tissue and Luminal microbiome"
-output: 
-  pdf_document:
-    fig_caption: yes
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 header-includes: 
@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/SupplFigure2/", ...)
+    })
 ---
 
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(
-  fig.path="./Figures/",
+  fig.path="./Suppl.Figures/",
+  fig.align = "center",
   fig.process = function(filename){
@@
 })
-#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
 ```
 
-```{r echo=FALSE}
+```{r Load data, message=FALSE, warning=FALSE}
 ##################
@@
 #############
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
+datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+metadata <- read.csv("../../data/metadata.csv", row.names = 1, stringsAsFactors = F)
+
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+# datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
+# metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
 group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
@@
 
-```{r Suppl.Fig.2 Diff_bact_abundance, eval=T,warning=FALSE, message=FALSE,echo = FALSE,fig.width=10,fig.height=9}
-all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
-datasets <- factor(c( rep("Tissue",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("Luminal",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
+```{r Suppl.Fig.2 Diff_bact_abundance, warning=FALSE, message=FALSE, fig.width=10, fig.height=9}
+# all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
+#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
+# datasets <- factor(c( rep("Tissue",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("Luminal",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
+# datasets <- datasets[ colSums(all_microbiome)>0 ]
+# all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
+
+all_microbiome <- cbind(datasets_all_samples[["ASV_Tissue_normalized"]],
+                        datasets_all_samples[["ASV_Luminal_normalized"]])
+datasets <- factor(c( rep("Tissue",ncol(datasets_all_samples[["ASV_Tissue_normalized"]])) , rep("Luminal",ncol(datasets_all_samples[["ASV_Luminal_normalized"]])) ))
 datasets <- datasets[ colSums(all_microbiome)>0 ]

### 2022-08-25 reports/manuscript/Suppl.Figure3-4.Rmd
cahange figure output format

--- 2022-02-10 reports/rmarkdown/manuscript/Suppl.Figure3-4.Rmd
@@
 title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
-output: 
-  pdf_document:
-    fig_caption: yes
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 header-includes: 
@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/SupplFigure3_4/", ...)
+    })
 ---
 
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(
-  fig.path="./Figures/",
+  fig.path="./Suppl.Figures/",
+  fig.align = "center",
   fig.process = function(filename){
@@
 })
-#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
 ```
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r message=FALSE, warning=FALSE}
 ##################
@@
 
+#########
+# PATHS #
+#########
+PATH <- "../../.."
+reprod_files <- "../../data/reprod.Suppl.Fig.3_4/"
+
 #############
@@
 #############
-#Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv") 
-datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-PATH <- "../../.."
-SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
+datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+SNN_bacteria <- read.csv("../../results/04_clustering_output/bacteria_SNN_graph.csv",row.names = 1)
+# setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
+#datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
+#SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
 bac_communities <- read.xlsx("./Suppl.Tbl/Suppl.Tbl.01.xlsx", sheet = "Bacterial Communities")
@@
 
-```{r echo=FALSE}
+```{r eval=FALSE, include=FALSE}
 counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
@@
 
-```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, echo = FALSE,fig.width=12,fig.height=13}
+```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, fig.width=12, fig.height=13}
 myletter <- letters
@@
 
-ASV <- list(c("Luminal","ASV_CVL_V3_normalized_batch_corrected"),
-            c("Tissue","ASV_tissue_V3_normalized_batch_corrected")
+# ASV <- list(c("Luminal","ASV_CVL_V3_normalized_batch_corrected"),
+#             c("Tissue","ASV_tissue_V3_normalized_batch_corrected")
+#             )
+ASV <- list(c("Luminal","ASV_Luminal_normalized"),
+            c("Tissue","ASV_Tissue_normalized")
             )
@@
   if(reprod. == TRUE){
-    NESes <- read.csv2(paste0("../../../results/",i[1],"_TRX_Bact_norm_enrichment_scores",".csv"), row.names = 1)
+    NESes <- read.csv2(paste0(reprod_files,i[1],"_TRX_Bact_norm_enrichment_scores",".csv"), row.names = 1)
   }else{

### 2022-08-25 reports/01_taxonomy_improved.Rmd

@@
 title: "Improve Taxonomy Assignment"
-editor_options:
-  chunk_output_type: console
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
   html_document:
-    code_folding: hide
+    code_folding: show
     toc: yes
@@
       collapsed: no
+editor_options:
+  chunk_output_type: console
 ---
@@
 
-```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data, message=FALSE, warning=FALSE}
 ##################
@@
 
-```{r SeqID dict, include=FALSE, results='hide'}
+```{r SeqID dict, results='hide'}
 ########################

### 2022-08-25 reports/03_normalize_data.Rmd

---
title: "Normalization"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
```

# Load libraries and other scripts

```{r Load data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
library(tidyverse)
})  })  })

#########
# PATHS #
#########
input_dir <- "../results/02_data_preprocessing_output/"
result_dir <- "../results/03_normalize_data_output/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#################
# DOWNLOAD DATA #
#################
## TRX DATA:
GEO_url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE194276&format=file&file=GSE194276%5FRaw%5Fgene%5Fcounts%5Fmatrix%2Ecsv%2Egz"
# download.file(GEO_url, paste0("../data/"," Raw_gene_counts_matrix.csv.gz"), method="auto")
# system("gunzip ../data/*.gz")

## MICROBIOME DATA:
# if(params$run=="Boston_run1"){
#   url <- c(
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/001/SRR9198521/SRR9198521.fastq.gz",
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/007/SRR9198517/SRR9198517.fastq.gz",
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/008/SRR9198518/SRR9198518.fastq.gz"
#     ) }

# purrr::walk(url, ~download.file(.x, file.path("../data/Fastq/", basename(.x)), method="auto"))
# system("gunzip ../data/*.gz")
#

#############
# LODA DATA #
#############
trx <- read.csv(paste0("../data/","Raw_gene_counts_matrix.csv"))

dataset_names <- c("ASV_Luminal_raw_counts.csv", # Tissue, Boston run 1 (108 samples)
                   "ASV_Tissue_raw_counts.csv")  # Tissue, Boston run 2+1 (93 sample)

datasets <- map(dataset_names, ~read.csv(paste0(input_dir,.x),row.names = 1)) %>% set_names(., dataset_names)
```

# normalize and save microbiome datasets

```{r save raw and norm microbiome files, fig.height=5, fig.width=5}
# normalized files
datasets <- datasets %>%
  map(., ~round( t(log2( t(.x) / colSums(.x) * 1000 + 1 )),4))

write.csv( datasets[["ASV_Luminal_raw_counts.csv"]] , paste0(result_dir,"ASV_tissue_normalized.csv"),row.names = T )
write.csv( datasets[["ASV_Tissue_raw_counts.csv"]] , paste0(result_dir,"ASV_luminal_normalized.csv"),row.names = T )
```

# normalize and save trx datasets

```{r save raw and norm trx files}
trx <- rowsum(trx[,-c(1:3)],trx$symbol)
trx  <- trx[ rownames(trx) != "" , ]
trx <- trx[ rownames(trx) != "NA" , ]
trx <- trx[ rowSums(trx > 5) >= 3 , ]
trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )

write.csv( trx , paste0(result_dir,"Tissue_RNAseq_normalized.csv"),row.names = T )
```

```{r save RDS}
normalized_data <- list( datasets[["ASV_Luminal_raw_counts.csv"]], 
                         datasets[["ASV_Tissue_raw_counts.csv"]], 
                         trx) %>% 
  set_names(., c("ASV_Luminal_normalized", "ASV_Tissue_normalized", "Tissue_RNAseq_V3_normalized"))

saveRDS(normalized_data, paste0(result_dir, "datasets_all_samples.RDS"))
```

### 2022-08-25 reports/manuscript/Figure2.Rmd

@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/Figure2/", ...)
+    })
 ---
@@
 
-```{r Figure 2,echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
+```{r Figure 2, fig.height=8}
 
@@
 
-```{r Suppl. Table 2., include=FALSE}
+```{r Suppl. Table 2.}
 #saveRDS( y , "../../../results/Picrust_EdgeR_estimations.rds" )

### 2022-08-25 reports/manuscript/Figure3.Rmd

@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/Figure3/", ...)
+    })
 ---

### 2022-08-25 reports/manuscript/Suppl.Figure5.Rmd

@@
 title: "Suppl. Figure 5. Rarefaction curves"
-output: 
-  pdf_document:
-    fig_caption: yes
+geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
 header-includes: 
@@
   chunk_output_type: console
+knit: (function(inputFile, out_dir, ...) {
+    source("../../code/knit_function.R");
+    custom_knit(inputFile, "../../lab_book/SupplFigure5/", ...)
+    })
 ---
 
-```{r setup}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(
-  fig.path="./Figures/",
+  fig.path="./Suppl.Figures/",
+  fig.align = "center",
   fig.process = function(filename){
@@
              '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
-             '../../../data/phyloseq_Sthlm.RDS')       # CVL V2, (only v4 region)
+             '../../../data/phyloseq_Sthlm.RDS'        # CVL V2, (only v4 region)
+             )
 
@@
 
-```{r rarefaction curves, fig.height=6, fig.width=8, fig.asp=0.75,  include=FALSE}
+```{r rarefaction curves, fig.height=6, fig.width=8, fig.asp=0.75}
 ##################
@@
 
-```{r Suppl.Fig5 Rarefraction_plot, warning=FALSE, echo=FALSE, fig.align="center", fig.height=6, fig.width=8, fig.asp=0.75,}
+```{r Suppl.Fig5 Rarefraction_plot, warning=FALSE, fig.height=6, fig.width=8, fig.asp=0.75,}
 ### A & B

### 2022-08-29 reports/00_dada2_script.Rmd
intial commit

---
title: "dada2 Single-End Protocol - `r params$run`"
author: "Seth M. Bloom"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
  word_document: default
params:
  run: "Boston_run1"
---

### Researchers/collaborators:  
* Gabriella Edfeldt <gabriella.edfeldt@ki.se>  
* Jiawu Xu <jxu25@mgjh.harvard.edu>  
* Douglas S. Kwon <dkwon@mgh.harvard.edu>

## References:  
* Callahan, B. J., McMurdie, P. J., Rosen, M. J., Han, A. W., Johnson, A. J. A., & Holmes, S. P. (2016). DADA2: High-resolution sample inference from Illumina amplicon data. Nature Methods, 13(7), 581–3.  
    + https://doi.org/10.1038/nmeth.3869  
* McMurdie, P. J., & Holmes, S. (2013). phyloseq: an R package for reproducible interactive analysis and graphics of microbiome census data. PloS One, 8(4), e61217.  
    + https://doi.org/10.1371/journal.pone.0061217  
* DADA2 tutorial:           http://benjjneb.github.io/dada2/tutorial.html  
* DADA2 big data tutorial:  http://benjjneb.github.io/dada2/bigdata.html  
* Phyloseq:                 https://joey711.github.io/phyloseq/index.html  
* Qiime 1 Mapping files:  http://qiime.org/documentation/file_formats.html#mapping-file-overview

### Description of project and source of data

DESCRIPTION OF PROJECT(S) INCLUDED IN SEQUENCING RUN:

* Project 1:  
    + Researchers:  Gabriella Edfeldt, Jiawu Xu   
    + Number of Samples:  201 + ctrl.
* TECHNICAL CONTROLS    
    + Zymo DNA Standard: 0   
    + Zymo Community DNA Standard: 0    
    + Water_pcrneg: 2

* Sequence files were demultiplexed and split on sample name using Qiime1.  
* This script assigns taxonomy using a variety of different 16S gene reference databases including RDP, GreenGenes, Silva, and HitDB.  
* For archiving purposes, the demultiplexed and split files are stored on O2 in /n/groups/kwon/data1/sequencing_run_archive_2019_02_25_MiSeq/Demultiplexed

### Sequencing run parameters and metrics (Boston run 1 only) 
**RUN PARAMETERS (from RunParameters.xml MiSeq output file)**

* Sequencing Date:          **2019_02_25**  
* Run ID:                   **190225M05907 0033 000000000-C8CH9 **   
* Sequencer:                **Instrument M05907**   
* Run type:                 **Single-end V2 300x1 kit (e.g. Single-end V2 300x1 kit)**   
* Flow cell information:    **Serial number: 000000000-C8CH9 Exp.Date: 2019_10_26 **   
* PR2 Buffer information:   **Serial number: MS7625117-00PR2 Exp.Date: 2019_11_04**   
* Reagent kit information:  **Serial number: MS7731647-300V2 Exp.Date: 2019_10_05**

**RUN METRICS (From Illumina BaseSpace record of the run, https://basespace.illumina.com/home/index )**

* Flow cell status:           **QC passed**   
* Target PhiX conc.:          **10%**   
* Read 1 aligned (%):         **4.87%**   
* Read 2 (I) aligned (%)      **%**   
* Clusters passing filter:    **75.16% +/- %**  
* Read 1 % >=Q30:             **82.52%**   
* Read 2 (I) % >=Q30:         **76.26%**   
* Read 1 yield:               **4.85Gbp**   
* Read 2 (I) yield:           **178.59Mbp**   
* Read 1 error rate:          **1.84% +/-0.08%**   
* Read 2 error rate:          **0.00%**   
* Reads passing filter:       **16,235,386**   
* Cluster density:            **1072 +/- 145k/mm^2 (MiSeq target is 865 to 965)**   
* Tiles:                      **28**   
* Legacy Phas/Prephas Read 1: **0.078  /0.154**   
* Read 1 intensity:           **18 +/-0**   
* Read 2 (I) intensity:       **524 +/- 86**

## Expected output  
1.  A knitted R markdown document (default format is HTML, but can also select options including PDF or MSWord) that contains all comments, code, and figures produced. Key highlights include:  
    + The document returns the names of any samples that failed to pass filter and were eliminated from the final output tables and phyloseq output.  
    + The document calculates and returns cumulative processing time at various steps along the analysis.  
2.  Filtered reads with quality information for each sample stored in newly created subdirectory /Data/filtered.  
    + If sequence files were originally named **samplename**.fastq, the corresponding filtered files will be named **samplename**.filt.fastq  
3.  A newly created subdirectory named /Output containing:  
    + Plot of aggregated read quality scores from the {plot-unfiltered-run-quality} chunk  
    + Plot(s) of read quality scores for individual sample(s) (if code for these plots is included in the {plot-unfiltered-run-quality} chunk)  
    + Plot of error rates from the {learn-errors} chunk  
    + Table of read counts and percentages retained after each sequence analysis step for each sample, saved in a .txt file in TSV format (can be opened in Excel)  
    + Plot of read percentages retained after each sequence analysis step for each sample  
4.  A newly created subdirectory named /ps_objects containing a different phyloseq object for each taxonomic database used. Each phyloseq object is created in the {create-phyloseq} chunk and contains:  
    + An otu_table (ASV table) based on amplicon sequence variants (ASVs) after dada sequence inference and chimera removal  
    + A tax_table based on taxonomic assignments performed in the {assign-taxonomy} chunk
    + sample_data derived from the mapping file.  
    + Phylogenetic trees are not constructed because depending on the type of sample sequenced, they may not be relevant (e.g. in vitro synthetic mixtures of laboratory bacterial strains for QC)  
5. A newly created subdirectory named /RDS containing:  
    + An RDS file containing the sequence table after chimera removal.  
    + An RData workspaced imaging saving the final workspace for the analysis.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = getwd())
```

```{r start-tracking-processing-time, results = "hold"}
#Define variable with start time of running script.
start_time <- proc.time()
```

```{r initiate-environment, message=FALSE, collapse=TRUE}
# Load libraries
library(ShortRead)
packageVersion("ShortRead")
library(dada2)
packageVersion("dada2")
library("phangorn")
packageVersion("phangorn")
library("phyloseq")
packageVersion("phyloseq")
library("ips")
packageVersion("ips")
library("tidyverse")
packageVersion("tidyverse")

# Set random seed for reproducibility purposes
set.seed(100)

```

```{r define-sequencing-run-date-and-version, results = "hold"}
if(params$run=="Boston_run1"){
  run.name <- "Boston_run1"
  run.date <-  "2019_02_25_v1" 
  mapping_file <- "Mapping_info_Boston_run1.csv"
  # Trim read lenght left and right
  FwdTrimLeft <- 10   
  FwdTrimRight <- 225  
}

if(params$run=="Boston_run2"){
  run.name <- "Boston_run2"
  run.date <-  "2019_05_25" 
  mapping_file <- "Mapping_info_Boston_run2.csv"
  # Trim read lenght left and right
  FwdTrimLeft <- 0   
  FwdTrimRight <- 252  
}

# Sequencing run date in format YYYY_MM_DD
print("Sequencing run date and version:")
run.date
```

```{r Donload data from ENA}
# if(params$run=="Boston_run1"){
#   url <- c(
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/001/SRR9198521/SRR9198521.fastq.gz",
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/007/SRR9198517/SRR9198517.fastq.gz",
#     "ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR919/008/SRR9198518/SRR9198518.fastq.gz"
#     ) }
# dir <- "../data/Fastq"
# dir.create(dir)
# purrr::walk(url, ~download.file(.x, file.path("../data/Fastq/", basename(.x)), method="auto"))
# system("gunzip ../data/Fastq/*.gz")
#

if(params$run == "Boston_run1"){dir  <- "/Users/vilkal/Raw_data/Microbiome/Run1_Boston_CVLv3_March2019"}
if(params$run == "Boston_run2"){dir <- "/Users/vilkal/Raw_data/Microbiome/Run2_Boston_Tissuev3_June2019"}

files <- list.files(path = dir, pattern =".fastq$", full.names = T, recursive = T)

```

```{r eval=FALSE, include=FALSE}
ENA_info <- read_csv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/data/samples-2022-01-30T12 42 43.csv")

mapping_br1 <- readxl::read_xlsx("/Users/vilkal/work/Brolidens_work/Projects/Run1_Boston_CVLv3_March2019/Mapping/Mapping_info_Boston_run1.xlsx")

ENA <- ENA_info %>%
  select(id, SampleID="alias", secondaryId) %>%
  left_join(., mapping_br1, by="SampleID") %>%
  na.omit()
intersect(mapping_br1$SampleID, ENA$SampleID)

# Update again:
# write_csv(ENA, "/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/data/Mapping/Mapping_info_Boston_run1.csv")

# ENA_run1 <- read_tsv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/results/ENA_files_submission/2022-01-30/ENA_Run_template_Run1.tsv")
# ENA_run2 <- read_tsv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/results/ENA_files_submission/2022-01-30/ENA_Run_template_Run2.tsv")
# ENA_T <-read_tsv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/results/ENA_files_submission/2022-01-30/ENA_Sample_template_T.tsv")
# ENA_L <-read_tsv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/results/ENA_files_submission/2022-01-30/ENA_Sample_template_T.tsv")
# 
# ENA_samples_report <- read_csv("/Users/vilkal/work/Brolidens_work/Projects/Microbiom_PSC/data/samples-2022-01-30T12 42 43.csv")
# 
# ENA_LT <- bind_rows(ENA_L, ENA_T)
# ENA <- bind_rows(ENA_run1, ENA_run2) %>%
#   rename(id="sample") %>%
#   left_join(., select(ENA_samples_report, id, secondaryId, alias), by="id") %>%
#   select(id, study, forward_file_name)
# 
# mapping_1 <- read_csv(file ="../data/Mapping/Mapping_info_Boston_run1.csv", col_names = TRUE) %>%
#   left_join(., ENA, by="id") %>%
#   select(id, study, everything(),-PatID_correct)
# 
# mapping_2 <- read_csv(file ="../data/Mapping/Mapping_info_Boston_run2.csv", col_names = TRUE) %>%
#   left_join(., ENA, by="id") %>%
#   select(id, study, everything(),-PatID_correct)

# cahnge again
# mapping_1 <- read_csv(file ="../data/Mapping/Mapping_info_Boston_run1.csv", col_names = TRUE) %>%
#   mutate(SampleID = str_replace(.$forward_file_name, ".fastq","")) %>%
#   select(-forward_file_name)
# 
# mapping_2 <- read_csv(file ="../data/Mapping/Mapping_info_Boston_run2.csv", col_names = TRUE) %>%
#   mutate(SampleID = str_replace(.$forward_file_name, ".fastq","")) %>%
#   select(-forward_file_name)
# 
#  write_csv(mapping_1, "../data/Mapping/Mapping_info_Boston_run1.csv")
#  write_csv(mapping_2, "../data/Mapping/Mapping_info_Boston_run2.csv")
```

### Filter and trim reads

```{r establish-directory-structure, results = "hold", warning=FALSE}
#Define and print path to working directory
pathwd <- getwd()
print("Working directory:")
pathwd

# Define path to subdirectory containing demultiplexed forward-read fastq files
pathF <- file.path("../data/Fastq")
print("Directory containing raw demultiplexed .fastq sequence files:")
pathF

#Define path to directory containing 16S databases
print("Path to 16S databases:")
pathdb <- file.path("../Resources/RDP_database")
pathdb

#Define path to directory containing mapping file
print("Path to mapping file:")
pathMap <- file.path("../data/Mapping")
pathMap

#Create and define path to subdirectory in which filtered files will be stored
filtpathF <- file.path("../data/Fastq/filtered") 
print("Directory that will contain filtered .fastq files:")
filtpathF
dir.create(filtpathF)

#Create and define path to subdirectory in which to store output figures and tables
print("Directory in which to store output figures and tables:")
pathOut <- file.path("../results/00_dada2_output")
pathOut
dir.create(pathOut)

#Create and define path to subdirectory in which RDS files will be stored
print("Path to directory for saving RDS files:")
pathF.RDS <- file.path(pathOut, "RDS")
pathF.RDS
dir.create(pathF.RDS)

#Create and define path to subdirectory in which RDS files containing phyloseq objects will be stored
print("Path to sub-directory in which to save phyloseq objects:")
pathps0 <- file.path(pathOut, "ps_objects")
pathps0
dir.create(pathps0)
```

```{r Define-sequence-file-names-and-sample-names, results = "hold"}
# File parsing
#Create vector of fastq files in pathF.directory
#If analyzing paired end reads, the search pattern below would need to be modified (see DADA2 tutorial)

print("Sequence files to be analyzed:")
fnFs <- sort(files)
head(fnFs)

#Extract sample names, assuming filenames have format: SAMPLENAME.fastq or SAMPLENAME.fastq.gz
sample.names <- sapply(strsplit(basename(fnFs), ".fast"), `[`, 1)
head(sample.names)
```

```{r plot-unfiltered-run-quality, results = "hold"}
#Plot forward read quality aggregated across all forward read files
p.qual.f.1.agg <- plotQualityProfile(fnFs, aggregate = TRUE) + 
  ggtitle("Fwd read quality aggregated across samples")
p.qual.f.1.agg

#Save figure to file
ggsave(filename = file.path(pathOut, paste0("Read_quality_aggregate_fwd_", run.date, ".pdf")), 
       plot = p.qual.f.1.agg, device = "pdf", width = 8, height = 6, units = "in")

```

```{r calculate-quality-plot-processing-time, results = "hold"}
#Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

## Define positions at which to trim filtered data

```{r define-trim-positions, results = "hold"}
## ## ## ## ## ## ## ## ## ##
## USER INPUT REQUIRED HERE ##
## ## ## ## ## ## ## ## ## ##

# Define base positions at which to trim the filtered forward read sequences
# Decisions about where to trim should be based on the quality profile above.
# See DADA2 tutorial for additional information.
# The two variable defined here will then be supplied to the filterAndTrim function
  # FwdTrimLeft gives the posion from the left margin at which to trim
  # FwdTrimRight gives the position from the right margin at which to trim
# For example if FwdTrimLeft is 10 and FwdTrimRight is 240, sequence lengths will be 230 bp.

print(paste0("Trimming forward reads on left at base position ", FwdTrimLeft, 
       " and on right at base position ", FwdTrimRight))
```

## Quality filter and trim
```{r quality-filter-and-trim, results = "hold"}
# Create vector of modified names of fastq files after filtering in
# which the modified file name changes from "samplename.fastq" 
# to "samplename.filt.fastq"

filtFs <- file.path(filtpathF, str_replace(basename(fnFs), "(.fastq)", ".filt\\1"))
head(filtFs)

# Filter and trim sequences listed in fnFs and store them in filtFs, 
# while storing summary in dataframe out

out <- filterAndTrim(fwd = fnFs, filt = filtFs, rev = NULL, filt.rev = NULL,
              trimLeft = FwdTrimLeft, truncLen = c(FwdTrimRight), maxEE = 2, truncQ = 11, maxN = 0, 
              rm.phix = TRUE, compress = TRUE, verbose = TRUE, multithread = TRUE)
head(out)
#truncate file extension (".fastq" or ".fastq.gz") from rownames
rownames(out) <- rownames(out) %>% str_replace(".fast.+", "")
head(out)
```

```{r calculate-filter-trim-processing-time, results = "hold"}
# Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

## Sequence inference

```{r sample-inference-1, results = "hold"}
## Set paramaters
# Create character of all files in directory filtpathF.1 (i.e. all filtered files)
# with names including string "fastq" (i.e. all *.fastq and *.fastq.gz files)
filtFs.pass <- list.files(filtpathF, pattern="fastq", full.names = TRUE)
head(filtFs.pass)

# Create character vector of sample names from the file names in filtFs.1
# This code assumes all filename = "samplename.filt.fastq"" or "samplename.filt.fastq.gz""
sample.names.filt.pass <- sapply(strsplit(basename(filtFs.pass), ".filt.fast"), `[`, 1) 
head(sample.names.filt.pass)

# Name the elements in filtFs according to the elements in sample.names.filt
names(filtFs.pass) <- sample.names.filt.pass

# Samples for which no sequences passed the filtering step
print("Samples for which no sequences passed filter criteria (i.e. samples excluded from filtered dataset):")
sample.names.filt.fail <- setdiff(sample.names, sample.names.filt.pass)
sample.names.filt.fail

```

```{r learn-errors, results = "hold"}
# Set random seed for reproducibility purposes
set.seed(100)

# Learn forward read error rates
errF <- learnErrors(filtFs.pass, nread=1e6, multithread=TRUE, randomize = "TRUE")

# Plot convergence of error rate computation
plot(dada2:::checkConvergence(errF), type = "o", col = "firebrick3", main = "Convergence")

# Plot calculated errors
p.err.F <- plotErrors(errF, nominalQ = TRUE)
p.err.F

# Save figure to file
ggsave(filename = file.path(pathOut, paste0("Error_rates_fwd_", run.date, ".pdf")),
       plot = p.err.F, device = "pdf", width = 8, height = 6, units = "in")
```

```{r calculate-learn-error-processing-time, results = "hold"}
#Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

```{r dereplication-and-sample-inference, results = "hold"}
# Dereplicate and apply error rate to resolve sequence variants

## ## ## ## ## ## ## ## ## ## ## ##
## This chunk gives two different options for analyzing the data, depending on the size of the dataset
## and availability of memory for analysis. Use only Option 1 or Option 2 for analysis, not both.
## Comment (or uncomment) the relevant option by highlighting and hitting <command> + <shift> + C
## ## ## ## ## ## ## ## ## ## ## ##

## ## ## ## ## ## ## ## ## ## ## ## ## ## ##
## OPTION 1: **************************** ##
## ## ## ## ## ## ## ## ## ## ## ## ## ## ##

### Code for smaller datasets (e.g. most MiSeq) for which memory is adequate for parallelization ###
### This code analyzes the samples in parallel, which is more efficient but more memory-intensive ###
### Derived from https://benjjneb.github.io/dada2/tutorial.html ###

# Dereplicate the filtered forward sequences by sample
derepFs <- derepFastq(filtFs.pass, verbose = FALSE)

# Set random seed for reproducibility purposes
set.seed(100)

# Run dada on the dereplicated forward sequences to resolve sequence variants
# based on the forward read error model calculated in errF
dadaFs <- dada(derepFs, err = errF, multithread = TRUE)

# ## ## ## ## ## ## ## ## ## ## ## ## ## ## ##
# ## OPTION 2: *************************** ##
# ## ## ## ## ## ## ## ## ## ## ## ## ## ## ##

# ### Code for very large (e.g. HiSeq) datasets if parallelization would exceed available memory ###
# ### This code analyzes the samples in series, which is less efficient but less memory-intensive ###
# ### Derived from https://benjjneb.github.io/dada2/bigdata.html ###
# 
# #Create lists derepFs and dadaFs of same length and element names as sample.names.filt.pass
# derepFs <- vector("list", length(sample.names.filt.pass))
# names(derepFs) <- sample.names.filt.pass
# dadaFs <- vector("list", length(sample.names.filt.pass))
# names(dadaFs) <- sample.names.filt.pass
# 
# #For each sequence file in filtFs.1, dereplicate the sequences from the file
# #then run dada to resolve sequence variants based on the error model errF.1
# #store results in the corresponding named element of singles.1
# for(sam in sample.names.filt.pass) {
#   cat("Processing:", sam, "\n")
#   derepFs[[sam]] <- derepFastq(filtFs.pass[[sam]])
#   # Set random seed for reproducibility purposes
#   set.seed(100)
#   dadaFs[[sam]] <- dada(derepFs[[sam]], err=errF, multithread=TRUE)
# }
# rm(sam)
```

```{r calculate-sequence-inference-processing-time}
# Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

## Construct sequence table and remove chimeras

```{r seqtab-chimera-removal-1, results = "hold"}
# Make sequence table and review dimensions
seqtab.prechimeraremoval <- makeSequenceTable(dadaFs)
print("Sequence table dimensions (samples, resolved sequences) before chimera removal:")
dim(seqtab.prechimeraremoval)

# Inspect distribution of sequence lengths
print("Distribution of sequence lengths before chimera removal:")
table(nchar(getSequences(seqtab.prechimeraremoval)))

# Remove chimeras
seqtab.nochim <- removeBimeraDenovo(seqtab.prechimeraremoval,
                                    method = "consensus",
                                    multithread = TRUE,
                                    verbose = TRUE)

print("Sequence table dimensions (samples, resolved sequences) after chimera removal:")
dim(seqtab.nochim)

# Inspect distribution of sequence lengths
print("Distribution of sequence lengths after chimera removal:")
table(nchar(getSequences(seqtab.nochim)))
print("Proportion of reads eliminated by chimera removal:")
sum(seqtab.nochim) / sum(seqtab.prechimeraremoval)
```

```{r save-RDS-and-image, results = "hold"}
# Save seqtab.nochim RDS file
print("Save seqtab.nochim as RDS file named:")
f.seqtab.nochim.RDS <- file.path(pathF.RDS, paste0("MiSeq_", run.date, "_preprocessing_single_nochim.RDS"))
f.seqtab.nochim.RDS
saveRDS(seqtab.nochim, f.seqtab.nochim.RDS)

```

```{r calculate-chimera-removal-processing-time, results = "hold"}
# Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

## Track reads through pipeline
The code in the chunk below is substantially altered from the code suggested in the DADA2 tutorial because  
the tutorial code can lead to errors if any samples are eliminated at a step (e.g. if no reads pass filter).

```{r track-reads}
# Function to calculate reads for each element in dada output
getN <- function(x) sum(getUniques(x))

# Read count remaining at each step of processing. Would need additional columns if analyzing
# paired-end data.

# Create interim dataframes listing the sequence count after each step of processing per sample
print("Read count remaining at each step of processing:")
outdf <- out %>% 
  as.data.frame() %>%
  tibble::rownames_to_column() %>% 
  dplyr::rename(Sample = "rowname", input = "reads.in", filtered = "reads.out")

dadacountdf <- sapply(dadaFs, getN) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>%
  dplyr::rename(Sample = "rowname", denoisedF = ".")

nonchimdf <- rowSums(seqtab.nochim) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>%
  dplyr::rename(Sample = "rowname", nonchim = ".")

# Join dataframes by Sample column to track counts for each sample at each step of processing
track <- dplyr::left_join(outdf, dadacountdf, by = c("Sample" = "Sample")) %>%
  dplyr::left_join(nonchimdf, by = c("Sample" = "Sample"))
head(track)

# Percentage of original unfiltered reads remaining at each step of processing
print("Percentage of original unfiltered foward reads remaining at each step of processing:")
track.percent <- dplyr::mutate_at(track, vars(-matches("Sample")), funs(. * 100 / track[["input"]]))
track.percent <- rename_at(track.percent, vars(-matches("Sample")), funs(paste0(., "_percent")))
head(track.percent)

# Create a "long" form version of track.percent for use with plotting
track.percent.long <- track.percent %>% 
  tidyr::gather(key = "Step", value = "Percent", -Sample)

# Modify track.percent.long so that Step column is a factor with levels in the correct order
track.percent.long[["Step"]] <- track.percent.long[["Step"]] %>% forcats::fct_relevel(colnames(track.percent[-1]))

# Plot percentages of input forward reads remaining after each step on a per sample basis.
p.track <- ggplot(track.percent.long, aes(x = Step, y = Percent)) + 
  geom_line(aes(group = Sample, colour = Sample)) +
  geom_point() +
  labs(title = "Percent of input forward reads remaining\nafter each processing step per sample",
       subtitle = paste0("Fwd read left trim position: ", FwdTrimLeft, "\nFwd read right trim position: ", FwdTrimRight),
       x = "Processing Step", y = "% of Input Reads Remaining") +
  theme(plot.title = element_text(hjust = 0.5, size = 18), 
        plot.subtitle = element_text(hjust = 0.5, size = 16),
        legend.position = "none",
        axis.text = element_text(size = 12), axis.title = element_text(size = 16))
p.track

# Save figure to file
ggsave(filename = file.path(pathOut, paste0("read_track_fwd_L", FwdTrimLeft, "_R", FwdTrimRight, "_", run.date, ".pdf")), 
       plot = p.track, device = "pdf", width = 8, height = 6, units = "in")

# Merge track and track.percent dataframes
track.merged <- dplyr::left_join(track, track.percent, by = c("Sample" = "Sample"))
head(track.merged)

# Save log file of tracking reads to file.
print("Save track.merged dataframe tracking read counts as TSV file named:")
f.track.merged <- file.path(pathOut, paste0("MiSeq_", run.date, "_fwd_read_tracking_log.txt"))
head(f.track.merged)
readr::write_tsv(track.merged, f.track.merged)

```

## Assign taxonomy

```{r assign-taxonomy, results = "hold"}
# RDP
# Record start time for this taxonomy assignment step
tax_start_time = proc.time()
print(paste0("RDP training database: ", "rdp_train_set_16.fa.gz"))
taxa.rdp <- assignTaxonomy(seqtab.nochim, file.path(pathdb, "rdp_train_set_16.fa.gz"), multithread = TRUE)

# Add RDP species assignment
print(paste0("RDP species assignment database: ", "rdp_species_assignment_16.fa.gz"))
taxa.rdp.plus <- addSpecies(taxa.rdp, file.path(pathdb, "rdp_species_assignment_16.fa.gz"))
writeLines("RDP taxonomy assignment step processing time (seconds):")
proc.time() - tax_start_time

```

```{r calculate-taxonomy-assignment-processing-time, results = "hold"}
# Calculate cumulative processing time
print("Cumulative processing time (seconds):")
proc.time() - start_time
```

## Generate PhyloSeq Objects and final save

```{r create-phyloseq, echo=FALSE, results = "hold"}
# Create PhyloSeq objects

# Load mapping file for sequencing run
print("Sequencing run mapping file:")
f.mapping <- file.path(pathMap, list.files(path = pathMap, pattern = mapping_file))
f.mapping
mapping <- readr::read_csv(file = f.mapping, col_names = TRUE)

names <- mapping$`SampleID`
# Remove any rows with "#SampleID" column value NA and convert to dataframe
mapping <- mapping   %>%
  dplyr::filter(!is.na(`SampleID`)) %>%
  as.data.frame(mapping)

# Convert mapping$#SampleID column to rownames
mapping <- tibble::column_to_rownames(mapping, var = "SampleID")

# RDP
ps0.rdp <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows = FALSE), 
                    tax_table(taxa.rdp.plus),
                    sample_data(mapping))
ps0.rdp

# Rename samples
sample_names(ps0.rdp) <- sample_data(ps0.rdp)[["Sample_ID"]]

# Sanity checks
get_taxa_unique(ps0.rdp, "Phylum")

# Save RDS files for downstream analysis
saveRDS(ps0.rdp, file = file.path(pathps0, paste0("phyloseq_",run.name, ".RDS")))

# Save current R image (workspace) as .RData file
print("Save R image (workspace) as .RData file named:")
f.image <- file.path(pathF.RDS, paste0("MiSeq_", run.date, "_preprocessing_image.RData"))
f.image
save.image(f.image)
```

```{r calculate-final-processing-time, results = "hold"}
# Calculate total processing time
print("Total processing time (seconds):")
proc.time() - start_time
```

```{r open R image, eval=FALSE, include=FALSE}
load("../results/00_dada2_output/RDS/MiSeq_2019_05_25_preprocessing_image.RData")
load("../results/00_dada2_output/RDS/MiSeq_2019_05_25_preprocessing_image.RData")
```

### 2022-09-30 reports/manuscript/Figure4-5.Rmd
Add library info

@@
   library(scales)
+  library(fgsea)
   library(RColorBrewer)
@@
 return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS))
+#return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS, results=results))
 }
@@
 pages <- list(list(Across=dge_T$top), pairwise=pairwise_dge_T$top) %>% flatten()
+write.xlsx(pages, file=paste0(result_dir,DEG_files[2]))
+
+#Files with filtered p-value < 0.01 
+# Have to update the summary table, however the decideTests
+# function does not provide support for p-value cutoffs only FDR
+# therefore I need to write code to get the correct summary
+
+p_L <- map(pairwise_dge_L$top[-1], ~filter(.x, `PValue`<0.01))
+p_T <- map(pairwise_dge_T$top[-1], ~filter(.x, `PValue`<0.01))
+
+pages <- list(list(Across=dge_L$top), pairwise=p_L) %>% flatten()
+write.xlsx(pages, file=paste0(result_dir,DEG_files[1]))
+
+pages <- list(list(Across=dge_T$top), pairwise=p_T) %>% flatten()
 write.xlsx(pages, file=paste0(result_dir,DEG_files[2]))

### 2022-09-30 reports/manuscript/Figure3.Rmd
add library info

@@
   library(scales)
+  library(vegan)
   library(RColorBrewer)

### 2022-09-30 reports/manuscript/Figure6.Rmd
Add scale legend

@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r Load data, message=FALSE, warning=FALSE}
 ##################
@@
 # top_PROT_mfi <- t
+# sacale = T
 # gr <- "Luminal_gr"
@@
       add_letter("c")
-    }else{height <- round(max(h$height))
+    }else{
+      add_scale_legend(x = par("usr")[1]-0.1, height = 1.5,#diff(par("usr")[3:4])-20,
+                     pal = colorRampPalette(c("navy","grey95","firebrick"))(91),
+                     labels = c("-5","0","5"))
+      height <- round(max(h$height))
       add_letter("a")}

### 2022-09-30 reports/04_clustering.Rmd

@@
 knitr::opts_chunk$set(fig.width = 6.6929133858,
+                      results    = "hold",
                       message    = FALSE,
@@
   library(rafalib)
+  library(RANN)
+  library(uwot)
   #remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)

### 2022-10-04 reports/manuscript/Suppl.Figure4-5.Rmd

--- 2022-08-25 reports/manuscript/Suppl.Figure3-4.Rmd
@@
   library(Rcpp)
+  library(igraph)
   #remotes::install_github("czarnewski/niceRplots",force=T)
   library(niceRplots)
+  library(parallel)
 })  })  })

### 2022-10-11 reports/manuscript/Suppl.Figure4-5.Rmd
change name

@@
     source("../../code/knit_function.R");
-    custom_knit(inputFile, "../../lab_book/SupplFigure3_4/", ...)
+    custom_knit(inputFile, "../../lab_book/SupplFigure4_5/", ...)
     })
@@
 PATH <- "../../.."
-reprod_files <- "../../data/reprod.Suppl.Fig.3_4/"
+reprod_files <- "../../data/reprod.Suppl.Fig.4_5/"
 
@@
 SNN_bacteria <- read.csv("../../results/04_clustering_output/bacteria_SNN_graph.csv",row.names = 1)
-# setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
-#datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
-#SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
 bac_communities <- read.xlsx("./Suppl.Tbl/Suppl.Tbl.01.xlsx", sheet = "Bacterial Communities")
@@
 
-# ASV <- list(c("Luminal","ASV_CVL_V3_normalized_batch_corrected"),
-#             c("Tissue","ASV_tissue_V3_normalized_batch_corrected")
-#             )
 ASV <- list(c("Luminal","ASV_Luminal_normalized"),

### 2022-10-11 reports/manuscript/Suppl.Figure6.Rmd
change name

--- 2022-08-25 reports/manuscript/Suppl.Figure5.Rmd
@@
     source("../../code/knit_function.R");
-    custom_knit(inputFile, "../../lab_book/SupplFigure5/", ...)
+    custom_knit(inputFile, "../../lab_book/SupplFigure6/", ...)
     })
@@
 #############
-paths_p <- c('../../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVL V2)
-             '../../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
-             '../../../data/phyloseq_Sthlm.RDS'        # CVL V2, (only v4 region)
+paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVL V2)
+             '../../data/phyloseq_boston_r2.RDS',   # Tissue V3 & CVL V2
+             '../../data/phyloseq_Sthlm.RDS'        # CVL V2, (only v4 region)
              )

### 2022-10-11 reports/manuscript/Figure6.Rmd
add download link

@@
 
+prot_url <- "https://doi.org/10.1371/journal.ppat.1010494.s017"
+
+############
+# DOWNLOAD #
+############
+if(isFALSE(file.exists("../../data/Protein_norm_MFI.xlsx"))){download.file(prot_url, paste0("../../data/", "Protein_norm_MFI.xlsx"), method="auto")}
+
 #############
@@
 datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
-Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI_Bradley_et.al.xlsx")
+Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI.xlsx", skip = 2) %>% rename(ID="Patient/Sample ID")
 cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")

### 2023-02-24 reports/02_data_preprocessing.Rmd

@@
 datasets <- map(dataset_names, ~read.csv(paste0(input_dir,.x))) %>% set_names(., dataset_names)
-metadata <- read.csv(paste0("../data/", "metadata.csv"),row.names = 1)
 trx <- read.csv(paste0("../data/","Raw_gene_counts_matrix.csv"),row.names = 1)

### 2023-02-24 reports/manuscript/Figure1.Rmd

@@
   library(tidyverse)
+  library(magick)
+  library(uwot)
   library(scales)
@@
 write.xlsx(pages, file=paste0(result_dir,"Suppl.Tbl.01",".xlsx"))
-alpa_div <- set_names(div_list, c("Luminal","Tissue"))
+alpa_div <- set_names(div_list, c("Luminal","Tissue")) %>%
+  map2(., c("Luminal_gr", "Tissue_gr"), ~left_join(.x, select(metadata, ID, .y), by="ID")) %>%
+  map2(., c("Luminal_gr", "Tissue_gr"), ~select(.x, ID, .y, everything()))
 write.xlsx(alpa_div, file=paste0(result_dir,"alpha_diversity",".xlsx"))

### 2023-02-24 reports/manuscript/Figure2.Rmd

@@
   library(igraph)
+  library(fgsea)
   library(RColorBrewer)

### 2023-02-24 reports/manuscript/Suppl.Figure2.Rmd

@@
 ```{r Suppl.Fig.2 Diff_bact_abundance, warning=FALSE, message=FALSE, fig.width=10, fig.height=9}
-# all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
-#                         datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
-# datasets <- factor(c( rep("Tissue",ncol(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]])) , rep("Luminal",ncol(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])) ))
-# datasets <- datasets[ colSums(all_microbiome)>0 ]
-# all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]
-
 all_microbiome <- cbind(datasets_all_samples[["ASV_Tissue_normalized"]],

### 2023-03-08 code/knit_function.R
update pdf settings

@@ -1,4 +1,3 @@
-
 #################
 # KNIT FUNCTION #
 #################
@@ -12,6 +11,7 @@ custom_knit <- function(inputFile, out_dir, ...) {
         dev = c("jpeg", "tiff", "pdf"),
         dpi = 300,
         echo=TRUE,
+        #keep_md = TRUE,
         toc = FALSE,                             
         toc_float = FALSE,
         code_folding = "show",
@@ -19,9 +19,7 @@ custom_knit <- function(inputFile, out_dir, ...) {
       )#;
     # rmarkdown::render(
     #   inputFile,
-    #   output_format = rmarkdown::pdf_document(
-    #     dev = c("pdf", "png"),
-    #     dpi = 300)
+    #   output_format = rmarkdown::pdf_document()
     #   )
 }
 

### 2023-03-08 reports/03_normalize_data.Rmd
change GEO url

@@
 ## TRX DATA:
-GEO_url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE194276&format=file&file=GSE194276%5FRaw%5Fgene%5Fcounts%5Fmatrix%2Ecsv%2Egz"
+GEO_url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE217237"
 # download.file(GEO_url, paste0("../data/"," Raw_gene_counts_matrix.csv.gz"), method="auto")

### 2023-03-08 reports/manuscript/Figure1.Rmd
Add filtering
Add figure legend

@@
   fig.width  = 6.6929133858,
-  fig.path="./Figures/",
+  fig.align  = "center",
+  fig.path   = "./Figures/",
   fig.process = function(filename){
@@
 
-```{r Load data}
+```{r Load-data}
 ##################
@@
 taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
-# show_col(pal)
+# scales::show_col(taxa_pal)
 ```
@@
 
+  dim(x)
+  x <- x[ rowSums(x>0) >= 3 ,  ]
+  #x_ <- x[ rowSums(x>0) ,  ]
+  dim(x)
   temp <- t(t(2^x-1)/colSums(2^x-1))
@@
 
+**Figure 1.
+Characterization of a highly diversemicrobiome in cervicovaginal (luminal) samples from Kenyan sex workers.** Cervicovaginal lavage (luminal) and ectocervical tissue study samples were assessed by 16S rRNA sequencing, gene expression, and protein profiling.
+**a** Schematic drawing depicting the sampling scheme and the resulting omics datasets.
+**b** Bar plots of alpha diversity indices and taxonomy profiles for each individual luminal sample.
+Color-coded squares above the stacked bar plots indicate bacterial vaginosis (BV, binned Nugent’s scores): Gray: negative, orange: intermediate, red: positive; and HIV diagnosis: Gray: HIV seronegative, red: HIV seropositive.
+Two12-SNN graphs were constructed using:
+**c** Louvain community detection algorithm, and d Uniform Manifold Approximation (UMAP).
+The two graphs were overlayed in color with the predefined luminal study groups.
+The undirected edges are included in gray connecting the nodes
+
 ```{r Suppl. Table 1.}

### 2023-03-08 reports/manuscript/Figure1.Rmd
fix space in title

@@
 ---
-title: "Figure 1.Luminal Study Groups"
+title: "Figure 1. Luminal Study Groups"
 geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"

### 2023-03-08 reports/manuscript/Figure2.Rmd
add figure legend

--- 2022-08-25 reports/manuscript/Figure2.Rmd
@@
 
-```{r Load data, message=FALSE, warning=FALSE}
+```{r Load-data, message=FALSE, warning=FALSE}
 ##################
@@
 
+**Figure 2.
+Identification of bacterial communities and functional profiles in the luminal samples.** The luminal study samples were assessed for bacterial communities and functional profiles.
+**a** Bacterial community embedding of 5-SNN graph clustered using Louvain community detection algorithm based on luminal bacterial abundances.
+**b** Wet smear counts of the genera Lactobacillus, Mobiluncus, and Gardnerella with corresponding 16S read counts.
+**c** Differential expression analysis was applied to PICRUSt2 predicted KO terms across the five luminal study groups.
+Resulting significant (FDR < 1 × 10−5) KO terms were divided into seven modules by hierarchical agglomerative clustering using inverse Pearson’s correlation as distance measure and Ward’s method (“ward.
+D2”) for linkage.
+Enrichment analysis was performed on each module and the three most significant KEGG pathways were included in the heatmap.
+d Uniform Manifold Approximation (UMAP) of the predicted KO terms
+
 ```{r Suppl. Table 2.}

### 2023-03-08 reports/manuscript/Figure3.Rmd
add figure legend

--- 2022-08-25 reports/manuscript/Figure3.Rmd
@@
   fig.width  = 6.6929133858,
-  fig.path="./Figures/",
   fig.align  = "center",
+  fig.path   ="./Figures/",
   fig.process = function(filename){
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r Load-data, message=FALSE, warning=FALSE}
 ##################
@@
 ```
+
+**Fig.
+3 Identification of a distinct ectocervical tissue-adherent microbiome.** Ectocervical tissue samples were assessed for presence of a tissue-adherent microbiome.
+a Bar plots of alpha diversity indices and taxonomy profiles for each individual tissue sample.
+Color-coded squares above the stacked bar plots show bacterial vaginosis (BV, binned Nugent’s scores) and HIV diagnosis, respectively.
+Gray: negative, orange: intermediate, red: positive BV; Gray: HIV seronegative, red: HIV seropositive.
+b Total relative abundance in the luminal and tissue microbiome datasets.
+All taxa with a total relative abundance < 0.55 are included in the “other” category.
+c Microbiome profile shift between luminal and tissue

### 2023-03-08 reports/manuscript/Figure4-5.Rmd
add figure legend

@@
   fig.width  = 6.6929133858,
-  fig.path="./Figures/",
   fig.align  = "center",
+  fig.path   = "./Figures/",
   fig.process = function(filename){
@@
 
-```{r Pairwise functions, message=FALSE, warning=FALSE}
+```{r Pairwise-functions, message=FALSE, warning=FALSE}
 ###########################
@@
 
-```{r Across tables, fig.height=4, fig.width=7, warning=F}
+```{r Across-tables, fig.height=4, fig.width=7, warning=F}
 TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
@@
 
-```{r Pairwise tables}
+```{r Pairwise-tables}
 ##################
@@
 
-```{r TF PPI Analysis function, include=FALSE}
+```{r TF-PPI-Analysis-function, include=FALSE}
 # filename <- DEG_files[1]
@@
 
-```{r TF PPI Analysis, eval=FALSE, include=FALSE}
+```{r TF-PPI-Analysis, eval=FALSE, include=FALSE}
 TF_PPI_analysis.fun(result_dir, DEG_files[1], gr_name[1], enrichR_files[1]) # writes Suppl.Tbl.07
@@
 
-```{r plot functions, warning=FALSE, include=FALSE}
+```{r plot-functions, warning=FALSE, include=FALSE}
 ############################
@@
 
+**Figure.
+4-5.
+Characterization of the host transcriptome as stratified by the luminal microbiome study groups.** The luminal samples were assessed for differential gene expression across the study groups.
+**a** Differential gene expression analysis was applied across the five luminal study groups.
+Significant DEGs (p-value < 0.01) were divided into six modules by hierarchical agglomerative clustering using inverse Pearson’s correlation as distance measure and Ward’s method (“ward.
+D2”) for linkage.
+Enrichment analysis was performed on each module using both the KEGG and GO databases.
+The three most significant terms were included in the heatmap.
+**b** Pairwise enrichment analysis of protein-protein interactions of transcription factors (TF-PPI).
+Top-10 up- and down regulated transcription factors with p-value < 0.01 were included in the bar plots
+
 ```{r Suppl. Tbl. 12}
@@
 ## igraph network plot
-```{r igraph network, eval=FALSE, warning=FALSE}
+```{r igraph-network, eval=FALSE, warning=FALSE}
 # 1. Change the groups of interest by assigning condition_use

### 2023-03-08 reports/manuscript/Figure6.Rmd
reorder cytokine plot for proof

--- 2022-09-30 reports/manuscript/Figure6.Rmd
@@
 
+prot_url <- "https://doi.org/10.1371/journal.ppat.1010494.s017"
+#cyt_url <- ""
+#meta_url <- ""
+
+############
+# DOWNLOAD #
+############
+if(isFALSE(file.exists("../../data/Protein_norm_MFI.xlsx"))){download.file(prot_url, paste0("../../data/", "Protein_norm_MFI.xlsx"), method="auto")}
+#if(isFALSE(file.exists("../../data/Cytokine_data.xlsx"))){download.file(cyt_url, paste0("../../data/", "Cytokine_data.xlsx"), method="auto")}
+#if(isFALSE(file.exists("../../data/metadata.xlsx"))){download.file(meta_url, paste0("../../data/", "metadata.xlsx"), method="auto")}
+
 #############
@@
 #############
-datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
-Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI_Bradley_et.al.xlsx")
-cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
+# datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
+Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI.xlsx", skip = 2) %>% rename(ID="Patient/Sample ID")
+# cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
 metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)
 
+cytokine_data <- read_xlsx(paste0("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/data/", "Cytokine_data.xlsx"), na = "NA", sheet = "Cytokine_data") 
 #################
@@
 
-```{r DPE Functions, message=FALSE, warning=FALSE, include=FALSE}
+```{r DPE-Functions, message=FALSE, warning=FALSE, include=FALSE}
 ################################
@@
 
-```{r format data, echo=FALSE, warning=FALSE, message=FALSE}
+```{r format-data, echo=FALSE, warning=FALSE, message=FALSE}
 to_remove <- c("CSTB_HPA017380","SPINK5_HPA009067","SPRR3_HPA024330","FGA_HPA064755","S100A12_HPA002881","KRT1_HPA019797","KRT1_HPA062908","LGMN_HPA001426","TACSTD2_HPA043104","FLNA_HPA001115","GPX3_HPA071520","CSTA_HPA001031","SERPINB5_HPA019132","ITIH2_HPA059150","NCF2_HPA002327","ANXA3_HPA013431")
@@
 
-```{r Across and Pairwise tables, include=FALSE}
+```{r Across-and-Pairwise-tables, include=FALSE}
 # Samples to use
-TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
-sample_use <- intersect(metadata$ID, colnames(TRX)[colSums(TRX)!=0])
+# TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
+# sample_use <- intersect(metadata$ID, colnames(TRX)[colSums(TRX)!=0])
+sample_use <- metadata$ID
 sample_use <- intersect(colnames(Protein), sample_use)
@@
 
-```{r heatmap functions, echo=FALSE}
+```{r heatmap-functions, echo=FALSE}
 top_dpe.fun <- function(top){
@@
                     7,7,8,8,
+                    7,7,9,9,
                     3,3,4,4),
-           nrow = 3, ncol = 4,byrow = T),
-           widths = c(1,2,.8,2), heights = c(1,1,.01) ) }
+           nrow = 4, ncol = 4,byrow = T),
+           widths = c(1,2,.8,2), heights = c(1,.35,.35,.01) )
+    # layout.show(9)
+    }
   if(layout == "single") {
@@
     plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
-          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .7 )
+          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .8 )
     #title(xlab="height", line=3)
@@
     }else{
-      add_scale_legend(x = par("usr")[1]-0.1, height = 1.5,#diff(par("usr")[3:4])-20,
-                     pal = colorRampPalette(c("navy","grey95","firebrick"))(91),
-                     labels = c("-5","0","5"))
-      height <- round(max(h$height))
-      add_letter("a")}
+      # add_scale_legend(x = par("usr")[1]-0.1, height = 1.5,#diff(par("usr")[3:4])-20,
+      #                pal = colorRampPalette(c("navy","grey95","firebrick"))(91),
+      #                labels = c("-5","0","5"))
+      # height <- round(max(h$height))
+      # add_letter("a")
+      }
     #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob
@@
     ## Heatmap
-    par(mar=c(4,.2,2.5,2.5)) #(bottom, left, top, right) 
+    par(mar=c(4,.2,2.5,3.2)) #(bottom, left, top, right) 
     image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
@@
             pch=15,xpd=T,cex=size) 
-    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
+    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.020,
             col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
@@
 
-    mtext( paste0(" ", gsub("_.+","",rownames(top_PROT_mfi)[rev(h$order)])), side=4, las = 1, cex = .45,
+    mtext( paste0(" ", gsub("_.+","",rownames(top_PROT_mfi)[rev(h$order)])), side=4, las = 1, cex = .53,
            at=seq(0,1,length.out = length(rownames(top_PROT_mfi))), xpd=T)
@@
 
-```{r sig. cytokines function, include=FALSE}
+```{r sig-cytokines-function, include=FALSE}
 sig_stars.fun <- function(df_krusk, df_dunn, group, n = 6, order = NULL){
@@
   # select only sig. cytokines 
-  sig <- df_krusk %>% filter(p.adj<0.05) %>% pull(Cytokines)
+  sig <- df_krusk %>% 
+    filter(p.adj<0.05) %>%
+    pull(Cytokines)
   sig_df <- cytokine_data_log[sig,samples_use$ID]
@@
 
-```{r Cytokine stats tests, include=FALSE}
-cytokine_data$ID <- metadata [ match( cytokine_data$PatID , as.character(metadata$PatID)  ) ,"ID" ]
-cytokine_data <- cytokine_data[ !is.na(cytokine_data$ID) , ]
-rownames(cytokine_data) <- cytokine_data$ID
-cytokine_data <- cytokine_data[,grepl("CVL",colnames(cytokine_data))]
-colnames(cytokine_data) <- sub("CVL_","",colnames(cytokine_data))
+```{r Cytokine-stats-tests, include=TRUE}
+cytokine_data <- cytokine_data %>% 
+  column_to_rownames(var = "ID") %>% 
+  as.matrix()
+
 cytokine_data <- t(cytokine_data)
@@
 
-```{r Suppl. Tbl. 15 16 & 17, include=FALSE}
+```{r Suppl. Tbl. 15 16 & 17, include=TRUE, results = 'hide'}
 names <- c("15 Luminal", "16 Tissue") %>% set_names()
@@
 
-```{r Figure 6, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
+```{r Figure 6, echo = TRUE, eval = TRUE, fig.height=10, fig.width=7}
+# dev.new(height=10, width=6.6929133858, noRStudioGD = TRUE)
+figlabels <- letters
+####################################
+# TOP DIFF. EXPRESSED PROT HEATMAP #
+####################################
+### A
+# Luminal groups
+t <- top_dpe.fun(dpe_L$top)
+plot_heatmap.fun(t, "Luminal_gr", scale = T, layout = "multi")
+
+### C
+# Tissue groups
+t <- top_dpe.fun(dpe_T$top)
+par(mfg = c(2,1)) # adds the remaning plots to the previous layout
+plot_heatmap.fun(t, "Tissue_gr", scale = T)
+
+##############################
+# SIG. CYTOKINES VIOLIN PLOT #
+##############################
+### B
+# Luminal groups
+g <- c("MIG", "IL1b", "IL12_p70", "MCP1", "IL1a", "IP10") 
+#g <- c("MIG", "IL12_p70","IL1a", "IP10", "IL1b", "MCP1" )
+lab <- c("MIG", "IL-1\u03b2", "IL-12-p70", "MCP-1", "IL-1\u03b1", "IP-10" )
+#lab <- c("IP-10", "IL-1\u03b2", "IL-12-p70", "MCP-1","IL-1\u03b1", "MIG")
+Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g[1:4])
+rownames(Luminal$sig_df) <- lab
+stars <- Luminal$stars_df
+
+par( mar=c(1.4,5,1.5,2)) #b,l,t,r 
+violist(data=Luminal$sig_df,
+          srt=45, y_padding = 2, smooth = .2,
+          genes= lab[1:4],
+          #cex.axis = 1.1,
+          #add_ylims = T, #gives starting value of y-axis and max value of all data points
+          col = pal, transparency = 50,
+          clustering=c(Luminal$groupings))
+segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
+text( stars$x , stars$y_+.05,
+        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
+title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = 1.1)
+
+add_letter("b")
+
+Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g[5:6])
+rownames(Luminal$sig_df) <- lab
+stars <- Luminal$stars_df 
+par(mar=c(1.4,5,1.5,2) ) #b,l,t,r
+violist(data=Luminal$sig_df,
+          srt=45, y_padding = 2, smooth = .2,
+          genes= lab[5:6],
+          #add_ylims = T, #gives starting value of y-axis and max value of all data points
+          col = pal, transparency = 50,
+          clustering=c(Luminal$groupings))
+segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_)
+text( stars$x , stars$y_+.05,
+        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
+title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = 1.1)
+
+### D
+# Tissue groups
+lab <- c("IL-1\u03b1", "IP-10")
+Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
+rownames(Tissue$sig_df) <- lab
+stars <- Tissue$stars_df
+
+#par(mar=c(20.4,6,1.5,3)) #b,l,t,r # ps = 20, point size
+par(mar=c(1,5,1.5,5)) #b,l,t,r
+violist(data=Tissue$sig_df,
+          srt=45, y_padding = 2, smooth = .2,
+          genes= rownames(Tissue$sig_df),
+          col = pal, transparency = 50,
+          # add_ylims = T,
+          clustering=Tissue$groupings )
+segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_)
+text( stars$x , stars$y_+.05,
+        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
+title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = 1.1)
+
+add_letter("d")
+
+```
+
+**Figure 6.
+Characterization of the host protein profile as stratified by the luminal and tissue-adherent microbiome study groups.** All study samples were assessed for significant differences in protein levels across the study groups.
+**a**,**b** The five luminal study groups and **c**,**d** the five corresponding tissue groups.
+a, c Proteins with a p-value > 0.05 were omitted from the heatmap.
+Proteins with significantly different levels across the groups were clustered by hierarchical agglomerative clustering using inverse Pearson’s correlation as distance measure and Ward’s method (“ward.
+D2”) for linkage.
+Color-coded rows below the heatmaps show clinical diagnosis of bacterial vaginosis (BV, binned Nugent’s scores) and HIV diagnosis, respectively.
+Color coding for BV: Gray: negative, orange: intermediate, red: positive; and for HIV: Gray: HIV seronegative, red: HIV seropositive.
+b, d Violin plots of log2 transformed cytokine levels across the luminal and tissue study groups (as indicated on the x-axis), respectively.
+The asterisk and lines indicate statistically significant results of Dunn’s test with Benjamini Hochberg’s correction analysis.
+Adjusted p-values: * < 0.05, ** < 0.01, *** < 0.001, **** < 0.0001
+
+```{r old-Figure 6, echo = FALSE, eval = FALSE, fig.height=10, fig.width=7}
+# old layout
+    # layout(matrix(c(1,2,5,6,
+    #                 7,7,8,8,
+    #                 3,3,4,4),
+    #        nrow = 3, ncol = 4,byrow = T),
+    #        widths = c(1,2,.8,2), heights = c(1,1,.01) )
 figlabels <- letters
@@
 g <- c("IL1a", "MIG", "IP10", "IL1b", "IL12_p70", "MCP1")
+lab <- c("IL-1\u03b1", "MIG", "IP-10", "IL-1\u03b2", "IL-12-p70", "MCP-1")
 Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g)
+rownames(Luminal$sig_df) <- lab
 stars <- Luminal$stars_df
@@
           srt=45, y_padding = 2, smooth = .2,
-          genes= g, #add_ylims = T, #gives starting value of y-axis and max value of all data points
+          genes= lab,
+          #add_ylims = T, #gives starting value of y-axis and max value of all data points
           col = pal, transparency = 50,
@@
 # Tissue groups
+lab <- c("IL-1\u03b1", "IP-10")
 Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
+rownames(Tissue$sig_df) <- lab
 stars <- Tissue$stars_df
 
-par( mar=c(20.4,6,1.5,3)) #b,l,t,r
+par(mar=c(20.4,6,1.5,3)) #b,l,t,r # ps = 20, point size
 violist(data=Tissue$sig_df,
@@
           genes= rownames(Tissue$sig_df), 
-          col = pal, transparency = 50, #add_ylims = T,
+          col = pal, transparency = 50, 
+          # add_ylims = T,
           clustering=Tissue$groupings )
@@
 
-**Figure 6.
-Protein Heatmap**.
-**a** and **c** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively.
-(p-value < 0.05) **b** and **d** sig.
-cytokines for Luminal and Tissue respectively.
-
-```{r All dpe plot, eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
+```{r All-dpe-plot, eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
 library(cowplot)

### 2023-03-08 reports/manuscript/Figure3.Rmd
show code in output document

@@
 
-```{r Figure 3, echo=FALSE, warning=FALSE, message=FALSE, fig.pos="t"}
+```{r Figure 3, warning=FALSE, message=FALSE, fig.pos="t"}
 # fig.asp=1, fig.height=7,

### 2023-03-08 reports/manuscript/Suppl.Figure1.Rmd
remove old version of figure

@@
 knitr::opts_chunk$set(
-  fig.path="./Suppl.Figures/",
+  fig.path    = "./Suppl.Figures/",
   fig.align   = "center",
@@
 #################
-Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
+Set1 <- c("#4DAF4A", "#E41A1C", "#984EA3", "#377EB8","#A6D854", "#FFFF33", "#FF7F00", "#A65628", "#F781BF", "#999999")
 pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
          RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
+
 ```
 
-```{r Suppl.Fig.1 Bacterial_abundance, fig.asp=.5}
-# test to get Paulos code to work with the new files who has less taxa due to filtering after normalization with only selected samples
-# it does not currently work
-# top_all_ <- lapply( datasets_all_samples_[1:2] , function(x){
-#    sort(apply(x,1,var),decreasing = T)[1:20]} )
-# top_ <- unique( sub( ".*[.]" , "" , names(unlist(top_all_)) ) )
-# 
-# x_ <- t(sapply(datasets_all_samples_[1:2],function(x){
-#   c(t(x))
-# }))
-# colnames(x_) <- c(sapply(rownames(datasets_all_samples_[[1]]),
-#                         function(i){rep(i,ncol(datasets_all_samples_[[1]]))}))
-# rownames(datasets_all_samples_[[1]]) == rownames(datasets_all_samples_[[2]])
+### Violin plot
+```{r Suppl.Fig.1, fig.asp=.5}
+x <- datasets_all_samples_[[1]]
+top <- t(t(2^x-1)/colSums(2^x-1)) %>%
+  as_tibble(., rownames ="tax") %>%
+  pivot_longer(cols = -tax, names_to = 'ID') %>%
+  group_by(tax) %>%
+  summarise(mean = mean(value)) %>% 
+  arrange(desc(mean)) %>%
+  filter(!(tax == "Not assigned")) %>%
+  pluck(., "tax") %>%
+  .[1:30]
 
-top_all <- lapply( datasets_all_samples[2:3] , function(x){
-   sort(apply(x,1,var),decreasing = T)[1:20]} )
-top <- unique( sub( ".*[.]" , "" , names(unlist(top_all)) ) )
-
-x <- t(sapply(datasets_all_samples[2:3],function(x){
-  c(t(x))
-}))
-colnames(x) <- c(sapply(rownames(datasets_all_samples[[2]]),
-                        function(i){rep(i,ncol(datasets_all_samples[[2]]))}))
-
-rownames(x) <- c("Tissue", "Luminal")
-
-top <- sort(sapply( unique(colnames(x)),
-               function(i) { mean( x[,colnames(x)==i ] ) } ),decreasing = T)
-top <- names(top)[1:30]
-
-###################################
-# BACTERIAL ABUNDANCE IN DATASETS #
-###################################
-par(mar=c(6.5,2.5,0,0)) #b,l,t,r
-violist(data = x[,colnames(x)%in%top],
-        genes = rownames(x),
-        clustering = factor( colnames(x)[colnames(x)%in%top],levels = top) ,
-        srt=45, cex.axis=0.7, pt.cex=0.2,  smooth = .1,
-        transparency=50,col = pal)
-```
-
-### Paulos code with tidyverse
-```{r New_files, fig.asp=.5}
 Lum <- datasets_all_samples_[[1]][rownames(datasets_all_samples_[[1]])%in%top,]
@@
 
-I belive that the reson for the diference seen in the tissue distribution is that some of the samples was removed later on.
+**Supplementary Figure 1.
+Abundance distribution of individual taxa in the luminal and tissue microbiome data sets.** Violin plots showing the distribution of relative abundance of the top 30 most abundant taxa in the luminal and tissue-adherent data sets.

### 2023-03-08 reports/manuscript/Suppl.Figure2.Rmd
add figure legend

@@
 
-```{r Suppl.Fig.2 Diff_bact_abundance, warning=FALSE, message=FALSE, fig.width=10, fig.height=9}
+### Dot plot and bar plot
+```{r Suppl.Fig.2, warning=FALSE, message=FALSE, fig.width=10, fig.height=9}
 all_microbiome <- cbind(datasets_all_samples[["ASV_Tissue_normalized"]],
@@
 **Suppl.
-Figure 2.** Differential bacterial abundance across tissue and CVL microbiome week3 datasets.
-The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` (from the wilcoxon test) were considered significant and were sorted by the highest expression.
-The colour scale indicates the difference in total abundance between the datsets as a proportion, where Max is the higest abundace of the two datasets and the other becomes a proportion of this value.The size of the dots indicates the average abundance of the given bacteria in the given dataset.
+Figure 2.
+Differential bacterial abundance across the luminal and tissue microbiome datasets.** Differential bacterial abundance was compared between the luminal and tissue-adherent microbiome data sets.
+The results are shown as **a)** dot plots, and **b)** bar plots, respectively.
+Bacteria with log2FC above `r fct` and p-value `r pvt` (from the Wilcoxon’s test) were considered significantly different and were sorted by the highest expression.
+The colour scale indicates the difference in total abundance between the datasets as a proportion, where “max” is the highest abundance of the two datasets, and the other becomes a proportion of this value.
+The size of the dots indicates the average abundance of the given bacteria in the given data set.

### 2023-03-08 reports/manuscript/Suppl.Figure6.Rmd
rename to suppl fig 6

--- 2022-10-11 reports/manuscript/Suppl.Figure6.Rmd
@@
 ---
-title: "Suppl. Figure 5. Rarefaction curves"
+title: "Suppl. Figure 6. Rarefaction curves"
 geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
@@
     new_filename <- stringr::str_remove(string = filename, 
-                                        pattern = "-1")
+                                        pattern = "-.")
     fs::file_move(path = filename, new_path = new_filename)
@@
 #############
+data_folder <- "/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/"
 paths_p <- c('../../data/phyloseq_boston_r1.RDS',   # CVL V3 (2 replicates of one tissue + 27 CVL V2)
@@
 n <- c("boston_r1", "boston_r2", "Sthlm")
-ps_list <- map(paths_p, ~readRDS(.x)) %>% set_names(n)
+ps_list <- map(paths_p, ~readRDS(paste0(data_folder, .x))) %>% set_names(n)
 
@@
 
-```{r rarefaction curves, fig.height=6, fig.width=8, fig.asp=0.75}
+```{r rarefaction curves, results = 'hide', fig.asp=0.75, fig.height=6, fig.width=8}
 ##################
@@
 limit <- c(65000, 65000, 65000, 65000, 65000, 10000)
-title <- c("Run B1. CVLv3 + CVLv2 rep.", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B1. CVLv3", "Run B2. CVLv2", "Run B2. TISSUEv3")
+title <- c("Run B1. CVLv3 + CVLv2 rep.", "Run B2. CVLv2 + TISSUEv3",
+           "Run Sthlm CVLv2", "Run B1. CVLv3", "Run B2. CVLv2", "Run B2. TISSUEv3")
 title <- c("Run CVL", "Run B2. CVLv2 + TISSUEv3","Run Sthlm CVLv2", "Run B2. CVLv2", "Run Tissue")
@@
 
-```{r Suppl.Fig5 Rarefraction_plot, warning=FALSE, fig.height=6, fig.width=8, fig.asp=0.75,}
+### Rarefraction plot
+```{r Suppl.Fig.6, warning=FALSE, fig.height=6, fig.width=8, fig.asp=0.75,}
 ### A & B
@@
 
-ggsave(filename = paste0('Rarefaction_plot.pdf'), plot = plot,
-       path = paste0("../../../results/Plots"),
-       width = 8, height = 6
-)
+# ggsave(filename = paste0('Suppl.Figures6.pdf'), plot = plot,
+#        path = paste0("./Suppl.Figures/"),
+#        width = 8, height = 6
+# )
 ```
@@
 **Supp.
-Figure 1.
+Figure 6.
 Rarefraction curves**.
@@
 **a)** The Luminal sequencing run and **b)** the Tissue sequencing run.
-The sequencing depth was > `r as.character(count[4])` reads in all but `r filter(read_counts$CVLv3, Counts == "< 40 000") %>% nrow()` samples for the Luminal dataset while `r filter(read_counts$tissue, Counts == "< 2 500") %>% nrow()` samples had fewer than `r as.character(count[6])` reads in the tissue dataset.
+The sequencing depth was > `r as.character(count[4])` reads in all but `r filter(read_counts$CVLv3, Counts == "< 40 000") %>% nrow()` samples for the Luminal dataset while `r filter(read_counts$tissue, Counts == "< 2 500") %>% nrow()` samples had fewer than `r as.character(count[6])` reads in the tissue-adherent microbiome dataset.

### 2023-03-08 code/knit_function.R
add gfm function

@@ -23,6 +23,49 @@ custom_knit <- function(inputFile, out_dir, ...) {
     #   )
 }
 
+##################
+# KNIT RMARKDOWN #
+#################
+#### html ####
+knit_gfm_with_job <- function() {
+  rstudioapi::verifyAvailable()
+  
+  job_file <- tempfile(fileext = ".R")
+  
+  active_doc_ctx <- rstudioapi::getSourceEditorContext()
+  rmd_path <- active_doc_ctx$path
+  
+  if (identical(rmd_path, "")) {
+    rstudioapi::showDialog(
+      "Cannot Render Unsaved R Markdown Document",
+      "Please save the current document before rendering."
+    )
+    return(invisible())
+  }
+  
+  rstudioapi::documentSave(active_doc_ctx$id)
+  
+  rmd_path <- normalizePath(rmd_path, mustWork = TRUE)
+  out_file <- paste0(basename(xfun::sans_ext(rmd_path)), '.md')
+  
+  cat(
+    'res <- rmarkdown::render("', rmd_path, '",',
+    '"', out_file, '",',
+    ' output_format = rmarkdown::github_document(html_preview = FALSE, dev = "jpeg")', ')\n',
+    'unlink("', job_file, '")\n',
+    'rstudioapi::viewer(res)',
+    sep = "",
+    file = job_file
+  )
+  
+  rstudioapi::jobRunScript(
+    path = job_file,
+    name = basename(rmd_path),
+    workingDir = dirname(rmd_path),#out_dir,
+    importEnv = FALSE
+  )
+} 
+
 #params = list(print=TRUE, setup=TRUE),
 
 # https://stackoverflow.com/questions/66620582/customize-the-knit-button-with-source

### 2023-03-08 reports/manuscript/Figure4-5.Rmd

@@
 ## Heatmap DGEs and enrichment
-```{r Across functions, message=FALSE, warning=FALSE}
+```{r Across-functions, message=FALSE, warning=FALSE}
 ##################

### 2023-03-08 reports/manuscript/Suppl.Figure4-5.Rmd

@@
 ---
-title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
+title: "Suppl. Figure 4. & 5. Mirobiome-KEGG correlations."
 geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
@@
 knitr::opts_chunk$set(
-  fig.path="./Suppl.Figures/",
+  fig.path    = "./Suppl.Figures/",
   fig.align   = "center",
   fig.process = function(filename){
-    new_filename <- stringr::str_remove(string = filename, 
-                                        pattern = "-1")
+    new_filename <- stringr::str_replace_all(string = filename, 
+                                        c("4-5-1"="4", "4-5-2"="5"))
     fs::file_move(path = filename, new_path = new_filename)
@@
 
-```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, fig.width=12, fig.height=13}
+### Correlation heatmap
+```{r Suppl.Fig.4-5, warning=FALSE, fig.width=12, fig.height=13}
 myletter <- letters
@@
 
-  ### Suppl. Figure 3-4
+  ### Suppl. Figure 4-5
   #######################################
@@
 **Suppl.
-Figure 3-4.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset.
-Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes.
-Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database.
-This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue.
+Figure 4-5.
+Functional associations of the luminal microbiome with host tissue gene expression profiles.** Bacterial abundances in the luminal samples were correlated with gene expression of the top 5,000 highly variable genes from the RNAseq dataset.
+This generated a correlation matrix between bacteria and genes.
+For each bacteria, genes were ranked based on their correlation to that bacteria, followed by gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database.
+The resulting matrix display associations between individual bacterial taxa and corresponding KEGG term as defined in the host tissue sample.
 The heatmap shows the normalized enrichment score (NES).
-Only enrichments with pvalue below 0.05 are shown.Only bacterium and pathways with at least 10 significant NES scores were included in the heatmap.
+Only enrichments with p-value < 0.05 are shown.
+Bacterium and pathways with less than 10 significant NES scores were omitted from the heatmap.
+Bacteria are grouped according to anatomical/functional activity and marked with different colours per category.
 
